### 2017-10-12 paper/sections/1_introduction.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Introduction}

% What is the problem?
File system metadata services in HPC have scalability problems. It has been
shown that HPC workloads are metadata resource intensive because administrative
tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
system~\cite{zheng:pdsw2014-batchfs}, on large data sets lead to contention
for the same directories and inodes ({\it
e.g.}, path traversal). Applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every client is unreasonable. This problem is exacerbated
by current trends in HPC, where architectures are transitioning from complex
storage stacks with burst buffer, file system, object store, and tape tiers to
more simplified stacks with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
because more requests end up hitting the same layer and latencies cannot be 
hidden while data migrates across tiers.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
\caption{With CudeleFS, users can assign weaker consistency and durability
policies to subtrees to get the same performance benefits of state-of-the-art
HPC architectures, such as those shown in the embedded table, while
applications that require the stronger guarantees of POSIX can still reside in
the same namespace.
}\label{fig:subtree-policies}
\end{figure}

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, batch style jobs often do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
patterns).

% One example
One popular approach for relaxing consistency and durability is to ``decouple
the namespace", where clients lock the subtree they want exclusive access to as
a way to tell the file system that the subtree is important or may cause
resource contention in the near-future~\cite{grider:pdsw2015-marfs,
zheng:pdsw2015-deltafs, zheng:pdsw2014-batchfs, ren:sc2014-indexfs,
bent:slides-twotiers}. Then the file system can change its internal structure
to optimize performance. For example, the file system could enter a mode that
prevents other clients from interfering with the decoupled directory.  This
delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
durability improves performance and scalability by avoiding the costs of remote procedure calls (RPCs),
synchronization, false sharing, and serialization.  While the performance
benefits of decoupling the namespace are obvious, applications that rely on the
file system's guarantees must be deployed on an entirely different system or
re-written to coordinate strong consistency/durability themselves.

%\begin{table}
%\begin{tabular}{ r | l }
%  Subtree         & Example \\\hline
%  (1)   & \{Index, Batch\}FS~\cite{ren:sc2014-indexfs, zheng:pdsw2014-batchfs} \\
%  (2)   & \{Index, Ceph\}FS~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} \\
%  (3)   & RAMDisk \\
%  (4)   & DeltaFS~\cite{zheng:pdsw2015-deltafs} \\
%\end{tabular}
%
%\caption{State-of-the-art systems in HPC improve file system metadata
%performance by relaxing consistency and durability guarantees. Note that
%IndexFS also supports weak consistency with bulk inserts.
%\label{table:namespaces}} \end{table}

% What did we do
To address this problem, we propose a subtree policy API that lets developers dynamically control the
consistency and durability guarantees for subtrees in the file system
namespace. For example, one subtree can adopt weaker consistency semantics
while another subtree can retain the rigidity of POSIX's strong consistency.
The subtrees in Figure~\ref{fig:subtree-policies} show a setup in our proposed system where a single global
namespace has subtrees for applications optimized with techniques from different
state-of-the-art HPC architectures.

In this work, we present CudeleFS, a prototype programmable file system that
supports different degrees of consistency and durability in a global namespace.
CudeleFS achieves this by exposing ``mechanisms" that developers use to specify
their preferred semantics.  CudeleFS supports 3 forms of consistency
(invisible, weak, and strong) and 3 degrees of durability (none, local, and
global) giving the user a wide range of policies and optimizations that can be
custom fit to an application. We make the following contributions:

\begin{enumerate}

  \item a prototype that lets developers choose from a range of
  consistency and durability semantics (9 permutations), allowing them to dynamically custom
  fit the storage system to the application.

  \item an API for selecting consistency/durability policies and assigning
  them to subtrees in the file system namespace.

  \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
  previously unexplored metadata designs, all implemented using CudeleFS.

\end{enumerate}

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.
In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the CudeleFS
prototype and API. Section~\ref{sec:implementation} describes the CudeleFS
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places CudeleFS in the context of other related
work.

### 2017-10-12 paper/sections/2_related_work.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Related Work} 
\label{sec:related-work}

% General
The bottlenecks associated with accessing POSIX file system metadata are not limited
to HPC workloads and the same challenges that plagued these systems for years are
finding their way into the cloud. Workloads that deal with many small files
({\it e.g.}, log processing and database
queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
simultaneous clients ({\it e.g.}, MapReduce
jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
the metadata service. The biggest challenge is that whenever a file
is touched the client must access the file's metadata and maintaining a file
system namespace imposes small, frequent accesses on the underlying storage
system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
system metadata is a well-known problem and solutions for scaling data IO do
not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
abad:techreport2012-fstrace, abad:ucc2012-mimesis,
alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
for improving the performance of metadata access.

\subsection{Metadata Load Balancing}

% approaches to load balancing (strong = more resuorces per unit work, weak =
% fixed resource per work unit)
One approach for improving metadata performance and scalability is to alleviate
overloaded servers by load balancing metadata IO across a cluster. Common
techniques include partitioning metadata when there are many writes and
replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
directories and clients write to different partitions by grabbing leases and
caching ancestor metadata for path traversal; it does well for strong scaling
because servers can keep more inodes in the cache which results in less RPCs.
Alternatively, ShardFS replicates directory state so servers do not need to
contact peers for path traversal; it does well for read workloads because all
file operations only require 1 RPC and for weak scaling because requests will
never incur extra RPCs due to a full cache.  CephFS employs both techniques to
a lesser extent; directories can be replicated or sharded but the caching and
replication policies do not change depending on the balancing technique.
Despite the performance benefits these techniques add complexity and jeopardize
the robustness and performance characteristics of the metadata service because
the systems now need (1) policies to guide the migration decisions and (2)
mechanisms to address inconsistent states across servers.

% policies
Setting policies for migrations is arguably more difficult than adding the
migration mechanisms themselves.  For example, IndexFS and CephFS use the
GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
threshold and using lazy synchronization to redirect queries to the server that
``owns" the targeted metadata.  Determining when to partition directories and
when to migrate the directory fragments are policies that vary between systems:
GIGA+ partitions directories when the size reaches a certain number of files
and migrates directory fragments immediately; CephFS partitions directories
when they reach a threshold size or when the write temperature reaches a
certain value and migrates directory fragments when the hosting server has more
load than the other servers in the metadata cluster. Another policy is when and
how to replicate directory state; ShardFS replicates immediately and
pessimistically while CephFS replicates only when the read temperature reaches
a threshold.  There is a wide range of policies and it is difficult to maneuver
tunables and hard-coded design decisions.

% addressing inconsistency
In addition to the policies, distributing metadata across a cluster requires
distributed transactions and cache coherence protocols to ensure strong
consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
directory state and uses optimistic concurrency control for conflicts; namely
it does the operation and if there is a conflict at verification time it falls
back to two-phase locking.  Another example is IndexFS's inode cache which
reduces RPCs by caching ancestor paths -- the locality of this cache can be
thrashed by random reads but performs well for metadata writes. For
consistency, writes to directories in IndexFS block until the lease expires
while writes to directories in ShardFS are slow for everyone as it either
requires serialization or locking with many servers; reads in IndexFS are
subject to cache locality while reads in ShardFS always resolve to 1 RPC.
Another example of the overheads of addressing inconsistency is how CephFS
maintains client sessions and inode caches for capabilities (which in turn make
metadata access faster). When metadata is exchanged between metadata servers
these sessions/caches must be flushed and new statistics exchanged with a
scatter-gather process; this halts updates on the directories and blocks until
the authoritative metadata server responds.  These protocols are discussed in
more detail in the next section but their inclusion here is a testament to the
complexity of migrating metadata.

%For metadata writes it does a distributed transaction; monotonic writes with
%concurrent clients fail and do pessimistic locking through a mediated lock
%server to ensure strong consistency, non-monotonic writes grab locks at every
%server. Zooming in on monotonic writes: if permissions increase it executes on
%a primary then non-primary, if permissions decrease it executes on all
%non-primary then on primary.

The conclusion we have drawn from this related work is that metadata protocols
have a bigger impact on performance and scalability than load balancing.  
Understanding these protocols helps load balancing and gives us a better
understanding of the metrics we should use to make migration decisions ({\it
e.g.}, which operations reflect the state of the system), what types of
requests cause the most load, and how an overloaded system reacts ({\it e.g.},
increasing latencies, lower throughput, etc.).

\subsection{Relaxing POSIX}

% POSIX 
POSIX workloads require strong consistency and many file systems improve
performance by reducing the number of remote calls per operation ({\it i.e.}
RPC amplification). As discussed in the previous section, caching with leases and
replication are popular approaches to reducing the overheads of path traversals
but their performance is subject to cache locality and the amount of available
resources, respectively; for random workloads larger than the cache extra RPCs
hurt performance~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} and for write heavy workloads with more
resources the RPCs for invalidations are harmful. Another approach to reducing
RPCs is to use leases or capabilities.

%IndexFS aggressively caches paths and
%handles permissions by handing out leases for metadata writes; metadata may
%only be modified when all leases have expired. 
%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
%permissions on the client servers. Modifications to metadata cached by clients
%is delayed until all client leases have expired. This reduces the RPC
%amplification to 1 when mutatating directory metadata ({\it e.g.,}
%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying metadata servers for
%path traversals. The disadvantage of this approach is the high latency of
%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
%replicates metadata (specifically directory lookup state) across the metadata server
%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
%directory lookup state are done with optimistic concurrecny control and fall
%back to retry if verification fails. The disadvantage of this appraoch is the
%high number of RPCs for maintaining directory metadata mutations (writes to
%directories).  CephFS also maintains an inode cache but its notion of leases
%are much shorter, on the order of microseconds.

% Non POSIX
High performance computing has unique requirements for file systems ({\it
e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
make relaxing POSIX sensible.  BatchFS assumes the application coordinates
accesses to the namespace, so the clients can batch local operations and merge
with a global namespace image lazily. Similarly, DeltaFS eliminates RPC
traffic using subtree snapshots for non-conflicting workloads and middleware
for conflicting workloads. MarFS gives users the ability to lock
``project directories" and allocate GPFS clusters for demanding metadata
workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
tier; applications lock the namespace so that metadata can be accessed more quickly.
Unfortunately, decoupling the namespace has costs: (1) merging metadata state
back into the global namespace is slow; (2) failures are local to the failing
node; and (3) the systems are not backwards compatible.

% NON POSIX
%Decoupling the namespaces has many advantages, including improved scalability,
%higher resource utilization, and better performance.  BatchFS and DeltaFS are
%near-POSIX filesystems that give clients the ability to decouple subtrees from
%the namespace so that the applications can execute metadata operations without
%synchronization and serialization.  These operations are applied to a local
%snapshot of the file system namespace and conflicts are resolved either by the
%application or by an external service . Applications link into a metadata
%server library to reduce resource utilization and code paths (e.g., no daemons
%and less interprocess communication).

For (1), state-of-the-art systems manage consistency in non-traditional ways:
IndexFS maintains the global namespace but blocks operations from other clients
until the first client drops the lease, BatchFS does operations on a snapshot
of the namespace and merges batches of operations into the global namespace,
and DeltaFS never merges back into the global namespace. The merging for
BatchFS is done by an auxiliary metadata server running on the client and
conflicts are resolved by the application. Although DeltaFS never explicitly
merges, applications needing some degree of ground truth can either manage
consistency themselves on a read or add a bolt-on service to manage the
consistency.

For (2), if the client fails and stays down, all metadata operations on the
decoupled namespace are lost. If the client recovers, the on-disk structures
(for BatchFS and DeltaFS this is the SSTables used in TableFS) can be
recovered. In other words, the clients have state that cannot be recovered if
the node stays failed and any progress will be lost. This scenario is a
disaster for checkpoint-restart where missed cycles may cause the checkpoint to
bleed over into computation time.

For (3), decoupled namespace approaches sacrifice POSIX going as far as
requiring the application to link against the systems they want to talk to. In
today's world of software defined caching, this can be a problem for large data
centers with many types and tiers of storage. Despite well-known performance
problems POSIX and REST are the dominant APIs for data transfer.

### 2017-10-12 paper/sections/5_implementation.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Implementation}
\label{sec:implementation}

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
CudeleFS; namely, we try to leverage components inside Ceph to inherit the
robustness and correctness of the internal subsystems. Using this `dirty-slate'
approach, we only had to implement 4 of the 6 mechanisms from
Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
system itself.  In this section, we first describe a Ceph internal subsystem or
component and then we show how we use it in CudeleFS.

%- no change (rpcs, stream)s, library(create, nonvolatile apply, local/global
%persist),  change to metadata server (nonvolatile apply)

% how it works in CephFS, how I used it
\subsection{Metadata Store}
\label{sec:metadata-store}

% how it works in CephFS
In CephFS, the metadata store is a data structure that represents the file
system namespace. This data structure is stored in two places: in memory ({\it
i.e.} in the collective memory of the metadata server cluster) and as objects
in the object store. In the object store, directories and their inodes are
stored together in objects to improve the performance of scans.  The metadata
store data structure is structured as a tree of directory fragments making it
easier to read and traverse.\\

% how i used it
\noindent\textbf{CudeleFS}: the ``RPCs" mechanism uses the metadata store to
service requests. Using code designed for recovery, ``Nonvolatile Apply" and
``Volatile Apply" replay updates onto the metadata store in memory and in the
object store, respectively.  When the clients are ready to merge their updates
back into the global namespace, they pass a binary file of  metadata updates to
the metadata server.

%\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply

\subsection{Journal}
\label{sec:journal}

The journal is the second way that CephFS represents the file system namespace;
it is a log of metadata updates that can materialize the namespace when the
updates are replayed onto the metadata store. The journal is a ``pile system";
writes are fast but reads are slow because state must be reconstructed.
Specifically, reads are slow because there is more state to read, it is
unorganized, and many of the updates may be redundant.\\

\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Append Client Journal",
``Local Persist", and ``Global Persist".  ``Stream" is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemented by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, CudeleFS leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

\subsection{Journal Tool}
\label{sec:journal-tool}

The journal tool is used for disaster recovery and lets administrators view and
modify the journal. It can read the journal, export the journal as a file,
erase events, and apply updates to the metadata store.  To apply journal
updates to the metadata store, the journal tool reads the journal from object
store objects and replays the updates on the metadata store in the object
store.\\

\noindent\textbf{CudeleFS}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
``Append Client Journal", ``Volatile Apply", and ``Nonvolatile Apply".

% Journal import
%The journal tool imports journals from binary files stored on disk.  First the
%header of the dump is sanity checked and written into RADOS to the ``header"
%object.  The ``header" object has metadata about the journal as well as the
%locations of all the journal pointers (e.g., where the tail of the journal is,
%where we are currently trimming, etc.).  Then the journal events are cleaned
%(erasing trailing events that are not part of the header) and written as
%objects into RADOS.  Note that while the journal is in RADOS, the metadata
%servers do do not have the namespace reconstructed in memory so the metadata
%cluster will not service requests relating to the journal of imported events.
%To construct the namespace in the collective memory of the metadata servers we
%need to first construct the namespace in RADOS. The journal tool can explicitly
%do this by  applying the journal to the metadata store in RAODS. This will pull
%the objects containing journal segments and replay them on the metadata store.
%Finally, we delete the journal in RADOS and restart the metadata servers so
%they rebuild their caches.

% Journal export
%First the
%journal is scanned for the header and then journal is recovered. To recover the
%journal the ``header" object is read off disk and then objects are probed in
%order and starting from the write position saved in the header. Probing will
%update the write position if it finds objects with data in them.

% Journal export
%When exporting a journal of events, the journal tool first scans the journal to
%check for corruption. Then it recovers the journal by reading the ``header"
%object out of RADOS.  After reading the header, the journal tool can pull
%journal segments from RADOS because it knows how many objects to pull and how
%far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

\subsection{Inode Cache}
\label{sec:inode-cache}

In CephFS, the inode cache reduces the number of RPCs between clients and
metadata servers. Without contention clients can resolve metadata reads locally
thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For example, if a client or
metadata server is not caching the directory inode, all creates within that
directory will result in a lookup and a create request. If the directory inode
is cached then only the create needs to be sent.  The size of the inode cache
is configurable so as not to saturate the memory on the metadata server --
inodes in CephFS are about 1400
bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
The inode cache has code for manipulating inode numbers, such as pre-allocating
inodes to clients.\\

\noindent\textbf{CudeleFS}: ``Nonvolatile Apply" uses the internal inode cache
code to allocate inodes to clients that decouple parts of the namespace and to
skip inodes used by the client at merge time.\\

\subsection{Large Inodes}
\label{sec:large-inodes}

In CephFS, inodes already store policies, like how the file is striped
across the object store or for managing subtrees for load balancing. These
policies adhere to logical partitionings of metadata or data, like Ceph pools and file
system namespace subtrees. To implement this, the namespace data structure has
the ability to recursively apply policies to subtrees and to isolate subtrees
from each other.  \\

\noindent\textbf{CudeleFS}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different CudeleFS mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

% General information about the journal
%The journal segments are saved as objects in RADOS.  The journal has 4
%pointers, described in `osdc/Journaler.h`:
%
%\begin{itemize}
%  \item write position: tail of the journal; points to the current session where we are appending events
%  \item unused field: where someone is reading
%  \item expire position: old journal segments
%  \item trimmed position: where daemon is expiring old items
%\end{itemize}
%
%% How the journal tool works
%Journal segments in RADDS have a header followed by serialized log events. The
%log events are read by hopping over objects using the read offset and object
%size pulled from the journal header.  After decoding them, we can examine the
%metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
%that the event touches.
%
%% The metablobs
%The metadata for inodes that the event touches are called metadata blobs and
%the ones associated with events are \textbf{unordered}; this layout makes
%writing journal updates fast but the cost is realized when reading the metadata
%blobs.  It makes sense to optimize for writing since reading only occurs on
%failures. To reconstruct the namespace for the metadata blob, the journal tool
%iterates over each metadata blob in the events and builds mappings of inodes to
%directory entries (for directories) and parent inodes to child inodes/directory
%entries.

%Reads/Writes Journal Events}

% How it works
% Why we re-use stuff
%Step 3 is the most complicated and requires understanding how the snapshot is
%materialized in memory. 
%
%\subsection{Operating on Snapshots} 
%
%Our first implementation attempted to re-create journal events using the same
%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
%to instantiate a Ceph inode and directory entry for the current file/dir and
%its parent.  This is too hard because there are too many moving parts in the
%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
%traverse up and down namespace, etc.). So when I tried add dentries and inodes
%it was trying to traverse up/down and it would almost always segfault when it
%was looking for something. These metablobs are supposed to be self container --
%the problem is I do not know what is supposed to go *inside* them. 
%
%Our second idea was to copy the metadata blog and change just what we needed.
%For example, we would save a binary dump of a generic \texttt{mkdir} event on
%disk. When the application makes a directory, this dump would be loaded and the
%fields would be changed before being written back to disk. Rather than
%traversing up and down a namespace in memory of a metadata server, we should
%traverse up and down the namespace *inside* the metadata blob. This
%implementation requires disk IO and editing the log event is non-trivial for
%two reasons:
%
%\begin{itemize}
%
%  \item methods do not edit events; they just write them
%
%  \item the metadata that the event touches (e.g., the metablob) is unorganized
%  on disk for performance -- it is trade-off for writing data faster serially and
%  reconstructing information slowly since failure is not considered the norm
%
%\end{itemize}
%
%Faced with these challenges we landed on our final implementation: load the
%snapshot into the the data structures used to examine and replay journals, edit
%those data structures, and write them out to disk as binary.

% - how it actually works
% - bugs I fixed in the last couple of commits

%The metadata objects are located with naming schemes (200.000* for journal
%objects and 1.inode for metadata storage objects).

% How it works: socket for changing daemon's internal state (debugging, logging, behaviour)
% 1. API for putting state into the daemon dynamically
% 2. Hooks directly into daemon code so we can use any parsing functionality in there
% 3. Documentation all the tunables

% 1. read journal of updates from file
% 2. call replay (uses same code as when an metadata server comes back) on each event
% 3. skip inodes so metadata server doesn't hand out those new inodes

### 2017-10-12 paper/sections/6_results.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Evaluation}
\label{sec:evaluation}

We evaluate CudeleFS on a 15 node cluster, partitioned into 8 object storage
servers, 3 metadata servers, and 2 monitor servers. The object storage servers
double as clients which is fine because clients are CPU and memory bound while
object storage servers are disk IO bound. All daemons run as a single process
which is the default setting for Ceph and the nodes have 2 dual core 2GHz
processors with 8GB of RAM. There are three daemons per object storage server
(one for each disk formatted with XFS) and they share an SSD for the journal.
The nodes are running Ubuntu 12.04.4, kernel version 3.2.0-63 but all
experiments run in Docker containers; this makes it easier to tear down and
re-initialize the cluster ({\it e.g.}, dropping the kernel cache) between
experiments.

This paper adheres to The Popper
Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}, so
experiments presented here are available in the repository for this
article\footnote{Links have been removed for double-blind submission}.
Experiments can be examined in more detail, or even re-run, by visiting the
\texttt{[source]} link next to each figure. That link points to a Jupyter
notebook that shows the analysis and source code for that graph, which points
to an experiment and its artifacts.

%\begin{figure*}[tb]
%\centering
%\includegraphics[width=180mm]{graphs/slowdown.png}
%\caption{Compared to the \texttt{create} phase, saving and persisting
%updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
%a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
%In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
%slower.  The disadvantage of decoupling the namespace is the merge phase where
%updates are applied to the metadata store (\texttt{create+apply}, resulting in a
%905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
%\end{figure*}

\subsection{CudeleFS Mechanism Performance}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{
[\href{https://...}{source}]
The performance of the CudeleFS mechanisms normalized to the runtime of the
``Append Client Journal" mechanism (the runtime of writing \(n\) file creates
to the client's in-memory journal).  \label{fig:slowdown-mechanisms}}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
\caption{
[\href{https://...}{source}]
The size of the client's journal scales with the number of
updates.\label{fig:behavior-journal-size}}
\end{figure}

Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
mechanisms, normalized to the time it takes to write \(n\) file create updates
to the client's in-memory journal ({\it i.e.} the ``Append Client Journal"
mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal"
mechanism.  ``Stream" is an approximation of the
overhead and is calculated by subtracting the runtime of the job with the
journal turned off from the runtime with the journal turned on.  The largest
workload we tested is 100K file creates in the same directory.  This is the
maximum recommended size of a directory in CephFS; preliminary experiments with
larger directory sizes show memory problems.

\subsubsection{Poorly Scaling Data Structures} Despite doing the same amount of
work, mechanisms that rely on poorly scaling data structures have large
slowdowns for the larger number of creates. For example, ``RPCs" which relies
on the internal CephFS directory structures, goes from a \(40\times\) slowdown
for 10K files to a \(90\times\) slowdown for 100K files. It is a well-known
problem that directory data structures do not scale when creating files in the
same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
structures will experience similar slowdowns. Other mechanisms that write
events to a journal ({\it e.g.} the persists, ``Volatile Apply") experience a
much less drastic slowdown because the journal data structure does not need to
be scanned for every operation. Events are written to the end of the journal without
even checking the validity ({\it e.g.}, if the file already exists for a create),
which is another form of relaxed consistency because the file system assumes the
application has resolved conflicting updates in a different way.

% RPCs vs. apply: calls to metadata server vs. RADOS
\subsubsection{Overhead of RPCs} ``RPCs" is \(66\times\) slower than ``Volatile
Apply" because sending individual metadata updates over the network is costly.
While ``RPCs" sends a request for every file create, ``Nonvolatile Apply"
writes all the updates to the in-memory journal and applies them to the
in-memory data structures in the metadata server. While communicating the
decoupled namespace directly to the metadata server is faster, communicating
through the object store (``Nonvolatile Apply") is \(10\times\) slower.

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
\subsubsection{Overhead of ``Nonvolatile Apply"} The cost of ``Nonvolatile
apply" is much larger than all the other mechanisms.  That mechanism was not
implemented as part of CudeleFS -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communicating through the object store is shown
by comparing the runtime of ``Volatile apply" + ``Global persist" to
``Nonvolatile Apply". These two operations end up with the same final metadata
state but using ``Nonvolatile Apply" is clearly inferior.

% persist vs. save: one disk vs. many
\subsubsection{Parallelism of the Object Store} Comparing ``Local" and ``Global
persist" demonstrates the bandwidth advantages of storing the journal in a
distributed object store. For 100K file creates, the ``Global Persist"
performance is \(1.5\times\) faster because the object store is leveraging the
collective bandwidth of the disks in the cluster. This benefit comes from the
object store itself but should be acknowledged when making decisions for the
application; the size of the object store can help mitigate the overheads of
globally persisting metadata updates.

\subsubsection{Journal Size} Figure~\ref{fig:behavior-journal-size} shows the
amount of storage per journal update (\(y\) axis) for the range of file creates
we tested (\(x\) axis). The increase in file size is linear with the number of
metadata creates and suggests that updates for a million files would be
\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
payloads of this size on an HPC network are reasonable.\\

\noindent\textbf{Takeaway}: the CudeleFS mechanisms have overheads and costs
that can differ {\it by orders of magnitude}. CudeleFS gives users the ability
to compose these mechanisms based on their application's correctness
requirements and performance goals.

\subsection{Weak vs. Strong Consistency}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
\caption{
[\href{https://...}{source}]
The RPC per metadata update of ``Strong Consistency" has a large
overhead compared to the decoupled namespace strategy of ``Weak
Consistency".\label{fig:slowdown-strong-weak}}
\end{figure}

Figure~\ref{fig:slowdown-strong-weak} shows the runtime of systems employing
weak and strong consistency, normalized to the runtime of the ``Append Client Journal"
mechanism (again, just creating files in the client's in-memory journal).  We
use the following compositions from the mechanisms in
Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal" +
``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal" +
``Local Persist" + ``Volatile Apply".

% Final file system metadata states are equivalent No guarantees while
% transitioning mechanisms
We compare these semantics because the final metadata states are equivalent.
CudeleFS makes no guarantees during execution of the mechanisms or when
transitioning semantics -- the semantics are guaranteed {\it once the mechanism
completes}. So if servers fail during a mechanism, metadata or data may be
lost.

% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
% (5-7x) vs. (90-104x) slower
% Does not scale?
% Strong Consistency > 10X slower than Weak Consistency 
\subsubsection{Speedups of Decoupled Namespaces} Weak consistency uses the
decoupled namespace strategy and shows up to a \(20\times\) speedup over the
traditional namespaces that use RPCs. Compared to the baseline the slowdown is
\(5-7\times\) for Strong Consistency, which emulates BatchFS and
\(90-104\times\) for Weak Consistency, which emulates DeltaFS.

% POSIX (no stream) vs. POSIX: cost of durability < consistency
\subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
Strong Consistency compared to Strong Consistency, No Fault Tolerance for
100K files is negligible. It suggests that the overhead of consistency is much
larger than the overhead of durability. This conclusion should be stronger as
we scale the number of files because the cost of streaming the journal into the
object store is constant. We omit the same analysis for Weak Consistency
because the runtimes are so short that the normalized slowdowns are misleading.

% Weak Consistency benefit not due to metadata format
\subsubsection{Metadata Formats} Because the metadata formats are the same for
all schemes we argue that the performance gain for decoupled namespaces comes
from relaxing the consistency guarantees and not from the metadata formats.\\

\noindent\textbf{Takeaway}: CudeleFS shows the true benefit of eventual
consistency, where we see over a \(100\times\) slowdown for achieving strong
consistency, in the worst case.

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: weak vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: weak, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

\subsection{Benefits of Isolated Subtree Policies}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
\caption{
[\href{https://...}{source}]
Using the ``allow" and ``block" API, users can isolate directories from
interfering clients. Variance with blocking turned on is the same as
``isolated" from Figure~\ref{fig:overhead-b}.
\label{fig:slowdown-allow-block}}
\end{figure}

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
Next we show how CudeleFS can be programmed to block interfering clients. We
use the same problematic workload from Figure~\ref{fig:overhead-b}, where
clients write to their own private directories and another client interferes with a stream of creates at
30 seconds.  We also have another client write to a decoupled namespace and
merge its updates at 90 seconds.  We equip each client directory with the
configuration:

\begin{listing}[tb]
\begin{minted}[xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\end{listing}

Each directory functions with ``RPCs"
and ``Stream" enabled -- which is the default implementation of CephFS. The only
difference is that we enable blocking on each subtree so while the tree behaves
like CephFS, all interfering operations will be returned with \texttt{-EBUSY}.
Note that IndexFS does a similar operation with leases except clients block.

% results
To show the benefits of this isolations, our results in
Figure~\ref{fig:slowdown-allow-block} are plotted alongside the variance bars from
Figure~\ref{fig:overhead-b}. Because ``allow"/``interfere and
``block"/``isolated" have the same variability we draw the following three
conclusions: (1) clients that use the API to block interfering clients  get
the same performance as isolated clients, (2) there is a negligible effect on
performance for the extra work the metadata server does to return
\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.\\

\noindent\textbf{Takeaway}: the API lets users isolate directories when
applications need better and more reliable performance. This is a way of
controlling consistency.

\subsection{Scaling Serialized Merges}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
      \caption{Merging journals serially.} \label{fig:merge-a}
  \end{subfigure}\\
  ~
  \begin{subfigure}[b]{\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
      \caption{Merging a journal w/ 100K create updates.}
      \label{fig:merge-b}
  \end{subfigure}
  \caption{Merging updates has linear scalability.\label{fig:merge}}
\end{figure}

CudeleFS serializes merge requests at the metadata server before applying the
updates to the in-memory metadata store. Merging concurrently is an
optimization for future work, but merges for the largest journal size we tested
were quite fast.  Furthermore, since we place no restrictions on the validity
of metadata inserted into the journal, we can avoid touching poorly scaling
data structures. In other words, we can scale linearly if we weaken our
consistency by never checking for the existence of files before creating files.

Figure~\ref{fig:merge-a} demonstrates this scalability by showing the runtime
(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
of the runtime is over 20 serial merges, each hitting a different directory in
the namespace.  The runtime scales linearly and with low variance because the
``Create" mechanism that was used to create the journal only appends
\texttt{open()} requests to the journal. When the updates are merged by the
metadata server into the in-memory metadata store they never scan growing data
structures. Had we implemented the ``Create" mechanism to include
\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
the poor scaling that we see with the ``RPCs" mechanism.
Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
structure because the runtimes do not increase as we add more metadata. \\

\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
({\it i.e.} not checking data structures for the validity of an update or
existence of a file) shows linear scalability and stable performance.

### 2017-10-12 paper/sections/7_conclusion.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%\section{Discussion}
%
%% Could we have implemented this on something other than CephFS?
%One question that comes up often in our work with Ceph is: ``Why do we always
%choose Ceph and can we implement this on another storage system?" Ceph is
%popular, robust, and open source -- getting it merged into mainline has a
%better chance of getting used than if we had built the system `clean-slate'
%from the ground up. Using Ceph, we can also leverage its robustness using an
%approach called ``programmability".
%
%% How awesome programmability is
%``Programmability" is a way of designing storage systems that encourages code
%re-use and composition.  A programmable storage system exposes internal
%subsystem as building blocks for higher level services. This `dirty-slate'
%approach limits redundant code and leverages the robustness of the underlying
%storage system. CudeleFS uses this approach and re-uses some of the building
%blocks from the Malacology programmable storage system~\cite{sevilla:eurosys17}
%to great success and requires only:
%
%\begin{itemize}
%
%  \item 354 lines of library code
%
%  \item 219 lines of non-destructive metadata server code, which is not used
%  unless it is turned on
%
%  \item 4 lines of destructive client/server code to check whether a namespace
%  is decoupled
%
%\end{itemize}

\section{Conclusion}

Relaxing consistency and durability semantics in the file system is a
double-edged sword. While it achieves better performance and scalability, it
alienates applications that rely on strong consistency and durability. CudeleFS
lets users assign consistency and durability guarantees to subtrees in
the global namespace, giving users a wide range of policies and
optimizations that can be custom fit to the application. Using CudeleFS, we
compare related work on the same file system and conclude that strong
consistency can cause a 104\(\times\) slow down while merging updates and
maintaining durability only have between a \(7-10\times\) slow down.

% Implied Namespaces

% Intermediate Update Bursts

% See if this helps load balancing

% executing mechanisms in parallel

### 2017-10-12 paper/build.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
+set -ex
+
+# Build the paper
+docker run --rm \
+  --name latex \
+  --entrypoint=/bin/bash \
+  -v `pwd`/:/mnt \
+  michaelsevilla/texlive:latest -c \
+    "cd /mnt ; \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
+     bibtex paper && \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper" &> build.log
+
+ERR=$?
+if [ $ERR != "0" ] ; then
+  echo "ERROR: $ERR"
+  cat build.log
+  exit 1
+fi
+
+set +e
+rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
+
+echo "SUCCESS"
+

### 2017-10-12 paper/sections/0_abstract.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}

HPC developers are abandoning POSIX because the synchronization and
serialization overheads of providing strong consistency and durability are too
costly -- and often unnecessary -- for their applications.  Unfortunately,
designing near-POSIX file systems excludes applications that rely on strong
consistency or durability, forcing developers to re-write their applications or
deploy them on a different system.  We present a file system and API that lets
clients specify their consistency/durability requirements and assign them to
subtrees in the namespace, allowing users to optimize subtrees within the same
namespace for different workloads.  We draw conclusions about the performance
impact of unexplored consistency/durability metadata designs and show that
maintaining strong consistency can cause about a 100\(\times\) slow down
compared to relaxed consistency and no durability. Comparatively, merging
updates after a period of relaxed consistency (less than a 10\(\times\) slow
down) and maintaining durability (about a 10\(\times\) slow down) have more
reasonable costs.

\end{abstract}

### 2017-10-12 paper/sections/3_posix_overheads.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{POSIX Overheads}
\label{sec:posix-overheads}

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
\caption{Create-heavy workloads (such as \texttt{untar}) incur the highest disk, network, and
CPU utilization because of the consistency and durability demands of
CephFS.}\label{fig:overhead-creates}
\end{figure}

In our examination of the overheads of POSIX we benchmark and analyze CephFS,
the file system that uses the Ceph's object store ({\it i.e.} RADOS) to store
its data and metadata. We choose CephFS because it is an open-source production
quality system. This file system is an implementation of one set of design
decisions and our goal is to highlight the effect that those decisions have on
performance. During this process we discovered, based on the analysis and
breakdown of costs, that durability and consistency have high overhead.

To show how the file system behaves under high metadata load we use a
create-heavy workload. Create-heavy workloads are studied the most in HPC
research because of the checkpoint/restart use case but they also happen to
stress the underlying storage system the most.
Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
the Linux kernel.  The \texttt{untar} phase, which is characterized by many creates, has
the highest resource usage which suggests that it is stressing the consistency
and journaling subsystems of the metadata server the most. Traditional file
system techniques for improving performance, such as caching inodes, do not
help for create-heavy workloads.

In this section, we quantify the costs of strong consistency and global
durability in CephFS. At the end of each subsection we compare the approach to
``decoupled namespaces", the technique in related work that detaches subtrees
from the global namespace to relax consistency/durability guarantees.  We use
the kernel client so that we can find the true create speed of the server; our
experiments show a low CPU utilization for the clients which indicates that we
are stressing the servers more.

\subsection{Durability}
\label{sec:durability}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
      \caption{Interference forces \texttt{lookup()}s}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
      \caption{Interference hurts variability}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{The overhead of global durability and strong consistency in CephFS.
  (a) shows the effect of different journal segment sizes, which are streamed
  into the object store for fault tolerance. (b) and (c) show that when a second
  client ``interferes", capabilities are revoked and metadata servers do more
  work.  \label{fig:overhead}}
\end{figure*}

% what is durability
While durability is not specified by POSIX, users expect that files they create
or modify survive failures.  We define three types of durability: global,
local, and none.  Global
durability means that the client or server can fail at any time and metadata
will not be lost. Local durability means that metadata can be lost if the
client or server stays down after a failure. None means that metadata is volatile
and that the system provides no guarantees when clients or servers fail. 
None is different than local durability because regardless of the type failure,
metadata will be lost when components die in a None configuration.\\

% - sequential IO, trim redundant operations
\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
ensures (1) sequential writes into the object store and (2) the ability for daemons to
trim redundant or irrelevant journal entries.  The journal is striped over
objects where multiple journal updates can reside on the same object. There are
two tunables for controlling the journal: the segment size and the number of
parallel segments that can be written in parallel. Unless the journal saturates 
memory or CPU resources, larger values for these tunables results in better
performance.

% purpose of the journal
As shown in Figure~\ref{fig:journal}, in addition to the metadata journal,
CephFS also represents metadata in RADOS as a metadata store, where directories
and their file inodes are stored as objects.  The metadata server applies the
updates in the journal to the metadata store when the journal reaches a certain
size. The metadata store is optimized for recovery ({\it i.e.} reading) while
the metadata journal is write-optimized.

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./figures/journal.png} 
\caption{CephFS has two views of the file system namespace: a journal of
metadata updates and a metadata store. For fault tolerance, they are stored in
the object store as segments and fragments, respectively.
\label{fig:journal}}
\end{figure}

% Effects on performance
Figure~\ref{fig:overhead-a} shows the
effect of journaling of different journal segment sizes; the larger the segment
size the bigger that the writes into the object store are. The trade-off for
better performance is memory consumption because larger segments take up
more space for buffering. When journaling is on, 
the metadata server periodically stops serving requests to flush ({\it i.e.}
apply journal updates) to the metadata store.  The journal overhead is
sufficient enough to slow down metadata throughput but not so much as to
overwhelm the bandwidth of the object store. We measured our peak bandwidth to
be 100MB/s, which is the speed of our network link.\\

\noindent\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
to the best of our knowledge, when a client or server fails there is no recovery
scheme. For BatchFS, if a client fails when it is writing to the local
log-structured merged tree (implemented as an SSTable) then those batched
metadata operations are lost. For DeltaFS, if the client fails then, on restart,
the computation does the work again -- since the snapshots of the namespace are
never globally consistent and there is no ground truth.  On the server side,
BatchFS and DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which
initially reside in memory but are later flushed to the underlying
distributed file system.

\subsection{Strong Consistency}
\label{sec:strong-consistency}

Access to metadata in a POSIX-compliant file system is strongly consistent, so
reads and writes to the same inode or directory are globally ordered.  The
synchronization and serialization machinery needed to ensure that all clients
see the same state has high overhead.\\

\noindent\textbf{CephFS Design}: capabilities keep metadata strongly
consistent. To reduce the number of RPCs needed for consistency, clients can
obtain capabilities for reading, creating and updating inodes, as well as caching reads,
buffering writes, changing the file size, and performing lazy IO.

% inode cache - reduces RPCs (lookups for create, readdirs for stats)
To keep track of the read caching and write buffering capabilities, the clients
and metadata servers agree on the state of each inode using an inode cache.  If
a client has the directory inode cached it can do metadata writes ({\it e.g.},
create) with a single RPC. If the client is not caching the directory inode
then it must do an extra RPC to determine if the file exists.  Unless the
client immediately reads all the inodes in the cache ({\it i.e.} \texttt{ls
-alR}), the inode cache is less useful for create-heavy workloads.

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but performance suffers.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
increased variability. The results are normalized to a single
isolated client and the \(y\) axis is the standard deviation of the runtime of
all clients.  For the ``interfere" bars, each client creates files in private
directories and at 30 seconds we launch another process that creates files in
those directories.  For less than 7 clients, the runtime and deviation is worse
when clients interfere. At 7 clients, the metadata server is overloaded so the
directory caching in the ``isolated" bars has no benefit.\\

% TODO: what is the cost of trimming the cache?
% TODO: does CephFS still cache inodes when I turn off caching? Why is still keeping inodes in memory? Gah!

\noindent\textbf{Comparison to decoupled namespaces}: Decoupled namespaces
merge batches of metadata operations into the global namespaces when the job
completes.  In BatchFS the merge is delayed by the application using an API to
switch between asynchronous to synchronous mode. The merge itself is explicitly
managed by the application but future work looks at more automated
methodologies. In DeltaFS snapshots of the metadata subtrees stays on the client
machines; there is no ground truth and consistent namespaces are constructed
and resolved at application read time or when a 3rd party system ({\it e.g.},
middleware, scheduler, etc.) needs a view of the metadata. As a result, all the
overheads of maintaining consistency that we showed above are delayed until the
merge phase.

### 2017-10-12 paper/sections/4_methodology.tex
paper: add template
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Methodology: Global Namespace with Per-Subtree Consistency/Durability}
\label{sec:methodology-decoupled-namespaces}

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications decouple the namespace, write updates to a local journal,
and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
\end{figure}

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{The CudeleFS mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

In this section we describe CudeleFS, our file system that lets users
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. CudeleFS exposes these
mechanisms and the user composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

\subsection{The CudeleFS Mechanisms}
\label{sec:the-cudelesfs-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Append Client Journal"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisfied by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
client and applies the updates directly to the in-memory metadata store maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- CudeleFS makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

% difference between apply and volatile apply
The biggest difference between ``Volatile Apply" and ``Nonvolatile Apply" is
the medium they use to communicate. ``Volatile Apply" applies updates directly
to the metadata servers' metadata store while ``Nonvolatile Apply" uses the
object store to communicate the journal of updates from the client to the
metadata servers.  ``Nonvolatile Apply" is safer but has a large performance
overhead because objects in the metadata store need to be read from and written
back to the object store.

%The metadata store and journal
%are different ways of representing the namespace.  CudeleFS presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Append Client Journal" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
mechanisms used by default in CephFS.  Using existing configuration settings in
Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
will not save journals in the object store. For ``Local Persist", clients write
serialized log events to a file on local disk and for ``Global Persist",
clients push the journal into the objects store. The overheads for both ``Local
Persist" and ``Global Persist" is the write bandwidth of the local disk and
object store, respectively.  These persist mechanisms are part of the library that
links into the client.

\subsection{Defining Policies in CudeleFS}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & append client journal            & append client journal          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & append client journal            & append client journal          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & append client journal            & append client journal          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}

\caption{Users can explore the consistency (C) and
durability (D) spectrums by composing CudeleFS mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that users can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``weak" merges updates at some time in the
future ({\it e.g.}, when the system has time, when the number of updates reaches a
certain threshold, when the client is done writing, etc.); and updates in
``strong" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability ~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{append client journal+RPCs} since
both mechanisms do the same thing or \texttt{stream+save} since ``global"
durability is stronger and has more overhead than ``local" durability.

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed to have ``global" durability until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that changes a subtree from ``local
persist" to ``global persist", CudeleFS makes no guarantee until the mechanisms are
complete.

\subsection{CudeleFS Namespace API}
\label{sec:cudelefs-namespace-api}

% the interface
Users control consistency and durability for subtrees by contacting the monitor
daemon with a directory path and presenting a policies configuration file.
For example, (msevilla/mydir, policies.yml) would decouple the path
``msevilla/mydir" and would apply the policies in ``policies.yml".  The
policies file supports the following parameters (default values are in parenthesis):

\begin{itemize}

  \item \texttt{consistency}: which consistency model to use (\texttt{RPCs})

  \item \texttt{durability}: which durability model to use (\texttt{stream})

  \item \texttt{allocated\_inodes}: the number of inodes to provision to the
  decoupled namespace (100)

  \item \texttt{interfere\_policy}: how to handle a request from another
  client targeted at this subtree (\texttt{allow}))

\end{itemize}

The ``Consistency" and ``Durability" parameters are set to the CudeleFS
mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
``Allocated Inodes" is a way for the application to specify how many files it
intends to create. It is a contract so that the file system can provision
enough resources for the incumbent merge and so it can give valid inodes to
other clients. The inodes can be used anywhere within the decoupled namespace
({i.e.} at any depth in the subtree).

% description
``Interfere Policy" has two settings: \texttt{block} and \texttt{allow}.
For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources
for updates that may get overwritten. If the application does not mind losing
updates, for example it wants approximations for results that take too long to
compute, it can select \texttt{allow}. In this case, metadata from the interfering client will be
written and the computation from the decoupled namespace will take priority at
merge time because the results are more accurate.

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
would use the configuration:

\begin{listing}[h]
\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "append_client_journal"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

BatchFS merges updates into the global namespace after the job is complete.
This can be achieved with by composing two mechanisms:

\begin{listing}[h]
\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
  "allocated_inodes": "100000"
  "interfere_policy": "block"
  "consistency": "append_client_journal+volatile_apply"
  "durability": "local_persist"
}
\end{minted}
\end{listing}

### 2017-10-12 experiments/baseline/run.sh
benchmark: port Luminous to CloudLab
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,9 +34,9 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-$DOCKER ceph.yml monitor.yml \
-        /workloads/radosbench.yml \
-        /workloads/netbench.yml \
-        /workloads/osdbench.yml
+$DOCKER ceph.yml monitor.yml #\
+#        /workloads/radosbench.yml \
+#        /workloads/netbench.yml \
+#        /workloads/osdbench.yml
 
 exit 0

### 2017-10-13 experiments/baseline-osds/docker/build.sh
benchmark: figure out how many OSDs to use on CloudLab
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,50 @@
+#!/bin/bash
+
+set -ex
+
+# horrible hack
+HASH=`cat vars.yml | grep "ceph_docker_version: " | grep -v "#" | awk '{print $2}'`
+GITR=`cat vars.yml | grep "ceph_repository: " | grep -v "#" | awk '{print $2}'`
+USERNAME=`cat vars.yml | grep "ceph_docker_username: " | grep -v "#" | awk '{print $2}'`
+IMAGENAME=`cat vars.yml | grep "ceph_docker_imagename: " | grep -v "#" | awk '{print $2}'`
+if [ -z $HASH ]; then exit 1; fi
+if [ -z $GITR ]; then exit 1; fi
+if [ -z $USERNAME ]; then exit 1; fi
+if [ -z $IMAGENAME ]; then exit 1; fi
+TAG="$USERNAME/$IMAGENAME:$HASH"
+echo "building Docker image with Ceph version $HASH from git repo $GITR tagged with $TAG"
+
+# create the image
+SRC="/tmp/ceph-daemon"
+mkdir $SRC || true
+cd $SRC
+
+# pull base image from ceph (we will layer on top of this)
+wget https://raw.githubusercontent.com/systemslab/docker-cephdev/master/aliases.sh
+. aliases.sh && rm aliases.sh
+docker pull ceph/daemon:tag-build-master-jewel-ubuntu-14.04
+docker tag ceph/daemon:tag-build-master-jewel-ubuntu-14.04 ceph/daemon:jewel
+docker pull cephbuilder/ceph:latest
+
+dmake \
+  -e SHA1_OR_REF="$HASH" \
+  -e GIT_URL="$GITR" \
+  -e BUILD_THREADS=`grep processor /proc/cpuinfo | wc -l` \
+  -e CONFIGURE_FLAGS="-DWITH_TESTS=OFF" \
+  -e RECONFIGURE="true" \
+  cephbuilder/ceph:latest build-cmake
+cd -
+
+docker tag ceph-$HASH ceph/daemon:$HASH
+
+docker run \
+  --name copier \
+  --entrypoint=/bin/bash \
+  -v $SRC:/ceph \
+  ceph/daemon:jewel -c "cp -r /ceph/build/install/usr/* /usr"
+docker commit --change='ENTRYPOINT ["/entrypoint.sh"]' copier $USERNAME/$IMAGENAME:$HASH &> /dev/null
+docker stop copier
[…]

### 2017-10-13 experiments/baseline-osds/run.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+#MSEVILLA
+for nosd in 7 6 5 4 3; do
+  mkdir results
+  cp host_setup/osds-$nosd hosts
+  ./teardown.sh
+  $DOCKER ceph.yml monitor.yml \
+          /workloads/radosbench.yml \
+          /workloads/netbench.yml \
+          /workloads/osdbench.yml
+  mv results results-osds-$nosd
+done
+
+exit 0

### 2017-10-13 experiments/baseline-osds/teardown.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+# Put all your cleanup tasks here.
+set -ex
+
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml -e @/popper/ansible/vars.yml -i /etc/ansible/hosts"
+ROOT=`dirname $PWD | xargs dirname`
+docker run -it --rm \
+  --net host \
+  -v $HOME/.ssh:/root/.ssh \
+  -v `pwd`/ansible/group_vars:/root/group_vars \
+  -v `pwd`/hosts:/etc/ansible/hosts \
+  -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+  -v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+  -v `pwd`:/popper \
+  -v $ROOT/ansible/ceph:/root -w /root \
+  -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+  -e ANSIBLE_CONFIG="/etc/ansible/ansible.cfg" \
+  michaelsevilla/ansible $ARGS $VARS /popper/ansible/cleanup.yml
+  #--entrypoint=/bin/bash michaelsevilla/ansible -c "$@"
+ 
+exit 0

### 2017-10-13 experiments/baseline-creates/run_cloudlab.sh
benchmark: cloudlab try to scale clients
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,45 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/cloudlab_configs/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+for clients in 5 4 3 2 1; do
+  mkdir results
+  ./teardown_cloudlab.sh
+  cp cloudlab_configs/clients$clients hosts
+  $DOCKER ceph.yml monitor.yml
+  $DOCKER -e nfiles=98000 /workloads/creates.yml
+  mv results results-cloudlab-clients$clients
+done
+
+exit 0

### 2017-10-13 experiments/baseline-creates/run_piha.sh

@@ -1,0 +1,45 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+for clients in 8 7 6 5 4 3 2 1; do
+  mkdir results
+  ./teardown.sh
+  cp hosts_piha/clients$clients hosts
+  $DOCKER ceph.yml monitor.yml
+  $DOCKER -e nfiles=98000 /workloads/creates.yml
+  mv results results-piha-clients$clients
+done
+
+exit 0

### 2017-10-13 experiments/baseline-creates/teardown_cloudlab.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+# Put all your cleanup tasks here.
+set -ex
+
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/cloudlab_configs/vars.yml -e @/popper/ansible/vars.yml -i /etc/ansible/hosts"
+ROOT=`dirname $PWD | xargs dirname`
+docker run -it --rm \
+  --net host \
+  -v $HOME/.ssh:/root/.ssh \
+  -v `pwd`/ansible/group_vars:/root/group_vars \
+  -v `pwd`/hosts:/etc/ansible/hosts \
+  -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+  -v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+  -v `pwd`:/popper \
+  -v $ROOT/ansible/ceph:/root -w /root \
+  -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+  -e ANSIBLE_CONFIG="/etc/ansible/ansible.cfg" \
+  michaelsevilla/ansible $ARGS $VARS /popper/ansible/cleanup.yml
+  #--entrypoint=/bin/bash michaelsevilla/ansible -c "$@"
+ 
+exit 0

### 2017-10-14 experiments/baseline-creates/run.sh
benchmark: port back to cloudlab
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -8,6 +8,7 @@ rm -fr results || true
 
 # if you know Ansible and Docker, the below should make sense
 # - we attach ceph-ansible to root because they expect us to be in that dir
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
 ROOT=`dirname $PWD | xargs dirname`
 NETW="--net host -v $HOME/.ssh:/root/.ssh"
 DIRS="-v `pwd`:/popper \
@@ -22,7 +23,7 @@ CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
       -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
 WORK="-v `pwd`/ansible/workloads:/workloads"
 ARGS="--forks 50 --skip-tags package-install,with_pkg"
-VARS="-e @/popper/cloudlab_configs/vars.yml \
+VARS="-e @/popper/vars.yml \
       -e @/popper/ansible/vars.yml \
       -i /etc/ansible/hosts"
 DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
@@ -33,13 +34,14 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-for clients in 5 4 3 2 1; do
+for clients in 1; do
   mkdir results
-  ./teardown_cloudlab.sh
-  cp cloudlab_configs/clients$clients hosts
+  ./teardown.sh
+  cp configs_$SITE/clients$clients hosts
   $DOCKER ceph.yml monitor.yml
+  exit
   $DOCKER -e nfiles=98000 /workloads/creates.yml
-  mv results results-cloudlab-clients$clients
+  mv results results-$SITE-clients$clients
 done
 
 exit 0

### 2017-10-14 experiments/baseline-creates/run_piha.sh

@@ -1,45 +1,0 @@
-#!/bin/bash
-# This file should contain the series of steps that are required to execute 
-# the experiment. Any non-zero exit code will be interpreted as a failure
-# by the 'popper check' command.
-set -ex
-
-rm -fr results || true
-
-# if you know Ansible and Docker, the below should make sense
-# - we attach ceph-ansible to root because they expect us to be in that dir
-ROOT=`dirname $PWD | xargs dirname`
-NETW="--net host -v $HOME/.ssh:/root/.ssh"
-DIRS="-v `pwd`:/popper \
-      -v $ROOT/ansible/ceph:/root \
-      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
-      -w /root "
-ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
-      -v `pwd`/hosts:/etc/ansible/hosts \
-      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
-      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
-CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
-      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
-WORK="-v `pwd`/ansible/workloads:/workloads"
-ARGS="--forks 50 --skip-tags package-install,with_pkg"
-VARS="-e @/popper/vars.yml \
-      -e @/popper/ansible/vars.yml \
-      -i /etc/ansible/hosts"
-DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
-
-# debug mode
-if [ ! -z $1 ]; then
-  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
-  exit
-fi
-
-for clients in 8 7 6 5 4 3 2 1; do
-  mkdir results
-  ./teardown.sh
-  cp hosts_piha/clients$clients hosts
-  $DOCKER ceph.yml monitor.yml
-  $DOCKER -e nfiles=98000 /workloads/creates.yml
-  mv results results-piha-clients$clients
-done
-
-exit 0

### 2017-10-14 experiments/baseline-creates/teardown.sh

@@ -2,8 +2,9 @@
 # Put all your cleanup tasks here.
 set -ex
 
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
 ARGS="--forks 50 --skip-tags package-install,with_pkg"
-VARS="-e @/popper/cloudlab_configs/vars.yml -e @/popper/ansible/vars.yml -i /etc/ansible/hosts"
+VARS="-e @/popper/vars.yml -e @/popper/ansible/vars.yml -i /etc/ansible/hosts"
 ROOT=`dirname $PWD | xargs dirname`
 docker run -it --rm \
   --net host \

### 2017-10-14 experiments/baseline-creates/run.sh
benchmark: partial resutls for cloudlab
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,12 +34,11 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-for clients in 1; do
+for clients in 10 11 12 13 14 15 16 17 18 19 20; do
   mkdir results
   ./teardown.sh
   cp configs_$SITE/clients$clients hosts
   $DOCKER ceph.yml monitor.yml
-  exit
   $DOCKER -e nfiles=98000 /workloads/creates.yml
   mv results results-$SITE-clients$clients
 done

### 2017-10-15 experiments/baseline/run.sh
benchmark: add interfere experiments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,9 +34,9 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-$DOCKER ceph.yml monitor.yml #\
-#        /workloads/radosbench.yml \
-#        /workloads/netbench.yml \
-#        /workloads/osdbench.yml
+$DOCKER ceph.yml monitor.yml \
+        /workloads/radosbench.yml \
+        /workloads/netbench.yml \
+        /workloads/osdbench.yml
 
 exit 0

### 2017-10-15 experiments/baseline-creates/run.sh
benchmark: more partial results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,14 +34,14 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-for clients in 35 30 25 20 15 10 5 1; do
-  mkdir results
-  ./teardown.sh
+./teardown.sh; rm -rf results
+for clients in 15 20 25 30 35; do
+  mkdir results || true
   cp configs_$SITE/clients$clients hosts
-  $DOCKER -e nfiles=98000 \
-          ceph.yml monitor.yml \
-          /workloads/creates-touchstream.yml \
-          /workloads/creates.yml
+  for job in "creates" "creates-touchstream"; do
+    ./teardown.sh
+    $DOCKER -e nfiles=98000 ceph.yml /workloads/${job}.yml
+  done
   mv results results-$SITE-clients$clients
 done
 

### 2017-10-15 experiments/cudele-mechanisms/run.sh
benchmark: mechanism microbenchmark
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,47 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+for run in 0 1 2; do
+  for nfiles in 10 100 1000 10000 100000 1000000; do
+    mkdir results || true
+    ./teardown.sh
+    cp configs_$SITE/hosts hosts
+    $DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
+  done
+  mv results results-$SITE-clients$clients-run$run
+done
+
+exit 0

### 2017-10-15 experiments/cudele-mechanisms/visualize/jupyter.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+set -e -x
+docker run --rm -v `dirname $PWD`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93

### 2017-10-16 experiments/baseline-creates/run.sh
benchmark: nclients sweep
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -35,14 +35,16 @@ if [ ! -z $1 ]; then
 fi
 
 ./teardown.sh; rm -rf results
-for clients in 15 20 25 30 35; do
-  mkdir results || true
-  cp configs_$SITE/clients$clients hosts
-  for job in "creates" "creates-touchstream"; do
-    ./teardown.sh
-    $DOCKER -e nfiles=98000 ceph.yml /workloads/${job}.yml
+for log in "nolog" "log"; do
+  cp configs_$SITE/all-$log ansible/group_vars/all 
+  for clients in 1 5 10 15 20 25 30; do
+    mkdir results || true
+    cp configs_$SITE/clients$clients hosts
+    for job in "creates" "creates-touchstream"; do
+      ./teardown.sh
+      $DOCKER -e nfiles=98000 ceph.yml /workloads/${job}.yml
+    done
+    mv results results-$SITE-clients$clients-$log
   done
-  mv results results-$SITE-clients$clients
 done
-
 exit 0

### 2017-10-17 experiments/cudele-mechanisms/run.sh
benchmarK: add mechanisms
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -35,7 +35,7 @@ if [ ! -z $1 ]; then
 fi
 
 for run in 0 1 2; do
-  for nfiles in 10 100 1000 10000 100000 1000000; do
+  for nfiles in 10 100 1000 10000 100000 1000000 10000000; do
     mkdir results || true
     ./teardown.sh
     cp configs_$SITE/hosts hosts

### 2017-10-17 paper/sections/1_introduction.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % What is the problem?
-File system metadata services in HPC have scalability problems. It has been
-shown that HPC workloads are metadata resource intensive because administrative
+File system metadata services in HPC have scalability problems. Administrative
 tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
-system~\cite{zheng:pdsw2014-batchfs}, on large data sets lead to contention
-for the same directories and inodes ({\it
-e.g.}, path traversal). Applications perform better with dedicated metadata
-servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
-metadata server for every client is unreasonable. This problem is exacerbated
-by current trends in HPC, where architectures are transitioning from complex
-storage stacks with burst buffer, file system, object store, and tape tiers to
-more simplified stacks with just a burst buffer and object
-store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
-because more requests end up hitting the same layer and latencies cannot be 
-hidden while data migrates across tiers.
+system~\cite{zheng:pdsw2014-batchfs} lead to contention for the same
+directories and inodes ({\it e.g.}, path traversal). Applications perform
+better with dedicated metadata servers~\cite{sevilla:sc15-mantle,
+ren:sc2014-indexfs} but provisioning a metadata server for every client is
+unreasonable. This problem is exacerbated by current trends in HPC, where
+architectures are transitioning from complex storage stacks with burst buffer,
+file system, object store, and tape tiers to more simplified stacks with just a
+burst buffer and object store~\cite{bent:login16-hpc-trends}; this puts more
+pressure on data access because more requests end up hitting the same layer and
+latencies cannot be hidden while data migrates across tiers.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
-\caption{With CudeleFS, users can assign weaker consistency and durability
-policies to subtrees to get the same performance benefits of state-of-the-art
-HPC architectures, such as those shown in the embedded table, while
-applications that require the stronger guarantees of POSIX can still reside in
-the same namespace.
+\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
+\caption{ Clients that need better performance can relax consistency on their
+subtree ({e.g.}, HDFS) or even decouple the subtree and move it locally
+({e.g.}, BatchFS, RAMDisk). Decoupled subtrees can further relax durability
+semantics to achieve even better performance. Meanwhile, clients that require
+stronger guarantees ({\it e.g.,} POSIX) can still reside in the same namespace.
 }\label{fig:subtree-policies}
@@
 semantics in the file system because weaker guarantees are sufficient for their
-applications. For example, batch style jobs often do not need the strong
+applications. For example, many batch style jobs do not need the strong
 consistency that the file system provides, so
@@
 delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
-durability improves performance and scalability by avoiding the costs of remote procedure calls (RPCs),
-synchronization, false sharing, and serialization.  While the performance
-benefits of decoupling the namespace are obvious, applications that rely on the
-file system's guarantees must be deployed on an entirely different system or
-re-written to coordinate strong consistency/durability themselves.
+durability improves performance and scalability by avoiding the costs of remote
+procedure calls (RPCs), synchronization, false sharing, and serialization.
+While the performance benefits of decoupling the namespace are obvious,
+applications that rely on the file system's guarantees must be deployed on an
+entirely different system or re-written to coordinate strong
+consistency/durability themselves.
 
@@
 % What did we do
-To address this problem, we propose a subtree policy API that lets developers dynamically control the
-consistency and durability guarantees for subtrees in the file system
-namespace. For example, one subtree can adopt weaker consistency semantics
-while another subtree can retain the rigidity of POSIX's strong consistency.
-The subtrees in Figure~\ref{fig:subtree-policies} show a setup in our proposed system where a single global
-namespace has subtrees for applications optimized with techniques from different
-state-of-the-art HPC architectures.
+To address this problem, we present an API and framework that lets developers
+dynamically control the consistency and durability guarantees for subtrees in
+the file system namespace.  Figure~\ref{fig:subtree-policies} shows a setup in
+our proposed system where a single global namespace has subtrees for
+applications optimized with techniques from different state-of-the-art HPC
+architectures.  The HDFS\footnote{This is example is not directly evaluated in
+this paper, although a similiar use case is explored in~\S\ref{use-case-1}}
+subtree has weaker consistency than strong consistency because it  does not
+guarantee that all updates are not immediately seen by all servers; the BatchFS
+and RAMDisk subtrees are decoupled from the global namespace and have similiar
+consistency/durability behavior to those systems; and the POSIX subtree retains
+the rigidity of POSIX's strong consistency.
 
-In this work, we present CudeleFS, a prototype programmable file system that
-supports different degrees of consistency and durability in a global namespace.
-CudeleFS achieves this by exposing ``mechanisms" that developers use to specify
-their preferred semantics.  CudeleFS supports 3 forms of consistency
-(invisible, weak, and strong) and 3 degrees of durability (none, local, and
-global) giving the user a wide range of policies and optimizations that can be
-custom fit to an application. We make the following contributions:
+Our protoype system, Cudele, achieves this by exposing ``mechanisms" that
+developers use to specify their preferred semantics.  Cudele supports 3 forms
+of consistency (invisible, weak, and strong) and 3 degrees of durability (none,
+local, and global) giving the user a wide range of policies and optimizations
+that can be custom fit to an application. We make the following contributions:
 
@@
   \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
-  previously unexplored metadata designs, all implemented using CudeleFS.
+  previously unexplored metadata designs, all implemented using Cudele.
 
@@
 (\(10\times\) slow down). We also show the effect of having a metadata specific
-file format in systems that are based on in-memory data structures.
-In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
-consistency and system-defined durability and
-Section~\ref{sec:methodology-decoupled-namespaces} presents the CudeleFS
-prototype and API. Section~\ref{sec:implementation} describes the CudeleFS
+file format in systems that are based on in-memory data structures. Because
+this paper focuses on file system metadata protocols, we only scale up to 20
+clients executing an extremely metdata intensive workload; this load is the
+maximum capacity of a single metadata server for the platform we chose to
+implement Cudele on.  Partitioning and replicating metadata across a cluster
+can be controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in
+this paper we show the capacity and performance of a metadata server with
+superior metadata protocols, which should be used to inform load balancing
+decisions.
+
+In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
+cost of POSIX consistency and system-defined durability and
+Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
+prototype and API. Section~\ref{sec:implementation} describes the Cudele
 mechanisms and shows how re-using internal subsystems results in an
@@
 explored and previously unexplored metadata designs.
-Section~\ref{sec:related-work} places CudeleFS in the context of other related
+Section~\ref{sec:related-work} places Cudele in the context of other related
 work.

### 2017-10-17 paper/sections/4_methodology.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \caption{Applications decouple the namespace, write updates to a local journal,
-and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
+and delay metadata updates using the Cudele Mechanisms }\label{fig:decouple}
 \end{figure}
@@
 \end{tabular}
-\caption{The CudeleFS mechanisms are composed together to form consistency and
+\caption{The Cudele mechanisms are composed together to form consistency and
 durability semantics.\label{table:mechanisms}} 
@@
 
-In this section we describe CudeleFS, our file system that lets users
+In this section we describe Cudele, our file system that lets users
 assign consistency and durability semantics to subtrees in the global
 namespace. A \textbf{mechanism} is an abstraction and basic building block for
-constructing consistency and durability guarantees. CudeleFS exposes these
+constructing consistency and durability guarantees. Cudele exposes these
 mechanisms and the user composes them together to construct
@@
 
-\subsection{The CudeleFS Mechanisms}
+\subsection{The Cudele Mechanisms}
 \label{sec:the-cudelesfs-mechanisms}
@@
 % describe the figure
-Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
+Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
 which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
@@
 mechanism to append metadata updates to a local, in-memory journal. Once the
-job is complete, the system calls CudeleFS mechanisms to achieve the desired
-consistency/durability semantics.  CudeleFS provides a library for clients to
+job is complete, the system calls Cudele mechanisms to achieve the desired
+consistency/durability semantics.  Cudele provides a library for clients to
 link into and all operations are performed by the client.
@@
 by the metadata servers. We say volatile because -- in exchange for peak
-performance -- CudeleFS makes no consistency or durability guarantees while
+performance -- Cudele makes no consistency or durability guarantees while
 ``Volatile Apply" is executing.  If a concurrent update from a client occurs
@@
 %The metadata store and journal
-%are different ways of representing the namespace.  CudeleFS presents 6
+%are different ways of representing the namespace.  Cudele presents 6
 %mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
@@
 
-\subsection{Defining Policies in CudeleFS}
+\subsection{Defining Policies in Cudele}
 \label{sec:setting-policies-with-cudele}
@@
 \caption{Users can explore the consistency (C) and
-durability (D) spectrums by composing CudeleFS mechanisms. 
+durability (D) spectrums by composing Cudele mechanisms. 
 \label{table:spectrum}}
@@
 persist" or if we inject a new policy that changes a subtree from ``local
-persist" to ``global persist", CudeleFS makes no guarantee until the mechanisms are
+persist" to ``global persist", Cudele makes no guarantee until the mechanisms are
 complete.
 
-\subsection{CudeleFS Namespace API}
+\subsection{Cudele Namespace API}
 \label{sec:cudelefs-namespace-api}
@@
 
-The ``Consistency" and ``Durability" parameters are set to the CudeleFS
+The ``Consistency" and ``Durability" parameters are set to the Cudele
 mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
@@
 would give the application 100 inodes but the subtree would behave like the
-existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
+existing CephFS implementation. To implement DeltaFS on Cudele, the user
 would use the configuration:

### 2017-10-17 paper/sections/5_implementation.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 We use a programmable storage approach~\cite{sevilla:eurosys17} to design
-CudeleFS; namely, we try to leverage components inside Ceph to inherit the
+Cudele; namely, we try to leverage components inside Ceph to inherit the
 robustness and correctness of the internal subsystems. Using this `dirty-slate'
@@
 system itself.  In this section, we first describe a Ceph internal subsystem or
-component and then we show how we use it in CudeleFS.
+component and then we show how we use it in Cudele.
 
@@
 % how i used it
-\noindent\textbf{CudeleFS}: the ``RPCs" mechanism uses the metadata store to
+\noindent\textbf{Cudele}: the ``RPCs" mechanism uses the metadata store to
 service requests. Using code designed for recovery, ``Nonvolatile Apply" and
@@
 
-\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Append Client Journal",
+\noindent\textbf{Cudele}: the journal format is used by ``Stream", ``Append Client Journal",
 ``Local Persist", and ``Global Persist".  ``Stream" is the default
@@
 servers know how to read the events the library is writing.  By re-using the
-journal subsystem to implement the namespace decoupling, CudeleFS leverages the
+journal subsystem to implement the namespace decoupling, Cudele leverages the
 write/read optimized data structures, the formats for persisting events
@@
 
-\noindent\textbf{CudeleFS}: the external library the clients link into is based
+\noindent\textbf{Cudele}: the external library the clients link into is based
 on the journal tool. It already had functions for importing, exporting, and
@@
 
-\noindent\textbf{CudeleFS}: ``Nonvolatile Apply" uses the internal inode cache
+\noindent\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
 code to allocate inodes to clients that decouple parts of the namespace and to
@@
 
-\noindent\textbf{CudeleFS}: uses the large inodes to store consistency and
+\noindent\textbf{Cudele}: uses the large inodes to store consistency and
 durability policies in the directory inode. This approach uses the File Type
@@
 underlying metadata. The underlying implementation stores executable code in
-the inode that calls the different CudeleFS mechanisms. Of course, there are many
+the inode that calls the different Cudele mechanisms. Of course, there are many
 security and access control aspects of this approach but that is beyond the

### 2017-10-17 paper/sections/6_results.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-We evaluate CudeleFS on a 15 node cluster, partitioned into 8 object storage
+We evaluate Cudele on a 15 node cluster, partitioned into 8 object storage
 servers, 3 metadata servers, and 2 monitor servers. The object storage servers
@@
 re-initialize the cluster ({\it e.g.}, dropping the kernel cache) between
-experiments.
+experiments. For these experiments, we compare CephFS to our programmable
+consistency and durability modifications, which we call Cudele.
 
@@
 
-\subsection{CudeleFS Mechanism Performance}
-
+\subsection{Microbenchmarks}
 \begin{figure}[tb]
@@
 [\href{https://...}{source}]
-The performance of the CudeleFS mechanisms normalized to the runtime of the
+The performance of the Cudele mechanisms normalized to the runtime of the
 ``Append Client Journal" mechanism (the runtime of writing \(n\) file creates
@@
 
-Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
-mechanisms, normalized to the time it takes to write \(n\) file create updates
-to the client's in-memory journal ({\it i.e.} the ``Append Client Journal"
-mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal"
-mechanism.  ``Stream" is an approximation of the
-overhead and is calculated by subtracting the runtime of the job with the
-journal turned off from the runtime with the journal turned on.  The largest
-workload we tested is 100K file creates in the same directory.  This is the
-maximum recommended size of a directory in CephFS; preliminary experiments with
-larger directory sizes show memory problems.
+Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the Cudele
+mechanisms for a single client creating files in the same directory, normalized
+to the time it takes to write \(n\) file create updates to the client's
+in-memory journal ({\it i.e.} the ``Append Client Journal" mechanism). Bars
+above \(10^0\) are slower than the ``Append Client Journal" mechanism.
+``Stream" is an approximation of the overhead and is calculated by subtracting
+the runtime of the job with the journal turned off from the runtime with the
+journal turned on.  The largest workload we tested is 100K file creates in the
+same directory.  This is the maximum recommended size of a directory in CephFS;
+preliminary experiments with larger directory sizes show memory problems.
 
-\subsubsection{Poorly Scaling Data Structures} Despite doing the same amount of
+{\it Poorly Scaling Data Structures:} Despite doing the same amount of
 work, mechanisms that rely on poorly scaling data structures have large
@@
 % RPCs vs. apply: calls to metadata server vs. RADOS
-\subsubsection{Overhead of RPCs} ``RPCs" is \(66\times\) slower than ``Volatile
+{\it Overhead of RPCs:} ``RPCs" is \(66\times\) slower than ``Volatile
 Apply" because sending individual metadata updates over the network is costly.
@@
 % v_apply vs. apply/persist: communicating through RADOS
-\subsubsection{Overhead of ``Nonvolatile Apply"} The cost of ``Nonvolatile
+{\it Overhead of ``Nonvolatile Apply":} The cost of ``Nonvolatile
 apply" is much larger than all the other mechanisms.  That mechanism was not
-implemented as part of CudeleFS -- it was a debugging and recovery tool packaged
+implemented as part of Cudele -- it was a debugging and recovery tool packaged
 with CephFS. It works by iterating over the updates in the journal and pulling
@@
 % persist vs. save: one disk vs. many
-\subsubsection{Parallelism of the Object Store} Comparing ``Local" and ``Global
+{\it Parallelism of the Object Store:} Comparing ``Local" and ``Global
 persist" demonstrates the bandwidth advantages of storing the journal in a
@@
 
-\subsubsection{Journal Size} Figure~\ref{fig:behavior-journal-size} shows the
+{\it Journal Size:} Figure~\ref{fig:behavior-journal-size} shows the
 amount of storage per journal update (\(y\) axis) for the range of file creates
@@
 
-\noindent\textbf{Takeaway}: the CudeleFS mechanisms have overheads and costs
-that can differ {\it by orders of magnitude}. CudeleFS gives users the ability
+\noindent\textbf{Takeaway}: the Cudele mechanisms have overheads and costs
+that can differ {\it by orders of magnitude}. Cudele gives users the ability
 to compose these mechanisms based on their application's correctness
@@
 
-\subsection{Weak vs. Strong Consistency}
+\subsection{Use Case 1: Creates in the Same Directory}
+\label{use-case-1}
+
+Clients creating files in private directories is heavily
+studied in HPC~\cite{plfs, cephfs04, indexfs, giga+, batchfs, deltafs,
+sevilla:eurosys, shardfs, mantle}, mostly due to checkpoint-restart~\cite{}.
+But the workload also appears in cloud workloads, where systems like Hadoop use
+the file abstraction to exchange work units to workers or to indicate when jobs
+complete~\cite{}. A more familiar example is uncompressing an archive ({\it
+e.g.}, \texttt{tar xzf}), where the file system services a flash crowd of
+creates across all directories as shown in Figure~\ref{fig:overhead-creates}.
+We use this as a microbenchmark because it allows us to control the size of the
+log, since each create creates a single journal event.
+
 \begin{figure}[tb]
@@
 We compare these semantics because the final metadata states are equivalent.
-CudeleFS makes no guarantees during execution of the mechanisms or when
+Cudele makes no guarantees during execution of the mechanisms or when
 transitioning semantics -- the semantics are guaranteed {\it once the mechanism
 completes}. So if servers fail during a mechanism, metadata or data may be
-lost.
+lost.\\
 
@@
 % Strong Consistency > 10X slower than Weak Consistency 
-\subsubsection{Speedups of Decoupled Namespaces} Weak consistency uses the
+{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
 decoupled namespace strategy and shows up to a \(20\times\) speedup over the
@@
 \(5-7\times\) for Strong Consistency, which emulates BatchFS and
-\(90-104\times\) for Weak Consistency, which emulates DeltaFS.
+\(90-104\times\) for Weak Consistency, which emulates DeltaFS.\\
 
 % POSIX (no stream) vs. POSIX: cost of durability < consistency
-\subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
+{\it Durability \(<<\) Consistency:} The \(1.15\times\) overhead of
 Strong Consistency compared to Strong Consistency, No Fault Tolerance for
@@
 object store is constant. We omit the same analysis for Weak Consistency
-because the runtimes are so short that the normalized slowdowns are misleading.
+because the runtimes are so short that the normalized slowdowns are misleading.\\
 
 % Weak Consistency benefit not due to metadata format
-\subsubsection{Metadata Formats} Because the metadata formats are the same for
+{\it Metadata Formats:} Because the metadata formats are the same for
 all schemes we argue that the performance gain for decoupled namespaces comes
@@
 
-\noindent\textbf{Takeaway}: CudeleFS shows the true benefit of eventual
+\noindent\textbf{Takeaway}: Cudele shows the true benefit of eventual
 consistency, where we see over a \(100\times\) slowdown for achieving strong
 consistency, in the worst case.
+
+\subsubsection{Network Bandwidth}
+% 1G - 10G
+\subsubsection{Number of Clients}
+% up to 30 clients
+\subsubsection{Number of Serialized Merges}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
+      \caption{Merging journals serially.} \label{fig:merge-a}
+  \end{subfigure}\\
+  ~
+  \begin{subfigure}[b]{\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
+      \caption{Merging a journal w/ 100K create updates.}
+      \label{fig:merge-b}
+  \end{subfigure}
+  \caption{Merging updates has linear scalability.\label{fig:merge}}
+\end{figure}
+
+Cudele serializes merge requests at the metadata server before applying the
+updates to the in-memory metadata store. Merging concurrently is an
+optimization for future work, but merges for the largest journal size we tested
+were quite fast.  Furthermore, since we place no restrictions on the validity
+of metadata inserted into the journal, we can avoid touching poorly scaling
+data structures. In other words, we can scale linearly if we weaken our
+consistency by never checking for the existence of files before creating files.
+
+Figure~\ref{fig:merge-a} demonstrates this scalability by showing the runtime
+(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
+of the runtime is over 20 serial merges, each hitting a different directory in
+the namespace.  The runtime scales linearly and with low variance because the
+``Create" mechanism that was used to create the journal only appends
+\texttt{open()} requests to the journal. When the updates are merged by the
+metadata server into the in-memory metadata store they never scan growing data
+structures. Had we implemented the ``Create" mechanism to include
+\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
+the poor scaling that we see with the ``RPCs" mechanism.
+Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
+structure because the runtimes do not increase as we add more metadata. \\
+
+\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
+({\it i.e.} not checking data structures for the validity of an update or
+existence of a file) shows linear scalability and stable performance.
 
@@
 
-\subsection{Benefits of Isolated Subtree Policies}
+\subsection{Use Case 2: Creates with Rogue Client}
+Clients create files in private directories and a separate client, which we
+call a ``rogue" client, creates files in each directory. This introduces false
+sharing and the metadata server revokes capabilities on directories touched by
+the rogue client. While HPC tries to avoid these situations with
+worklflows~\cite{batchfs, deltafs}, it stills happens in distributed file
+systems when users unintentionally access directories in a shared file system.
 
@@
 % setup: 2cs sep dirs, 2cs DN, 1c malicious write
-Next we show how CudeleFS can be programmed to block interfering clients. We
+Next we show how Cudele can be programmed to block interfering clients. We
 use the same problematic workload from Figure~\ref{fig:overhead-b}, where
@@
 
-\subsection{Scaling Serialized Merges}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
-      \caption{Merging journals serially.} \label{fig:merge-a}
-  \end{subfigure}\\
-  ~
-  \begin{subfigure}[b]{\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
-      \caption{Merging a journal w/ 100K create updates.}
-      \label{fig:merge-b}
-  \end{subfigure}
-  \caption{Merging updates has linear scalability.\label{fig:merge}}
-\end{figure}
-
-CudeleFS serializes merge requests at the metadata server before applying the
-updates to the in-memory metadata store. Merging concurrently is an
-optimization for future work, but merges for the largest journal size we tested
-were quite fast.  Furthermore, since we place no restrictions on the validity
-of metadata inserted into the journal, we can avoid touching poorly scaling
-data structures. In other words, we can scale linearly if we weaken our
-consistency by never checking for the existence of files before creating files.
-
-Figure~\ref{fig:merge-a} demonstrates this scalability by showing the runtime
-(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
-of the runtime is over 20 serial merges, each hitting a different directory in
-the namespace.  The runtime scales linearly and with low variance because the
-``Create" mechanism that was used to create the journal only appends
-\texttt{open()} requests to the journal. When the updates are merged by the
-metadata server into the in-memory metadata store they never scan growing data
-structures. Had we implemented the ``Create" mechanism to include
-\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
-the poor scaling that we see with the ``RPCs" mechanism.
-Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
-structure because the runtimes do not increase as we add more metadata. \\
-
-\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
-({\it i.e.} not checking data structures for the validity of an update or
-existence of a file) shows linear scalability and stable performance.
+\subsection{Use Case 3: Read while Writing}

### 2017-10-17 paper/sections/7_conclusion.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 %approach limits redundant code and leverages the robustness of the underlying
-%storage system. CudeleFS uses this approach and re-uses some of the building
+%storage system. Cudele uses this approach and re-uses some of the building
 %blocks from the Malacology programmable storage system~\cite{sevilla:eurosys17}
@@
 double-edged sword. While it achieves better performance and scalability, it
-alienates applications that rely on strong consistency and durability. CudeleFS
+alienates applications that rely on strong consistency and durability. Cudele
 lets users assign consistency and durability guarantees to subtrees in
 the global namespace, giving users a wide range of policies and
-optimizations that can be custom fit to the application. Using CudeleFS, we
+optimizations that can be custom fit to the application. Using Cudele, we
 compare related work on the same file system and conclude that strong

### 2017-10-17 paper/sections/1_introduction.tex
paper: draft 0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % What is the problem?
-File system metadata services in HPC have scalability problems. Administrative
-tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
-system~\cite{zheng:pdsw2014-batchfs} lead to contention for the same
-directories and inodes ({\it e.g.}, path traversal). Applications perform
-better with dedicated metadata servers~\cite{sevilla:sc15-mantle,
-ren:sc2014-indexfs} but provisioning a metadata server for every client is
-unreasonable. This problem is exacerbated by current trends in HPC, where
-architectures are transitioning from complex storage stacks with burst buffer,
-file system, object store, and tape tiers to more simplified stacks with just a
-burst buffer and object store~\cite{bent:login16-hpc-trends}; this puts more
-pressure on data access because more requests end up hitting the same layer and
-latencies cannot be hidden while data migrates across tiers.
+File system metadata services in HPC have scalability problems because
+administrative tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the
+file system~\cite{zheng:pdsw2014-batchfs}, contend for the same directories and
+inodes. Applications perform better with dedicated metadata
+servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
+metadata server for every client is unreasonable. This problem is exacerbated
+by current trends in HPC, where architectures are transitioning from complex
+storage stacks with burst buffer, file system, object store, and tape tiers to
+more simplified stacks with just a burst buffer and object
+store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
+because more requests end up hitting the same layer and latencies cannot be
+hidden while data migrates across tiers.
 
@@
 \includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
-\caption{ Clients that need better performance can relax consistency on their
-subtree ({e.g.}, HDFS) or even decouple the subtree and move it locally
-({e.g.}, BatchFS, RAMDisk). Decoupled subtrees can further relax durability
-semantics to achieve even better performance. Meanwhile, clients that require
-stronger guarantees ({\it e.g.,} POSIX) can still reside in the same namespace.
+
+\caption{For performance, clients can relax consistency on their subtree
+({e.g.}, HDFS) or even decouple the subtree and move it locally ({e.g.},
+BatchFS, RAMDisk). Decoupled subtrees can further relax durability semantics
+for even better performance. Clients that require stronger guarantees ({\it
+e.g.,} POSIX) can still reside in the same namespace.
 }\label{fig:subtree-policies}
+
 \end{figure}
@@
 (\(10\times\) slow down). We also show the effect of having a metadata specific
-file format in systems that are based on in-memory data structures. Because
-this paper focuses on file system metadata protocols, we only scale up to 20
-clients executing an extremely metdata intensive workload; this load is the
-maximum capacity of a single metadata server for the platform we chose to
-implement Cudele on.  Partitioning and replicating metadata across a cluster
-can be controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in
-this paper we show the capacity and performance of a metadata server with
-superior metadata protocols, which should be used to inform load balancing
-decisions.
+file format in systems that are based on in-memory data structures.  We scale
+up to 20 clients executing an extremely metadata intensive job because this
+workload exhausts the resources of a single metadata server for the codebase we
+built on.  Partitioning and replicating metadata across a cluster can be
+controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in this
+paper we show the capacity and performance of a metadata server with superior
+metadata protocols, which should be used to inform load balancing decisions.
 

### 2017-10-17 paper/sections/3_posix_overheads.tex
paper: draft 0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
       \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-scale.png}
       \caption{Interference hurts variability}
@@
 extra load but performance suffers.  This degradation is shown in
-Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
-increased variability. The results are normalized to a single
-isolated client and the \(y\) axis is the standard deviation of the runtime of
-all clients.  For the ``interfere" bars, each client creates files in private
-directories and at 30 seconds we launch another process that creates files in
-those directories.  For less than 7 clients, the runtime and deviation is worse
-when clients interfere. At 7 clients, the metadata server is overloaded so the
-directory caching in the ``isolated" bars has no benefit.\\
+Figure~\ref{fig:overhead-c}, where we scale the number of clients and show the
+slowdown of the slowest client.  The results are normalized to a single
+isolated client without a metadata server journal.  For the ``interfere" bars,
+each client creates files in private directories and at 30 seconds we launch
+another process that creates files in those directories. 18 clients is the
+maximum number the metadata server could handle for this metadata-intensive
+workload; at higher client load, the metadata server started complaining about
+laggy and unresponsive requests.\\
 

### 2017-10-17 paper/sections/6_results.tex
paper: draft 0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 to compose these mechanisms based on their application's correctness
-requirements and performance goals.
+requirements and performance goals.\\
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
+\caption{ [\href{https://...}{source}] Merging journals serially.}
+\label{fig:merge}
+\end{figure}
+
+%\begin{figure}[t]
+%  \centering
+%  \begin{subfigure}[b]{\linewidth}
+%      \centering
+%      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
+%      \caption{Merging journals serially.} \label{fig:merge-a}
+%  \end{subfigure}\\
+%  ~
+%  \begin{subfigure}[b]{\linewidth}
+%      \centering
+%      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
+%      \caption{Merging a journal w/ 100K create updates.}
+%      \label{fig:merge-b}
+%  \end{subfigure}
+%  \caption{Merging updates has linear scalability.\label{fig:merge}}
+%\end{figure}
+
+{\it Number of Serialized Merges:} Cudele serializes merge requests at the
+metadata server before applying the updates to the in-memory metadata store.
+Merging concurrently is an optimization for future work, but merges for the
+largest journal size we tested were quite fast.  Furthermore, since we place no
+restrictions on the validity of metadata inserted into the journal, we can
+avoid touching poorly scaling data structures. In other words, we can scale
+linearly if we weaken our consistency by never checking for the existence of
+files before creating files.
+
+Figure~\ref{fig:merge} demonstrates this scalability by showing the runtime
+(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
+of the runtime is over 20 serial merges, each hitting a different directory in
+the namespace.  The runtime scales linearly and with low variance because the
+``Create" mechanism that was used to create the journal only appends
+\texttt{open()} requests to the journal. When the updates are merged by the
+metadata server into the in-memory metadata store they never scan growing data
+structures. Had we implemented the ``Create" mechanism to include
+\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
+the poor scaling that we see with the ``RPCs" mechanism.\\
+%Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
+%structure because the runtimes do not increase as we add more metadata. \\
+
+\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
+({\it i.e.} not checking data structures for the validity of an update or
+existence of a file) shows linear scalability and stable performance.
 
@@
 
-Clients creating files in private directories is heavily
-studied in HPC~\cite{plfs, cephfs04, indexfs, giga+, batchfs, deltafs,
-sevilla:eurosys, shardfs, mantle}, mostly due to checkpoint-restart~\cite{}.
-But the workload also appears in cloud workloads, where systems like Hadoop use
-the file abstraction to exchange work units to workers or to indicate when jobs
-complete~\cite{}. A more familiar example is uncompressing an archive ({\it
-e.g.}, \texttt{tar xzf}), where the file system services a flash crowd of
-creates across all directories as shown in Figure~\ref{fig:overhead-creates}.
-We use this as a microbenchmark because it allows us to control the size of the
-log, since each create creates a single journal event.
+% CITEME xiao:socc2015-shardfs CITEME HADOOP
+Clients creating files in private directories is heavily studied in
+HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
+zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
+checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
+cloud workloads, where systems like Hadoop use the file abstraction to exchange
+work units to workers or to indicate when jobs
+complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
+uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
+services a flash crowd of creates across all directories as shown in
+Figure~\ref{fig:overhead-creates}.  We use this as a microbenchmark because it
+allows us to control the size of the log, since each create creates a single
+journal event.
 
@@
 completes}. So if servers fail during a mechanism, metadata or data may be
-lost.\\
+lost.
 
@@
 \(5-7\times\) for Strong Consistency, which emulates BatchFS and
-\(90-104\times\) for Weak Consistency, which emulates DeltaFS.\\
+\(90-104\times\) for Weak Consistency, which emulates DeltaFS.
 
@@
 object store is constant. We omit the same analysis for Weak Consistency
-because the runtimes are so short that the normalized slowdowns are misleading.\\
+because the runtimes are so short that the normalized slowdowns are misleading.
 
@@
 consistency, in the worst case.
-
-\subsubsection{Network Bandwidth}
-% 1G - 10G
-\subsubsection{Number of Clients}
-% up to 30 clients
-\subsubsection{Number of Serialized Merges}
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}[b]{\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
-      \caption{Merging journals serially.} \label{fig:merge-a}
-  \end{subfigure}\\
-  ~
-  \begin{subfigure}[b]{\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
-      \caption{Merging a journal w/ 100K create updates.}
-      \label{fig:merge-b}
-  \end{subfigure}
-  \caption{Merging updates has linear scalability.\label{fig:merge}}
-\end{figure}
-
-Cudele serializes merge requests at the metadata server before applying the
-updates to the in-memory metadata store. Merging concurrently is an
-optimization for future work, but merges for the largest journal size we tested
-were quite fast.  Furthermore, since we place no restrictions on the validity
-of metadata inserted into the journal, we can avoid touching poorly scaling
-data structures. In other words, we can scale linearly if we weaken our
-consistency by never checking for the existence of files before creating files.
-
-Figure~\ref{fig:merge-a} demonstrates this scalability by showing the runtime
-(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
-of the runtime is over 20 serial merges, each hitting a different directory in
-the namespace.  The runtime scales linearly and with low variance because the
-``Create" mechanism that was used to create the journal only appends
-\texttt{open()} requests to the journal. When the updates are merged by the
-metadata server into the in-memory metadata store they never scan growing data
-structures. Had we implemented the ``Create" mechanism to include
-\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
-the poor scaling that we see with the ``RPCs" mechanism.
-Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
-structure because the runtimes do not increase as we add more metadata. \\
-
-\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
-({\it i.e.} not checking data structures for the validity of an update or
-existence of a file) shows linear scalability and stable performance.
 
@@
 \subsection{Use Case 2: Creates with Rogue Client}
+
 Clients create files in private directories and a separate client, which we
@@
 the rogue client. While HPC tries to avoid these situations with
-worklflows~\cite{batchfs, deltafs}, it stills happens in distributed file
-systems when users unintentionally access directories in a shared file system.
+worklflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it stills
+happens in distributed file systems when users unintentionally access
+directories in a shared file system.
 
@@
 \subsection{Use Case 3: Read while Writing}
+
+% CITEME
+Scientists use the file system to check the progress of jobs using
+\texttt{ls}~\cite{CITEME}. The number of files or size of the files is
+indicative of the progress. This practice is not too different from systems
+that use the file system to manage the progress of jobs; Hadoop writes to
+temporary files, renames them when complete, and creates a ``DONE" file to
+indicate to the runtime that the task did not fail and should not be
+re-scheduled on another node. In this scenario, Cudele users will not see the
+progress of decoupled namespace since their updates are not globally visible.
+To help scientists judge the progress of their jobs, Cudele has a ``namespace
+sync" that sends batches of updates back to the global namespace at regular
+intervals.
+
+\subsection{Use Case 4: Reading Large Directories}
+
+% CITEME
+The final uses case is reading large directories. At job completion, scientists
+might use \texttt{ls} again to see the results. This causes great strain on the
+file system as paths need to traversed and the entire directory, with all its
+entries, must be transferred back to the client. Here we show how decoupling a
+large namespace and materializing the view in memory on the client is faster
+than doing RPCs for walks of the file system namespace.

### 2017-10-17 paper/sections/1_introduction.tex
paper: draft 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 architectures.  The HDFS\footnote{This is example is not directly evaluated in
-this paper, although a similiar use case is explored in~\S\ref{use-case-1}}
+this paper, although a similar use case is explored in~\S\ref{use-case-1}}
 subtree has weaker consistency than strong consistency because it  does not
 guarantee that all updates are not immediately seen by all servers; the BatchFS
-and RAMDisk subtrees are decoupled from the global namespace and have similiar
+and RAMDisk subtrees are decoupled from the global namespace and have similar
 consistency/durability behavior to those systems; and the POSIX subtree retains
@@
 
-Our protoype system, Cudele, achieves this by exposing ``mechanisms" that
+Our prototype system, Cudele, achieves this by exposing ``mechanisms" that
 developers use to specify their preferred semantics.  Cudele supports 3 forms
@@
 up to 20 clients executing an extremely metadata intensive job because this
-workload exhausts the resources of a single metadata server for the codebase we
+workload exhausts the resources of a single metadata server for the code base we
 built on.  Partitioning and replicating metadata across a cluster can be

### 2017-10-17 paper/sections/3_posix_overheads.tex
paper: draft 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+% CITEME PDSW paper
 In our examination of the overheads of POSIX we benchmark and analyze CephFS,
-the file system that uses the Ceph's object store ({\it i.e.} RADOS) to store
-its data and metadata. We choose CephFS because it is an open-source production
-quality system. This file system is an implementation of one set of design
-decisions and our goal is to highlight the effect that those decisions have on
-performance. During this process we discovered, based on the analysis and
-breakdown of costs, that durability and consistency have high overhead.
+the file system that uses Ceph's object store ({\it i.e.} RADOS) to store its
+data and metadata. Ceph was originally designed for HPC and has proven to be
+useful for HPC workloads~\cite{}. Several features were motivated by HPC
+requirements~\cite{weil:osdi2006-ceph}, so the lessons are useful (and can be
+implemented) in HPC environments.  This file system is an implementation of one
+set of design decisions and our goal is to highlight the effect that those
+decisions have on performance. During this process we discovered, based on the
+analysis and breakdown of costs, that durability and consistency have high
+overhead.
 

### 2017-10-17 paper/sections/6_results.tex
paper: draft 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-We evaluate Cudele on a 15 node cluster, partitioned into 8 object storage
-servers, 3 metadata servers, and 2 monitor servers. The object storage servers
-double as clients which is fine because clients are CPU and memory bound while
-object storage servers are disk IO bound. All daemons run as a single process
-which is the default setting for Ceph and the nodes have 2 dual core 2GHz
-processors with 8GB of RAM. There are three daemons per object storage server
-(one for each disk formatted with XFS) and they share an SSD for the journal.
-The nodes are running Ubuntu 12.04.4, kernel version 3.2.0-63 but all
-experiments run in Docker containers; this makes it easier to tear down and
-re-initialize the cluster ({\it e.g.}, dropping the kernel cache) between
-experiments. For these experiments, we compare CephFS to our programmable
-consistency and durability modifications, which we call Cudele.
+We did not compare to other HPC systems because we are not claiming that our
+file system metadata protocols are superior. Instead, we use Ceph as a platform
+for exploring related work over the same storage system, which facilitates an
+apples-to-apples comparison of the strategies themselves.  So we run
+experiments on the same cluster and normalize results to control for hardware
+differences and to make our results more generally applicable.  Our goal is not
+to build a file system optimized for HPC workloads and our prototype is not a
+competitor to Lustre. Rather it is a solution for letting the techniques of a
+system like Lustre live in the same namespace as techniques optimized for file
+creates, like BatchFS.
+
+We evaluate our prototype, which we call Cudele, on two clusters: a 15 node
+in-house cluster with 1 monitor server, 1 metadata server, 8 object storage
+servers, and 8 clients (co-located with object servers); and a 34 node CloudLab
+cluster with 1 monitor server, 1 metadata server, 3 object storage servers, and
+up to 29 clients. For the in-house cluster, We need more object storage servers
+and we have to co-locate clients on the same node because of resource
+contention \footnote{The bandwidth to RADOS is only sufficient with 8 nodes on
+this older hardware, which uses SSDs and nodes that are almost 10 years old.
+The CloudLab cluster performs just as well with and without a highly
+provisioned object storage server cluster.}. All daemons run as a single
+process which is the default setting for Ceph. We develop on Ceph's Jewel
+release, version 10.2.1, which was released in May 2016.
+
+%The in-house cluster has  2 dual core 2GHz processors with 8GB of RAM wh
+%three daemons per object storage server (one for each disk formatted with XFS)
+%and they share an SSD for the journal.  The nodes are running Ubuntu 12.04.4,
+%kernel version 3.2.0-63 but all experiments run in Docker containers; this
+%makes it easier to tear down and re-initialize the cluster ({\it e.g.},
+%dropping the kernel cache) between experiments.
 

### 2017-10-18 experiments/cudele-mechanisms/run.sh
benchmark: more mechanism results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -35,13 +35,13 @@ if [ ! -z $1 ]; then
 fi
 
 for run in 0 1 2; do
-  for nfiles in 10 100 1000 10000 100000 1000000 10000000; do
+  for nfiles in 10 100 1000 10000 100000 200000 400000 600000 800000 1000000 10000000; do
     mkdir results || true
     ./teardown.sh
     cp configs_$SITE/hosts hosts
     $DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
   done
-  mv results results-$SITE-clients$clients-run$run
+  mv results results-$SITE-clients$clients-run$run-parmsweep
 done
 
 exit 0

### 2017-10-18 paper/sections/4_methodology.tex
paper: draft 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: more eval organization
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{Methodology: Global Namespace with Per-Subtree Consistency/Durability}
+\section{Methodology: Global Namespace, Subtree Consistency/Durability}
 \label{sec:methodology-decoupled-namespaces}
@@
 \end{tabular}
-\caption{The Cudele mechanisms are composed together to form consistency and
+\caption{Mechanisms are composed together to form consistency and
 durability semantics.\label{table:mechanisms}} 
@@
 
-In this section we describe Cudele, our file system that lets users
+In this section we describe Cudele, our API and framework that lets users
 assign consistency and durability semantics to subtrees in the global
@@
 
-\subsection{The Cudele Mechanisms}
+\subsection{Mechanisms: Building Guarantees}
 \label{sec:the-cudelesfs-mechanisms}
@@
 \caption{Users can explore the consistency (C) and
-durability (D) spectrums by composing Cudele mechanisms. 
+durability (D) spectrum by composing Cudele mechanisms. 
 \label{table:spectrum}}

### 2017-10-18 paper/sections/6_results.tex
paper: more eval organization
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-We evaluate our prototype, which we call Cudele, on two clusters: a 15 node
-in-house cluster with 1 monitor server, 1 metadata server, 8 object storage
-servers, and 8 clients (co-located with object servers); and a 34 node CloudLab
-cluster with 1 monitor server, 1 metadata server, 3 object storage servers, and
-up to 29 clients. For the in-house cluster, We need more object storage servers
-and we have to co-locate clients on the same node because of resource
-contention \footnote{The bandwidth to RADOS is only sufficient with 8 nodes on
-this older hardware, which uses SSDs and nodes that are almost 10 years old.
-The CloudLab cluster performs just as well with and without a highly
-provisioned object storage server cluster.}. All daemons run as a single
-process which is the default setting for Ceph. We develop on Ceph's Jewel
-release, version 10.2.1, which was released in May 2016.
+\begin{table}
+\begin{tabular}{ r | l | l}
+  Cluster  & Hardware               & Software \\\hline
+  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
+           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
+  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
+           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
+\end{tabular}
+\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
+monitor daemon, MDS is metadata server daemon, and OSDs are object storage
+daemons.\label{table:clusters}} 
+\end{table}
+
+We evaluate our prototype on the two clusters in Table~\ref{table:clusters}.
+Microbenchmarks were run on the in-house clusters while client scalability
+tests were run on CloudLab. We only scale up to 29 clients because that is the
+maximum capacity of a single metadata server. For the in-house cluster, we need
+more object storage servers and we have to co-locate clients on the same node
+because of resource contention \footnote{The bandwidth to RADOS is only
+sufficient with 8 nodes on this older hardware, which uses SSDs and nodes that
+are almost 10 years old.  The CloudLab cluster performs just as well with and
+without a highly provisioned object storage server cluster.}. All daemons run
+as a single process which is the default setting for Ceph. We develop on Ceph's
+Jewel release, version 10.2.1, which was released in May 2016.
 
@@
 progress of decoupled namespace since their updates are not globally visible.
-To help scientists judge the progress of their jobs, Cudele has a ``namespace
-sync" that sends batches of updates back to the global namespace at regular
-intervals.
+To help scientists judge the progress of their jobs, Cudele clients have a
+``namespace sync" that sends batches of updates back to the global namespace at
+regular intervals.
+
+% Does this implementation need to be described in the implementation?
+FigureX shows the performance degradation of a single client writing updates to
+a decoupled namespace and pausing to send updates to the metadata server. We
+scale the namespace sync interval to show the trade-off of frequently pausing
+or writing large logs of updates.  We use an idle core to log the updates and
+to do the network transfer. The client only pauses to fork off a background
+process, which is expensive as the address space needs to copied. The
+alternative is to pause the client completely and write the update to disk but
+since this implementation is limited by the speed of the disk, we choose the
+memory-to-memory copy of the fork appraoch.`
+
+To show the contention at the metadata server, we scale the number of
+concurrent journal merge events in FigureX. The size of each journal is 100K
+files, the maximum recommended directory size for CephFS. We also plot the cost
+of RPCs with and without journaling.
 
@@
 % CITEME
-The final uses case is reading large directories. At job completion, scientists
+The final use case is reading large directories. At job completion, scientists
 might use \texttt{ls} again to see the results. This causes great strain on the
@@
 than doing RPCs for walks of the file system namespace.
+
+FigureX shows how fast a client can materialize decoupled namespaces in memory.
+This takes the on-disk file system metadata format and parses them into events
+that can be manipulated and appended to. Again, we scale the number of up

### 2017-10-19 experiments/cudele-partialreads/run.sh
benchmark: partial reads
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,48 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+#for run in 0 1 2; do
+  #for nfiles in 10 100 1000 10000 100000 200000 400000 600000 800000 1000000 10000000; do
+  for nfiles in 1000000; do
+    mkdir results || true
+    ./teardown.sh
+    cp configs_$SITE/hosts hosts
+    $DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
+  done
+  mv results results-$SITE
+#done
+
+exit 0

### 2017-10-19 experiments/cudele-mergescale/docker/cloudlab/pushkeys.sh
benchmark: partial updates complete
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+set -ex
+
+for i in $CLOUDLAB_NODES; do
+  echo "---- " $i
+  scp ~/.ssh/id_rsa.pub ${USER}@$i:~/.ssh/id_rsa.pub
+  scp ~/.ssh/id_rsa ${USER}@$i:~/.ssh/id_rsa
+  ssh ${USER}@$i "sudo usermod -aG docker ${USER}"
+done
+
+set +x
+echo "DONE!"

### 2017-10-19 experiments/cudele-mergescale/run.sh

@@ -1,0 +1,49 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+#for run in 0 1 2; do
+  #for nfiles in 10 100 1000 10000 100000 200000 400000 600000 800000 1000000 10000000; do
+  for nfiles in 10; do
+    mkdir results || true
+    #./teardown.sh
+    cp configs_$SITE/hosts hosts
+    #$DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
[…]

### 2017-10-19 experiments/cudele-mergescale/teardown.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+# Put all your cleanup tasks here.
+set -ex
+
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml -e @/popper/ansible/vars.yml -i /etc/ansible/hosts"
+ROOT=`dirname $PWD | xargs dirname`
+docker run -it --rm \
+  --net host \
+  -v $HOME/.ssh:/root/.ssh \
+  -v `pwd`/ansible/group_vars:/root/group_vars \
+  -v `pwd`/hosts:/etc/ansible/hosts \
+  -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+  -v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+  -v `pwd`:/popper \
+  -v $ROOT/ansible/ceph:/root -w /root \
+  -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+  -e ANSIBLE_CONFIG="/etc/ansible/ansible.cfg" \
+  michaelsevilla/ansible $ARGS $VARS /popper/ansible/cleanup.yml
+  #--entrypoint=/bin/bash michaelsevilla/ansible -c "$@"
+ 
+exit 0

### 2017-10-19 paper/sections/6_results.tex
paper: draft 2 with partial reads
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+\begin{figure}[tb] \centering
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png} \caption{
+[\href{https://...}{source}] Slowdown of a single client periodically syncing
+updates to the gobal namespace. The crossover point represents the trade-off of
+frequent updates of small journal files and infrequent updates of large journal
+files.\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}} \end{figure}
+
 % Does this implementation need to be described in the implementation?
-FigureX shows the performance degradation of a single client writing updates to
-a decoupled namespace and pausing to send updates to the metadata server. We
-scale the namespace sync interval to show the trade-off of frequently pausing
-or writing large logs of updates.  We use an idle core to log the updates and
-to do the network transfer. The client only pauses to fork off a background
-process, which is expensive as the address space needs to copied. The
-alternative is to pause the client completely and write the update to disk but
-since this implementation is limited by the speed of the disk, we choose the
-memory-to-memory copy of the fork appraoch.`
+Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
+client writing updates to a decoupled namespace and pausing to send updates to
+the metadata server. We scale the namespace sync interval to show the trade-off
+of frequently pausing or writing large logs of updates.  We use an idle core to
+log the updates and to do the network transfer. The client only pauses to fork
+off a background process, which is expensive as the address space needs to
+copied. The alternative is to pause the client completely and write the update
+to disk but since this implementation is limited by the speed of the disk, we
+choose the memory-to-memory copy of the fork appraoch.`
 

### 2017-10-19 experiments/cudele-partialreads/run.sh
paper: run partial reads multiple times
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,15 +34,16 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-#for run in 0 1 2; do
+#./teardown.sh
+#$DOCKER ceph.yml
+for run in 0 1 2 3 4 5 6 7 8 9 10; do
   #for nfiles in 10 100 1000 10000 100000 200000 400000 600000 800000 1000000 10000000; do
   for nfiles in 1000000; do
     mkdir results || true
-    ./teardown.sh
     cp configs_$SITE/hosts hosts
-    $DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
+    $DOCKER -e nfiles=$nfiles /workloads/vapply.yml
   done
-  mv results results-$SITE
-#done
+  mv results results-$SITE-run$run
+done
 
 exit 0

### 2017-10-19 paper/sections/6_results.tex
paper: add graph for partial updates
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 updates to the gobal namespace. The crossover point represents the trade-off of
-frequent updates of small journal files and infrequent updates of large journal
-files.\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}} \end{figure}
+frequent updates (with small journal files) and infrequent updates (with larger journal
+files).\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}} \end{figure}
 
@@
 to disk but since this implementation is limited by the speed of the disk, we
-choose the memory-to-memory copy of the fork appraoch.`
+choose the memory-to-memory copy of the fork approach.
 
-To show the contention at the metadata server, we scale the number of
-concurrent journal merge events in FigureX. The size of each journal is 100K
-files, the maximum recommended directory size for CephFS. We also plot the cost
-of RPCs with and without journaling.
+To show the contention at the metadata server of this approach, we scale the
+number of concurrent journal merge events in FigureX. The size of each journal
+is 100K files, the maximum recommended directory size for CephFS. We also plot
+the cost of RPCs with and without journaling.
 

### 2017-10-20 experiments/cudele-mergescale/ansible/workloads/generate_inos.py
benchmark: scale merge fails for 10 clients
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,11 @@
+#!/usr/bin/python
+
+inos = []
+# default start ino 1099511627780
+start = 1099511627780 + 100000
+for i in range(0, 30):
+  inos.append(i*100000000 + start)
+
+for i in inos:
+  print i
+print inos 

### 2017-10-20 experiments/cudele-mergescale/ansible/workloads/parallel_merge.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+if [ $# -lt 2 ]; then
+  echo "USAGE: $0 <mds name> <number of journals>"
+  exit 1
+fi
+
+set -e
+for i in `seq 0 $2`; do
+  ceph daemon mds.$1 merge /etc/ceph/events-${i}.bin &
+done
+
+echo "waiting..."
+wait

### 2017-10-20 experiments/cudele-mergescale/run.sh

@@ -35,15 +35,15 @@ if [ ! -z $1 ]; then
 fi
 
 #for run in 0 1 2; do
-  #for nfiles in 10 100 1000 10000 100000 200000 400000 600000 800000 1000000 10000000; do
-  for nfiles in 10; do
+  for nclients in 10 15 20 25 30; do
     mkdir results || true
-    #./teardown.sh
+    ./teardown.sh
     cp configs_$SITE/hosts hosts
-    #$DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
-    $DOCKER -e nfiles=$nfiles /workloads/vapply.yml
+    $DOCKER -e nfiles=100000 ceph.yml /workloads/clients${nclients}.yml
+    #$DOCKER -e nfiles=$nfiles /workloads/vapply.yml
+    break
+    mv results results-$SITE-${nclients}clients
   done
-  mv results results-$SITE
 #done
 
 exit 0

### 2017-10-20 experiments/cudele-mergescale/ansible/workloads/parallel_merge.sh
benchmark: scale merges
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,13 +1,14 @@
 #!/bin/bash
 
-if [ $# -lt 2 ]; then
-  echo "USAGE: $0 <mds name> <number of journals>"
+if [ $# -lt 3 ]; then
+  echo "USAGE: $0 <mds name> <number of journals> <interval>"
   exit 1
 fi
 
 set -e
 for i in `seq 0 $2`; do
-  ceph daemon mds.$1 merge /etc/ceph/events-${i}.bin &
+  ceph daemon mds.$1 merge /etc/ceph/events-${i}.bin
+  sleep $3
 done
 
 echo "waiting..."

### 2017-10-20 experiments/cudele-mergescale/run.sh

@@ -34,16 +34,11 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-#for run in 0 1 2; do
-  for nclients in 10 15 20 25 30; do
-    mkdir results || true
-    ./teardown.sh
-    cp configs_$SITE/hosts hosts
-    $DOCKER -e nfiles=100000 ceph.yml /workloads/clients${nclients}.yml
-    #$DOCKER -e nfiles=$nfiles /workloads/vapply.yml
-    break
-    mv results results-$SITE-${nclients}clients
-  done
-#done
+mkdir results || true
+for nclients in 30 25 20 15 10 5; do
+  ./teardown.sh
+  cp configs_$SITE/hosts hosts
+  $DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/merge.yml #/workloads/clients${nclients}.yml
+done
 
 exit 0

### 2017-10-20 paper/sections/6_results.tex
paper: add some figs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
-\caption{
-[\href{https://...}{source}]
-The performance of the Cudele mechanisms normalized to the runtime of the
-``Append Client Journal" mechanism (the runtime of writing \(n\) file creates
-to the client's in-memory journal).  \label{fig:slowdown-mechanisms}}
+\caption{ [\href{https://...}{source}] Microbenchmark: performance of each
+mechanism. Results are normalized to the runtime of the ``Append Client
+Journal" mechanism (the runtime of writing \(n\) file creates to the client's
+in-memory journal).  \label{fig:slowdown-mechanisms}}
 \end{figure}
@@
 \includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
-\caption{
-[\href{https://...}{source}]
-The size of the client's journal scales with the number of
+\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
+client. The size of the client's journal scales with the number of
 updates.\label{fig:behavior-journal-size}}
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
-\caption{ [\href{https://...}{source}] Merging journals serially.}
-\label{fig:merge}
-\end{figure}
-
-%\begin{figure}[t]
-%  \centering
-%  \begin{subfigure}[b]{\linewidth}
-%      \centering
-%      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
-%      \caption{Merging journals serially.} \label{fig:merge-a}
-%  \end{subfigure}\\
-%  ~
-%  \begin{subfigure}[b]{\linewidth}
-%      \centering
-%      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
-%      \caption{Merging a journal w/ 100K create updates.}
-%      \label{fig:merge-b}
-%  \end{subfigure}
-%  \caption{Merging updates has linear scalability.\label{fig:merge}}
-%\end{figure}
-
-{\it Number of Serialized Merges:} Cudele serializes merge requests at the
-metadata server before applying the updates to the in-memory metadata store.
-Merging concurrently is an optimization for future work, but merges for the
-largest journal size we tested were quite fast.  Furthermore, since we place no
-restrictions on the validity of metadata inserted into the journal, we can
-avoid touching poorly scaling data structures. In other words, we can scale
-linearly if we weaken our consistency by never checking for the existence of
-files before creating files.
-
-Figure~\ref{fig:merge} demonstrates this scalability by showing the runtime
-(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
-of the runtime is over 20 serial merges, each hitting a different directory in
-the namespace.  The runtime scales linearly and with low variance because the
-``Create" mechanism that was used to create the journal only appends
-\texttt{open()} requests to the journal. When the updates are merged by the
-metadata server into the in-memory metadata store they never scan growing data
-structures. Had we implemented the ``Create" mechanism to include
-\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
-the poor scaling that we see with the ``RPCs" mechanism.\\
-%Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
-%structure because the runtimes do not increase as we add more metadata. \\
-
-\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
-({\it i.e.} not checking data structures for the validity of an update or
-existence of a file) shows linear scalability and stable performance.
-
 \subsection{Use Case 1: Creates in the Same Directory}
@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
-\caption{
-[\href{https://...}{source}]
-The RPC per metadata update of ``Strong Consistency" has a large
-overhead compared to the decoupled namespace strategy of ``Weak
-Consistency".\label{fig:slowdown-strong-weak}}
+\caption{ [\href{https://...}{source}] Use Case 1: the RPC per metadata update
+(strong consistency) has a large overhead compared to decoupled namespaces
+(weak consistency.)\label{fig:slowdown-strong-weak}}
 \end{figure}
@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
-\caption{
-[\href{https://...}{source}]
-Using the ``allow" and ``block" API, users can isolate directories from
-interfering clients. Variance with blocking turned on is the same as
-``isolated" from Figure~\ref{fig:overhead-b}.
+\caption{ [\href{https://...}{source}] Use Case 2: using the ``allow" and
+``block" API, users can isolate directories from interfering clients. Variance
+with blocking turned on is the same as ``isolated" from
+Figure~\ref{fig:overhead-b}.
 \label{fig:slowdown-allow-block}}
+
 \end{figure}
@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png} \caption{
-[\href{https://...}{source}] Slowdown of a single client periodically syncing
+[\href{https://...}{source}] Use Case 3: slowdown of a single client periodically syncing
 updates to the gobal namespace. The crossover point represents the trade-off of
@@
 To show the contention at the metadata server of this approach, we scale the
-number of concurrent journal merge events in FigureX. The size of each journal
-is 100K files, the maximum recommended directory size for CephFS. We also plot
-the cost of RPCs with and without journaling.
+number of journal merge events, each hitting different directories
+in the namespace, in FigureX. The size of each journal is 100K files, the
+maximum recommended directory size for CephFS. We also plot the cost of RPCs
+with and without journaling.
+
+Merge requests are serialized at the metadata server; merging concurrently is
+an optimization for future work. The current implementation has race conditions
+because the metadata server recovery code was not designed to run in parallel
+({\it i.e.} the metadata server never recovers with multiple threads). As a
+result, the metadata server versions inodes and directory entries to ensure
+that the recovered metadata is consistent.
+
+Performance is better than RPCs because we place no restrictions on the
+validity of metadata inserted into the journal and avoid touching poorly
+scaling data structures. In other words, we can scale linearly if we weaken our
+consistency by never checking for the existence of files before creating files.
+only appends \texttt{open()} requests to the journal. When the updates are
+merged by the metadata server into the in-memory metadata store they never scan
+growing data structures.  Had we implemented the ``Create" mechanism to include
+\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
+the poor scaling that we see with the ``RPCs" mechanism.\\
+
+\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
+({\it i.e.} not checking data structures for the validity of an update or
+existence of a file) shows linear scalability and stable performance.
 

### 2017-10-20 experiments/cudele-mergescale/run.sh
benchmark: finish scalable merge
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -35,7 +35,7 @@ if [ ! -z $1 ]; then
 fi
 
 mkdir results || true
-for nclients in 30 25 20 15 10 5; do
+for nclients in 30 25 20 15 10 5 1; do
   ./teardown.sh
   cp configs_$SITE/hosts hosts
   $DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/merge.yml #/workloads/clients${nclients}.yml

### 2017-10-20 experiments/cudele-mergescale/visualize/jupyter.sh
paper: add some slowdown merge scale graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,4 +1,4 @@
 #!/bin/bash
 
 set -e -x
-docker run --rm -v `dirname $PWD`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93
+docker run --rm -v  `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93

### 2017-10-21 paper/sections/6_results.tex
paper: draft 3 with merge scale
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 frequent updates (with small journal files) and infrequent updates (with larger journal
-files).\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}} \end{figure}
+files).\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}}
+\end{figure}
+
+\begin{figure}[tb] \centering
+\includegraphics[width=1.0\linewidth]{graphs/mergescale.png} \caption{
+[\href{https://...}{source}] Use Case 3: performance of merging client journals
+at the metadata server. Performance is better than RPCs because (1) there are
+less messages on the network and (2) the consistency/durability code paths in
+the metadata server are bypassed. Merges are serialized and read from the local
+disk, so keeping client journals in memory or on SSD will improve the
+performance of ``decoupled" even more.  \ref{fig:mergescale}.
+\label{fig:mergescale-sync}}
+\end{figure}
 
@@
 To show the contention at the metadata server of this approach, we scale the
-number of journal merge events, each hitting different directories
-in the namespace, in FigureX. The size of each journal is 100K files, the
-maximum recommended directory size for CephFS. We also plot the cost of RPCs
-with and without journaling.
+number of journal merge events, each hitting different directories in the
+namespace, in Figure~\ref{fig:mergescale}. The size of each journal is 100K
+files, the maximum recommended directory size for CephFS. We also plot the cost
+of RPCs with and without journaling; recall from Figure~\ref{fig:overhead-c}
+that the scalability that max number of clients that the default metadata
+server could support with RPCs is 18.
 
@@
 that can be manipulated and appended to. Again, we scale the number of up
+
+\subsection{Discussion}
+\begin{itemize}
+  \item Composable guarantees: applications control their performance (knobs you didn't have before)
+  \item Hardware features: adding hardware in strategic places
+  \item Performance Prediction: quantifies the cost of each mechanism
+\end{itemize}

### 2017-10-21 paper/sections/6_results.tex
paper: fix microbenchmark section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-%The in-house cluster has  2 dual core 2GHz processors with 8GB of RAM wh
-%three daemons per object storage server (one for each disk formatted with XFS)
-%and they share an SSD for the journal.  The nodes are running Ubuntu 12.04.4,
-%kernel version 3.2.0-63 but all experiments run in Docker containers; this
-%makes it easier to tear down and re-initialize the cluster ({\it e.g.},
-%dropping the kernel cache) between experiments.
-
 This paper adheres to The Popper
@@
 
-%\begin{figure*}[tb]
-%\centering
-%\includegraphics[width=180mm]{graphs/slowdown.png}
-%\caption{Compared to the \texttt{create} phase, saving and persisting
-%updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
-%a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
-%In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
-%slower.  The disadvantage of decoupling the namespace is the merge phase where
-%updates are applied to the metadata store (\texttt{create+apply}, resulting in a
-%905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
-%\end{figure*}
-
 \subsection{Microbenchmarks}
+\label{sec:microbenchmarks}
 \begin{figure}[tb]
 \centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
+\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
 \caption{ [\href{https://...}{source}] Microbenchmark: performance of each
-mechanism. Results are normalized to the runtime of the ``Append Client
-Journal" mechanism (the runtime of writing \(n\) file creates to the client's
-in-memory journal).  \label{fig:slowdown-mechanisms}}
+mechanism (left) and building consistency/durability guarantees (right) for
+100K files on a single client. Results are normalized to the runtime of writing
+files creates to the client's in-memory journal.
+\label{fig:composable-mechanisms}}
 \end{figure}
@@
 
-Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the Cudele
+Figure~\ref{fig:composable-mechanisms} shows the runtime of the Cudele
 mechanisms for a single client creating files in the same directory, normalized
-to the time it takes to write \(n\) file create updates to the client's
-in-memory journal ({\it i.e.} the ``Append Client Journal" mechanism). Bars
-above \(10^0\) are slower than the ``Append Client Journal" mechanism.
+to the time it takes to write 100K file create updates to the client's
+in-memory journal ({\it i.e.} the ``Append Client Journal" mechanism).
 ``Stream" is an approximation of the overhead and is calculated by subtracting
 the runtime of the job with the journal turned off from the runtime with the
-journal turned on.  The largest workload we tested is 100K file creates in the
-same directory.  This is the maximum recommended size of a directory in CephFS;
-preliminary experiments with larger directory sizes show memory problems.
+journal turned on.  100K is the maximum recommended size of a directory in
+CephFS; preliminary experiments with larger directory sizes show memory
+problems.
 
-{\it Poorly Scaling Data Structures:} Despite doing the same amount of
-work, mechanisms that rely on poorly scaling data structures have large
-slowdowns for the larger number of creates. For example, ``RPCs" which relies
-on the internal CephFS directory structures, goes from a \(40\times\) slowdown
-for 10K files to a \(90\times\) slowdown for 100K files. It is a well-known
-problem that directory data structures do not scale when creating files in the
-same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
-structures will experience similar slowdowns. Other mechanisms that write
-events to a journal ({\it e.g.} the persists, ``Volatile Apply") experience a
-much less drastic slowdown because the journal data structure does not need to
-be scanned for every operation. Events are written to the end of the journal without
-even checking the validity ({\it e.g.}, if the file already exists for a create),
-which is another form of relaxed consistency because the file system assumes the
-application has resolved conflicting updates in a different way.
+{\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
+mechanisms that rely on poorly scaling data structures have large slowdowns for
+the larger number of creates. For example, ``RPCs" which relies on the internal
+CephFS directory structures has a \(90\times\) slowdown for 100K files. It is a
+well-known problem that directory data structures do not scale when creating
+files in the same directory~\cite{ren:sc2014-indexfs} and any mechanism that
+uses these data structures will experience similar slowdowns. Other mechanisms
+that write events to a journal ({\it e.g.} the persists, ``Volatile Apply")
+experience a much less drastic slowdown because the journal data structure does
+not need to be scanned for every operation. Events are written to the end of
+the journal without even checking the validity ({\it e.g.}, if the file already
+exists for a create), which is another form of relaxed consistency because the
+file system assumes the application has resolved conflicting updates in a
+different way.
 
@@
 persist" demonstrates the bandwidth advantages of storing the journal in a
-distributed object store. For 100K file creates, the ``Global Persist"
+distributed object store. The ``Global Persist"
 performance is \(1.5\times\) faster because the object store is leveraging the
@@
 
-\noindent\textbf{Takeaway}: the Cudele mechanisms have overheads and costs
-that can differ {\it by orders of magnitude}. Cudele gives users the ability
-to compose these mechanisms based on their application's correctness
-requirements and performance goals.\\
+\noindent\textbf{Takeaway}: measuring the mechanisms individually shows that
+their overheads and costs can differ {\it by orders of magnitude}. Cudele gives
+users the ability to quantify the performance impact of different
+consistency/durability guarantees.\\
 
 \subsection{Use Case 1: Creates in the Same Directory}
-\label{use-case-1}
+\label{sec:use-case-1}
 
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
-\caption{ [\href{https://...}{source}] Use Case 1: the RPC per metadata update
-(strong consistency) has a large overhead compared to decoupled namespaces
-(weak consistency.)\label{fig:slowdown-strong-weak}}
-\end{figure}
+%\begin{figure}[tb]
+%\centering
+%\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
+%\caption{ [\href{https://...}{source}] Use Case 1: the RPC per metadata update
+%(strong consistency) has a large overhead compared to decoupled namespaces
+%(weak consistency.)\label{fig:slowdown-strong-weak}}
+%\end{figure}
 
-Figure~\ref{fig:slowdown-strong-weak} shows the runtime of systems employing
-weak and strong consistency, normalized to the runtime of the ``Append Client Journal"
-mechanism (again, just creating files in the client's in-memory journal).  We
-use the following compositions from the mechanisms in
-Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
-Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal" +
-``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal" +
-``Local Persist" + ``Volatile Apply".
+The graph on the right of Figure~\ref{fig:composable-mechanisms} shows how
+applications can compose mechanisms together to get the consistency/durability
+guarantees they need.  We label the \(x\)-axis with systems that employ these
+semantics, as described in Figure~\ref{fig:subtree-policies}.  Again, the
+runtime of different semantics for subtrees in the same global namespace
+normalized to the runtime of the ``Append Client Journal" mechanism (again,
+just creating files in the client's in-memory journal).  We makes no guarantees
+during execution of the mechanisms or when transitioning semantics -- the
+semantics are guaranteed {\it once the mechanism completes}. So if servers fail
+during a mechanism, metadata or data may be lost.
 
-% Final file system metadata states are equivalent No guarantees while
-% transitioning mechanisms
-We compare these semantics because the final metadata states are equivalent.
-Cudele makes no guarantees during execution of the mechanisms or when
-transitioning semantics -- the semantics are guaranteed {\it once the mechanism
-completes}. So if servers fail during a mechanism, metadata or data may be
-lost.
+%% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
+%% (5-7x) vs. (90-104x) slower
+%% Does not scale?
+%% Strong Consistency > 10X slower than Weak Consistency 
+%{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
+%decoupled namespace strategy and shows up to a \(20\times\) speedup over the
+%traditional namespaces that use RPCs. Compared to the baseline the slowdown is
+%\(5-7\times\) for Strong Consistency, which emulates BatchFS and
+%\(90-104\times\) for Weak Consistency, which emulates DeltaFS.
 
-% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
-% (5-7x) vs. (90-104x) slower
-% Does not scale?
-% Strong Consistency > 10X slower than Weak Consistency 
-{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
-decoupled namespace strategy and shows up to a \(20\times\) speedup over the
-traditional namespaces that use RPCs. Compared to the baseline the slowdown is
-\(5-7\times\) for Strong Consistency, which emulates BatchFS and
-\(90-104\times\) for Weak Consistency, which emulates DeltaFS.
-
-% POSIX (no stream) vs. POSIX: cost of durability < consistency
-{\it Durability \(<<\) Consistency:} The \(1.15\times\) overhead of
-Strong Consistency compared to Strong Consistency, No Fault Tolerance for
-100K files is negligible. It suggests that the overhead of consistency is much
-larger than the overhead of durability. This conclusion should be stronger as
-we scale the number of files because the cost of streaming the journal into the
-object store is constant. We omit the same analysis for Weak Consistency
-because the runtimes are so short that the normalized slowdowns are misleading.
+{\it Durability \(<<\) Consistency:} The performance of these semantics
+suggests that the overhead of consistency is much larger than the overhead of
+durability. This conclusion should be stronger as we scale the number of files
+because the cost of streaming the journal into the object store is constant.
 
@@
 
-\noindent\textbf{Takeaway}: Cudele shows the true benefit of eventual
-consistency, where we see over a \(100\times\) slowdown for achieving strong
-consistency, in the worst case.
+\noindent\textbf{Takeaway}: we confirm the performance benefits of other
+well-established research systems but, more importantly, we show that the
+Cudele mechanisms we propose are useful for building evaluating different
+consistency/durability semantics.
 

### 2017-10-22 paper/sections/1_introduction.tex
paper: figure 1 replace client with semantics
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 applications optimized with techniques from different state-of-the-art HPC
-architectures.  The HDFS\footnote{This is example is not directly evaluated in
-this paper, although a similar use case is explored in~\S\ref{use-case-1}}
-subtree has weaker consistency than strong consistency because it  does not
-guarantee that all updates are not immediately seen by all servers; the BatchFS
-and RAMDisk subtrees are decoupled from the global namespace and have similar
-consistency/durability behavior to those systems; and the POSIX subtree retains
-the rigidity of POSIX's strong consistency.
+architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
+paper, although the HDFS semantics and their performance is explored
+in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
+consistency because it  does not guarantee that all updates are not immediately
+seen by all servers; the BatchFS and RAMDisk subtrees are decoupled from the
+global namespace and have similar consistency/durability behavior to those
+systems; and the POSIX subtree retains the rigidity of POSIX's strong
+consistency.
 
@@
 (\(10\times\) slow down). We also show the effect of having a metadata specific
-file format in systems that are based on in-memory data structures.  We scale
-up to 20 clients executing an extremely metadata intensive job because this
-workload exhausts the resources of a single metadata server for the code base we
-built on.  Partitioning and replicating metadata across a cluster can be
-controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in this
-paper we show the capacity and performance of a metadata server with superior
-metadata protocols, which should be used to inform load balancing decisions.
+file format in systems that are based on in-memory data structures.  Our
+prototype is built on Ceph because it was originally designed for HPC and has
+proven to be useful for HPC workloads~\cite{}. Several features were motivated
+by HPC requirements~\cite{weil:osdi2006-ceph}, so the lessons are useful (and
+can be implemented) in HPC environments.
 

### 2017-10-22 paper/sections/3_posix_overheads.tex
paper: figure 1 replace client with semantics
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the file system that uses Ceph's object store ({\it i.e.} RADOS) to store its
-data and metadata. Ceph was originally designed for HPC and has proven to be
-useful for HPC workloads~\cite{}. Several features were motivated by HPC
-requirements~\cite{weil:osdi2006-ceph}, so the lessons are useful (and can be
-implemented) in HPC environments.  This file system is an implementation of one
-set of design decisions and our goal is to highlight the effect that those
-decisions have on performance. During this process we discovered, based on the
-analysis and breakdown of costs, that durability and consistency have high
-overhead.
+data and metadata. To show how the file system behaves under high metadata load
+we use a create-heavy workload. Create-heavy workloads are studied the most in
+HPC research because of the checkpoint/restart use case but they also happen to
+stress the underlying storage system the most.  During this process we
+discovered, based on the analysis and breakdown of costs, that durability and
+consistency have high overhead but we urge the read to keep in mind that this
+file system is an implementation of one set of design decisions and our goal
+here is to highlight the effect that those decisions have on performance.
 
-To show how the file system behaves under high metadata load we use a
-create-heavy workload. Create-heavy workloads are studied the most in HPC
-research because of the checkpoint/restart use case but they also happen to
-stress the underlying storage system the most.
 Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
-the Linux kernel.  The \texttt{untar} phase, which is characterized by many creates, has
-the highest resource usage which suggests that it is stressing the consistency
-and journaling subsystems of the metadata server the most. Traditional file
-system techniques for improving performance, such as caching inodes, do not
-help for create-heavy workloads.
+the Linux kernel.  The \texttt{untar} phase, which is characterized by many
+creates, has the highest resource usage which suggests that it is stressing the
+consistency and journaling subsystems of the metadata server the most.
+Traditional file system techniques for improving performance, such as caching
+inodes, do not help for create-heavy workloads.
 

### 2017-10-22 paper/sections/6_results.tex
paper: figure 1 replace client with semantics
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \label{sec:evaluation}
+
+We scale
+up to 20 clients executing an extremely metadata intensive job because this
+workload exhausts the resources of a single metadata server for the code base we
+built on.  Partitioning and replicating metadata across a cluster can be
+controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in this
+paper we show the capacity and performance of a metadata server with superior
+metadata protocols, which should be used to inform load balancing decisions.
 

### 2017-10-22 paper/sections/6_results.tex
paper: intro to the evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-21 paper/sections/6_results.tex
@@
 
-We did not compare to other HPC systems because we are not claiming that our
-file system metadata protocols are superior. Instead, we use Ceph as a platform
-for exploring related work over the same storage system, which facilitates an
-apples-to-apples comparison of the strategies themselves.  So we run
-experiments on the same cluster and normalize results to control for hardware
-differences and to make our results more generally applicable.  Our goal is not
-to build a file system optimized for HPC workloads and our prototype is not a
-competitor to Lustre. Rather it is a solution for letting the techniques of a
-system like Lustre live in the same namespace as techniques optimized for file
-creates, like BatchFS.
+Cudele lets users construct consistency/durability guarantees using
+well-established research techniques from other systems; so instead of
+evaluating the scalability and performance of the techniques
+themselves\footnote{ This is the reason we did not compare to other HPC or data
+center systems; we are not claiming that our file system metadata protocols are
+superior.}, we show that (1) the mechanisms we propose are useful for
+constructing semantics used by real systems, (2) the techniques can work
+side-by-side in the same namespace, and (3) the combination of these techniques
+can help the system scale farther when subtrees are coupled to the correct type
+of application.  We scope the evaluation to one metadata server and scale the
+clients, which execute an extremely metadata intensive job in parallel. This
+type of analysis shows the capacity and performance of a metadata server with
+superior metadata protocols, which should be used to inform load balancing
+across a metadata cluster. Load balancing with partitioning and replication can
+be explored with something like Mantle~\cite{sevilla:sc15-mantle}.
 
@@
 as a single process which is the default setting for Ceph. We develop on Ceph's
-Jewel release, version 10.2.1, which was released in May 2016.
+Jewel release, version 10.2.1, which was released in May 2016.  We run
+experiments on the same cluster and normalize results to control for hardware
+differences and to make our results more generally applicable.
 

### 2017-10-22 paper/build.sh
paper: switch to IEEE format
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
-set -ex
+set -x
 
 # Build the paper
 docker run --rm \
@@ -10,16 +10,15 @@ docker run --rm \
   -v `pwd`/:/mnt \
   michaelsevilla/texlive:latest -c \
     "cd /mnt ; \
-     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
-     bibtex paper && \
-     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
-     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper" &> build.log
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper; \
+     bibtex paper; \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper; \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper;" &> build.log
 
 ERR=$?
 if [ $ERR != "0" ] ; then
   echo "ERROR: $ERR"
   cat build.log
-  exit 1
 fi
 
 set +e

### 2017-10-22 experiments/cudele-blockapi/run.sh
benchmark: scale block/allow api
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,53 @@
+#!/bin/bash
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
+set -ex
+
+rm -fr results || true
+
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
+ROOT=`dirname $PWD | xargs dirname`
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
+DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
+
+# debug mode
+if [ ! -z $1 ]; then
+  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
+  exit
+fi
+
+./teardown.sh; rm -rf results
+for log in "log"; do
+  cp configs_$SITE/all-$log ansible/group_vars/all 
+  for clients in 1 5 10 15 16 17 18 19 20; do
+    mkdir results || true
+    cp configs_$SITE/clients$clients hosts
+    for job in "creates-touchstream" "creates"; do
+      ./teardown.sh
+      $DOCKER -e nfiles=98000 \
+              ceph.yml \
[…]

### 2017-10-22 experiments/cudele-blockapi/visualize/jupyter.sh
paper: add scalability block/allow
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,4 +1,4 @@
 #!/bin/bash
 
 set -e -x
-docker run --rm -v `dirname $PWD`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93
+docker run --rm -v `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93

### 2017-10-22 paper/sections/0_abstract.tex
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-HPC developers are abandoning POSIX because the synchronization and
-serialization overheads of providing strong consistency and durability are too
-costly -- and often unnecessary -- for their applications.  Unfortunately,
-designing near-POSIX file systems excludes applications that rely on strong
-consistency or durability, forcing developers to re-write their applications or
-deploy them on a different system.  We present a file system and API that lets
-clients specify their consistency/durability requirements and assign them to
-subtrees in the namespace, allowing users to optimize subtrees within the same
-namespace for different workloads.  We draw conclusions about the performance
-impact of unexplored consistency/durability metadata designs and show that
-maintaining strong consistency can cause about a 100\(\times\) slow down
-compared to relaxed consistency and no durability. Comparatively, merging
-updates after a period of relaxed consistency (less than a 10\(\times\) slow
-down) and maintaining durability (about a 10\(\times\) slow down) have more
-reasonable costs.
+Data center scale application developers are abandoning POSIX because the file
+system metadata synchronization and serialization overheads of providing strong
+consistency and durability are too costly -- and often unnecessary -- for their
+applications.  Unfortunately, designing near-POSIX file systems excludes
+applications that rely on strong consistency or durability, forcing developers
+to re-write their applications or deploy them on a different system. Mounting
+multiple systems in the global namespace forces users to (1) provision separate
+storage clusters and (2) manually move data across systems.  We present a
+framework and API that lets clients dynamically specify their
+consistency/durability requirements and assign them to subtrees in the
+namespace, allowing users to optimize subtrees within the same namespace for
+different workloads.  We confirm the performance benefits of techniques
+presented in related work but also explore new consistency/durability metadata
+designs, all over the same storage system.  By custom fitting a subtree to a
+create-heavy application, we show an 8\(\times\) speedup and scale further
+(twice as many servers) than the system we implemented our prototype on.
 

### 2017-10-22 paper/sections/1_introduction.tex
paper: intro to the evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
-
-\caption{For performance, clients can relax consistency on their subtree
-({e.g.}, HDFS) or even decouple the subtree and move it locally ({e.g.},
-BatchFS, RAMDisk). Decoupled subtrees can further relax durability semantics
-for even better performance. Clients that require stronger guarantees ({\it
-e.g.,} POSIX) can still reside in the same namespace.
-}\label{fig:subtree-policies}
-
-\end{figure}
-
 % What is HPC doing?
@@
 computation, etc.) and because these applications wreak havoc on file systems designed for
-general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
-patterns).
+general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N and N:1 create
+patterns~\cite{bent_plfs_2009}).
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
+\caption{ Illustration of subtrees with different semantics co-existing in a
+global namespace.  For performance, clients can relax consistency on their
+subtree (HDFS) or even decouple the subtree and move it locally (BatchFS,
+RAMDisk). Decoupled subtrees can further relax durability for even better
+performance.  Clients that require stronger guarantees (POSIX) can still reside
+in the same namespace.  }\label{fig:subtree-policies}
+\end{figure}
 
@@
 dynamically control the consistency and durability guarantees for subtrees in
-the file system namespace.  Figure~\ref{fig:subtree-policies} shows a setup in
-our proposed system where a single global namespace has subtrees for
-applications optimized with techniques from different state-of-the-art HPC
+the file system namespace.  Figure~\ref{fig:subtree-policies} shows a potential
+setup in our proposed system where a single global namespace has subtrees for
+applications optimized with techniques from different state-of-the-art
 architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
@@
 in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
-consistency because it  does not guarantee that all updates are not immediately
-seen by all servers; the BatchFS and RAMDisk subtrees are decoupled from the
-global namespace and have similar consistency/durability behavior to those
-systems; and the POSIX subtree retains the rigidity of POSIX's strong
-consistency.
+consistency because it does not guarantee that all updates are immediately seen
+by all servers; the BatchFS and RAMDisk subtrees are decoupled from the global
+namespace and have similar consistency/durability behavior to those systems;
+and the POSIX subtree retains the rigidity of POSIX's strong consistency.
+Subtrees without policies inherit the consistency/durability semantics of the
+parent and future work will examine embeddable or composable policies.
 
@@
 
-  \item a prototype that lets developers choose from a range of
-  consistency and durability semantics (9 permutations), allowing them to dynamically custom
-  fit the storage system to the application.
+  \item A framework and API for assigning consistency/durability policies 
+  to subtrees in the file system namespace; this lets users dynamically navigate
+  the tradeoffs of different metadata protocols.
 
-  \item an API for selecting consistency/durability policies and assigning
-  them to subtrees in the file system namespace.
+  \item This framework lets subtrees with different semantics co-exist in a
+  global namespace. This is different than multiple mount points because
+  there is no need to provision storage clusters for applications
+  ({\it e.g.}, HDFS, CephFS) or move data between these systems.
 
-  \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
-  previously unexplored metadata designs, all implemented using Cudele.
+  \item A prototype that lets developers custom fit subtrees to applications
+  dynamically; this helps data center systems scale further than systems that
+  use one strategy for the entire namespace.
 
@@
 the costs of merging updates (7\(\times\) slow down) and maintaining durability
-(\(10\times\) slow down). We also show the effect of having a metadata specific
-file format in systems that are based on in-memory data structures.  Our
-prototype is built on Ceph because it was originally designed for HPC and has
-proven to be useful for HPC workloads~\cite{}. Several features were motivated
-by HPC requirements~\cite{weil:osdi2006-ceph}, so the lessons are useful (and
-can be implemented) in HPC environments.
+(\(10\times\) slow down). In our prototype, we get an 8\(\times\) speedup and
+can scale to twice as many servers when we assign a more relaxed form of
+consistency and durability to a subtree with a create heavy workload.
+Meanwhile, other subtrees in the namespace maintain strong consistency and
+durability with RPCs per operation.
+
+We build on Ceph because it is used in cloud-based and data center
+systems and has a presence in HPC~\cite{}.  More importantly,  we use Ceph as a
+platform for exploring related work over the same storage system, which
+facilitates an apples-to-apples comparison of the strategies themselves.
 

### 2017-10-22 paper/sections/2_related_work.tex
paper: draft 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 abad:techreport2012-fstrace, abad:ucc2012-mimesis,
-alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
-for improving the performance of metadata access.
+alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. 
+%There are two approaches
+%for improving the performance of metadata access.
+%
+%\subsection{Metadata Load Balancing}
+%
+%% approaches to load balancing (strong = more resuorces per unit work, weak =
+%% fixed resource per work unit)
+%% CITEME
+%One approach for improving metadata performance and scalability is to alleviate
+%overloaded servers by load balancing metadata IO across a cluster. Common
+%techniques include partitioning metadata when there are many writes and
+%replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
+%directories and clients write to different partitions by grabbing leases and
+%caching ancestor metadata for path traversal; it does well for strong scaling
+%because servers can keep more inodes in the cache which results in less RPCs.
+%Alternatively, ShardFS replicates directory state so servers do not need to
+%contact peers for path traversal; it does well for read workloads because all
+%file operations only require 1 RPC and for weak scaling because requests will
+%never incur extra RPCs due to a full cache.  CephFS employs both techniques to
+%a lesser extent; directories can be replicated or sharded but the caching and
+%replication policies do not change depending on the balancing technique.
+%Despite the performance benefits these techniques add complexity and jeopardize
+%the robustness and performance characteristics of the metadata service because
+%the systems now need (1) policies to guide the migration decisions and (2)
+%mechanisms to address inconsistent states across servers.
+%
+%% policies
+%Setting policies for migrations is arguably more difficult than adding the
+%migration mechanisms themselves.  For example, IndexFS and CephFS use the
+%GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
+%threshold and using lazy synchronization to redirect queries to the server that
+%``owns" the targeted metadata.  Determining when to partition directories and
+%when to migrate the directory fragments are policies that vary between systems:
+%GIGA+ partitions directories when the size reaches a certain number of files
+%and migrates directory fragments immediately; CephFS partitions directories
+%when they reach a threshold size or when the write temperature reaches a
+%certain value and migrates directory fragments when the hosting server has more
+%load than the other servers in the metadata cluster. Another policy is when and
+%how to replicate directory state; ShardFS replicates immediately and
+%pessimistically while CephFS replicates only when the read temperature reaches
+%a threshold.  There is a wide range of policies and it is difficult to maneuver
+%tunables and hard-coded design decisions.
+%
+%% addressing inconsistency
+%In addition to the policies, distributing metadata across a cluster requires
+%distributed transactions and cache coherence protocols to ensure strong
+%consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
+%directory state and uses optimistic concurrency control for conflicts; namely
+%it does the operation and if there is a conflict at verification time it falls
+%back to two-phase locking.  Another example is IndexFS's inode cache which
+%reduces RPCs by caching ancestor paths -- the locality of this cache can be
+%thrashed by random reads but performs well for metadata writes. For
+%consistency, writes to directories in IndexFS block until the lease expires
+%while writes to directories in ShardFS are slow for everyone as it either
+%requires serialization or locking with many servers; reads in IndexFS are
+%subject to cache locality while reads in ShardFS always resolve to 1 RPC.
+%Another example of the overheads of addressing inconsistency is how CephFS
+%maintains client sessions and inode caches for capabilities (which in turn make
+%metadata access faster). When metadata is exchanged between metadata servers
+%these sessions/caches must be flushed and new statistics exchanged with a
+%scatter-gather process; this halts updates on the directories and blocks until
+%the authoritative metadata server responds.  These protocols are discussed in
+%more detail in the next section but their inclusion here is a testament to the
+%complexity of migrating metadata.
+%
+%%For metadata writes it does a distributed transaction; monotonic writes with
+%%concurrent clients fail and do pessimistic locking through a mediated lock
+%%server to ensure strong consistency, non-monotonic writes grab locks at every
+%%server. Zooming in on monotonic writes: if permissions increase it executes on
+%%a primary then non-primary, if permissions decrease it executes on all
+%%non-primary then on primary.
+%
+%The conclusion we have drawn from this related work is that metadata protocols
+%have a bigger impact on performance and scalability than load balancing.  
+%Understanding these protocols helps load balancing and gives us a better
+%understanding of the metrics we should use to make migration decisions ({\it
+%e.g.}, which operations reflect the state of the system), what types of
+%requests cause the most load, and how an overloaded system reacts ({\it e.g.},
+%increasing latencies, lower throughput, etc.).
 
-\subsection{Metadata Load Balancing}
-
-% approaches to load balancing (strong = more resuorces per unit work, weak =
-% fixed resource per work unit)
-One approach for improving metadata performance and scalability is to alleviate
-overloaded servers by load balancing metadata IO across a cluster. Common
-techniques include partitioning metadata when there are many writes and
-replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
-directories and clients write to different partitions by grabbing leases and
-caching ancestor metadata for path traversal; it does well for strong scaling
-because servers can keep more inodes in the cache which results in less RPCs.
-Alternatively, ShardFS replicates directory state so servers do not need to
-contact peers for path traversal; it does well for read workloads because all
-file operations only require 1 RPC and for weak scaling because requests will
-never incur extra RPCs due to a full cache.  CephFS employs both techniques to
-a lesser extent; directories can be replicated or sharded but the caching and
-replication policies do not change depending on the balancing technique.
-Despite the performance benefits these techniques add complexity and jeopardize
-the robustness and performance characteristics of the metadata service because
-the systems now need (1) policies to guide the migration decisions and (2)
-mechanisms to address inconsistent states across servers.
-
-% policies
-Setting policies for migrations is arguably more difficult than adding the
-migration mechanisms themselves.  For example, IndexFS and CephFS use the
-GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
-threshold and using lazy synchronization to redirect queries to the server that
-``owns" the targeted metadata.  Determining when to partition directories and
-when to migrate the directory fragments are policies that vary between systems:
-GIGA+ partitions directories when the size reaches a certain number of files
-and migrates directory fragments immediately; CephFS partitions directories
-when they reach a threshold size or when the write temperature reaches a
-certain value and migrates directory fragments when the hosting server has more
-load than the other servers in the metadata cluster. Another policy is when and
-how to replicate directory state; ShardFS replicates immediately and
-pessimistically while CephFS replicates only when the read temperature reaches
-a threshold.  There is a wide range of policies and it is difficult to maneuver
-tunables and hard-coded design decisions.
-
-% addressing inconsistency
-In addition to the policies, distributing metadata across a cluster requires
-distributed transactions and cache coherence protocols to ensure strong
-consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
-directory state and uses optimistic concurrency control for conflicts; namely
-it does the operation and if there is a conflict at verification time it falls
-back to two-phase locking.  Another example is IndexFS's inode cache which
-reduces RPCs by caching ancestor paths -- the locality of this cache can be
-thrashed by random reads but performs well for metadata writes. For
-consistency, writes to directories in IndexFS block until the lease expires
-while writes to directories in ShardFS are slow for everyone as it either
-requires serialization or locking with many servers; reads in IndexFS are
-subject to cache locality while reads in ShardFS always resolve to 1 RPC.
-Another example of the overheads of addressing inconsistency is how CephFS
-maintains client sessions and inode caches for capabilities (which in turn make
-metadata access faster). When metadata is exchanged between metadata servers
-these sessions/caches must be flushed and new statistics exchanged with a
-scatter-gather process; this halts updates on the directories and blocks until
-the authoritative metadata server responds.  These protocols are discussed in
-more detail in the next section but their inclusion here is a testament to the
-complexity of migrating metadata.
-
-%For metadata writes it does a distributed transaction; monotonic writes with
-%concurrent clients fail and do pessimistic locking through a mediated lock
-%server to ensure strong consistency, non-monotonic writes grab locks at every
-%server. Zooming in on monotonic writes: if permissions increase it executes on
-%a primary then non-primary, if permissions decrease it executes on all
-%non-primary then on primary.
-
-The conclusion we have drawn from this related work is that metadata protocols
-have a bigger impact on performance and scalability than load balancing.  
-Understanding these protocols helps load balancing and gives us a better
-understanding of the metrics we should use to make migration decisions ({\it
-e.g.}, which operations reflect the state of the system), what types of
-requests cause the most load, and how an overloaded system reacts ({\it e.g.},
-increasing latencies, lower throughput, etc.).
-
-\subsection{Relaxing POSIX}
+%\subsection{Relaxing POSIX}
 

### 2017-10-22 paper/sections/6_results.tex
paper: add scalability block/allow
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
-\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
-client. The size of the client's journal scales with the number of
-updates.\label{fig:behavior-journal-size}}
-\end{figure}
+%\begin{figure}[tb]
+%\centering
+%\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
+%\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
+%client. The size of the client's journal scales with the number of
+%updates.\label{fig:behavior-journal-size}}
+%\end{figure}
 
@@
 
-{\it Journal Size:} Figure~\ref{fig:behavior-journal-size} shows the
-amount of storage per journal update (\(y\) axis) for the range of file creates
-we tested (\(x\) axis). The increase in file size is linear with the number of
+{\it Journal Size:} We measure the amount of storage per journal update to be
+about 2.5\text{KB}. The storage footprint scales linearly with the number of
 metadata creates and suggests that updates for a million files would be
 \(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
-payloads of this size on an HPC network are reasonable.\\
+payloads of this size in most HPC/data cente networks are reasonable.\\
 
@@
 well-established research systems but, more importantly, we show that the
-Cudele mechanisms we propose are useful for building evaluating different
+Cudele mechanisms we propose are useful for building and evaluating different
 consistency/durability semantics.
@@
 \centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
+\includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
 \caption{ [\href{https://...}{source}] Use Case 2: using the ``allow" and

### 2017-10-22 experiments/cudele-blockapi/run.sh
benchmark: finsish block API
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -37,17 +37,19 @@ fi
 ./teardown.sh; rm -rf results
 for log in "log"; do
   cp configs_$SITE/all-$log ansible/group_vars/all 
-  for clients in 1 5 10 15 16 17 18 19 20; do
-    mkdir results || true
-    cp configs_$SITE/clients$clients hosts
-    for job in "creates-touchstream" "creates"; do
-      ./teardown.sh
-      $DOCKER -e nfiles=98000 \
-              ceph.yml \
-              /workloads/block.yml \
-              /workloads/${job}.yml
+  for clients in 1 5 10 15 18 19 20; do
+    for run in 0 1 2; do
+      mkdir results || true
+      cp configs_$SITE/clients$clients hosts
+      for job in "creates-touchstream" "creates"; do
+        ./teardown.sh
+        $DOCKER -e nfiles=98000 \
+                ceph.yml \
+                /workloads/block.yml \
+                /workloads/${job}.yml
+      done
+      mv results results-$SITE-clients$clients-$log-run$run
     done
-    mv results results-$SITE-clients$clients-$log
   done
 done
 exit 0

### 2017-10-22 paper/sections/4_methodology.tex
paper: fix microbenchmark section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: space reduction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   Append Client Journal & events appended to in-memory journal \\
-  Volatile Apply    & apply to metadata store in obj store \\
-  Nonvolatile Apply & apply to metadata store in memory \\
+  Volatile Apply & apply to metadata store in memory \\
+  Nonvolatile Apply    & apply to metadata store in obj store \\
   Local Persist     & journal saved to client's disk \\
@@
 Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
-have global consistency and durability ~\footnote{This is the normal case.
+have global consistency and durability~\footnote{This is the normal case.
 IndexFS also has bulk merge which would transition the system into ``weak
@@
 the different guarantees in Table~\ref{table:spectrum}, not all of them make
-sense. For example, it makes little sense to do \texttt{append client journal+RPCs} since
-both mechanisms do the same thing or \texttt{stream+save} since ``global"
-durability is stronger and has more overhead than ``local" durability.
+sense. For example, it makes little sense to do \texttt{append client
+journal+RPCs} since both mechanisms do the same thing or \texttt{stream+save}
+since ``global" durability is stronger and has more overhead than ``local"
+durability. The cost of each mechanism and the semantics described above are
+quantified in Sections~\S\ref{sec:microbenchmarks} and~\S\ref{sec:use-case-1}.
 
@@
 would give the application 100 inodes but the subtree would behave like the
-existing CephFS implementation. To implement DeltaFS on Cudele, the user
-would use the configuration:
-
-\begin{listing}[h]
-\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "100000"
-    "interfere_policy": "block"
-    "consistency": "append_client_journal"
-    "durability": "local_persist"
-}
-\end{minted}
-\end{listing}
-
-BatchFS merges updates into the global namespace after the job is complete.
-This can be achieved with by composing two mechanisms:
-
-\begin{listing}[h]
-\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
-               tabsize=4]{js}
-{     
-  "allocated_inodes": "100000"
-  "interfere_policy": "block"
-  "consistency": "append_client_journal+volatile_apply"
-  "durability": "local_persist"
-}
-\end{minted}
-\end{listing}
+existing CephFS implementation.  BatchFS merges updates into the global
+namespace after the job is complete.  This can be achieved with by composing
+two mechanisms:

### 2017-10-22 paper/sections/6_results.tex
paper: space reduction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 evaluating the scalability and performance of the techniques
-themselves\footnote{ This is the reason we did not compare to other HPC or data
+themselves\footnote{This is the reason we did not compare to other HPC or data
 center systems; we are not claiming that our file system metadata protocols are
@@
 side-by-side in the same namespace, and (3) the combination of these techniques
-can help the system scale farther when subtrees are coupled to the correct type
+can help the system scale further when subtrees are coupled to the correct type
 of application.  We scope the evaluation to one metadata server and scale the
@@
 
-\begin{table}
-\begin{tabular}{ r | l | l}
-  Cluster  & Hardware               & Software \\\hline
-  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
-           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
-  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
-           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
-\end{tabular}
-\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
-monitor daemon, MDS is metadata server daemon, and OSDs are object storage
-daemons.\label{table:clusters}} 
-\end{table}
+%\begin{table}
+%\begin{tabular}{ r | l | l}
+%  Cluster  & Hardware               & Software \\\hline
+%  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
+%           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
+%  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
+%           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
+%\end{tabular}
+%\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
+%monitor daemon, MDS is metadata server daemon, and OSDs are object storage
+%daemons.\label{table:clusters}} 
+%\end{table}
 
-We evaluate our prototype on the two clusters in Table~\ref{table:clusters}.
-Microbenchmarks were run on the in-house clusters while client scalability
-tests were run on CloudLab. We only scale up to 29 clients because that is the
-maximum capacity of a single metadata server. For the in-house cluster, we need
-more object storage servers and we have to co-locate clients on the same node
-because of resource contention \footnote{The bandwidth to RADOS is only
-sufficient with 8 nodes on this older hardware, which uses SSDs and nodes that
-are almost 10 years old.  The CloudLab cluster performs just as well with and
-without a highly provisioned object storage server cluster.}. All daemons run
-as a single process which is the default setting for Ceph. We develop on Ceph's
-Jewel release, version 10.2.1, which was released in May 2016.  We run
-experiments on the same cluster and normalize results to control for hardware
-differences and to make our results more generally applicable.
+We run experiments on the same cluster and normalize results to control for
+hardware differences and to make our results more generally applicable.  We use
+a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.
+Each node uses Ubuntu 14.04 and has 1 monitor server, 3 object storage daemons,
+1 metadata server, and 29 clients.  All daemons run as a single process which
+is the default setting for Ceph. We develop on Ceph's Jewel release, version
+10.2.1, which was released in May 2016.
 
-This paper adheres to The Popper
-Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}, so
-experiments presented here are available in the repository for this
-article\footnote{Links have been removed for double-blind submission}.
+This paper adheres to The Popper Convention~\cite{jimenez_popper_2016}, so
+experiments presented here are available in the repository for this article.
 Experiments can be examined in more detail, or even re-run, by visiting the
@@
 \includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
-\caption{ [\href{https://...}{source}] Microbenchmark: performance of each
+\caption{[\href{https://www.google.com}{source}] Microbenchmark: performance of each
 mechanism (left) and building consistency/durability guarantees (right) for
@@
 metadata creates and suggests that updates for a million files would be
-\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
-payloads of this size in most HPC/data cente networks are reasonable.\\
+\(2.5\text{KB}*1\texttt{ million files} = 2.38\text{GB}\). Transfer times for
+payloads of this size in most HPC/data cente networks are reasonable.
 
-\noindent\textbf{Takeaway}: measuring the mechanisms individually shows that
+\textbf{Takeaway}: measuring the mechanisms individually shows that
 their overheads and costs can differ {\it by orders of magnitude}. Cudele gives
@@
 \label{sec:use-case-1}
+
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
+      \caption{Use Case 1} \label{fig:mergescale}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
+      \caption{Use Case 2}
+      \label{fig:block-allow}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
+      \caption{Use Case 3}
+      \label{fig:slowdown-sync}
+  \end{subfigure}
+  \caption{The performance and features of Cudele. (a) shows the performance of
+merging client journals at the metadata server. Performance is better than RPCs
+because there are less messages and consistency/durability code paths are
+bypassed. (b) shows how the allow/block API can isolate directories from
+interfering clients.  (c) shows the slowdown of a single client periodically
+syncing updates to the gobal namespace. The inflection point is the trade-off
+of frequent updates vs. larger journal files.\label{fig:use-cases}}
+\end{figure*}
 
@@
 all schemes we argue that the performance gain for decoupled namespaces comes
-from relaxing the consistency guarantees and not from the metadata formats.\\
+from relaxing the consistency guarantees and not from the metadata formats.
 
-\noindent\textbf{Takeaway}: we confirm the performance benefits of other
+\textbf{Takeaway}: we confirm the performance benefits of other
 well-established research systems but, more importantly, we show that the
@@
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
-\caption{ [\href{https://...}{source}] Use Case 2: using the ``allow" and
-``block" API, users can isolate directories from interfering clients. Variance
-with blocking turned on is the same as ``isolated" from
-Figure~\ref{fig:overhead-b}.
-\label{fig:slowdown-allow-block}}
-
-\end{figure}
-
 % setup: 2cs sep dirs, 2cs DN, 1c malicious write
-Next we show how Cudele can be programmed to block interfering clients. We
-use the same problematic workload from Figure~\ref{fig:overhead-b}, where
-clients write to their own private directories and another client interferes with a stream of creates at
-30 seconds.  We also have another client write to a decoupled namespace and
-merge its updates at 90 seconds.  We equip each client directory with the
-configuration:
-
-\begin{listing}[tb]
-\begin{minted}[xleftmargin=21pt,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "100000"
-    "interfere_policy": "block"
-    "consistency": "RPCs"
-    "durability": "stream"
-}
-\end{minted}
-\end{listing}
+Next we show how Cudele can be programmed to block interfering clients. We use
+the same problematic workload from Figure~\ref{fig:overhead-b}, where clients
+write to their own private directories and another client interferes with a
+stream of creates at 30 seconds.  We also have another client write to a
+decoupled namespace and merge its updates at 90 seconds.  We equip each client
+directory with the configuration: allocated\_inodes=100K,
+interfere\_policy=block, consistency=RPCs, and durability=stream.
 
@@
 To show the benefits of this isolations, our results in
-Figure~\ref{fig:slowdown-allow-block} are plotted alongside the variance bars from
+Figure~\ref{fig:allow-block} are plotted alongside the variance bars from
 Figure~\ref{fig:overhead-b}. Because ``allow"/``interfere and
@@
 \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
-negligible effect on performance.\\
+negligible effect on performance.
 
-\noindent\textbf{Takeaway}: the API lets users isolate directories when
+\textbf{Takeaway}: the API lets users isolate directories when
 applications need better and more reliable performance. This is a way of
@@
 
-\begin{figure}[tb] \centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png} \caption{
-[\href{https://...}{source}] Use Case 3: slowdown of a single client periodically syncing
-updates to the gobal namespace. The crossover point represents the trade-off of
-frequent updates (with small journal files) and infrequent updates (with larger journal
-files).\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}}
-\end{figure}
-
-\begin{figure}[tb] \centering
-\includegraphics[width=1.0\linewidth]{graphs/mergescale.png} \caption{
-[\href{https://...}{source}] Use Case 3: performance of merging client journals
-at the metadata server. Performance is better than RPCs because (1) there are
-less messages on the network and (2) the consistency/durability code paths in
-the metadata server are bypassed. Merges are serialized and read from the local
-disk, so keeping client journals in memory or on SSD will improve the
-performance of ``decoupled" even more.  \ref{fig:mergescale}.
-\label{fig:mergescale-sync}}
-\end{figure}
-
 % Does this implementation need to be described in the implementation?
@@
 files, the maximum recommended directory size for CephFS. We also plot the cost
-of RPCs with and without journaling; recall from Figure~\ref{fig:overhead-c}
+of RPCs with and without journaling; recall from Figure~\ref{fig:overhead}
 that the scalability that max number of clients that the default metadata
@@
 \texttt{lookup()} commands before \texttt{open()} requests, we would have seen
-the poor scaling that we see with the ``RPCs" mechanism.\\
+the poor scaling that we see with the ``RPCs" mechanism.
 
-\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
+\textbf{Takeaway}: the weakest form of consistency for creating files
 ({\it i.e.} not checking data structures for the validity of an update or
@@
 
-\subsection{Use Case 4: Reading Large Directories}
+%\subsection{Use Case 4: Reading Large Directories} 
+%
+%% CITEME
+%The final use case is reading large directories. At job completion, scientists
+%might use \texttt{ls} again to see the results. This causes great strain on the
+%file system as paths need to traversed and the entire directory, with all its
+%entries, must be transferred back to the client. Here we show how decoupling a
+%large namespace and materializing the view in memory on the client is faster
+%than doing RPCs for walks of the file system namespace.
+%
+%FigureX shows how fast a client can materialize decoupled namespaces in memory.
+%This takes the on-disk file system metadata format and parses them into events
+%that can be manipulated and appended to. Again, we scale the number of up
 
-% CITEME
-The final use case is reading large directories. At job completion, scientists
-might use \texttt{ls} again to see the results. This causes great strain on the
-file system as paths need to traversed and the entire directory, with all its
-entries, must be transferred back to the client. Here we show how decoupling a
-large namespace and materializing the view in memory on the client is faster
-than doing RPCs for walks of the file system namespace.
-
-FigureX shows how fast a client can materialize decoupled namespaces in memory.
-This takes the on-disk file system metadata format and parses them into events
-that can be manipulated and appended to. Again, we scale the number of up
-
-\subsection{Discussion}
-\begin{itemize}
-  \item Composable guarantees: applications control their performance (knobs you didn't have before)
-  \item Hardware features: adding hardware in strategic places
-  \item Performance Prediction: quantifies the cost of each mechanism
-\end{itemize}
+%\subsection{Discussion}
+%
+%Cudele and the experiments we present here prompt many avenues for future work.
+%The first is embeddable policies, where child subtrees have specialized
+%features but still maintain the guarantees  of their parent subtrees. We
+%already show how applciations can compose guarantees and control their
+%performance, knobs that we did not have before, but embeddable policies are a
+%way to compose the policies themsleves.  For example, a RAMDisk subtree is
+%POSIX-compliant but relaxes durability constraints, so it can reside under a
+%POSIX subtree that is strongly consistent. These embeddable policies should be
+%controlled and enforced by Cudele.
+%
+%Second is the opportunity for adding hardware in strategic places to speed up
+%the mechanisms that the application needs. In Figure~\ref{fig:mergescale} the
+%performance of merging updates at the metadata server is limited by the speed
+%of the disk. So applications that rely on merging many updates should consider
+%equipping the metadata server with SSDs or more memory. This is a form of
+%co-design that takes a ``dirty-slate" approach.
+%
+%Finally, performance prediction benefits from our cost quantification of each
+%mechanism. Applications can use Cudele to microbenchmark their components and
+%software, similar to what we did in Secion~\S\ref{sec:microbenchmarks}. Using
+%those results, they can predict how much slower their system will be if they
+%adopt stronger consistency or durability. This can also be a verification tool
+%where performance that varies wildly from the predicted performance can be a
+%red flag that something is wrong or that the bottleneck is not in the
+%consistency or durability plane.

### 2017-10-22 paper/sections/0_abstract.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-Data center scale application developers are abandoning POSIX because the file
+Data center scale application developers are abandoning POSIX IO because the file
 system metadata synchronization and serialization overheads of providing strong
 consistency and durability are too costly -- and often unnecessary -- for their
-applications.  Unfortunately, designing near-POSIX file systems excludes
+applications.  Unfortunately, designing near-POSIX IO file systems excludes
 applications that rely on strong consistency or durability, forcing developers

### 2017-10-22 paper/sections/1_introduction.tex
paper: space reduction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 do more client-side processing and merge updates when the job is done. HPC
-developers are turning to these non-POSIX solutions because their applications
+developers are turning to these non-POSIX IO solutions because their applications
 are well-understood ({\it e.g.}, well-defined read/write phases,
@@
 RAMDisk). Decoupled subtrees can further relax durability for even better
-performance.  Clients that require stronger guarantees (POSIX) can still reside
+performance.  Clients that require stronger guarantees (POSIX IO) can still reside
 in the same namespace.  }\label{fig:subtree-policies}
@@
 namespace and have similar consistency/durability behavior to those systems;
-and the POSIX subtree retains the rigidity of POSIX's strong consistency.
+and the POSIX IO subtree retains the rigidity of POSIX IO's strong consistency.
 Subtrees without policies inherit the consistency/durability semantics of the
-parent and future work will examine embeddable or composable policies.
+parent and future work will examine embeddable or inheritable policies.
 
@@
 In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
-cost of POSIX consistency and system-defined durability and
+cost of POSIX IO consistency and system-defined durability and
 Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele

### 2017-10-22 paper/sections/2_related_work.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % General
-The bottlenecks associated with accessing POSIX file system metadata are not limited
+The bottlenecks associated with accessing POSIX IO file system metadata are not limited
 to HPC workloads and the same challenges that plagued these systems for years are
@@
 %distributed transactions and cache coherence protocols to ensure strong
-%consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
+%consistency ({e.g.}, POSIX IO).  For example, ShardFS pessimistically replicates
 %directory state and uses optimistic concurrency control for conflicts; namely
@@
 
-%\subsection{Relaxing POSIX}
+%\subsection{Relaxing POSIX IO}
 
-% POSIX 
-POSIX workloads require strong consistency and many file systems improve
+% POSIX IO 
+POSIX IO workloads require strong consistency and many file systems improve
 performance by reducing the number of remote calls per operation ({\it i.e.}
@@
 
-% Non POSIX
+% Non POSIX IO
 High performance computing has unique requirements for file systems ({\it
 e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
-make relaxing POSIX sensible.  BatchFS assumes the application coordinates
+make relaxing POSIX IO sensible.  BatchFS assumes the application coordinates
 accesses to the namespace, so the clients can batch local operations and merge
@@
 
-% NON POSIX
+% NON POSIX IO
 %Decoupling the namespaces has many advantages, including improved scalability,
 %higher resource utilization, and better performance.  BatchFS and DeltaFS are
-%near-POSIX filesystems that give clients the ability to decouple subtrees from
+%near-POSIX IO filesystems that give clients the ability to decouple subtrees from
 %the namespace so that the applications can execute metadata operations without
@@
 
-For (3), decoupled namespace approaches sacrifice POSIX going as far as
+For (3), decoupled namespace approaches sacrifice POSIX IO going as far as
 requiring the application to link against the systems they want to talk to. In
@@
 centers with many types and tiers of storage. Despite well-known performance
-problems POSIX and REST are the dominant APIs for data transfer.
+problems POSIX IO and REST are the dominant APIs for data transfer.

### 2017-10-22 paper/sections/3_posix_overheads.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{POSIX Overheads}
+\section{POSIX IO Overheads}
 \label{sec:posix-overheads}
@@
 % CITEME PDSW paper
-In our examination of the overheads of POSIX we benchmark and analyze CephFS,
+In our examination of the overheads of POSIX IO we benchmark and analyze CephFS,
 the file system that uses Ceph's object store ({\it i.e.} RADOS) to store its
@@
 % what is durability
-While durability is not specified by POSIX, users expect that files they create
+While durability is not specified by POSIX IO, users expect that files they create
 or modify survive failures.  We define three types of durability: global,
@@
 
-Access to metadata in a POSIX-compliant file system is strongly consistent, so
+Access to metadata in a POSIX IO-compliant file system is strongly consistent, so
 reads and writes to the same inode or directory are globally ordered.  The

### 2017-10-22 paper/sections/6_results.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-%% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
+%% DeltaFS vs. BatchFS vs. POSIX IO: decoupled is faster
 %% (5-7x) vs. (90-104x) slower
@@
 %  - e.g., BatchFS/DeltaFS: weak, local
-%  - e.g., POSIX: strong, global
+%  - e.g., POSIX IO: strong, global
 %  - e.g., PLFS: no consistency
@@
 the rogue client. While HPC tries to avoid these situations with
-worklflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it stills
+workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
 happens in distributed file systems when users unintentionally access
@@
 %way to compose the policies themsleves.  For example, a RAMDisk subtree is
-%POSIX-compliant but relaxes durability constraints, so it can reside under a
-%POSIX subtree that is strongly consistent. These embeddable policies should be
+%POSIX IO-compliant but relaxes durability constraints, so it can reside under a
+%POSIX IO subtree that is strongly consistent. These embeddable policies should be
 %controlled and enforced by Cudele.

### 2017-10-22 paper/sections/0_abstract.tex
paper: finish eval; touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-22 paper/sections/0_abstract.tex
@@
 
-Data center scale application developers are abandoning POSIX because the file
+HPC and data center scale developers are abandoning POSIX IO because the file
 system metadata synchronization and serialization overheads of providing strong
 consistency and durability are too costly -- and often unnecessary -- for their
-applications.  Unfortunately, designing near-POSIX file systems excludes
-applications that rely on strong consistency or durability, forcing developers
-to re-write their applications or deploy them on a different system. Mounting
-multiple systems in the global namespace forces users to (1) provision separate
-storage clusters and (2) manually move data across systems.  We present a
-framework and API that lets clients dynamically specify their
-consistency/durability requirements and assign them to subtrees in the
-namespace, allowing users to optimize subtrees within the same namespace for
+applications.  Unfortunately, designing file systems with weaker consistency or
+durability excludes applications that rely on stronger guarantees, forcing
+developers to re-write their applications or deploy them on a different system.
+Mounting multiple systems in the global namespace forces users to (1) provision
+separate storage clusters and (2) manually move data across systems.  We
+present a framework and API that lets clients specify their
+consistency/durability requirements and dynamically assign them to subtrees in
+the same namespace, allowing users to optimize subtrees over time and space for
 different workloads.  We confirm the performance benefits of techniques
 presented in related work but also explore new consistency/durability metadata
-designs, all over the same storage system.  By custom fitting a subtree to a
-create-heavy application, we show an 8\(\times\) speedup and scale further
-(twice as many servers) than the system we implemented our prototype on.
+designs, all integrated the same storage system.  By custom fitting a subtree
+to a create-heavy application, we show 8\(\times\) speedup and 2\(\times\)
+scale when compared to our baseline system.
 

### 2017-10-22 paper/sections/1_introduction.tex
paper: finish eval; touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   \item A framework and API for assigning consistency/durability policies 
-  to subtrees in the file system namespace; this lets users dynamically navigate
+  to subtrees in the file system namespace; this lets users navigate
   the tradeoffs of different metadata protocols.
@@
   \item This framework lets subtrees with different semantics co-exist in a
-  global namespace. This is different than multiple mount points because
-  there is no need to provision storage clusters for applications
-  ({\it e.g.}, HDFS, CephFS) or move data between these systems.
+  global namespace. We show how this scales further and performs better than
+  systems that use one strategy for the entire namespace
 
-  \item A prototype that lets developers custom fit subtrees to applications
-  dynamically; this helps data center systems scale further than systems that
-  use one strategy for the entire namespace.
+  \item A prototype that lets users custom fit subtrees to applications
+  dynamically.
 
@@
 
+The last contribution lays the groundwork for future work on our prototype. It
+improves the current practice of mounting different storage systems in a global
+namespace because there is no need to provision dedicated storage clusters to
+applications or move data between these systems.  For example, the results of a
+Hadoop job do not need to be migrated into CephFS for other processing; instead
+the user can change the semantics of the HDFS subtree into a CephFS subtree.
+This may cause metadata/data movement to make things strongly consistent again
+but this is superior to moving all data across file system boundaries. We
+enable studies that adjust these semantics over {\it time and space}, where
+subtrees can change their semantics and migrate around the cluster without ever
+moving the data they reference.
+
 % Results
-Our results confirm the assertions of ``clean-state" research systems that
-decouple namespaces; specifically that the technique drastically improves
-performance (104\(\times\) speed up) but we go a step further by quantifying
-the costs of merging updates (7\(\times\) slow down) and maintaining durability
-(\(10\times\) slow down). In our prototype, we get an 8\(\times\) speedup and
-can scale to twice as many servers when we assign a more relaxed form of
-consistency and durability to a subtree with a create heavy workload.
-Meanwhile, other subtrees in the namespace maintain strong consistency and
-durability with RPCs per operation.
+The results in this paper lay the groundwork for such a system and confirm the
+assertions of ``clean-state" research that decouple namespaces; specifically
+that the technique drastically improves performance (104\(\times\) speed up).
+We go a step further by quantifying the costs of merging updates (7\(\times\)
+slow down) and maintaining durability (\(10\times\) slow down). In our
+prototype, we get an 8\(\times\) speedup and can scale to twice as many servers
+when we assign a more relaxed form of consistency and durability to a subtree
+with a create-heavy workload.  We build on Ceph because it is used in
+cloud-based and data center systems and has a presence in HPC~\cite{}.  More
+importantly,  we use Ceph as a platform for exploring related work over the
+same storage system, which facilitates an apples-to-apples comparison of the
+strategies themselves.
 
-We build on Ceph because it is used in cloud-based and data center
-systems and has a presence in HPC~\cite{}.  More importantly,  we use Ceph as a
-platform for exploring related work over the same storage system, which
-facilitates an apples-to-apples comparison of the strategies themselves.
-
-In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
-cost of POSIX IO consistency and system-defined durability and
-Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
-prototype and API. Section~\ref{sec:implementation} describes the Cudele
-mechanisms and shows how re-using internal subsystems results in an
-implementation of less than 500 lines of code. The evaluation in
-Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
-explored and previously unexplored metadata designs.
-Section~\ref{sec:related-work} places Cudele in the context of other related
-work.
+%In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
+%cost of POSIX IO consistency and system-defined durability and
+%Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
+%prototype and API. Section~\ref{sec:implementation} describes the Cudele
+%mechanisms and shows how re-using internal subsystems results in an
+%implementation of less than 500 lines of code. The evaluation in
+%Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
+%explored and previously unexplored metadata designs.
+%Section~\ref{sec:related-work} places Cudele in the context of other related
+%work.

### 2017-10-22 paper/sections/3_posix_overheads.tex
paper: finish eval; touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % purpose of the journal
-As shown in Figure~\ref{fig:journal}, in addition to the metadata journal,
-CephFS also represents metadata in RADOS as a metadata store, where directories
-and their file inodes are stored as objects.  The metadata server applies the
-updates in the journal to the metadata store when the journal reaches a certain
-size. The metadata store is optimized for recovery ({\it i.e.} reading) while
-the metadata journal is write-optimized.
+In addition to the metadata journal, CephFS also represents metadata in RADOS
+as a metadata store, where directories and their file inodes are stored as
+objects.  The metadata server applies the updates in the journal to the
+metadata store when the journal reaches a certain size. The metadata store is
+optimized for recovery ({\it i.e.} reading) while the metadata journal is
+write-optimized.
 
-\begin{figure}[tb] \centering
-\includegraphics[width=1\linewidth]{./figures/journal.png} 
-\caption{CephFS has two views of the file system namespace: a journal of
-metadata updates and a metadata store. For fault tolerance, they are stored in
-the object store as segments and fragments, respectively.
-\label{fig:journal}}
-\end{figure}
+%\begin{figure}[tb] \centering
+%\includegraphics[width=1\linewidth]{./figures/journal.png} 
+%\caption{CephFS has two views of the file system namespace: a journal of
+%metadata updates and a metadata store. For fault tolerance, they are stored in
+%the object store as segments and fragments, respectively.
+%\label{fig:journal}}
+%\end{figure}
 

### 2017-10-22 paper/sections/5_implementation.tex
paper: finish eval; touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 store data structure is structured as a tree of directory fragments making it
-easier to read and traverse.\\
+easier to read and traverse.
 
 % how i used it
-\noindent\textbf{Cudele}: the ``RPCs" mechanism uses the metadata store to
+\textbf{Cudele}: the ``RPCs" mechanism uses the metadata store to
 service requests. Using code designed for recovery, ``Nonvolatile Apply" and
@@
 Specifically, reads are slow because there is more state to read, it is
-unorganized, and many of the updates may be redundant.\\
+unorganized, and many of the updates may be redundant.
 
-\noindent\textbf{Cudele}: the journal format is used by ``Stream", ``Append Client Journal",
+\textbf{Cudele}: the journal format is used by ``Stream", ``Append Client Journal",
 ``Local Persist", and ``Global Persist".  ``Stream" is the default
@@
 store objects and replays the updates on the metadata store in the object
-store.\\
+store.
 
-\noindent\textbf{Cudele}: the external library the clients link into is based
+\textbf{Cudele}: the external library the clients link into is based
 on the journal tool. It already had functions for importing, exporting, and
@@
 The inode cache has code for manipulating inode numbers, such as pre-allocating
-inodes to clients.\\
+inodes to clients.
 
-\noindent\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
+\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
 code to allocate inodes to clients that decouple parts of the namespace and to
@@
 the ability to recursively apply policies to subtrees and to isolate subtrees
-from each other.  \\
+from each other.
 
-\noindent\textbf{Cudele}: uses the large inodes to store consistency and
+\textbf{Cudele}: uses the large inodes to store consistency and
 durability policies in the directory inode. This approach uses the File Type

### 2017-10-22 paper/sections/6_results.tex
paper: finish eval; touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 well-established research techniques from other systems; so instead of
-evaluating the scalability and performance of the techniques
-themselves\footnote{This is the reason we did not compare to other HPC or data
-center systems; we are not claiming that our file system metadata protocols are
-superior.}, we show that (1) the mechanisms we propose are useful for
+evaluating the scalability and performance of the techniques themselves against
+other file systems, we show that (1) the mechanisms we propose are useful for
 constructing semantics used by real systems, (2) the techniques can work
@@
 of application.  We scope the evaluation to one metadata server and scale the
-clients, which execute an extremely metadata intensive job in parallel. This
-type of analysis shows the capacity and performance of a metadata server with
-superior metadata protocols, which should be used to inform load balancing
-across a metadata cluster. Load balancing with partitioning and replication can
-be explored with something like Mantle~\cite{sevilla:sc15-mantle}.
+number of parallel clients. This type of analysis shows the capacity and
+performance of a metadata server with superior metadata protocols, which should
+be used to inform load balancing across a metadata cluster. Load balancing
+across a cluster of metadata servers with partitioning and replication can be
+explored with something like Mantle~\cite{sevilla:sc15-mantle}.
 
@@
 We run experiments on the same cluster and normalize results to control for
-hardware differences and to make our results more generally applicable.  We use
-a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.
-Each node uses Ubuntu 14.04 and has 1 monitor server, 3 object storage daemons,
-1 metadata server, and 29 clients.  All daemons run as a single process which
-is the default setting for Ceph. We develop on Ceph's Jewel release, version
-10.2.1, which was released in May 2016.
-
-This paper adheres to The Popper Convention~\cite{jimenez_popper_2016}, so
-experiments presented here are available in the repository for this article.
-Experiments can be examined in more detail, or even re-run, by visiting the
-\texttt{[source]} link next to each figure. That link points to a Jupyter
-notebook that shows the analysis and source code for that graph, which points
-to an experiment and its artifacts.
+hardware differences and to make our results more generally applicable and
+reproducible\footnote{ This paper adheres to The Popper
+Convention~\cite{jimenez_popper_2016}, so experiments presented here are
+available in the repository for this article.  Experiments can be examined in
+more detail, or even re-run, by visiting the \texttt{[source]} link next to
+each figure. That link points to a Jupyter notebook that shows the analysis and
+source code for that graph, which points to an experiment and its artifacts.}.
+We use a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM,
+and SSDs.  Each node uses Ubuntu 14.04 and We develop on Ceph's Jewel release,
+version 10.2.1, which was released in May 2016. All daemons run as a single
+process (the default setting for Ceph) and the Ceph
+components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
+object storage daemons, 1 metadata server, and 29 clients.
 
@@
 about 2.5\text{KB}. The storage footprint scales linearly with the number of
-metadata creates and suggests that updates for a million files would be
-\(2.5\text{KB}*1\texttt{ million files} = 2.38\text{GB}\). Transfer times for
-payloads of this size in most HPC/data cente networks are reasonable.
+metadata creates and suggests that updates for a million files would be 2.5KB
+\(*\) 1 million files \(=\) 2.38GB. Transfer times for payloads of this size in
+most HPC/data cente networks are reasonable.
 
@@
 allows us to control the size of the log, since each create creates a single
-journal event.
+journal event. The workload is clients creating 100K files in private
+directories in the same global namespace.
 
@@
 applications can compose mechanisms together to get the consistency/durability
-guarantees they need.  We label the \(x\)-axis with systems that employ these
-semantics, as described in Figure~\ref{fig:subtree-policies}.  Again, the
-runtime of different semantics for subtrees in the same global namespace
-normalized to the runtime of the ``Append Client Journal" mechanism (again,
-just creating files in the client's in-memory journal).  We makes no guarantees
+guarantees they need in a global namespace.  We label the \(x\)-axis with
+systems that employ these semantics, as described in
+Figure~\ref{fig:subtree-policies}.  Again, the runtime is normalized to
+creating files in the client's in-memory journal.  We make no guarantees
 during execution of the mechanisms or when transitioning semantics -- the
@@
 during a mechanism, metadata or data may be lost.
+
+\textbf{Takeaway}: we confirm the performance benefits of other
+well-established research systems but, more importantly, we show that the
+Cudele mechanisms we propose are useful for building and evaluating different
+consistency/durability semantics.\\
+
+To show the contention at the metadata server, in Figure~\ref{fig:mergescale}
+we scale the umber of clients for the RPC per operation strategy used by many
+POSIX IO systems and the decoupled namespace strategy used by DeltaFS. For
+RPCs, we scale the number of clients making requests to the metadata server and
+for the decoupled bars, we scale the number of concurrent journal merge
+requests. So for the 30 clients data point, we are measuring the operations per
+second for 30 client journals that land on the metadata server at the same
+time. The ``create \& merge" bar adds the time it takes for clients to generate
+the journal of events in parallel.
+
+%Merge requests are serialized at the metadata server; merging concurrently is
+%an optimization for future work. The current implementation has race conditions
+%because the metadata server recovery code was not designed to run in parallel
+%({\it i.e.} the metadata server never recovers with multiple threads). As a
+%result, the metadata server versions inodes and directory entries to ensure
+%that the recovered metadata is consistent.
+
+This workload scales further with decoupled namespaces because the metadata
+server is not overwhelmed with RPC requests. Compared to
+Figure~\ref{fig:overhead-c} we can scale to 40 clients which is more than
+double the capacity of a single metadata server using the RPC strategy to
+maintain strong consistency. Related work has pointed to on-disk metadata
+format as the primary factor for improved scalability; but our results show
+that the weakened consistency and durability enables the scalability
+improvements. The on-disk metadata formats in this experiment are the same for
+all schemes.
+
+The performance of decoupled namespaces is better than RPCs because we place no
+restrictions on the validity of metadata inserted into the journal and avoid
+touching poorly scaling data structures. We can scale linearly if we weaken our
+consistency by never checking for the existence of files before creating files
+({\it i.e.} only append \texttt{open()} requests to the journal).  When the
+updates are merged by the metadata server into the in-memory metadata store
+they never scan growing data structures.  Had we implemented the ``Create"
+mechanism to include \texttt{lookup()} commands before \texttt{open()}
+requests, we would have seen the poor scaling that we see with the ``RPCs"
+mechanism.
 
@@
 
-{\it Durability \(<<\) Consistency:} The performance of these semantics
-suggests that the overhead of consistency is much larger than the overhead of
-durability. This conclusion should be stronger as we scale the number of files
-because the cost of streaming the journal into the object store is constant.
+%{\it Durability \(<<\) Consistency:} The performance of these semantics
+%suggests that the overhead of consistency is much larger than the overhead of
+%durability. This conclusion should be stronger as we scale the number of files
+%because the cost of streaming the journal into the object store is constant.
+%
+% Weak Consistency benefit not due to metadata format
 
-% Weak Consistency benefit not due to metadata format
-{\it Metadata Formats:} Because the metadata formats are the same for
-all schemes we argue that the performance gain for decoupled namespaces comes
-from relaxing the consistency guarantees and not from the metadata formats.
-
-\textbf{Takeaway}: we confirm the performance benefits of other
-well-established research systems but, more importantly, we show that the
-Cudele mechanisms we propose are useful for building and evaluating different
-consistency/durability semantics.
+\textbf{Takeaway}: the ability to couple well-established techniques to
+specific applications allows the global namespace to scale further and perform
+better.
 
@@
 
-\subsection{Use Case 2: Creates with Rogue Client}
+\subsection{Use Case 2: Creates with Interferring Client}
 
-Clients create files in private directories and a separate client, which we
-call a ``rogue" client, creates files in each directory. This introduces false
-sharing and the metadata server revokes capabilities on directories touched by
-the rogue client. While HPC tries to avoid these situations with
+Next we show how Cudele can be programmed to block interfering clients using
+the same problematic workload from Figure~\ref{fig:overhead-b}. In this
+workload, clients create files in private directories and a separate client
+creates files in each directory. This introduces false sharing and the metadata
+server revokes capabilities on directories touched by the interfering client. While
+HPC tries to avoid these situations with
 workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
 happens in distributed file systems when users unintentionally access
-directories in a shared file system.
+directories in a shared file system.  We also have another client write to a
+decoupled namespace and merge its updates at 90 seconds.  We only scale to 15
+clients because the performance variability of a nearly overloaded metadata
+server, in our case 18 clients, is too high.
 
-% setup: 2cs sep dirs, 2cs DN, 1c malicious write
-Next we show how Cudele can be programmed to block interfering clients. We use
-the same problematic workload from Figure~\ref{fig:overhead-b}, where clients
-write to their own private directories and another client interferes with a
-stream of creates at 30 seconds.  We also have another client write to a
-decoupled namespace and merge its updates at 90 seconds.  We equip each client
-directory with the configuration: allocated\_inodes=100K,
-interfere\_policy=block, consistency=RPCs, and durability=stream.
+Figure~\ref{fig:block-allow} plots the overhead of the slowest client and the
+error bars are the standard deviation of 3 runs.  ``Interfere (Fig 3c)" is the
+result from Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API
+to return \texttt{-EBUSY} to interferring clients; and ``Isolated" is the
+baseline performance without an interferring Each subtree has ``RPCs" and
+``Stream" enabled.  Results results are normalized to the runtime of a single
+client creating files in a directory. Note that IndexFS has a similar behavior
+to ``interfere" with leases, except clients block.
 
-Each directory functions with ``RPCs"
-and ``Stream" enabled -- which is the default implementation of CephFS. The only
-difference is that we enable blocking on each subtree so while the tree behaves
-like CephFS, all interfering operations will be returned with \texttt{-EBUSY}.
-Note that IndexFS does a similar operation with leases except clients block.
-
-% results
-To show the benefits of this isolations, our results in
-Figure~\ref{fig:allow-block} are plotted alongside the variance bars from
-Figure~\ref{fig:overhead-b}. Because ``allow"/``interfere and
-``block"/``isolated" have the same variability we draw the following three
-conclusions: (1) clients that use the API to block interfering clients  get
-the same performance as isolated clients, (2) there is a negligible effect on
-performance for the extra work the metadata server does to return
-\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
+We draw three conclusions: (1) clients that use the API to block interfering
+clients  get the same performance as isolated clients, (2) there is a
+negligible effect on performance for the extra work the metadata server does to
+return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
 negligible effect on performance.
 
-\textbf{Takeaway}: the API lets users isolate directories when
-applications need better and more reliable performance. This is a way of
-controlling consistency.
+\textbf{Takeaway}: the API lets users isolate directories when applications
+need better and more reliable performance. This is a way of controlling
+consistency.
 
@@
 re-scheduled on another node. In this scenario, Cudele users will not see the
-progress of decoupled namespace since their updates are not globally visible.
+progress of decoupled namespaces since their updates are not globally visible.
 To help scientists judge the progress of their jobs, Cudele clients have a
@@
 
-To show the contention at the metadata server of this approach, we scale the
-number of journal merge events, each hitting different directories in the
-namespace, in Figure~\ref{fig:mergescale}. The size of each journal is 100K
-files, the maximum recommended directory size for CephFS. We also plot the cost
-of RPCs with and without journaling; recall from Figure~\ref{fig:overhead}
-that the scalability that max number of clients that the default metadata
-server could support with RPCs is 18.
-
-Merge requests are serialized at the metadata server; merging concurrently is
-an optimization for future work. The current implementation has race conditions
-because the metadata server recovery code was not designed to run in parallel
-({\it i.e.} the metadata server never recovers with multiple threads). As a
-result, the metadata server versions inodes and directory entries to ensure
-that the recovered metadata is consistent.
-
-Performance is better than RPCs because we place no restrictions on the
-validity of metadata inserted into the journal and avoid touching poorly
-scaling data structures. In other words, we can scale linearly if we weaken our
-consistency by never checking for the existence of files before creating files.
-only appends \texttt{open()} requests to the journal. When the updates are
-merged by the metadata server into the in-memory metadata store they never scan
-growing data structures.  Had we implemented the ``Create" mechanism to include
-\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
-the poor scaling that we see with the ``RPCs" mechanism.
-
-\textbf{Takeaway}: the weakest form of consistency for creating files
-({\it i.e.} not checking data structures for the validity of an update or
-existence of a file) shows linear scalability and stable performance.
+\textbf{Takeaway}: syncing namespace updates has up to a 9\% overhead and can
+be tuned depending on the users preference but, more importantly, the API gives
+users fine-grain control of their consistency/durability to support current
+practices or experimental workflows.
 
@@
 
-%\subsection{Discussion}
-%
-%Cudele and the experiments we present here prompt many avenues for future work.
-%The first is embeddable policies, where child subtrees have specialized
-%features but still maintain the guarantees  of their parent subtrees. We
-%already show how applciations can compose guarantees and control their
-%performance, knobs that we did not have before, but embeddable policies are a
-%way to compose the policies themsleves.  For example, a RAMDisk subtree is
-%POSIX IO-compliant but relaxes durability constraints, so it can reside under a
-%POSIX IO subtree that is strongly consistent. These embeddable policies should be
-%controlled and enforced by Cudele.
-%
+\subsection{Discussion and Future Work}
+
+Cudele and the experiments we present here prompt many avenues for future work.
+First is quantifying the benefits of dynamically changing the semantics of a
+subtree from stronger to weaker guarantees (or vice versa), as described in the
+introduction. Second is embeddable policies, where child subtrees have
+specialized features but still maintain the guarantees  of their parent
+subtrees. We already show how applciations can compose guarantees and control
+their performance, knobs that we did not have before, but embeddable policies
+are a way to compose the policies themsleves.  For example, a RAMDisk subtree
+is POSIX IO-compliant but relaxes durability constraints, so it can reside
+under a POSIX IO subtree that is strongly consistent. These embeddable policies
+should be controlled and enforced by Cudele.  Finally, performance prediction
+benefits from our cost quantification of each mechanism. Applications can use
+Cudele to microbenchmark their components and software, similar to what we did
+in Secion~\S\ref{sec:microbenchmarks}. Using those results, they can predict
+how much slower their system will be if they adopt stronger consistency or
+durability.  This is a form of co-design that takes a ``dirty-slate" approach
+but building just the guarantees the application needs from existing
+implementations.  This can also be a verification tool where performance that
+varies wildly from the predicted performance can be a red flag that something
+is wrong or that the bottleneck is not in the consistency or durability plane.
+
 %Second is the opportunity for adding hardware in strategic places to speed up
@@
 %of the disk. So applications that rely on merging many updates should consider
-%equipping the metadata server with SSDs or more memory. This is a form of
-%co-design that takes a ``dirty-slate" approach.
+%equipping the metadata server with SSDs or more memory. 
 %
-%Finally, performance prediction benefits from our cost quantification of each
-%mechanism. Applications can use Cudele to microbenchmark their components and
-%software, similar to what we did in Secion~\S\ref{sec:microbenchmarks}. Using
-%those results, they can predict how much slower their system will be if they
-%adopt stronger consistency or durability. This can also be a verification tool
-%where performance that varies wildly from the predicted performance can be a
-%red flag that something is wrong or that the bottleneck is not in the
-%consistency or durability plane.

### 2017-10-22 paper/sections/7_conclusion.tex
paper: finish eval; touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 double-edged sword. While it achieves better performance and scalability, it
-alienates applications that rely on strong consistency and durability. Cudele
-lets users assign consistency and durability guarantees to subtrees in
-the global namespace, giving users a wide range of policies and
-optimizations that can be custom fit to the application. Using Cudele, we
-compare related work on the same file system and conclude that strong
-consistency can cause a 104\(\times\) slow down while merging updates and
-maintaining durability only have between a \(7-10\times\) slow down.
+alienates applications that rely on strong consistency and durability. Mounting
+other systems to the global namespace is convenient but wastes resources and
+incurs data movement.  Cudele lets users assign consistency and durability
+guarantees to subtrees in the global namespace, which can be custom fit to the
+application. Using Cudele, we show how applications can co-exist and perform
+well in a global namespace and lay the groundwork for new systems that
+dynamically change consistency/durability guarantees to avoid provisioning
+dedicated storage clusters and moving large amounts of data.
 

### 2017-10-23 experiments/cudele-mergescale/ansible/workloads/generate_inos.py
benchmark: bigger mergescale results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -3,9 +3,10 @@
 inos = []
 # default start ino 1099511627780
 start = 1099511627780 + 100000
-for i in range(0, 30):
+for i in range(0, 40):
   inos.append(i*100000000 + start)
 
 for i in inos:
   print i
 print inos 
+print len(inos)

### 2017-10-23 experiments/cudele-mergescale/run.sh

@@ -35,10 +35,14 @@ if [ ! -z $1 ]; then
 fi
 
 mkdir results || true
-for nclients in 30 25 20 15 10 5 1; do
-  ./teardown.sh
-  cp configs_$SITE/hosts hosts
-  $DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/merge.yml #/workloads/clients${nclients}.yml
+#./teardown.sh
+#$DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/generate_journals.yml
+for run in 3 4 5; do
+  for nclients in 40 30 20 10 1; do
+    ./teardown.sh
+    cp configs_$SITE/hosts hosts
+    $DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/merge.yml #/workloads/clients${nclients}.yml
+  done
+  mv results results-run$run
 done
-
 exit 0

### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: add last graph
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 None is different than local durability because regardless of the type failure,
-metadata will be lost when components die in a None configuration.\\
+metadata will be lost when components die in a None configuration.
 
 % - sequential IO, trim redundant operations
-\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
+\textbf{CephFS Design}: a journal of metadata updates that streams into
 the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
@@
 overwhelm the bandwidth of the object store. We measured our peak bandwidth to
-be 100MB/s, which is the speed of our network link.\\
+be 100MB/s, which is the speed of our network link.
 
-\noindent\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
+\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
 to the best of our knowledge, when a client or server fails there is no recovery
@@
 synchronization and serialization machinery needed to ensure that all clients
-see the same state has high overhead.\\
+see the same state has high overhead.
 
-\noindent\textbf{CephFS Design}: capabilities keep metadata strongly
+\textbf{CephFS Design}: capabilities keep metadata strongly
 consistent. To reduce the number of RPCs needed for consistency, clients can
@@
 workload; at higher client load, the metadata server started complaining about
-laggy and unresponsive requests.\\
+laggy and unresponsive requests.
 
@@
 
-\noindent\textbf{Comparison to decoupled namespaces}: Decoupled namespaces
+\textbf{Comparison to decoupled namespaces}: Decoupled namespaces
 merge batches of metadata operations into the global namespaces when the job

### 2017-10-23 paper/sections/6_results.tex
paper: add last graph
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   \end{subfigure}
-  \caption{The performance and features of Cudele. (a) shows the performance of
-merging client journals at the metadata server. Performance is better than RPCs
+\caption{The performance and features of Cudele. (a) shows the cost of
+merging client journals at the metadata server. Merging is better than RPCs
 because there are less messages and consistency/durability code paths are
-bypassed. (b) shows how the allow/block API can isolate directories from
-interfering clients.  (c) shows the slowdown of a single client periodically
-syncing updates to the gobal namespace. The inflection point is the trade-off
-of frequent updates vs. larger journal files.\label{fig:use-cases}}
+bypassed. In (b), the allow/block API isolates directories from interfering
+clients. (c) is the slowdown of a single client syncing updates to the gobal
+namespace. The inflection point is the trade-off of frequent updates vs. larger
+journal files.\label{fig:use-cases}}
 \end{figure*}

### 2017-10-23 paper/sections/1_introduction.tex
paper: patrick feedback and spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-%\begin{table}
-%\begin{tabular}{ r | l }
-%  Subtree         & Example \\\hline
-%  (1)   & \{Index, Batch\}FS~\cite{ren:sc2014-indexfs, zheng:pdsw2014-batchfs} \\
-%  (2)   & \{Index, Ceph\}FS~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} \\
-%  (3)   & RAMDisk \\
-%  (4)   & DeltaFS~\cite{zheng:pdsw2015-deltafs} \\
-%\end{tabular}
-%
-%\caption{State-of-the-art systems in HPC improve file system metadata
-%performance by relaxing consistency and durability guarantees. Note that
-%IndexFS also supports weak consistency with bulk inserts.
-%\label{table:namespaces}} \end{table}
-
 % What did we do
@@
   to subtrees in the file system namespace; this lets users navigate
-  the tradeoffs of different metadata protocols.
+  the trade-offs of different metadata protocols.
 

### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: patrick feedback and spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+% CITE ME Lustre paralle
 Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
 the Linux kernel.  The \texttt{untar} phase, which is characterized by many
-creates, has the highest resource usage which suggests that it is stressing the
-consistency and journaling subsystems of the metadata server the most.
-Traditional file system techniques for improving performance, such as caching
-inodes, do not help for create-heavy workloads.
+creates, has the highest resource usage (combined CPU, network, and disk)
+because of the number of RPCs needed for consistency and durability. The RPCs are
+also serialized because the metadata server is single threaded; although naive,
+this design is common because of the complexity of multi-threaded metadata
+servers.  Traditional file system techniques for improving performance, such as
+caching inodes, do not help for create-heavy workloads.
 
@@
       \centering
-      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
+      \includegraphics[width=1.2\linewidth]{graphs/behavior-interfere.png}
       \caption{Interference forces \texttt{lookup()}s}
@@
 % what is durability
-While durability is not specified by POSIX IO, users expect that files they create
-or modify survive failures.  We define three types of durability: global,
-local, and none.  Global
-durability means that the client or server can fail at any time and metadata
-will not be lost. Local durability means that metadata can be lost if the
-client or server stays down after a failure. None means that metadata is volatile
-and that the system provides no guarantees when clients or servers fail. 
-None is different than local durability because regardless of the type failure,
-metadata will be lost when components die in a None configuration.
+While durability is not specified by POSIX IO, users expect that files they
+create or modify survive failures.  We define three types of durability:
+global, local, and none.  Global durability means that the client or server can
+fail at any time and metadata will not be lost because it is ``safe" ({\it
+i.e.} striped or replicated across a cluster). Local durability means that
+metadata can be lost if the client or server stays down after a failure. None
+means that metadata is volatile and that the system provides no guarantees when
+clients or servers fail.  None is different than local durability because
+regardless of the type failure, metadata will be lost when components die in a
+None configuration.
 
 % - sequential IO, trim redundant operations
-\textbf{CephFS Design}: a journal of metadata updates that streams into
-the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
-ensures (1) sequential writes into the object store and (2) the ability for daemons to
-trim redundant or irrelevant journal entries.  The journal is striped over
-objects where multiple journal updates can reside on the same object. There are
-two tunables for controlling the journal: the segment size and the number of
-parallel segments that can be written in parallel. Unless the journal saturates 
-memory or CPU resources, larger values for these tunables results in better
-performance.
+\textbf{CephFS Design}: a journal of metadata updates that streams into the
+resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
+WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow large
+(on the order of MBs) which ensures (1) sequential writes into the object store
+and (2) the ability for daemons to trim redundant or irrelevant journal
+entries.  The journal is striped over objects where multiple journal updates
+can reside on the same object. There are two tunables for controlling the
+journal: the segment size and the number of segments that can be written in
+parallel. Unless the journal saturates memory or CPU resources, larger values
+for these tunables results in better performance.
 
@@
 
+% CITEME: SSTables
 \textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
@@
 The benefits of caching the directory inode when creating files is shown in
-Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
-directory (``isolated" curve on \(y1\) axis) then that client can lookup the
-existence of new files locally before issuing a create request to the metadata
-server. If another client starts creating files in the same directory
-(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
-read caching and the first client must send \texttt{lookup()}s to the metadata
-server (``interfere" curve on \(y2\) axis). These extra requests increase the
+Figure~\ref{fig:overhead-b}. The colors show the behavior of the client for two
+different runs.  If only one client is creating files in a directory
+(``isolated" curve on \(y1\) axis) then that client can lookup the existence of
+new files locally before issuing a create request to the metadata server. If
+another client starts creating files in the same directory (``interfere" curve
+on \(y1\) axis) then the directory inode transitions out of read caching and
+the first client must send \texttt{lookup()}s to the metadata server
+(``interfere" curve on \(y2\) axis). These extra requests increase the
 throughput of the ``interfere" curve because the metadata server can handle the

### 2017-10-23 paper/sections/4_methodology.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: patrick feedback and spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \caption{Applications decouple the namespace, write updates to a local journal,
-and delay metadata updates using the Cudele Mechanisms }\label{fig:decouple}
+and delay updates using the \textbf{mechanisms} shown here; they are described
+in Table~\ref{table:mechanisms}.}\label{fig:decouple}
 \end{figure}
@@
 constructing consistency and durability guarantees. Cudele exposes these
-mechanisms and the user composes them together to construct
-\textbf{policies}. These policies are assigned to subtrees and they dictate how
-the file system should handle operations within that subtree.  Below, we
-describe the mechanisms, the policies, and the API for assigning policies to
-subtrees.
+mechanisms and the user composes them together to construct \textbf{policies}.
+These policies are assigned to subtrees and they dictate how the file system
+should handle operations within that subtree.  Below, we describe the
+mechanisms (which are \underline{underlined}), the policies, and the API for assigning
+policies to subtrees.
 
@@
 which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
-description of what each mechanism does.  Decoupled clients use the ``Append Client Journal"
-mechanism to append metadata updates to a local, in-memory journal. Once the
-job is complete, the system calls Cudele mechanisms to achieve the desired
-consistency/durability semantics.  Cudele provides a library for clients to
-link into and all operations are performed by the client.
+description of what each mechanism does.  Decoupled clients use the
+\underline{Append Client Journal} mechanism to append metadata updates to a
+local, in-memory journal. This is fast because clients do not need to check for
+consistency when writing events and the metadata server blindly applies the
+updates because it assumes the events were already checked for consistency. The
+trade-off here is performance; it is a dangerous approach but could be
+implemented safely if the clients or metadata server are configured to check
+the validity of events before writing them.  Once the job is complete, the
+system calls Cudele mechanisms to achieve the desired consistency/durability
+semantics.  Cudele provides a library for clients to link into and all
+operations are performed by the client.
 
-\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
+\subsubsection{Mechanisms Used for Consistency} \underline{RPCs} send remote procedure
 calls for every metadata operation from the client to the metadata server,
@@
 part of the default CephFS implementation and is the strongest form of
-consistency because clients see metadata updates right away.  ``Nonvolatile
-Apply" replays the client's journal onto the metadata cluster's metadata store.
+consistency because clients see metadata updates right away.  \underline{Nonvolatile
+Apply} replays the client's journal onto the metadata cluster's metadata store.
 The client's in-memory journal is written into the object store and the
@@
 notice new journal updates in the object store and replay the events onto their
-in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
+in-memory metadata stores.  \underline{Volatile Apply} takes the client's in-memory journal on the
 client and applies the updates directly to the in-memory metadata store maintained
@@
 performance -- Cudele makes no consistency or durability guarantees while
-``Volatile Apply" is executing.  If a concurrent update from a client occurs
+Volatile Apply is executing.  If a concurrent update from a client occurs
 there is no rule for resolving conflicts and if the client or metadata server
@@
 % difference between apply and volatile apply
-The biggest difference between ``Volatile Apply" and ``Nonvolatile Apply" is
-the medium they use to communicate. ``Volatile Apply" applies updates directly
-to the metadata servers' metadata store while ``Nonvolatile Apply" uses the
+The biggest difference between Volatile Apply and Nonvolatile Apply is
+the medium they use to communicate. Volatile Apply applies updates directly
+to the metadata servers' metadata store while Nonvolatile Apply uses the
 object store to communicate the journal of updates from the client to the
-metadata servers.  ``Nonvolatile Apply" is safer but has a large performance
+metadata servers.  Nonvolatile Apply is safer but has a large performance
 overhead because objects in the metadata store need to be read from and written
@@
 %Persist. ``RPCs" does round trip remote procedure calls to establish
-%consistency; it is the default implementation for complying with POSIX in
+%consistency; it is the default implementation for complying with POSIX IO in
 %CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
@@
 
-\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
+\subsubsection{Mechanisms Used for Durability} \underline{Stream} is one of the
 mechanisms used by default in CephFS.  Using existing configuration settings in
-Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
-will not save journals in the object store. For ``Local Persist", clients write
-serialized log events to a file on local disk and for ``Global Persist",
-clients push the journal into the objects store. The overheads for both ``Local
-Persist" and ``Global Persist" is the write bandwidth of the local disk and
+Ceph we can turn Stream on and off.  If it is off, then the metadata servers
+will not save journals in the object store. For \underline{Local Persist}, clients write
+serialized log events to a file on local disk and for \underline{Global Persist},
+clients push the journal into the objects store. The overheads for both Local
+Persist and Global Persist is the write bandwidth of the local disk and
 object store, respectively.  These persist mechanisms are part of the library that
@@
 % describe table
-The spectrum of consistency and durability guarantees that users can
-construct is shown in Table~\ref{table:spectrum}. The columns are the different
+The spectrum of consistency and durability guarantees that users can construct
+is shown in Table~\ref{table:spectrum}. The columns are the different
 consistency semantics and the rows cover the spectrum of durability guarantees.
@@
 into a global namespace and it is assumed that middleware or the application
-manages consistency lazily; ``weak" merges updates at some time in the
-future ({\it e.g.}, when the system has time, when the number of updates reaches a
+manages consistency lazily; ``weak" merges updates at some time in the future
+({\it e.g.}, when the system has time, when the number of updates reaches a
 certain threshold, when the client is done writing, etc.); and updates in
@@
 guarantees are made with ``local", which means updates will be retained if the
-client node recovers, and ``global", where all updates are always recoverable.
+client node recovers and reads the updates from local storage, and ``global",
+where all updates are always recoverable.
 
@@
 Existing, state-of-the-art systems in HPC can be represented by the cells in
-Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
+Table~\ref{table:spectrum}.  POSIX IO-compliant systems like CephFS and IndexFS
 have global consistency and durability~\footnote{This is the normal case.
@@
 should be considered atomic and there are no guarantees while transitioning
-between policies. For example, updates are not deemed to have ``global" durability until
-they are safely saved in the object store. If a failure occurs during ``global
-persist" or if we inject a new policy that changes a subtree from ``local
-persist" to ``global persist", Cudele makes no guarantee until the mechanisms are
-complete.
+between policies. For example, updates are not deemed to have ``global"
+durability until they are safely saved in the object store. If a failure occurs
+during Global Persist or if we inject a new policy that changes a subtree from
+Local Persist to Global Persist, Cudele makes no guarantee until the mechanisms
+are complete. Despite this, production systems that use Cudele should state
+up-front what the transition guarantees are for subtrees.
 
@@
 % the interface
-Users control consistency and durability for subtrees by contacting the monitor
-daemon with a directory path and presenting a policies configuration file.
-For example, (msevilla/mydir, policies.yml) would decouple the path
-``msevilla/mydir" and would apply the policies in ``policies.yml".  The
-policies file supports the following parameters (default values are in parenthesis):
+Users control consistency and durability for subtrees by contacting a daemon in
+the system called a monitor, which manages cluster state changes.  Users
+present a directory path and a policies configuration that gets distributed and
+versioned by the monitor to all daemons in the system. For example,
+(msevilla/mydir, policies.yml) would decouple the path ``msevilla/mydir" and
+would apply the policies in ``policies.yml".  The policies file supports the
+following parameters (default values are in parenthesis):
 

### 2017-10-23 paper/sections/6_results.tex
paper: patrick feedback and spellcheck
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
-\caption{[\href{https://www.google.com}{source}] Microbenchmark: performance of each
+\caption{[\href{https://www.google.com}{source}] Performance of each
 mechanism (left) and building consistency/durability guarantees (right) for
-100K files on a single client. Results are normalized to the runtime of writing
-files creates to the client's in-memory journal.
+100K files creates from a single client. Results are normalized to the runtime of writing
+events to the client's in-memory journal.
 \label{fig:composable-mechanisms}}
@@
 \(*\) 1 million files \(=\) 2.38GB. Transfer times for payloads of this size in
-most HPC/data cente networks are reasonable.
+most HPC/data center networks are reasonable.
 
@@
 bypassed. In (b), the allow/block API isolates directories from interfering
-clients. (c) is the slowdown of a single client syncing updates to the gobal
+clients. (c) is the slowdown of a single client syncing updates to the global
 namespace. The inflection point is the trade-off of frequent updates vs. larger
@@
 % CITEME xiao:socc2015-shardfs CITEME HADOOP
-Clients creating files in private directories is heavily studied in
-HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
-zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
-checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
-cloud workloads, where systems like Hadoop use the file abstraction to exchange
-work units to workers or to indicate when jobs
+First we show how we can build application-specific subtrees by composing
+mechanisms and that this approach results in more a scalable global
+namespace. We start with clients creating files in private directories because
+this workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
+ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
+sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
+But the workload also appears in cloud workloads, where systems like Hadoop use
+the file abstraction to exchange work units to workers or to indicate when jobs
 complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
@@
 
-\subsection{Use Case 2: Creates with Interferring Client}
+\subsection{Use Case 2: Creates with Interfering Client}
 
@@
 result from Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API
-to return \texttt{-EBUSY} to interferring clients; and ``Isolated" is the
-baseline performance without an interferring Each subtree has ``RPCs" and
+to return \texttt{-EBUSY} to interfering clients; and ``Isolated" is the
+baseline performance without an interfering. Each subtree has ``RPCs" and
 ``Stream" enabled.  Results results are normalized to the runtime of a single
@@
 % CITEME
-Scientists use the file system to check the progress of jobs using
-\texttt{ls}~\cite{CITEME}. The number of files or size of the files is
-indicative of the progress. This practice is not too different from systems
-that use the file system to manage the progress of jobs; Hadoop writes to
-temporary files, renames them when complete, and creates a ``DONE" file to
-indicate to the runtime that the task did not fail and should not be
-re-scheduled on another node. In this scenario, Cudele users will not see the
-progress of decoupled namespaces since their updates are not globally visible.
-To help scientists judge the progress of their jobs, Cudele clients have a
-``namespace sync" that sends batches of updates back to the global namespace at
-regular intervals.
+The final use case shows how the API gives users fine-grained control of the
+consistency semantics for niche or {\it ad-hoc} workflows.  Scientists use the
+file system to check the progress of jobs using \texttt{ls}~\cite{CITEME}. The
+number of files or size of the files is indicative of the progress. This
+practice is not too different from systems that use the file system to manage
+the progress of jobs; Hadoop writes to temporary files, renames them when
+complete, and creates a ``DONE" file to indicate to the runtime that the task
+did not fail and should not be re-scheduled on another node. In this scenario,
+Cudele users will not see the progress of decoupled namespaces since their
+updates are not globally visible.  To help scientists judge the progress of
+their jobs, Cudele clients have a ``namespace sync" that sends batches of
+updates back to the global namespace at regular intervals.
 
@@
 specialized features but still maintain the guarantees  of their parent
-subtrees. We already show how applciations can compose guarantees and control
+subtrees. We already show how applications can compose guarantees and control
 their performance, knobs that we did not have before, but embeddable policies
-are a way to compose the policies themsleves.  For example, a RAMDisk subtree
+are a way to compose the policies themselves.  For example, a RAMDisk subtree
 is POSIX IO-compliant but relaxes durability constraints, so it can reside
@@
 Cudele to microbenchmark their components and software, similar to what we did
-in Secion~\S\ref{sec:microbenchmarks}. Using those results, they can predict
+in Section~\S\ref{sec:microbenchmarks}. Using those results, they can predict
 how much slower their system will be if they adopt stronger consistency or

### 2017-10-23 paper/sections/5_implementation.tex
paper: remove popper :'(
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % how it works in CephFS, how I used it
-\subsection{Metadata Store}
-\label{sec:metadata-store}
+%\subsection{Metadata Store}
+%\label{sec:metadata-store}
 
 % how it works in CephFS
-In CephFS, the metadata store is a data structure that represents the file
+\textbf{Metadata Store}: In CephFS, the metadata store is a data structure that represents the file
 system namespace. This data structure is stored in two places: in memory ({\it
@@
 
-\subsection{Journal}
-\label{sec:journal}
+%\subsection{Journal}
+%\label{sec:journal}
 
-The journal is the second way that CephFS represents the file system namespace;
+\textbf{Journal}: The journal is the second way that CephFS represents the file system namespace;
 it is a log of metadata updates that can materialize the namespace when the
@@
 
-\subsection{Journal Tool}
-\label{sec:journal-tool}
+%\subsection{Journal Tool}
+%\label{sec:journal-tool}
 
-The journal tool is used for disaster recovery and lets administrators view and
+\textbf{Journal Tool}: The journal tool is used for disaster recovery and lets administrators view and
 modify the journal. It can read the journal, export the journal as a file,
@@
 
-\subsection{Inode Cache}
-\label{sec:inode-cache}
+%\subsection{Inode Cache}
+%\label{sec:inode-cache}
 
-In CephFS, the inode cache reduces the number of RPCs between clients and
+\textbf{Inode Cache}: In CephFS, the inode cache reduces the number of RPCs between clients and
 metadata servers. Without contention clients can resolve metadata reads locally
@@
 code to allocate inodes to clients that decouple parts of the namespace and to
-skip inodes used by the client at merge time.\\
+skip inodes used by the client at merge time.
 
-\subsection{Large Inodes}
-\label{sec:large-inodes}
+%\subsection{Large Inodes}
+%\label{sec:large-inodes}
 
-In CephFS, inodes already store policies, like how the file is striped
+\textbf{Large Inodes}: In CephFS, inodes already store policies, like how the file is striped
 across the object store or for managing subtrees for load balancing. These

### 2017-10-23 paper/sections/6_results.tex
paper: remove popper :'(
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-We run experiments on the same cluster and normalize results to control for
-hardware differences and to make our results more generally applicable and
-reproducible\footnote{ This paper adheres to The Popper
-Convention~\cite{jimenez_popper_2016}, so experiments presented here are
-available in the repository for this article.  Experiments can be examined in
-more detail, or even re-run, by visiting the \texttt{[source]} link next to
-each figure. That link points to a Jupyter notebook that shows the analysis and
-source code for that graph, which points to an experiment and its artifacts.}.
-We use a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM,
-and SSDs.  Each node uses Ubuntu 14.04 and We develop on Ceph's Jewel release,
-version 10.2.1, which was released in May 2016. All daemons run as a single
-process (the default setting for Ceph) and the Ceph
-components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
-object storage daemons, 1 metadata server, and 29 clients.
+%We run experiments on the same cluster and normalize results to control for
+%hardware differences and to make our results more generally applicable and
+%reproducible\footnote{ This paper adheres to The Popper
+%Convention~\cite{jimenez_popper_2016}, so experiments presented here are
+%available in the repository for this article.  Experiments can be examined in
+%more detail, or even re-run, by visiting the \texttt{[source]} link next to
+%each figure. That link points to a Jupyter notebook that shows the analysis and
+%source code for that graph, which points to an experiment and its artifacts.}.
+%We use a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM,
+%and SSDs.  Each node uses Ubuntu 14.04 and We develop on Ceph's Jewel release,
+%version 10.2.1, which was released in May 2016. All daemons run as a single
+%process (the default setting for Ceph) and the Ceph
+%components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
+%object storage daemons, 1 metadata server, and 29 clients.
 
@@
 \includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
-\caption{[\href{https://www.google.com}{source}] Performance of each
-mechanism (left) and building consistency/durability guarantees (right) for
-100K files creates from a single client. Results are normalized to the runtime of writing
-events to the client's in-memory journal.
-\label{fig:composable-mechanisms}}
+\caption{Performance of each mechanism (left) and building
+consistency/durability guarantees (right) for 100K files creates from a single
+client. Results are normalized to the runtime of writing events to the client's
+in-memory journal.  \label{fig:composable-mechanisms}}
 \end{figure}

### 2017-10-23 paper/build.sh

@@ -2,6 +2,7 @@
 
 rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
 set -x
+cp biblio/references.bib paper.bib
 
 # Build the paper
 docker run --rm \

### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: peter edits
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 creates, has the highest resource usage (combined CPU, network, and disk)
-because of the number of RPCs needed for consistency and durability. The RPCs are
-also serialized because the metadata server is single threaded; although naive,
-this design is common because of the complexity of multi-threaded metadata
-servers.  Traditional file system techniques for improving performance, such as
-caching inodes, do not help for create-heavy workloads.
-
-In this section, we quantify the costs of strong consistency and global
-durability in CephFS. At the end of each subsection we compare the approach to
-``decoupled namespaces", the technique in related work that detaches subtrees
-from the global namespace to relax consistency/durability guarantees.  We use
-the kernel client so that we can find the true create speed of the server; our
-experiments show a low CPU utilization for the clients which indicates that we
-are stressing the servers more.
+because of the number of RPCs needed for consistency and durability. The RPCs
+are also serialized because the metadata server is single threaded; although
+naive, this design is common because of the complexity of multi-threaded
+metadata servers.  Traditional file system techniques for improving
+performance, such as caching inodes, do not help for create-heavy workloads.
+In this section, we quantify the costs of consistency and durability in CephFS.
+At the end of each subsection we compare the approach to ``decoupled
+namespaces", the technique in related work that detaches subtrees from the
+global namespace to relax consistency/durability guarantees.
 
@@
 resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow large
+WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to be large
 (on the order of MBs) which ensures (1) sequential writes into the object store

### 2017-10-23 paper/sections/4_methodology.tex
paper: peter edits
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Existing, state-of-the-art systems in HPC can be represented by the cells in
-Table~\ref{table:spectrum}.  POSIX IO-compliant systems like CephFS and IndexFS
-have global consistency and durability~\footnote{This is the normal case.
-IndexFS also has bulk merge which would transition the system into ``weak
-consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
-and BatchFS uses ``weak" consistency and ``local" durability. These systems
-have other features that could push them into different semantics but we assign
-labels here based on the points emphasized in the papers.  To compose the
-mechanisms users inject which mechanisms to run and which to use in parallel
-using a domain specific language.  Although we can achieve all permutations of
-the different guarantees in Table~\ref{table:spectrum}, not all of them make
-sense. For example, it makes little sense to do \texttt{append client
-journal+RPCs} since both mechanisms do the same thing or \texttt{stream+save}
-since ``global" durability is stronger and has more overhead than ``local"
-durability. The cost of each mechanism and the semantics described above are
-quantified in Sections~\S\ref{sec:microbenchmarks} and~\S\ref{sec:use-case-1}.
+Table~\ref{table:spectrum} (we construct the semantics for these systems later
+in Section~\S\ref{sec:microbenchmarks}).  POSIX IO-compliant systems like
+CephFS and IndexFS have global consistency and durability~\footnote{This is the
+normal case.  IndexFS also has bulk merge which would transition the system
+into ``weak consistency"}; DeltaFS uses ``invisible" consistency and ``local"
+durability and BatchFS uses ``weak" consistency and ``local" durability. These
+systems have other features that could push them into different semantics but
+we assign labels here based on the points emphasized in the papers.  To compose
+the mechanisms users inject which mechanisms to run and which to use in
+parallel using a domain specific language.  Although we can achieve all
+permutations of the different guarantees in Table~\ref{table:spectrum}, not all
+of them make sense. For example, it makes little sense to do \texttt{append
+client journal+RPCs} since both mechanisms do the same thing or
+\texttt{stream+save} since ``global" durability is stronger and has more
+overhead than ``local" durability. The cost of each mechanism and the semantics
+described above are quantified in Sections~\S\ref{sec:microbenchmarks}
+and~\S\ref{sec:use-case-1}.
 

### 2017-10-23 paper/sections/6_results.tex
paper: peter edits
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\textbf{Takeaway}: measuring the mechanisms individually shows that
-their overheads and costs can differ {\it by orders of magnitude}. Cudele gives
-users the ability to quantify the performance impact of different
-consistency/durability guarantees.\\
+\textbf{Takeaway}: measuring the mechanisms individually shows that their
+overheads and costs can differ {\it by orders of magnitude}. We also show that
+some mechanisms, like Nonvolatile Apply, are not worthwhile as currently
+implemented.
 
@@
 \textbf{Takeaway}: the API lets users isolate directories when applications
-need better and more reliable performance. This is a way of controlling
-consistency.
+need better and more reliable performance. Blocking updates is an effective way of
+controlling consistency.
 

### 2017-10-23 paper/sections/6_results.tex
paper: ken edits
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 uses these data structures will experience similar slowdowns. Other mechanisms
-that write events to a journal ({\it e.g.} the persists, ``Volatile Apply")
-experience a much less drastic slowdown because the journal data structure does
-not need to be scanned for every operation. Events are written to the end of
-the journal without even checking the validity ({\it e.g.}, if the file already
-exists for a create), which is another form of relaxed consistency because the
-file system assumes the application has resolved conflicting updates in a
-different way.
+that write events to a journal) experience a much less drastic slowdown because
+the journal data structure does not need to be scanned for every operation.
+Events are written to the end of the journal without even checking the validity
+({\it e.g.}, if the file already exists for a create), which is another form of
+relaxed consistency because the file system assumes the application has
+resolved conflicting updates in a different way.
 
@@
 Apply" because sending individual metadata updates over the network is costly.
-While ``RPCs" sends a request for every file create, ``Nonvolatile Apply"
+While ``RPCs" sends a request for every file create, ``Volatile Apply"
 writes all the updates to the in-memory journal and applies them to the

### 2017-10-23 paper/sections/1_introduction.tex
paper: jeff comments/citations and fix figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
-paper, although the HDFS semantics and their performance is explored
-in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
-consistency because it does not guarantee that all updates are immediately seen
-by all servers; the BatchFS and RAMDisk subtrees are decoupled from the global
-namespace and have similar consistency/durability behavior to those systems;
-and the POSIX IO subtree retains the rigidity of POSIX IO's strong consistency.
-Subtrees without policies inherit the consistency/durability semantics of the
-parent and future work will examine embeddable or inheritable policies.
+paper, although the semantics and their performance is explored
+in~\S\ref{sec:use-case-1}.} subtree has weaker than strong consistency because
+for performance it lets clients read files opened for
+writing~\cite{hakimzadeh:dais14-hdfs-consistency}, which means that not all
+updates are immediately seen by all servers; the BatchFS and RAMDisk subtrees
+are decoupled from the global namespace and have similar consistency/durability
+behavior to those systems; and the POSIX IO subtree retains the rigidity of
+POSIX IO's strong consistency.  Subtrees without policies inherit the
+consistency/durability semantics of the parent and future work will examine
+embeddable or inheritable policies.
 
@@
   to subtrees in the file system namespace; this lets users navigate
-  the trade-offs of different metadata protocols.
+  the trade-offs of different metadata protocols on the same storage system.
 
@@
   global namespace. We show how this scales further and performs better than
-  systems that use one strategy for the entire namespace
+  systems that use one strategy for the entire namespace .
 
@@
 The last contribution lays the groundwork for future work on our prototype. It
-improves the current practice of mounting different storage systems in a global
+is better than the current practice of mounting different storage systems in a global
 namespace because there is no need to provision dedicated storage clusters to
@@
 slow down) and maintaining durability (\(10\times\) slow down). In our
-prototype, we get an 8\(\times\) speedup and can scale to twice as many servers
+prototype, we get an 8\(\times\) speedup and can scale to twice as many clients
 when we assign a more relaxed form of consistency and durability to a subtree
-with a create-heavy workload.  We build on Ceph because it is used in
-cloud-based and data center systems and has a presence in HPC~\cite{}.  More
-importantly,  we use Ceph as a platform for exploring related work over the
-same storage system, which facilitates an apples-to-apples comparison of the
-strategies themselves.
+with a create-heavy workload.  We use Ceph as a prototyping platform because it
+is used in cloud-based and data center systems and has a presence in
+HPC~\cite{wang:pdsw13-ceph-hpc}.
 

### 2017-10-23 paper/sections/2_related_work.tex
paper: jeff comments/citations and fix figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % General
-The bottlenecks associated with accessing POSIX IO file system metadata are not limited
-to HPC workloads and the same challenges that plagued these systems for years are
-finding their way into the cloud. Workloads that deal with many small files
-({\it e.g.}, log processing and database
+The bottlenecks associated with accessing POSIX IO file system metadata are not
+limited to HPC workloads and the same challenges that plagued these systems for
+years are finding their way into the cloud. Workloads that deal with many small
+files ({\it e.g.}, log processing and database
 queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
@@
 jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
-the metadata service. The biggest challenge is that whenever a file
-is touched the client must access the file's metadata and maintaining a file
-system namespace imposes small, frequent accesses on the underlying storage
+the metadata service. The biggest challenge is that whenever a file is touched
+the client must access the file's metadata and maintaining a file system
+namespace imposes small, frequent accesses on the underlying storage
 system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
@@
 not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
-abad:techreport2012-fstrace, abad:ucc2012-mimesis,
-alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. 
-%There are two approaches
-%for improving the performance of metadata access.
+abad:ucc2012-mimesis, alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. 
+%There are two approaches for improving the performance of metadata access.
 %
@@
 %% approaches to load balancing (strong = more resuorces per unit work, weak =
-%% fixed resource per work unit)
-%% CITEME
-%One approach for improving metadata performance and scalability is to alleviate
-%overloaded servers by load balancing metadata IO across a cluster. Common
-%techniques include partitioning metadata when there are many writes and
-%replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
+%% fixed resource per work unit) % CITEME One approach for improving metadata
+%performance and scalability is to alleviate overloaded servers by load
+%balancing metadata IO across a cluster. Common techniques include partitioning
+%metadata when there are many writes and replicating metadata when there are
+%many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
 %directories and clients write to different partitions by grabbing leases and
@@
 %replication policies do not change depending on the balancing technique.
-%Despite the performance benefits these techniques add complexity and jeopardize
-%the robustness and performance characteristics of the metadata service because
-%the systems now need (1) policies to guide the migration decisions and (2)
-%mechanisms to address inconsistent states across servers.
+%Despite the performance benefits these techniques add complexity and
+%jeopardize the robustness and performance characteristics of the metadata
+%service because the systems now need (1) policies to guide the migration
+%decisions and (2) mechanisms to address inconsistent states across servers.
 %
-%% policies
-%Setting policies for migrations is arguably more difficult than adding the
-%migration mechanisms themselves.  For example, IndexFS and CephFS use the
-%GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
-%threshold and using lazy synchronization to redirect queries to the server that
-%``owns" the targeted metadata.  Determining when to partition directories and
-%when to migrate the directory fragments are policies that vary between systems:
-%GIGA+ partitions directories when the size reaches a certain number of files
-%and migrates directory fragments immediately; CephFS partitions directories
-%when they reach a threshold size or when the write temperature reaches a
-%certain value and migrates directory fragments when the hosting server has more
-%load than the other servers in the metadata cluster. Another policy is when and
-%how to replicate directory state; ShardFS replicates immediately and
-%pessimistically while CephFS replicates only when the read temperature reaches
-%a threshold.  There is a wide range of policies and it is difficult to maneuver
-%tunables and hard-coded design decisions.
+%% policies Setting policies for migrations is arguably more difficult than
+%adding the migration mechanisms themselves.  For example, IndexFS and CephFS
+%use the GIGA+~\cite{patil:fast2011-giga} technique for partitioning
+%directories at a predefined threshold and using lazy synchronization to
+%redirect queries to the server that ``owns" the targeted metadata.
+%Determining when to partition directories and when to migrate the directory
+%fragments are policies that vary between systems: GIGA+ partitions directories
+%when the size reaches a certain number of files and migrates directory
+%fragments immediately; CephFS partitions directories when they reach a
+%threshold size or when the write temperature reaches a certain value and
+%migrates directory fragments when the hosting server has more load than the
+%other servers in the metadata cluster. Another policy is when and how to
+%replicate directory state; ShardFS replicates immediately and pessimistically
+%while CephFS replicates only when the read temperature reaches a threshold.
+%There is a wide range of policies and it is difficult to maneuver tunables and
+%hard-coded design decisions.
 %
-%% addressing inconsistency
-%In addition to the policies, distributing metadata across a cluster requires
-%distributed transactions and cache coherence protocols to ensure strong
-%consistency ({e.g.}, POSIX IO).  For example, ShardFS pessimistically replicates
-%directory state and uses optimistic concurrency control for conflicts; namely
-%it does the operation and if there is a conflict at verification time it falls
-%back to two-phase locking.  Another example is IndexFS's inode cache which
-%reduces RPCs by caching ancestor paths -- the locality of this cache can be
-%thrashed by random reads but performs well for metadata writes. For
-%consistency, writes to directories in IndexFS block until the lease expires
-%while writes to directories in ShardFS are slow for everyone as it either
-%requires serialization or locking with many servers; reads in IndexFS are
-%subject to cache locality while reads in ShardFS always resolve to 1 RPC.
-%Another example of the overheads of addressing inconsistency is how CephFS
-%maintains client sessions and inode caches for capabilities (which in turn make
-%metadata access faster). When metadata is exchanged between metadata servers
-%these sessions/caches must be flushed and new statistics exchanged with a
-%scatter-gather process; this halts updates on the directories and blocks until
-%the authoritative metadata server responds.  These protocols are discussed in
-%more detail in the next section but their inclusion here is a testament to the
-%complexity of migrating metadata.
+%% addressing inconsistency In addition to the policies, distributing metadata
+%across a cluster requires distributed transactions and cache coherence
+%protocols to ensure strong consistency ({e.g.}, POSIX IO).  For example,
+%ShardFS pessimistically replicates directory state and uses optimistic
+%concurrency control for conflicts; namely it does the operation and if there
+%is a conflict at verification time it falls back to two-phase locking.
+%Another example is IndexFS's inode cache which reduces RPCs by caching
+%ancestor paths -- the locality of this cache can be thrashed by random reads
+%but performs well for metadata writes. For consistency, writes to directories
+%in IndexFS block until the lease expires while writes to directories in
+%ShardFS are slow for everyone as it either requires serialization or locking
+%with many servers; reads in IndexFS are subject to cache locality while reads
+%in ShardFS always resolve to 1 RPC.  Another example of the overheads of
+%addressing inconsistency is how CephFS maintains client sessions and inode
+%caches for capabilities (which in turn make metadata access faster). When
+%metadata is exchanged between metadata servers these sessions/caches must be
+%flushed and new statistics exchanged with a scatter-gather process; this halts
+%updates on the directories and blocks until the authoritative metadata server
+%responds.  These protocols are discussed in more detail in the next section
+%but their inclusion here is a testament to the complexity of migrating
+%metadata.
 %
@@
 %%server to ensure strong consistency, non-monotonic writes grab locks at every
-%%server. Zooming in on monotonic writes: if permissions increase it executes on
-%%a primary then non-primary, if permissions decrease it executes on all
+%%server. Zooming in on monotonic writes: if permissions increase it executes
+%on %a primary then non-primary, if permissions decrease it executes on all
 %%non-primary then on primary.
@@
 consistency themselves on a read or add a bolt-on service to manage the
-consistency.
-
-For (2), if the client fails and stays down, all metadata operations on the
-decoupled namespace are lost. If the client recovers, the on-disk structures
-(for BatchFS and DeltaFS this is the SSTables used in TableFS) can be
-recovered. In other words, the clients have state that cannot be recovered if
-the node stays failed and any progress will be lost. This scenario is a
-disaster for checkpoint-restart where missed cycles may cause the checkpoint to
-bleed over into computation time.
-
-For (3), decoupled namespace approaches sacrifice POSIX IO going as far as
-requiring the application to link against the systems they want to talk to. In
-today's world of software defined caching, this can be a problem for large data
-centers with many types and tiers of storage. Despite well-known performance
-problems POSIX IO and REST are the dominant APIs for data transfer.
+consistency. For (2), if the client fails and stays down, all metadata
+operations on the decoupled namespace are lost. If the client recovers, the
+on-disk structures (for BatchFS and DeltaFS this is the SSTables used in
+TableFS) can be recovered. In other words, the clients have state that cannot
+be recovered if the node stays failed and any progress will be lost. This
+scenario is a disaster for checkpoint-restart where missed cycles may cause the
+checkpoint to bleed over into computation time. For (3), decoupled namespace
+approaches sacrifice POSIX IO going as far as requiring the application to link
+against the systems they want to talk to. In today's world of software defined
+caching, this can be a problem for large data centers with many types and tiers
+of storage. Despite well-known performance problems POSIX IO and REST are the
+dominant APIs for data transfer.

### 2017-10-23 paper/sections/7_conclusion.tex
paper: jeff comments/citations and fix figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-Relaxing consistency and durability semantics in the file system is a
-double-edged sword. While it achieves better performance and scalability, it
-alienates applications that rely on strong consistency and durability. Mounting
-other systems to the global namespace is convenient but wastes resources and
-incurs data movement.  Cudele lets users assign consistency and durability
-guarantees to subtrees in the global namespace, which can be custom fit to the
-application. Using Cudele, we show how applications can co-exist and perform
-well in a global namespace and lay the groundwork for new systems that
-dynamically change consistency/durability guarantees to avoid provisioning
-dedicated storage clusters and moving large amounts of data.
+Relaxing consistency/durability semantics in the file system is a double-edged
+sword. While it performs and scales better, it alienates applications that rely
+on strong consistency and durability. Mounting other systems to the global
+namespace is convenient but wastes resources and incurs data movement.  Cudele
+lets users assign consistency/durability guarantees to subtrees in the global
+namespace, which can be custom fit to the application. Using Cudele, we show
+how applications can co-exist and perform well in a global namespace and lay
+the groundwork for new systems that dynamically change consistency/durability
+guarantees to avoid provisioning dedicated storage clusters and moving large
+amounts of data.
 

### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: carlos feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: fill in all CITEMEs, hammer on main points
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-% CITEME PDSW paper
-In our examination of the overheads of POSIX IO we benchmark and analyze CephFS,
-the file system that uses Ceph's object store ({\it i.e.} RADOS) to store its
-data and metadata. To show how the file system behaves under high metadata load
-we use a create-heavy workload. Create-heavy workloads are studied the most in
-HPC research because of the checkpoint/restart use case but they also happen to
-stress the underlying storage system the most.  During this process we
+In our examination of the overheads of POSIX IO we benchmark and analyze
+CephFS, the file system that uses Ceph's object store ({\it i.e.} RADOS) to
+store its data and metadata. To show how the file system behaves under high
+metadata load we use a create-heavy workload.  During this process we
 discovered, based on the analysis and breakdown of costs, that durability and
-consistency have high overhead but we urge the read to keep in mind that this
+consistency have high overhead but we urge the reader to keep in mind that this
 file system is an implementation of one set of design decisions and our goal
@@
 
-% CITE ME Lustre paralle
 Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
@@
 are also serialized because the metadata server is single threaded; although
-naive, this design is common because of the complexity of multi-threaded
-metadata servers.  Traditional file system techniques for improving
-performance, such as caching inodes, do not help for create-heavy workloads.
-In this section, we quantify the costs of consistency and durability in CephFS.
-At the end of each subsection we compare the approach to ``decoupled
-namespaces", the technique in related work that detaches subtrees from the
-global namespace to relax consistency/durability guarantees.
+na\"{\i}ve, this design is common because of the complexity of multi-threaded
+metadata servers~\cite{konstantinos:pdsw2014-lustre-metadata}.  Traditional
+file system techniques for improving performance, such as caching inodes, do
+not help for create-heavy workloads.  In this section, we quantify the costs of
+consistency and durability in CephFS.  At the end of each subsection we compare
+the approach to ``decoupled namespaces", the technique in related work that
+detaches subtrees from the global namespace to relax consistency/durability
+guarantees.
 
@@
   \end{subfigure}
-  \caption{The overhead of global durability and strong consistency in CephFS.
+  \caption{The overhead of durability and strong consistency in CephFS.
   (a) shows the effect of different journal segment sizes, which are streamed
@@
 
-% CITEME: SSTables
-\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
-to the best of our knowledge, when a client or server fails there is no recovery
+\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, to the
+best of our knowledge, when a client or server fails there is no recovery
 scheme. For BatchFS, if a client fails when it is writing to the local
-log-structured merged tree (implemented as an SSTable) then those batched
-metadata operations are lost. For DeltaFS, if the client fails then, on restart,
-the computation does the work again -- since the snapshots of the namespace are
-never globally consistent and there is no ground truth.  On the server side,
-BatchFS and DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which
-initially reside in memory but are later flushed to the underlying
-distributed file system.
+log-structured merged tree (implemented as an
+SSTable~\cite{ren:atc2013-tablefs}) then those batched metadata operations are
+lost. For DeltaFS, if the client fails then, on restart, the computation does
+the work again -- since the snapshots of the namespace are never globally
+consistent and there is no ground truth.  On the server side, BatchFS and
+DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which initially
+reside in memory but are later flushed to the underlying distributed file
+system.
 

### 2017-10-23 paper/sections/6_results.tex
paper: jeff comments/citations and fix figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: fill in all CITEMEs, hammer on main points
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-% CITEME xiao:socc2015-shardfs CITEME HADOOP
 First we show how we can build application-specific subtrees by composing
-mechanisms and that this approach results in more a scalable global
-namespace. We start with clients creating files in private directories because
-this workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
+mechanisms and that this approach results in more a scalable global namespace.
+We start with clients creating files in private directories because this
+workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
 ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
 sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
-But the workload also appears in cloud workloads, where systems like Hadoop use
-the file abstraction to exchange work units to workers or to indicate when jobs
+But the workload also appears in cloud workloads~\cite{xiao:socc15-shardfs},
+where systems like Hadoop use the file abstraction to exchange work units to
+workers or to indicate when jobs
 complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
@@
 specific applications allows the global namespace to scale further and perform
-better.
+better. In our case, RPCs overwhelm the system but decoupling/shipping a
+journal of updates improves scalability by 2\(\times\).
 
@@
 
-% CITEME
 The final use case shows how the API gives users fine-grained control of the
 consistency semantics for niche or {\it ad-hoc} workflows.  Scientists use the
-file system to check the progress of jobs using \texttt{ls}~\cite{CITEME}. The
-number of files or size of the files is indicative of the progress. This
-practice is not too different from systems that use the file system to manage
-the progress of jobs; Hadoop writes to temporary files, renames them when
-complete, and creates a ``DONE" file to indicate to the runtime that the task
-did not fail and should not be re-scheduled on another node. In this scenario,
-Cudele users will not see the progress of decoupled namespaces since their
-updates are not globally visible.  To help scientists judge the progress of
-their jobs, Cudele clients have a ``namespace sync" that sends batches of
-updates back to the global namespace at regular intervals.
+file system to check the progress of jobs using \texttt{ls} even though this
+operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
+eshel:fast10-panache} . The number of files or size of the files is indicative
+of the progress. This practice is not too different from systems that use the
+file system to manage the progress of jobs; Hadoop writes to temporary files,
+renames them when complete, and creates a ``DONE" file to indicate to the
+runtime that the task did not fail and should not be re-scheduled on another
+node. In this scenario, Cudele users will not see the progress of decoupled
+namespaces since their updates are not globally visible.  To help scientists
+judge the progress of their jobs, Cudele clients have a ``namespace sync" that
+sends batches of updates back to the global namespace at regular intervals.
 

### 2017-10-23 paper/sections/1_introduction.tex
paper: fill in all CITEMEs, hammer on main points
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: tone down HPCness and simplify captions
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % What is the problem?
-File system metadata services in HPC have scalability problems because
-administrative tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the
-file system~\cite{zheng:pdsw2014-batchfs}, contend for the same directories and
+File system metadata services in HPC and data center applications  have
+scalability problems because administrative tasks, like
+checkpointing~\cite{bent_plfs_2009} or scanning the file
+system~\cite{zheng:pdsw2014-batchfs}, contend for the same directories and
 inodes. Applications perform better with dedicated metadata
@@
 metadata server for every client is unreasonable. This problem is exacerbated
-by current trends in HPC, where architectures are transitioning from complex
-storage stacks with burst buffer, file system, object store, and tape tiers to
-more simplified stacks with just a burst buffer and object
-store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
-because more requests end up hitting the same layer and latencies cannot be
-hidden while data migrates across tiers.
+by current trends; for example, HPC architectures are transitioning from
+complex storage stacks with burst buffer, file system, object store, and tape
+tiers to more simplified stacks with just a burst buffer and object
+store~\cite{bent:login16-hpc-trends}. These types of trends put more pressure
+on data access because more requests end up hitting the same layer and
+latencies cannot be hidden while data migrates across tiers.
 
@@
 BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
-do more client-side processing and merge updates when the job is done. HPC
-developers are turning to these non-POSIX IO solutions because their applications
-are well-understood ({\it e.g.}, well-defined read/write phases,
-synchronization only needed during certain phases, workflows describing
-computation, etc.) and because these applications wreak havoc on file systems designed for
-general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N and N:1 create
-patterns~\cite{bent_plfs_2009}).
+do more client-side processing and merge updates when the job is done.
+Developers in these domains are turning to these non-POSIX IO solutions because
+their applications are well-understood ({\it e.g.}, well-defined read/write
+phases, synchronization only needed during certain phases, workflows describing
+computation, etc.) and because these applications wreak havoc on file systems
+designed for general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N
+and N:1 create patterns~\cite{bent_plfs_2009}).
 
-\begin{figure}[tb]
-\centering
+\begin{figure}[tb] \centering
 \includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
-\caption{ Illustration of subtrees with different semantics co-existing in a
+\caption{Illustration of subtrees with different semantics co-existing in a
 global namespace.  For performance, clients can relax consistency on their
-subtree (HDFS) or even decouple the subtree and move it locally (BatchFS,
-RAMDisk). Decoupled subtrees can further relax durability for even better
-performance.  Clients that require stronger guarantees (POSIX IO) can still reside
-in the same namespace.  }\label{fig:subtree-policies}
-\end{figure}
+subtree (HDFS) or decouple the subtree and move it locally (BatchFS, RAMDisk).
+Decoupled subtrees can relax durability for even better performance.
+}\label{fig:subtree-policies} \end{figure}
 

### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: tone down HPCness and simplify captions
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % - sequential IO, trim redundant operations
-\textbf{CephFS Design}: a journal of metadata updates that streams into the
+\textbf{CephFS Design}: A journal of metadata updates that streams into the
 resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
@@
 
-\textbf{CephFS Design}: capabilities keep metadata strongly
+\textbf{CephFS Design}: Capabilities keep metadata strongly
 consistent. To reduce the number of RPCs needed for consistency, clients can

### 2017-10-23 paper/sections/4_methodology.tex
paper: ken edits
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: tone down HPCness and simplify captions
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=90mm]{figures/fig-decouple.png}
-\caption{Applications decouple the namespace, write updates to a local journal,
-and delay updates using the \textbf{mechanisms} shown here; they are described
-in Table~\ref{table:mechanisms}.}\label{fig:decouple}
+\caption{Illustration of the \textbf{mechanisms} used by applications to build
+consistency/durability semantics. Descriptions are provided in
+Table~\ref{table:mechanisms} and the underlined words in
+Section~\S\ref{sec:the-cudelesfs-mechanisms}.}\label{fig:decouple}
 \end{figure}
@@
 \end{tabular}
-\caption{Mechanisms are composed together to form consistency and
-durability semantics.\label{table:mechanisms}} 
+\caption{Descriptions of the mechanisms. Example compositions are shown in Table~\ref{table:spectrum}.\label{table:mechanisms}} 
 \end{table}
@@
 client journal+RPCs} since both mechanisms do the same thing or
-\texttt{stream+save} since ``global" durability is stronger and has more
+\texttt{stream+local persist} since ``global" durability is stronger and has more
 overhead than ``local" durability. The cost of each mechanism and the semantics

### 2017-10-23 paper/sections/6_results.tex
paper: tone down HPCness and simplify captions
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
-\caption{Performance of each mechanism (left) and building
-consistency/durability guarantees (right) for 100K files creates from a single
-client. Results are normalized to the runtime of writing events to the client's
-in-memory journal.  \label{fig:composable-mechanisms}}
+\caption{Performance of each mechanism (left) and building the
+consistency/durability semantics of real-world systems (right) for 100K files
+creates from a single client. Results are normalized to the runtime of writing
+events to the client's in-memory journal.  \label{fig:composable-mechanisms}}
 \end{figure}
@@
 \caption{The performance and features of Cudele. (a) shows the cost of
-merging client journals at the metadata server. Merging is better than RPCs
+merging client journals at the metadata server. Shipping and merging journals of updates scales better than RPCs
 because there are less messages and consistency/durability code paths are

### 2017-10-23 paper/sections/0_abstract.tex
paper: final read-through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-HPC and data center scale developers are abandoning POSIX IO because the file
-system metadata synchronization and serialization overheads of providing strong
-consistency and durability are too costly -- and often unnecessary -- for their
-applications.  Unfortunately, designing file systems with weaker consistency or
-durability excludes applications that rely on stronger guarantees, forcing
-developers to re-write their applications or deploy them on a different system.
-Mounting multiple systems in the global namespace forces users to (1) provision
-separate storage clusters and (2) manually move data across systems.  We
-present a framework and API that lets clients specify their
-consistency/durability requirements and dynamically assign them to subtrees in
-the same namespace, allowing users to optimize subtrees over time and space for
-different workloads.  We confirm the performance benefits of techniques
-presented in related work but also explore new consistency/durability metadata
-designs, all integrated the same storage system.  By custom fitting a subtree
-to a create-heavy application, we show 8\(\times\) speedup and 2\(\times\)
-scale when compared to our baseline system.
+HPC and data center scale application developers are abandoning POSIX IO
+because the file system metadata synchronization and serialization overheads of
+providing strong consistency and durability are too costly -- and often
+unnecessary -- for their applications.  Unfortunately, designing file systems
+with weaker consistency or durability excludes applications that rely on
+stronger guarantees, forcing developers to re-write their applications or
+deploy them on a different system.  Users can mount multiple systems in the
+global namespace but this means (1) provisioning separate storage clusters and
+(2) manually moving data across system boundaries.  We present a framework and
+API that lets clients specify their consistency/durability requirements and
+dynamically assign them to subtrees in the same namespace, allowing users to
+optimize subtrees over time and space for different workloads.  We confirm the
+performance benefits of techniques presented in related work but also explore
+new consistency/durability metadata designs, all integrated over the same
+storage system.  By custom fitting a subtree to a create-heavy application, we
+show 8\(\times\) speedup and 2\(\times\) scale when compared to our baseline
+system.
 

### 2017-10-23 paper/sections/1_introduction.tex
paper: final read-through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 File system metadata services in HPC and data center applications  have
-scalability problems because administrative tasks, like
+scalability problems because common tasks, like
 checkpointing~\cite{bent_plfs_2009} or scanning the file
@@
 metadata server for every client is unreasonable. This problem is exacerbated
-by current trends; for example, HPC architectures are transitioning from
-complex storage stacks with burst buffer, file system, object store, and tape
-tiers to more simplified stacks with just a burst buffer and object
-store~\cite{bent:login16-hpc-trends}. These types of trends put more pressure
-on data access because more requests end up hitting the same layer and
+by current hardware and software trends; for example, HPC architectures are
+transitioning from complex storage stacks with burst buffer, file system,
+object store, and tape tiers to more simplified stacks with just a burst buffer
+and object store~\cite{bent:login16-hpc-trends}. These types of trends put
+pressure on data access because more requests end up hitting the same layer and
 latencies cannot be hidden while data migrates across tiers.
@@
 Our prototype system, Cudele, achieves this by exposing ``mechanisms" that
-developers use to specify their preferred semantics.  Cudele supports 3 forms
+users combine to specify their preferred semantics.  Cudele supports 3 forms
 of consistency (invisible, weak, and strong) and 3 degrees of durability (none,
@@
 The last contribution lays the groundwork for future work on our prototype. It
-is better than the current practice of mounting different storage systems in a global
+is more scalable than the current practice of mounting different storage systems in a global
 namespace because there is no need to provision dedicated storage clusters to
 applications or move data between these systems.  For example, the results of a
-Hadoop job do not need to be migrated into CephFS for other processing; instead
+Hadoop job do not need to be migrated into a Ceph file system (CephFS) for other processing; instead
 the user can change the semantics of the HDFS subtree into a CephFS subtree.
 This may cause metadata/data movement to make things strongly consistent again
-but this is superior to moving all data across file system boundaries. We
-enable studies that adjust these semantics over {\it time and space}, where
+but this is superior to moving all data across file system boundaries. Our prototype
+enables studies that adjust these semantics over {\it time and space}, where
 subtrees can change their semantics and migrate around the cluster without ever
@@
 % Results
-The results in this paper lay the groundwork for such a system and confirm the
+The results in this paper pave the way for such a system and confirm the
 assertions of ``clean-state" research that decouple namespaces; specifically
-that the technique drastically improves performance (104\(\times\) speed up).
+that these techniques drastically improve performance (we see 104\(\times\) speed up).
 We go a step further by quantifying the costs of merging updates (7\(\times\)

### 2017-10-23 paper/sections/3_posix_overheads.tex
paper: final read-through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
-the Linux kernel.  The \texttt{untar} phase, which is characterized by many
-creates, has the highest resource usage (combined CPU, network, and disk)
-because of the number of RPCs needed for consistency and durability. The RPCs
-are also serialized because the metadata server is single threaded; although
-na\"{\i}ve, this design is common because of the complexity of multi-threaded
-metadata servers~\cite{konstantinos:pdsw2014-lustre-metadata}.  Traditional
-file system techniques for improving performance, such as caching inodes, do
-not help for create-heavy workloads.  In this section, we quantify the costs of
-consistency and durability in CephFS.  At the end of each subsection we compare
-the approach to ``decoupled namespaces", the technique in related work that
-detaches subtrees from the global namespace to relax consistency/durability
-guarantees.
+the Linux kernel in a CephFS mount.  The \texttt{untar} phase, which is
+characterized by many creates, has the highest resource usage (combined CPU,
+network, and disk) because of the number of RPCs needed for consistency and
+durability. The RPCs are also serialized because the metadata server is single
+threaded; although na\"{\i}ve, this design is common because of the complexity
+of multi-threaded metadata
+servers~\cite{konstantinos:pdsw2014-lustre-metadata}.  In this section, we
+quantify the costs of consistency and durability in CephFS.  At the end of each
+subsection we compare the approach to ``decoupled namespaces", the technique in
+related work that detaches subtrees from the global namespace to relax
+consistency/durability guarantees.
 
@@
 clients or servers fail.  None is different than local durability because
-regardless of the type failure, metadata will be lost when components die in a
+regardless of the type of failure, metadata will be lost when components die in a
 None configuration.
@@
 % Effects on performance
-Figure~\ref{fig:overhead-a} shows the
-effect of journaling of different journal segment sizes; the larger the segment
-size the bigger that the writes into the object store are. The trade-off for
-better performance is memory consumption because larger segments take up
-more space for buffering. When journaling is on, 
-the metadata server periodically stops serving requests to flush ({\it i.e.}
-apply journal updates) to the metadata store.  The journal overhead is
+Figure~\ref{fig:overhead-a} shows the effect of journaling of with different
+segment sizes; the larger the segment size the bigger that the writes into the
+object store are. The trade-off for better performance is memory consumption
+because larger segments take up more space for buffering. When journaling is
+on, the metadata server periodically stops serving requests to flush ({\it
+i.e.} apply journal updates) to the metadata store.  The journal overhead is
 sufficient enough to slow down metadata throughput but not so much as to
-overwhelm the bandwidth of the object store. We measured our peak bandwidth to
-be 100MB/s, which is the speed of our network link.
+overwhelm the bandwidth of the object store.
 
-\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, to the
-best of our knowledge, when a client or server fails there is no recovery
-scheme. For BatchFS, if a client fails when it is writing to the local
-log-structured merged tree (implemented as an
-SSTable~\cite{ren:atc2013-tablefs}) then those batched metadata operations are
+\textbf{Comparison to decoupled namespaces}: For BatchFS, if a client fails
+when it is writing to the local log-structured merge tree (implemented as an
+SSTable~\cite{ren:atc2013-tablefs}) then unwritten metadata operations are
 lost. For DeltaFS, if the client fails then, on restart, the computation does
@@
 consistent and there is no ground truth.  On the server side, BatchFS and
-DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which initially
-reside in memory but are later flushed to the underlying distributed file
-system.
+DeltaFS use IndexFS~\cite{ren:sc2014-indexfs}. IndexFS writes metadata to
+SSTables, which initially reside in memory but are later flushed to the
+underlying distributed file system.
 
@@
 consistent. To reduce the number of RPCs needed for consistency, clients can
-obtain capabilities for reading, creating and updating inodes, as well as caching reads,
+obtain capabilities for reading and writing inodes, as well as caching reads,
 buffering writes, changing the file size, and performing lazy IO.
@@
 new files locally before issuing a create request to the metadata server. If
-another client starts creating files in the same directory (``interfere" curve
-on \(y1\) axis) then the directory inode transitions out of read caching and
-the first client must send \texttt{lookup()}s to the metadata server
-(``interfere" curve on \(y2\) axis). These extra requests increase the
-throughput of the ``interfere" curve because the metadata server can handle the
-extra load but performance suffers.  This degradation is shown in
-Figure~\ref{fig:overhead-c}, where we scale the number of clients and show the
-slowdown of the slowest client.  The results are normalized to a single
-isolated client without a metadata server journal.  For the ``interfere" bars,
-each client creates files in private directories and at 30 seconds we launch
-another process that creates files in those directories. 18 clients is the
-maximum number the metadata server could handle for this metadata-intensive
-workload; at higher client load, the metadata server started complaining about
-laggy and unresponsive requests.
+another client starts creating files in the same directory then the directory
+inode transitions out of read caching and the first client must send
+\texttt{lookup()}s to the metadata server (``interfere" curve on \(y2\) axis).
+These extra requests increase the throughput of the ``interfere" curve on the
+\(y1\) axis because the metadata server can handle the extra load but
+performance suffers.  This degradation is shown in Figure~\ref{fig:overhead-c},
+where we scale the number of clients and show the slowdown of the slowest
+client.  The results are normalized to a single isolated client without a
+metadata server journal and the error bars are the standard deviations of all
+client runtimes.  For the ``interfere" bars, each client creates files in
+private directories and at 30 seconds we launch another process that creates
+files in those directories. 18 is the maximum number of clients the metadata
+server can handle for this metadata-intensive workload; at higher client load,
+the metadata server complains about laggy and unresponsive requests.
 
@@
 completes.  In BatchFS the merge is delayed by the application using an API to
-switch between asynchronous to synchronous mode. The merge itself is explicitly
+switch between asynchronous and synchronous mode. The merge itself is explicitly
 managed by the application but future work looks at more automated

### 2017-10-23 paper/sections/4_methodology.tex
paper: final read-through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-In this section we describe Cudele, our API and framework that lets users
+In this section we describe our API and framework that lets users
 assign consistency and durability semantics to subtrees in the global
@@
 \underline{Append Client Journal} mechanism to append metadata updates to a
-local, in-memory journal. This is fast because clients do not need to check for
+local, in-memory journal. Clients do not need to check for
 consistency when writing events and the metadata server blindly applies the
 updates because it assumes the events were already checked for consistency. The
-trade-off here is performance; it is a dangerous approach but could be
+trade-off here is fast performance; it is a dangerous approach but could be
 implemented safely if the clients or metadata server are configured to check
 the validity of events before writing them.  Once the job is complete, the
-system calls Cudele mechanisms to achieve the desired consistency/durability
+system calls mechanisms to achieve the desired consistency/durability
 semantics.  Cudele provides a library for clients to link into and all
@@
 
-\subsubsection{Mechanisms Used for Consistency} \underline{RPCs} send remote procedure
-calls for every metadata operation from the client to the metadata server,
-assuming the request cannot be satisfied by the inode cache. This mechanism is
-part of the default CephFS implementation and is the strongest form of
-consistency because clients see metadata updates right away.  \underline{Nonvolatile
-Apply} replays the client's journal onto the metadata cluster's metadata store.
-The client's in-memory journal is written into the object store and the
-metadata servers are restarted. When the metadata servers re-initialize, they
-notice new journal updates in the object store and replay the events onto their
-in-memory metadata stores.  \underline{Volatile Apply} takes the client's in-memory journal on the
-client and applies the updates directly to the in-memory metadata store maintained
-by the metadata servers. We say volatile because -- in exchange for peak
-performance -- Cudele makes no consistency or durability guarantees while
-Volatile Apply is executing.  If a concurrent update from a client occurs
-there is no rule for resolving conflicts and if the client or metadata server
-crashes there may be no way to recover.
+\subsubsection{Mechanisms Used for Consistency} \underline{RPCs} send remote
+procedure calls for every metadata operation from the client to the metadata
+server, assuming the request cannot be satisfied by the inode cache. This
+mechanism is part of the default CephFS implementation and is the strongest
+form of consistency because clients see metadata updates right away.
+\underline{Nonvolatile Apply} replays the client's in-memory journal into the
+object store and restarts the metadata servers. When the metadata servers
+re-initialize, they notice new journal updates in the object store and replay
+the events onto their in-memory metadata stores.  \underline{Volatile Apply}
+takes the client's in-memory journal on the client and applies the updates
+directly to the in-memory metadata store maintained by the metadata servers. We
+say volatile because -- in exchange for peak performance -- Cudele makes no
+consistency or durability guarantees while Volatile Apply is executing.  If a
+concurrent update from a client occurs there is no rule for resolving conflicts
+and if the client or metadata server crashes there may be no way to recover.
 
@@
 serialized log events to a file on local disk and for \underline{Global Persist},
-clients push the journal into the objects store. The overheads for both Local
+clients push the journal into the object store. The overheads for both Local
 Persist and Global Persist is the write bandwidth of the local disk and
@@
   \item \texttt{interfere\_policy}: how to handle a request from another
-  client targeted at this subtree (\texttt{allow}))
+  client targeted at this subtree (\texttt{allow})
 
@@
 
-The ``Consistency" and ``Durability" parameters are set to the Cudele
-mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
-``Allocated Inodes" is a way for the application to specify how many files it
-intends to create. It is a contract so that the file system can provision
-enough resources for the incumbent merge and so it can give valid inodes to
-other clients. The inodes can be used anywhere within the decoupled namespace
-({i.e.} at any depth in the subtree).
-
-% description
-``Interfere Policy" has two settings: \texttt{block} and \texttt{allow}.
-For \texttt{block}, any requests to this part of the namespace returns with
-``Device is busy", which will spare the metadata server from wasting resources
-for updates that may get overwritten. If the application does not mind losing
-updates, for example it wants approximations for results that take too long to
-compute, it can select \texttt{allow}. In this case, metadata from the interfering client will be
-written and the computation from the decoupled namespace will take priority at
-merge time because the results are more accurate.
-
-% examples
-Given these default values decoupling the namespace with an empty policies file
-would give the application 100 inodes but the subtree would behave like the
-existing CephFS implementation.  BatchFS merges updates into the global
-namespace after the job is complete.  This can be achieved with by composing
-two mechanisms:
+The ``Consistency" and ``Durability" parameters are compositions of mechanisms;
+they can be serialized (\(+\)) or run in parallel (\(|\|\)).  ``Allocated
+Inodes" is a way for the application to specify how many files it intends to
+create. It is a contract so that the file system can provision enough resources
+for the incumbent merge and so it can give valid inodes to other clients. The
+inodes can be used anywhere within the decoupled namespace ({i.e.} at any depth
+in the subtree).  ``Interfere Policy" has two settings: \texttt{block} and
+\texttt{allow}.  For \texttt{block}, any requests to this part of the namespace
+returns with ``Device is busy", which will spare the metadata server from
+wasting resources for updates that may get overwritten. If the application does
+not mind losing updates, for example it wants approximations for results that
+take too long to compute, it can select \texttt{allow}. In this case, metadata
+from the interfering client will be written and the computation from the
+decoupled namespace will take priority at merge time because the results are
+more accurate.  Given these default values decoupling the namespace with an
+empty policies file would give the application 100 inodes but the subtree would
+behave like the existing CephFS implementation.

### 2017-10-23 paper/sections/5_implementation.tex
paper: final read-through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % how i used it
-\textbf{Cudele}: the ``RPCs" mechanism uses the metadata store to
-service requests. Using code designed for recovery, ``Nonvolatile Apply" and
-``Volatile Apply" replay updates onto the metadata store in memory and in the
+\textbf{Cudele}: the RPCs mechanism uses the metadata store to
+service requests. Using code designed for recovery, Nonvolatile Apply and
+Volatile Apply replay updates onto the metadata store in memory and in the
 object store, respectively.  When the clients are ready to merge their updates
@@
 
-\textbf{Cudele}: the journal format is used by ``Stream", ``Append Client Journal",
-``Local Persist", and ``Global Persist".  ``Stream" is the default
+\textbf{Cudele}: the journal format is used by Stream, Append Client Journal,
+Local Persist, and Global Persist.  Stream is the default
 implementation for achieving global durability in CephFS but the rest of the
@@
 modifying the updates in the journal so we re-purposed that code to implement
-``Append Client Journal", ``Volatile Apply", and ``Nonvolatile Apply".
+Append Client Journal, Volatile Apply, and Nonvolatile Apply.
 
@@
 
-\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
+\textbf{Cudele}: Nonvolatile Apply uses the internal inode cache
 code to allocate inodes to clients that decouple parts of the namespace and to

### 2017-10-23 paper/sections/6_results.tex
paper: final read-through
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 to the time it takes to write 100K file create updates to the client's
-in-memory journal ({\it i.e.} the ``Append Client Journal" mechanism).
-``Stream" is an approximation of the overhead and is calculated by subtracting
+in-memory journal ({\it i.e.} the Append Client Journal mechanism).
+Stream is an approximation of the overhead and is calculated by subtracting
 the runtime of the job with the journal turned off from the runtime with the
@@
 mechanisms that rely on poorly scaling data structures have large slowdowns for
-the larger number of creates. For example, ``RPCs" which relies on the internal
-CephFS directory structures has a \(90\times\) slowdown for 100K files. It is a
+the larger number of creates. For example, RPCs has a \(90\times\) slowdown
+because it relies on the internal CephFS directory data structures. It is a
 well-known problem that directory data structures do not scale when creating
@@
 % RPCs vs. apply: calls to metadata server vs. RADOS
-{\it Overhead of RPCs:} ``RPCs" is \(66\times\) slower than ``Volatile
-Apply" because sending individual metadata updates over the network is costly.
-While ``RPCs" sends a request for every file create, ``Volatile Apply"
+{\it Overhead of RPCs:} RPCs is \(66\times\) slower than Volatile
+Apply because sending individual metadata updates over the network is costly.
+While RPCs sends a request for every file create, Volatile Apply
 writes all the updates to the in-memory journal and applies them to the
@@
 decoupled namespace directly to the metadata server is faster, communicating
-through the object store (``Nonvolatile Apply") is \(10\times\) slower.
+through the object store (Nonvolatile Apply) is \(10\times\) slower.
 
@@
 % v_apply vs. apply/persist: communicating through RADOS
-{\it Overhead of ``Nonvolatile Apply":} The cost of ``Nonvolatile
-apply" is much larger than all the other mechanisms.  That mechanism was not
+{\it Overhead of Nonvolatile Apply:} The cost of Nonvolatile
+Apply is much larger than all the other mechanisms.  That mechanism was not
 implemented as part of Cudele -- it was a debugging and recovery tool packaged
@@
 i.e.} \texttt{/}).  The cost of communicating through the object store is shown
-by comparing the runtime of ``Volatile apply" + ``Global persist" to
-``Nonvolatile Apply". These two operations end up with the same final metadata
-state but using ``Nonvolatile Apply" is clearly inferior.
+by comparing the runtime of Volatile Apply + Global Persist to
+Nonvolatile Apply. These two operations end up with the same final metadata
+state but using Nonvolatile Apply is clearly inferior.
 
 % persist vs. save: one disk vs. many
-{\it Parallelism of the Object Store:} Comparing ``Local" and ``Global
-persist" demonstrates the bandwidth advantages of storing the journal in a
-distributed object store. The ``Global Persist"
+{\it Parallelism of the Object Store:} Comparing Local and Global
+Persist demonstrates the bandwidth advantages of storing the journal in a
+distributed object store. The Global Persist
 performance is \(1.5\times\) faster because the object store is leveraging the
@@
 object store itself but should be acknowledged when making decisions for the
-application; the size of the object store can help mitigate the overheads of
+application; the bandwidth of the object store can help mitigate the overheads of
 globally persisting metadata updates.
@@
 services a flash crowd of creates across all directories as shown in
-Figure~\ref{fig:overhead-creates}.  We use this as a microbenchmark because it
-allows us to control the size of the log, since each create creates a single
-journal event. The workload is clients creating 100K files in private
-directories in the same global namespace.
+Figure~\ref{fig:overhead-creates}.  The workload is clients creating 100K files
+in private directories in the same global namespace.
 
@@
 To show the contention at the metadata server, in Figure~\ref{fig:mergescale}
-we scale the umber of clients for the RPC per operation strategy used by many
-POSIX IO systems and the decoupled namespace strategy used by DeltaFS. For
-RPCs, we scale the number of clients making requests to the metadata server and
-for the decoupled bars, we scale the number of concurrent journal merge
-requests. So for the 30 clients data point, we are measuring the operations per
-second for 30 client journals that land on the metadata server at the same
-time. The ``create \& merge" bar adds the time it takes for clients to generate
-the journal of events in parallel.
+we scale the number of clients for the RPC per operation strategy and the
+decoupled namespace strategy. For RPCs, we increase the number of clients
+making requests to the metadata server and for decoupled (the merge bars), we
+increase the number of concurrent journal merge requests. So for the 30 clients
+data point, we are measuring the operations per second for 30 client journals
+that land on the metadata server at the same time. The ``create \& merge" bar
+adds the time it takes for clients to generate the journal of events in
+parallel. The experiment is run three times but the standard deviation is too
+small to see.
 
@@
 updates are merged by the metadata server into the in-memory metadata store
-they never scan growing data structures.  Had we implemented the ``Create"
-mechanism to include \texttt{lookup()} commands before \texttt{open()}
-requests, we would have seen the poor scaling that we see with the ``RPCs"
+they never scan growing data structures.  Had we implemented the Append Client
+Journal mechanism to include \texttt{lookup()} commands before \texttt{open()}
+requests, we would have seen the poor scaling that we see with the RPCs
 mechanism.
@@
 workload, clients create files in private directories and a separate client
-creates files in each directory. This introduces false sharing and the metadata
-server revokes capabilities on directories touched by the interfering client. While
-HPC tries to avoid these situations with
+interferes by creating more files in each directory. This introduces false
+sharing and the metadata server revokes capabilities on directories touched by
+the interfering client. While HPC tries to avoid these situations with
 workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
 happens in distributed file systems when users unintentionally access
-directories in a shared file system.  We also have another client write to a
-decoupled namespace and merge its updates at 90 seconds.  We only scale to 15
-clients because the performance variability of a nearly overloaded metadata
-server, in our case 18 clients, is too high.
+directories in a shared file system.  We only scale to 15 clients because the
+performance variability of a nearly overloaded metadata server, in our case 18
+clients, is too high.
 
@@
 to return \texttt{-EBUSY} to interfering clients; and ``Isolated" is the
-baseline performance without an interfering. Each subtree has ``RPCs" and
-``Stream" enabled.  Results results are normalized to the runtime of a single
+baseline performance without an interfering. Each subtree has RPCs and
+Stream enabled.  Results are normalized to the runtime of a single
 client creating files in a directory. Note that IndexFS has a similar behavior
@@
 The final use case shows how the API gives users fine-grained control of the
-consistency semantics for niche or {\it ad-hoc} workflows.  Scientists use the
-file system to check the progress of jobs using \texttt{ls} even though this
+consistency semantics.  The use case is that users often leverage the file
+system to check the progress of jobs using \texttt{ls} even though this
 operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
@@
 node. In this scenario, Cudele users will not see the progress of decoupled
-namespaces since their updates are not globally visible.  To help scientists
+namespaces since their updates are not globally visible.  To help users
 judge the progress of their jobs, Cudele clients have a ``namespace sync" that
@@
 client writing updates to a decoupled namespace and pausing to send updates to
-the metadata server. We scale the namespace sync interval to show the trade-off
-of frequently pausing or writing large logs of updates.  We use an idle core to
-log the updates and to do the network transfer. The client only pauses to fork
-off a background process, which is expensive as the address space needs to
-copied. The alternative is to pause the client completely and write the update
-to disk but since this implementation is limited by the speed of the disk, we
-choose the memory-to-memory copy of the fork approach.
+the metadata server. The error bars are the standard deviations of 3 runs. We
+scale the namespace sync interval to show the trade-off of frequently pausing
+or writing large logs of updates.  We use an idle core to log the updates and
+to do the network transfer. The client only pauses to fork off a background
+process, which is expensive as the address space needs to copied. The
+alternative is to pause the client completely and write the update to disk but
+since this implementation is limited by the speed of the disk, we choose the
+memory-to-memory copy of the fork approach.
 

### 2017-10-23 paper/sections/0_abstract.tex
paper: final version
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 storage system.  By custom fitting a subtree to a create-heavy application, we
-show 8\(\times\) speedup and 2\(\times\) scale when compared to our baseline
-system.
+show 8\(\times\) speedup and can scale to 2\(\times\) as many clients when
+compared to our baseline system.
 

### 2017-10-23 paper/sections/1_introduction.tex
paper: fix some typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % What is the problem?
-File system metadata services in HPC and data center applications  have
+File system metadata services in HPC and large-scale data centers have
 scalability problems because common tasks, like
@@
 in~\S\ref{sec:use-case-1}.} subtree has weaker than strong consistency because
-for performance it lets clients read files opened for
+it lets clients read files opened for
 writing~\cite{hakimzadeh:dais14-hdfs-consistency}, which means that not all
-updates are immediately seen by all servers; the BatchFS and RAMDisk subtrees
+updates are immediately seen by all clients; the BatchFS and RAMDisk subtrees
 are decoupled from the global namespace and have similar consistency/durability

### 2017-10-25 experiments/baseline-durability/results-done/journal-cache/plot.sh
benchmark: add experiments (with pointers) to baseline repo
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,35 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
+   "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
+  ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+for osd in 0 1 14 24 27 29 34 40; do
+  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+    ../../../../../visualize.sh ${m}
+    sed -i "s/:/,/g" `basename ${m}`.out
+    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+  done
+done
+

### 2017-10-25 experiments/baseline-durability/results-done/journal-nocache/plot.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
+  ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+for osd in 0 1 14 24 27 29 34 40; do
+  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+    ../../../../../visualize.sh ${m}
+    sed -i "s/:/,/g" `basename ${m}`.out
+    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+  done
+done
+

### 2017-10-25 experiments/baseline-durability/results-done/nojournal-nocache/plot.sh

@@ -1,0 +1,37 @@
+#!/bin/bash
+
+set -ex
+#for i in `ls *.tar.gz`; do
+#  tar xzf $i
+#done
+#
+## prepare as a csv
+#for m in \
+#   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
+#   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+#   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
+#   "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
+for m in \
+  "tmp/graphite/whisper/issdm-12/meminfo/used.wsp" \
+  ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+for osd in 0 1 14 24 27 29 34 40; do
+  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+    ../../../../../visualize.sh ${m}
+    sed -i "s/:/,/g" `basename ${m}`.out
+    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+  done
+done
+

### 2017-10-25 experiments/baseline-durability/run.sh

@@ -1,0 +1,38 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+#for site in "nojournal-cache" "nojournal-nocache" "journal-cache" "journal-nocache"; do
+for site in "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
+  nclients=3
+
+  # configure ceph and setup results directory
+  mkdir -p results/$site/logs || true
+  cp site/* site/roles/ceph-ansible || true
+  cp site_confs/${site}.yml site/group_vars/all
+  cp -r site/group_vars site/roles/ceph-ansible/
+  cp inventory/${nclients}client site/hosts
+  cp inventory/${nclients}client site/roles/ceph-ansible/hosts
+
+  # cleanup and start ceph
+  $SRL_ANSIBLE cleanup.yml
+  $CEPH_ANSIBLE ceph.yml cephfs.yml
+  $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+  
+  # warmup and get baseline
+  for i in `seq 0 2`; do
+    ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
+  done
+  
+  ## baseline the drop delays
+  #./ansible-playbook.sh -e site=$site -e nfiles=50000 -e drop_delay=15 ../workloads/stat.yml
+  #./ansible-playbook.sh -e site=$site -e nfiles=50000 -e drop_delay=15 ../workloads/touch.yml
+
+  ./ansible-playbook.sh -e site=$site collect.yml
+done

### 2017-10-25 experiments/baseline-durability/visualize/jupyter.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+set -e -x
+mkdir graphs || true
+docker run --rm -v  `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93
+

### 2017-10-25 experiments/baseline-interfere/ceph-mds-admin-daemon.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+set -x
+
+./ansible.sh mdss -m shell -a "docker exec ceph-\`hostname\`-mds ceph --admin-daemon /var/run/ceph/ceph-mds.\`hostname\`.asok $@"
+

### 2017-10-25 experiments/baseline-interfere/results-paper/nojournal-cache/plot.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+GRAPHITE="docker run -v `pwd`/tmp:/tmp --entrypoint=/bin/bash -it michaelsevilla/graphite -c"
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-18/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_server/req_lookup.wsp" \
+   ; do
+  $GRAPHITE "whisper-dump.py $m > tmp/`basename $m.out`"
+  $GRAPHITE "sed -i 's/:/,/g' tmp/`basename ${m}.out`"
+done

### 2017-10-25 experiments/baseline-interfere/results-paper/nojournal-nocache/plot.sh

@@ -1,0 +1,35 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-18/mds_mem/ino.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_mem/ino+.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_mem/ino-.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds/inodes.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds/inodes_bottom.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds/inodes_with_caps.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds/inodes_pin_tail.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds/inodes_pinned.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds/inodes_top.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-18/cputotals/user.wsp" \
+   "tmp/graphite/whisper/issdm-18/cputotals/sys.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_server/req_create.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_server/req_lookup.wsp" \
+   ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+for osd in 0 1 14 24 27 29 34 40; do
+  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+    ../../../../../visualize.sh ${m}
+    sed -i "s/:/,/g" `basename ${m}`.out
+    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+  done
+done

### 2017-10-25 experiments/baseline-interfere/run.sh

@@ -1,0 +1,37 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+for site in "nojournal-cache"; do
+  nclients=3
+
+  # configure ceph and setup results directory
+  mkdir -p results/$site/logs || true
+  cp site/* site/roles/ceph-ansible || true
+  cp site_confs/${site}.yml site/group_vars/all
+  cp -r site/group_vars site/roles/ceph-ansible/
+  cp inventory/${nclients}client site/hosts
+  cp inventory/${nclients}client site/roles/ceph-ansible/hosts
+
+  # cleanup and start ceph
+  $SRL_ANSIBLE cleanup.yml
+#  $CEPH_ANSIBLE ceph.yml cephfs.yml
+#  $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+#  
+#  # warmup and get baseline
+#  #for i in `seq 0 2`; do
+#  #  ./ansible-playbook.sh -e site=$site -e nfiles=98000 ../workloads/creates.yml
+#  #done
+#  
+#  # baseline the drop delays
+#  ./ansible-playbook.sh -e site=$site -e nfiles=98000 -e drop_delay=30 ../workloads/stat.yml
+#  ./ansible-playbook.sh -e site=$site -e nfiles=98000 -e drop_delay=30 ../workloads/touch.yml
+#
+#  ./ansible-playbook.sh -e site=$site collect.yml
+done

### 2017-10-25 experiments/baseline-interfere/visualize/jupyter.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+if [ $# -lt 1 ]; then
+  echo "USAGE <re-parse>"
+  echo "- where re-parse is 0 or 1"
+  exit 1
+fi
+
+set -e -x
+function uncompress() {
+  for i in `ls *.tar.gz`; do
+    tar xzf $i
+  done
+  
+  GRAPHITE="docker run -v `pwd`/tmp:/tmp --entrypoint=/bin/bash -it michaelsevilla/graphite -c"
+  
+  # prepare as a csv
+  for m in \
+     "tmp/graphite/whisper/issdm-18/mds_server/handle_client_request_tput.wsp" \
+     "tmp/graphite/whisper/issdm-18/mds_server/req_lookup.wsp" \
+     ; do
+    $GRAPHITE "whisper-dump.py $m > tmp/`basename $m.out`"
+    $GRAPHITE "sed -i 's/:/,/g' tmp/`basename ${m}.out`"
+  done
+}
+
+if [ "$1" -eq 1 ]; then
+  cd ../results-paper/nojournal-cache/
+  uncompress
+  cd -
+fi
+mkdir graphs || true
+docker run --rm -v  `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93
+

### 2017-12-10 paper/cover_letter.tex
paper: add cover letter skeleton (#25)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% Template: http://www.acm.org/publications/proceedings-template
\documentclass[onecolumn,conference]{IEEEtran}
\usepackage{booktabs} % For formal tables
\usepackage{minted}
\usepackage{graphicx}
\usepackage{arydshln}
\usepackage{subcaption}
\usepackage{times}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\begin{document}
\title{Cover Letter}
\maketitle

We thank the reviewers for their hard work and genuinely helpful suggestions.
In addition to this cover letter, we have posted a PDF online explicitly
showing additions in \textcolor{red}{\textbf{red}} and deletions in
\textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
feedback:

%Experimental Setup (R1, R2, R3)
%- code R1
%- servers, network, storage R1, R3
\section{No Experimental Setup (Reviewers 1, 2, and 3)}

We apologize for the omission of experimental setup and source code details.
For the experimental setup, we have added details about the cluster we ran on
in Table III as well as a description in the text (addresses Reviewers 1 and
3). For the source code, we have added code snippets to the Evaluation to
concretize what was actually done (addresses Reviewer 1).

Additionally, we have added paper artifacts online.  For the experimental
setup, we have made the infrastructure code available and added links after
each figure to show exactly how experiments are run.  This infrastructure code
contains scripts to deploy the system, run experiments, and gather results.
This process follows the Popper
Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
aims to make research reproducible.  We also make the source code available and
linked in the paper.

%Mixing contributions and future work (R1)
%Statements in Introduction do not align with Evaluation (R1, R2, R3)
%- quantify performance speedups R3
%Evaluation Structure (R2)
%- remove major takeaways and cross refercnes
%Will not address:
%- cost of dynamically changing consistency/durability not presented R1
\section{Structure and Layout of Evaluation (All Reviewers)}

One pervasive issue across reviews is that the results and contributions we
cite in the introduction are not validated, explained, or even mentioned in the
Evaluation.  To fix this, we separate out future work to make the contributions
of this paper more explicit (addresses Reviewer 1) and we connect all numbers
we cite in the introduction to the evaluation (addresses all reviewers).  To
further clarify the Evaluation, we:

\begin{itemize}

  \item added Table IV which includes raw baseline, Cudele, and hardware limit
numbers. This adds raw numbers and helps us make comparisons to hardware
speeds. Ultimately, we hope this adds insight into the results (addresses
Reviewers 2 nad 3).

  \item re-organized the section to make experiments more independent
and removed cross-references (addresses Reviewer 2). Experiments are now
self-contained so the reader can see the effects of different API
configurations (addresses Reviewer 2) and we do a better job of explaining how
results are derived (addresses Reviewer 3).

  \item removed the ``major takeaways". We deleted the headings but left the
conclusions we make as transitions (Reviewer 2). We also add insights into the
results by analyzing the raw numbers we observe.

\end{itemize}

%Use Cases R2, R4
%- Spark discussion R2
%- more emphasis on parallel and distributed computing themes R4
\section{Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}

We add a section to the introduction that scopes the importance of storage for
large scale runtimes and workflows. Parallel and distributed jobs (whether that
be MPI-based or more structured runtimes like Hadoop) need parallel and
distributed storage systems that keep up with the computation. Furthermore, we
highlight the parallel and distributed computing themes (addresses Reviewer 4)
and cast light on them from the context of our prototype. We also add the Spark
discussion proposed by Reviewer 2 to the Evaluation. The goal is to show how a
system such as Cudele would benefit more general runtimes. We also move the
discussion about the potential for HDFS and CephFS subtrees in the same
namespace and the cost of dynamically changing consistency and durability
semantics on a subtree.

\section*{Cosmetic Fixes}
%Terminology
%- users vs. clients R2

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,paper}
\end{document}

### 2017-12-10 paper/build_cover_letter.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+rm -r _minted-cover_letter cover_letter.aux cover_letter.bbl cover_letter.blg cover_letter.log cover_letter.out cover_letter.synctex.gz build.log >> /dev/null 2>&1
+set -x
+cp biblio/references.bib paper.bib
+
+# Build the cover_letter
+docker run --rm \
+  --name latex \
+  --entrypoint=/bin/bash \
+  -v `pwd`/:/mnt \
+  michaelsevilla/texlive:latest -c \
+    "cd /mnt ; \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape cover_letter; \
+     bibtex cover_letter; \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape cover_letter; \
+     pdflatex -synctex=1 -interaction=nonstopmode -shell-escape cover_letter;" &> build.log
+
+ERR=$?
+if [ $ERR != "0" ] ; then
+  echo "ERROR: $ERR"
+  cat build.log
+fi
+
+set +e
+rm -r _minted-cover_letter cover_letter.aux cover_letter.bbl cover_letter.blg cover_letter.log cover_letter.out cover_letter.synctex.gz build.log >> /dev/null 2>&1
+
+echo "SUCCESS"
+

### 2017-12-10 paper/sections/6_results.tex
paper: experimental setup, popper (#27, #30)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-%We run experiments on the same cluster and normalize results to control for
-%hardware differences and to make our results more generally applicable and
-%reproducible\footnote{ This paper adheres to The Popper
-%Convention~\cite{jimenez_popper_2016}, so experiments presented here are
-%available in the repository for this article.  Experiments can be examined in
-%more detail, or even re-run, by visiting the \texttt{[source]} link next to
-%each figure. That link points to a Jupyter notebook that shows the analysis and
-%source code for that graph, which points to an experiment and its artifacts.}.
-%We use a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM,
-%and SSDs.  Each node uses Ubuntu 14.04 and We develop on Ceph's Jewel release,
-%version 10.2.1, which was released in May 2016. All daemons run as a single
-%process (the default setting for Ceph) and the Ceph
-%components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
-%object storage daemons, 1 metadata server, and 29 clients.
+\newcomment{We run experiments on the same cluster and normalize results to
+control for hardware differences and to make our results more generally
+applicable. We use a CloudLab cluster of 34 nodes connected with 10Gbit
+ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.  Each node uses
+Ubuntu 14.04 and we develop on Ceph's Jewel release, version 10.2.1, which was
+released in May 2016. All daemons run as a single process (the default setting
+for Ceph) and the Ceph
+components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
+object storage daemons, 1 metadata server, and 29 clients.}
+
+\newcomment{We understand concerns about cluster size and network, especially
+for systems aimed at HPC but our evaluation focuses on verifying the behavior of
+different techniques in the same global namespace. These techniques have
+already proven to be effective and scalable in systems that specialize an
+entire namespace according to a single optimization strategy (e.g., Lustre,
+BatchFS, DeltaFS), so our evaluation explores the behavior of the
+consistency/durability mechanisms and their effect with/without isolation.}
+
+\newcomment{To make our results reproducible, this paper adheres to The Popper
+Convention~\cite{jimenez_popper_2016}, so experiments presented here are
+available in the repository for this article.  Experiments can be examined in
+more detail, or even re-run, by visiting the \texttt{[source]} link next to
+each figure. That link points to a Jupyter notebook that shows the analysis and
+source code for that graph, which points to an experiment and its artifacts.}
+\addressesissue{1}
 

### 2017-12-11 experiments/baseline-compile/visualize/plot.sh
change name from progFS to cudele
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -11,16 +11,27 @@ cd ../results
 #   "tmp/graphite/whisper/issdm-12/mds_server/req_create.wsp" \
 #   "tmp/graphite/whisper/issdm-12/mds_server/req_readdir.wsp" \
 #   "tmp/graphite/whisper/issdm-12/mds_server/req_lookup.wsp" \
-#   ; do
+#   "tmp/graphite/whisper/issdm-12/mds_server/req_unlink.wsp" \
+#   "tmp/graphite/whisper/issdm-12/mds_server/req_rmdir.wsp" \
+#    ; do
 for m in \
-    "tmp/graphite/whisper/issdm-12/mds_server/req_unlink.wsp" \
-    "tmp/graphite/whisper/issdm-12/mds_server/req_rmdir.wsp" \
+    "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+    "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
+    "tmp/graphite/whisper/issdm-12/nettotals/kbin.wsp" \
     ; do
+  # dump trace
   docker run \
     -v `pwd`/tmp:/tmp \
     --entrypoint=whisper-dump.py \
     michaelsevilla/graphite \
     $m > `basename $m`.out
-  sed -i "s/:/,/g" `basename ${m}`.out
+
+  # clean trace to prepare for pandas
+  docker run \
+    -v `pwd`:/root \
+    --entrypoint=/bin/bash \
+    michaelsevilla/graphite \
+    -c "sed -i \"s/:/,/g\" /root/`basename ${m}`.out"
+
 done
 cd -

### 2017-12-11 experiments/baseline-compile/visualize/clean.sh
tmp
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,6 @@
+#!/bin/bash
+set -x
+cd ../results >> /dev/null 2>&1
+rm *.out
+rm -r tmp >> /dev/null 2>&1
+cd -

### 2017-12-11 jupyter.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+set -e -x
+
+docker run --rm -p 81:8888 \
+  -v $PWD:/home/jovyan/work \
+  jupyter/scipy-notebook:c411f52fcc93

### 2017-12-11 experiments/baseline/docker/cloudlab/install.sh
benchmark: change name
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -36,4 +36,4 @@ sudo ifconfig eth2:0 $ip up
 echo $HOSTNAME | sed "s/\./ /g" | awk '{print $1}' | xargs sudo hostname
 
 # pull the code
-git clone --recursive https://github.com/michaelsevilla/progFS-popper.git /tmp/progFS-popper
+git clone --recursive https://github.com/michaelsevilla/cudele-popper.git /tmp/cudele-popper

### 2017-12-12 experiments/baseline-durability/results-done/journal-cache/plot.sh
benchmark: popperize durability (#27)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,35 +1,0 @@
-#!/bin/bash
-
-set -ex
-for i in `ls *.tar.gz`; do
-  tar xzf $i
-done
-
-# prepare as a csv
-for m in \
-   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
-   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
-   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
-   "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
-  ; do
-  ../../../../../visualize.sh ${m}
-  sed -i "s/:/,/g" `basename ${m}`.out
-done
-
-for osd in 0 1 14 24 27 29 34 40; do
-  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
-    ../../../../../visualize.sh ${m}
-    sed -i "s/:/,/g" `basename ${m}`.out
-    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
-  done
-done
-

### 2017-12-12 experiments/baseline-compile/visualize/plot.sh
benchmark: popperize durability (#27)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -2,23 +2,21 @@
 
 set -ex
 cd ../results
-#for i in `ls *.tar.gz`; do
-#  tar xzf $i
-#done
-#
-## prepare as a csv
-#for m in \
-#   "tmp/graphite/whisper/issdm-12/mds_server/req_create.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_server/req_readdir.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_server/req_lookup.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_server/req_unlink.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_server/req_rmdir.wsp" \
-#    ; do
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
 for m in \
-    "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
-    "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
-    "tmp/graphite/whisper/issdm-12/nettotals/kbin.wsp" \
-    ; do
+   "tmp/graphite/whisper/node-4/mds_server/req_create.wsp" \
+   "tmp/graphite/whisper/node-4/mds_server/req_readdir.wsp" \
+   "tmp/graphite/whisper/node-4/mds_server/req_lookup.wsp" \
+   "tmp/graphite/whisper/node-4/mds_server/req_unlink.wsp" \
+   "tmp/graphite/whisper/node-4/mds_server/req_rmdir.wsp" \
+   "tmp/graphite/whisper/node-4/cputotals/user.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbout.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbin.wsp" \
+   ; do
   # dump trace
   docker run \
     -v `pwd`/tmp:/tmp \

### 2017-12-12 experiments/baseline-durability/results-done/journal-nocache/plot.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-set -ex
-for i in `ls *.tar.gz`; do
-  tar xzf $i
-done
-
-# prepare as a csv
-for m in \
-   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
-   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
-   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
-   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
-  ; do
-  ../../../../../visualize.sh ${m}
-  sed -i "s/:/,/g" `basename ${m}`.out
-done
-
-for osd in 0 1 14 24 27 29 34 40; do
-  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
-    ../../../../../visualize.sh ${m}
-    sed -i "s/:/,/g" `basename ${m}`.out
-    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
-  done
-done
-

### 2017-12-12 experiments/baseline-durability/results-done/nojournal-nocache/plot.sh

@@ -1,37 +1,0 @@
-#!/bin/bash
-
-set -ex
-#for i in `ls *.tar.gz`; do
-#  tar xzf $i
-#done
-#
-## prepare as a csv
-#for m in \
-#   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
-#   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
-#   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
-#   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
-#   "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
-for m in \
-  "tmp/graphite/whisper/issdm-12/meminfo/used.wsp" \
-  ; do
-  ../../../../../visualize.sh ${m}
-  sed -i "s/:/,/g" `basename ${m}`.out
-done
-
-for osd in 0 1 14 24 27 29 34 40; do
-  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
-    ../../../../../visualize.sh ${m}
-    sed -i "s/:/,/g" `basename ${m}`.out
-    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
-  done
-done
-

### 2017-12-12 experiments/baseline-durability/run.sh

@@ -3,18 +3,18 @@
 set -ex
 
 # setup the docker command
+ROOT=`dirname $PWD | xargs dirname`
 RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
 ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
-CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
+CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
 SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
 
-#for site in "nojournal-cache" "nojournal-nocache" "journal-cache" "journal-nocache"; do
-for site in "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
+for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
   nclients=3
 
   # configure ceph and setup results directory
   mkdir -p results/$site/logs || true
-  cp site/* site/roles/ceph-ansible || true
+  cp site/* $ROOT/ansible/ceph || true
   cp site_confs/${site}.yml site/group_vars/all
   cp -r site/group_vars site/roles/ceph-ansible/
   cp inventory/${nclients}client site/hosts
@@ -30,9 +30,5 @@ for site in "journal120-cache" "journal150-cache" "journal180-cache" "journal30-
     ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
   done
   
-  ## baseline the drop delays
-  #./ansible-playbook.sh -e site=$site -e nfiles=50000 -e drop_delay=15 ../workloads/stat.yml
-  #./ansible-playbook.sh -e site=$site -e nfiles=50000 -e drop_delay=15 ../workloads/touch.yml
-
   ./ansible-playbook.sh -e site=$site collect.yml
 done

### 2017-12-12 experiments/baseline-durability/visualize/jupyter.sh

@@ -1,6 +1,0 @@
-#!/bin/bash
-
-set -e -x
-mkdir graphs || true
-docker run --rm -v  `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93
-

### 2017-12-12 experiments/baseline-interfere/visualize/jupyter.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-if [ $# -lt 1 ]; then
-  echo "USAGE <re-parse>"
-  echo "- where re-parse is 0 or 1"
-  exit 1
-fi
-
-set -e -x
-function uncompress() {
-  for i in `ls *.tar.gz`; do
-    tar xzf $i
-  done
-  
-  GRAPHITE="docker run -v `pwd`/tmp:/tmp --entrypoint=/bin/bash -it michaelsevilla/graphite -c"
-  
-  # prepare as a csv
-  for m in \
-     "tmp/graphite/whisper/issdm-18/mds_server/handle_client_request_tput.wsp" \
-     "tmp/graphite/whisper/issdm-18/mds_server/req_lookup.wsp" \
-     ; do
-    $GRAPHITE "whisper-dump.py $m > tmp/`basename $m.out`"
-    $GRAPHITE "sed -i 's/:/,/g' tmp/`basename ${m}.out`"
-  done
-}
-
-if [ "$1" -eq 1 ]; then
-  cd ../results-paper/nojournal-cache/
-  uncompress
-  cd -
-fi
-mkdir graphs || true
-docker run --rm -v  `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93
-

### 2017-12-12 experiments/baseline-interfere/visualize/plot.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+set -ex
+cd ../results-paper/nojournal-cache/
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+GRAPHITE="docker run -v `pwd`/tmp:/tmp --entrypoint=/bin/bash -it michaelsevilla/graphite -c"
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-18/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-18/mds_server/req_lookup.wsp" \
+   ; do
+  $GRAPHITE "whisper-dump.py $m > tmp/`basename $m.out`"
+  $GRAPHITE "sed -i 's/:/,/g' tmp/`basename ${m}.out`"
+done
+cd -

### 2017-12-12 experiments/baseline-compile/run.sh
benchmark: re-run compile experiment
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,9 +34,10 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-./teardown.sh
 sudo rm -rf results; mkdir results
 cp configs_$SITE/all ansible/group_vars/all 
+cp configs_$SITE/clients1 hosts
+./teardown.sh
 $DOCKER ceph.yml monitor.yml \
         /workloads/compile.yml \
         /popper/ansible/collect.yml

### 2017-12-12 experiments/baseline-interfere/visualize/clean.sh
tmp
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,6 @@
+#!/bin/bash
+set -x
+cd ../results-paper/nojournal-cache/
+rm *.out >> /dev/null 2>&1
+rm -r tmp >> /dev/null 2>&1
+cd -

### 2017-12-12 experiments/baseline-compile/visualize/plot.sh
benchmark: popperize compile/interfere, rerun compile (#38, #27)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -14,9 +14,11 @@ for m in \
    "tmp/graphite/whisper/node-4/mds_server/req_unlink.wsp" \
    "tmp/graphite/whisper/node-4/mds_server/req_rmdir.wsp" \
    "tmp/graphite/whisper/node-4/cputotals/user.wsp" \
+   "tmp/graphite/whisper/node-4/cpuload/avg1.wsp" \
    "tmp/graphite/whisper/node-4/nettotals/kbout.wsp" \
    "tmp/graphite/whisper/node-4/nettotals/kbin.wsp" \
-   ; do
+  ; do
+
   # dump trace
   docker run \
     -v `pwd`/tmp:/tmp \
@@ -32,4 +34,28 @@ for m in \
     -c "sed -i \"s/:/,/g\" /root/`basename ${m}`.out"
 
 done
+
+count=0
+for m in \
+  "tmp/graphite/whisper/node-1/diskinfo/writekbs/sdc.wsp" \
+  "tmp/graphite/whisper/node-2/diskinfo/writekbs/sdc.wsp" \
+  "tmp/graphite/whisper/node-3/diskinfo/writekbs/sdc.wsp" \
+  ; do
+
+  count=$((count+1))
+  # dump trace
+  docker run \
+    -v `pwd`/tmp:/tmp \
+    --entrypoint=whisper-dump.py \
+    michaelsevilla/graphite \
+    $m > node$count-`basename $m`.out
+
+  # clean trace to prepare for pandas
+  docker run \
+    -v `pwd`:/root \
+    --entrypoint=/bin/bash \
+    michaelsevilla/graphite \
+    -c "sed -i \"s/:/,/g\" /root/node$count-`basename ${m}`.out"
+done
+
 cd -

### 2017-12-12 paper/sections/3_posix_overheads.tex
paper: final version
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: popperize all graphs (#27, #28, #37, #40)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
-\caption{Create-heavy workloads (such as \texttt{untar}) incur the highest disk, network, and
-CPU utilization because of the consistency and durability demands of
-CephFS.}\label{fig:overhead-creates}
+
+\caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-compile/visualize/viz.ipynb}{source}]
+Create-heavy workloads (such as \texttt{untar}) incur the highest disk,
+network, and CPU utilization because of the consistency and durability demands
+of CephFS.}\label{fig:overhead-creates}
+
 \end{figure}
@@
       \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
-      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
+      \includegraphics[width=0.95\linewidth]{graphs/slowdown-journal.png}
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-durability/visualize/viz.ipynb}{source}]
+      Effect of Journal} \label{fig:overhead-a}
   \end{subfigure}
@@
       \centering
-      \includegraphics[width=1.2\linewidth]{graphs/behavior-interfere.png}
-      \caption{Interference forces \texttt{lookup()}s}
+      \includegraphics[width=1.1\linewidth]{graphs/behavior-interfere.png}
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
+      Interference increases RPCs}
       \label{fig:overhead-b}
@@
       \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-scale.png}
-      \caption{Interference hurts variability}
+      \includegraphics[width=1\linewidth]{graphs/slowdown-interfere-scale.png}
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-creates/visualize/viz.ipynb}{source}]
+      Interference hurts variability}
       \label{fig:overhead-c}
   \end{subfigure}
-  \caption{The overhead of durability and strong consistency in CephFS.
-  (a) shows the effect of different journal segment sizes, which are streamed
-  into the object store for fault tolerance. (b) and (c) show that when a second
-  client ``interferes", capabilities are revoked and metadata servers do more
-  work.  \label{fig:overhead}}
+  \caption{ The overhead of durability and strong consistency in CephFS.  (a)
+  shows the effect of different journal segment sizes, normalized to a setup
+  without a journal (``No Journal").  The journal is streamed into the object
+  store for fault tolerance. (b) and (c) show that when a second client
+  ``interferes", capabilities are revoked and metadata servers do more work.
+  \label{fig:overhead}}
 \end{figure*}

### 2017-12-12 paper/sections/6_results.tex
paper: popperize all graphs (#27, #28, #37, #40)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
-\caption{Performance of each mechanism (left) and building the
-consistency/durability semantics of real-world systems (right) for 100K files
-creates from a single client. Results are normalized to the runtime of writing
-events to the client's in-memory journal.  \label{fig:composable-mechanisms}}
+\caption{
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
+Performance of each mechanism (left) and building the consistency/durability
+semantics of real-world systems (right) for 100K files creates from a single
+client. Results are normalized to the runtime of writing events to the client's
+in-memory journal.  \label{fig:composable-mechanisms}}
 \end{figure}
@@
 {\it Journal Size:} We measure the amount of storage per journal update to be
-about 2.5\text{KB}. The storage footprint scales linearly with the number of
+about 2.5KB. The storage footprint scales linearly with the number of
 metadata creates and suggests that updates for a million files would be 2.5KB
@@
       \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
-      \caption{Use Case 1} \label{fig:mergescale}
+      \caption{
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
+      Use Case 1} \label{fig:mergescale}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
-      \caption{Use Case 2}
-      \label{fig:block-allow}
+      \caption{
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
+      Use Case 2} \label{fig:block-allow}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
-      \caption{Use Case 3}
-      \label{fig:slowdown-sync}
+      \caption{
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
+      Use Case 3} \label{fig:slowdown-sync}
   \end{subfigure}
-\caption{The performance and features of Cudele. (a) shows the cost of
-merging client journals at the metadata server. Shipping and merging journals of updates scales better than RPCs
-because there are less messages and consistency/durability code paths are
-bypassed. In (b), the allow/block API isolates directories from interfering
-clients. (c) is the slowdown of a single client syncing updates to the global
-namespace. The inflection point is the trade-off of frequent updates vs. larger
-journal files.\label{fig:use-cases}}
+\caption{
+The performance and features of Cudele. (a) shows the cost of merging client
+journals at the metadata server. Shipping and merging journals of updates
+scales better than RPCs because there are less messages and
+consistency/durability code paths are bypassed. In (b), the allow/block API
+isolates directories from interfering clients. (c) is the slowdown of a single
+client syncing updates to the global namespace. The inflection point is the
+trade-off of frequent updates vs. larger journal files.\label{fig:use-cases}}
 \end{figure*}
@@
 double the capacity of a single metadata server using the RPC strategy to
-maintain strong consistency. Related work has pointed to on-disk metadata
-format as the primary factor for improved scalability; but our results show
-that the weakened consistency and durability enables the scalability
-improvements. The on-disk metadata formats in this experiment are the same for
-all schemes.
+maintain strong consistency\footnote{\newcomment{Note that for the 20 clients
+bar in Figure~\ref{fig:mergescale}, we plot the throughput for using 18 clients
+(the maximum number of clients supported by RPCs)}}. Related work has pointed
+to on-disk metadata format as the primary factor for improved scalability; but
+our results show that the weakened consistency and durability enables the
+scalability improvements. The on-disk metadata formats in this experiment are
+the same for all schemes.
 

### 2017-12-12 experiments/cudele-partialreads/visualize/jupyter.sh

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-set -e -x
-docker run --rm -v `dirname $PWD`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93

### 2017-12-12 experiments/baseline-compile/visualize/jupyter.sh
paper: popperize all graphs (#27, #28, #37, #40)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-set -e -x
-docker run --rm -v `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93

### 2017-12-12 experiments/cudele-mergescale/visualize/jupyter.sh

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-set -e -x
-docker run --rm -v  `dirname $PWD | xargs dirname`:/home/jovyan/work -p 81:8888 jupyter/scipy-notebook:c411f52fcc93

### 2017-12-12 paper/build.sh

@@ -20,9 +20,10 @@ ERR=$?
 if [ $ERR != "0" ] ; then
   echo "ERROR: $ERR"
   cat build.log
+  exit 1
 fi
 
-set +e
+set +x
 rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
 
 echo "SUCCESS"

### 2017-12-12 paper/sections/6_results.tex
benchmark: make source available (#31)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 each figure. That link points to a Jupyter notebook that shows the analysis and
-source code for that graph, which points to an experiment and its artifacts.}
-\addressesissue{1}
+source code for that graph, which points to an experiment and its artifacts.
+The source code is available on the Cudele
+branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
+of our Ceph fork.} \addressesissue{1}
 

### 2017-12-13 experiments/cudele-mechanisms/run.sh
benchmark: re-run cudele mechanisms (#38)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -35,7 +35,7 @@ if [ ! -z $1 ]; then
 fi
 
 for run in 0 1 2; do
-  for nfiles in 10 100 1000 10000 100000 200000 400000 600000 800000 1000000 10000000; do
+  for nfiles in 100000; do
     mkdir results || true
     ./teardown.sh
     cp configs_$SITE/hosts hosts

### 2017-12-13 experiments/run.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+set -ex
+cd baseline-durability
+./run.sh
+cd -
+
+cd cudele-mechanisms
+./run.sh
+cd -

### 2017-12-13 paper/cover_letter.tex
paper: experimental setup, popper (#27, #30)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: fix graphs and add baselines (#33, #34)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
-feedback:
+feedback; Reviewer 1 and 3's comments are addressed in issues I and II,
+Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
+comments are addressed in issues II and III.
 
@@
 %- servers, network, storage R1, R3
-\section{No Experimental Setup (Reviewers 1, 2, and 3)}
+\section*{Issue I: No Experimental Setup (Reviewers 1, 2, and 3)}
 
-We apologize for the omission of experimental setup and source code details.
-For the experimental setup, we have added details about the cluster we ran on
-in Table III as well as a description in the text (addresses Reviewers 1 and
-3). For the source code, we have added code snippets to the Evaluation to
-concretize what was actually done (addresses Reviewer 1).
-
-Additionally, we have added paper artifacts online.  For the experimental
-setup, we have made the infrastructure code available and added links after
-each figure to show exactly how experiments are run.  This infrastructure code
+We apologize for the omission of experimental setup and source code details. To
+provide a more comprehensive view of our experiments, we added text to the
+paper and made paper artifacts available online. For the experimental setup, we
+describe the cluster setup (hardware, software, etc.) in Section~{\S}V.  We
+have also made the infrastructure code available and added links after each
+figure to show exactly how experiments are run.  This infrastructure code
 contains scripts to deploy the system, run experiments, and gather results.
@@
 Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
-aims to make research reproducible.  We also make the source code available and
-linked in the paper.
+aims to make research reproducible.  For the source code, we have added code
+snippets to the evaluation to concretize what was actually done and made the
+source code available online. A link is provided in a footnote in
+Section~{\S}V.
 
@@
 %- cost of dynamically changing consistency/durability not presented R1
-\section{Structure and Layout of Evaluation (All Reviewers)}
+\section*{Issue II: Structure and Layout of Evaluation (All Reviewers)}
 
-One pervasive issue across reviews is that the results and contributions we
-cite in the introduction are not validated, explained, or even mentioned in the
-Evaluation.  To fix this, we separate out future work to make the contributions
-of this paper more explicit (addresses Reviewer 1) and we connect all numbers
-we cite in the introduction to the evaluation (addresses all reviewers).  To
-further clarify the Evaluation, we:
+All reviewers note that the results and contributions we cite in the
+introduction are not validated, explained, or even mentioned in the evaluation.
+One confusing component of this issue is that we mixed future work with the
+contributions of this paper; to make the contributions more explicit we remove
+future work from the introduction and add it to the future work section in
+Section~{\S}VI. We also connect all numbers we cite in the introduction to the
+evaluation. To further clarify the evaluation, we:
 
@@
 
-  \item added Table IV which includes raw baseline, Cudele, and hardware limit
-numbers. This adds raw numbers and helps us make comparisons to hardware
-speeds. Ultimately, we hope this adds insight into the results (addresses
-Reviewers 2 nad 3).
+  \item added raw numbers (as annotations) to the baselines in each figure.
+This helps the reader calculate the raw numbers instead of just relying on
+overhead and speedup numbers.
 
-  \item re-organized the section to make experiments more independent
-and removed cross-references (addresses Reviewer 2). Experiments are now
-self-contained so the reader can see the effects of different API
-configurations (addresses Reviewer 2) and we do a better job of explaining how
-results are derived (addresses Reviewer 3).
+  \item re-organized the section to make experiments more independent and
+removed cross-references. Experiments are now self-contained so the reader can
+see the effects of different API configurations and we do a better job of
+explaining how results are derived.
 
   \item removed the ``major takeaways". We deleted the headings but left the
-conclusions we make as transitions (Reviewer 2). We also add insights into the
-results by analyzing the raw numbers we observe.
+conclusions we make as transitions. We also add insights into the results by
+analyzing the raw numbers we observe in comparison to hardware capabilities.
 
@@
 %- more emphasis on parallel and distributed computing themes R4
-\section{Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}
+\section*{Issue III: Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}
 
 We add a section to the introduction that scopes the importance of storage for
-large scale runtimes and workflows. Parallel and distributed jobs (whether that
-be MPI-based or more structured runtimes like Hadoop) need parallel and
-distributed storage systems that keep up with the computation. Furthermore, we
-highlight the parallel and distributed computing themes (addresses Reviewer 4)
-and cast light on them from the context of our prototype. We also add the Spark
-discussion proposed by Reviewer 2 to the Evaluation. The goal is to show how a
-system such as Cudele would benefit more general runtimes. We also move the
-discussion about the potential for HDFS and CephFS subtrees in the same
-namespace and the cost of dynamically changing consistency and durability
-semantics on a subtree.
+large scale runtimes and workflows. Parallel and distributed jobs (whether they
+are MPI-based, MapReduce-based, etc.) need parallel and distributed storage
+systems that keep up with the computation. Furthermore, we highlight the
+parallel and distributed computing themes and cast light on them from the
+context of our prototype. We also add the Spark discussion proposed by to the
+evaluation. The goal is to show how a system such as Cudele would benefit from
+more general runtimes. We also move the discussion about the potential for HDFS
+and CephFS subtrees in the same namespace and the cost of dynamically changing
+consistency and durability semantics on a subtree.
 
@@
 %- users vs. clients R2
+\begin{itemize}
+  \item Reviewer 1: fixed user vs. client terminology in Section~{\S}III
+  \item Reviewer 1: clarified that the cost of dynamically changing semantics is future work Sections~{\S}III and ~{\S}V.E
+  \item Reviewer 1: added source code pointer in Section~{\S}IV
+  \item Reviewer 2: remove major takeways in Section~{\S}V
+  \item Reviewer 2: remove major cross-references in Section~{\S}V
+  \item Reviewer 3: quantified speedups with figure annotations in Section~{\S}V
+  \item Reviewers 1 and 3: add servers, network, and storage setups in Section~{\S}V
+  \item Reviewer 4: add new section ({\S}V.F) describing how Cudele would work with a system like Spark
+\end{itemize}
 

### 2017-12-13 paper/sections/1_introduction.tex
paper: fix graphs and add baselines (#33, #34)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-The last contribution lays the groundwork for future work on our prototype. It
-is more scalable than the current practice of mounting different storage systems in a global
-namespace because there is no need to provision dedicated storage clusters to
-applications or move data between these systems.  For example, the results of a
-Hadoop job do not need to be migrated into a Ceph file system (CephFS) for other processing; instead
-the user can change the semantics of the HDFS subtree into a CephFS subtree.
-This may cause metadata/data movement to make things strongly consistent again
-but this is superior to moving all data across file system boundaries. Our prototype
+\oldcomment{The last contribution lays the groundwork for future work on our
+prototype. It is more scalable than the current practice of mounting different
+storage systems in a global namespace because there is no need to provision
+dedicated storage clusters to applications or move data between these systems.
+For example, the results of a Hadoop job do not need to be migrated into a Ceph
+file system (CephFS) for other processing; instead the user can change the
+semantics of the HDFS subtree into a CephFS subtree.  This may cause
+metadata/data movement to make things strongly consistent again but this is
+superior to moving all data across file system boundaries. Our prototype
 enables studies that adjust these semantics over {\it time and space}, where
 subtrees can change their semantics and migrate around the cluster without ever
-moving the data they reference.
+moving the data they reference.}
 

### 2017-12-13 paper/sections/6_results.tex
paper: fix graphs and add baselines (#33, #34)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
-Performance of each mechanism (left) and building the consistency/durability
+\newcomment{Overhead of processing 100K events for each mechanism in
+Table~\ref{table:mechanisms}. Results are normalized to the runtime of writing
+events to client memory. The bars on the far right show the overhead of
+building consistency/durability semantics of real world systems.}
+\oldcomment{Performance of each mechanism (left) and building the consistency/durability
 semantics of real-world systems (right) for 100K files creates from a single
 client. Results are normalized to the runtime of writing events to the client's
-in-memory journal.  \label{fig:composable-mechanisms}}
+in-memory journal.}\label{fig:composable-mechanisms}}
 \end{figure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
-      \caption{
-      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
-      Use Case 1} \label{fig:mergescale}
+      \caption{\texttt{create}: done in parallel on clients}
+      \label{fig:mergescale}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
-      \caption{
-      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
-      Use Case 2} \label{fig:block-allow}
+      \caption{Subtree semantics: block interference}
+      \label{fig:block-allow}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
-      \caption{
+      \caption{Syncing to global namespace}
+      \label{fig:slowdown-sync}
+  \end{subfigure}
+\caption{ The performance and features of Cudele. (a)
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
+shows the cost of merging client journals at the metadata server. Shipping and
+merging journals of updates scales better than RPCs because there are less
+messages and consistency/durability code paths are bypassed. (b)
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
+the allow/block API isolates directories from interfering clients. (c)
 [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
-      Use Case 3} \label{fig:slowdown-sync}
-  \end{subfigure}
-\caption{
-The performance and features of Cudele. (a) shows the cost of merging client
-journals at the metadata server. Shipping and merging journals of updates
-scales better than RPCs because there are less messages and
-consistency/durability code paths are bypassed. In (b), the allow/block API
-isolates directories from interfering clients. (c) is the slowdown of a single
-client syncing updates to the global namespace. The inflection point is the
-trade-off of frequent updates vs. larger journal files.\label{fig:use-cases}}
+is the slowdown of a single client syncing updates to the global namespace. The
+inflection point is the trade-off of frequent updates vs. larger journal
+files.} \label{fig:use-cases}
 \end{figure*}
@@
 
+\newcomment{We have shown that our API and framework is an effective
+abstraction for letting users custom fit subtrees to applications.  It is more
+scalable than the current practice of mounting different storage systems in a
+global namespace because there is no need to provision dedicated storage
+clusters to applications or move data between these systems.  For example, the
+results of a Hadoop job do not need to be migrated into a Ceph file system
+(CephFS) for other processing; instead the user can change the semantics of the
+HDFS subtree into a CephFS subtree.  This may cause metadata/data movement to
+make things strongly consistent again but this is superior to moving all data
+across file system boundaries. Our prototype enables studies that adjust these
+semantics over {\it time and space}, where subtrees can change their semantics
+and migrate around the cluster without ever moving the data they reference.}
+
 Cudele and the experiments we present here prompt many avenues for future work.

### 2017-12-13 paper/build.sh

@@ -18,6 +18,7 @@ docker run --rm \
 
 ERR=$?
 if [ $ERR != "0" ] ; then
+  set +x
   echo "ERROR: $ERR"
   cat build.log
   exit 1

### 2017-12-13 paper/build_cover_letter.sh

@@ -18,8 +18,10 @@ docker run --rm \
 
 ERR=$?
 if [ $ERR != "0" ] ; then
+  set +x
   echo "ERROR: $ERR"
   cat build.log
+  exit 1
 fi
 
 set +e

### 2017-12-14 experiments/baseline-durability/run.sh
benchmark: mech/dur results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -4,21 +4,22 @@ set -ex
 
 # setup the docker command
 ROOT=`dirname $PWD | xargs dirname`
-RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root -v $ROOT/ansible/srl:/popper/ansible/roles/srl/"
 ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
 CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
 SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
 
-for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
-  nclients=3
+for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache" "nojournal-cache"; do
+  #nclients=3
 
   # configure ceph and setup results directory
   mkdir -p results/$site/logs || true
   cp site/* $ROOT/ansible/ceph || true
   cp site_confs/${site}.yml site/group_vars/all
-  cp -r site/group_vars site/roles/ceph-ansible/
-  cp inventory/${nclients}client site/hosts
-  cp inventory/${nclients}client site/roles/ceph-ansible/hosts
+  cp -r site/group_vars $ROOT/ansible/ceph/
+  #cp inventory/${nclients}client site/hosts
+  #cp inventory/${nclients}client $ROOT/ansible/ceph//hosts
+  cp site/hosts $ROOT/ansible/ceph/hosts
 
   # cleanup and start ceph
   $SRL_ANSIBLE cleanup.yml

### 2017-12-14 experiments/cudele-test/ansible-playbook.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+if [ -z $1 ]; then
+  echo "huh? usage: $0 <command>"
+  exit 1
+fi
+
+set -ex
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root/site"
+ANSIBLE="michaelsevilla/ansible"
+
+# cleanup and start ceph
+$RUN -v `pwd`:/root $ANSIBLE $@

### 2017-12-14 experiments/cudele-test/ansible.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+if [ -z $1 ]; then
+  echo "huh? usage: $0 <command>"
+  exit 1
+fi
+
+set -e
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root --entrypoint=/usr/bin/ansible"
+ANSIBLE="michaelsevilla/ansible"
+
+# cleanup and start ceph
+$RUN -v `pwd`/site:/root $ANSIBLE "$@"

### 2017-12-14 experiments/cudele-test0/results0/journal180-cache/plot.sh

@@ -1,0 +1,35 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
+   "tmp/graphite/whisper/issdm-12/meminfo/used.wsp" \
+  ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+#for osd in 0 1 14 24 27 29 34 40; do
+#  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+#    ../../../../../visualize.sh ${m}
+#    sed -i "s/:/,/g" `basename ${m}`.out
+#    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+#  done
+#done
+

### 2017-12-14 experiments/cudele-test0/run.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+ROOT=`dirname $PWD | xargs dirname`
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root -v $ROOT/ansible/srl:/popper/ansible/roles/srl/"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+#for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
+for site in "nojournal-cache"; do
+  #nclients=3
+
+  # configure ceph and setup results directory
+  mkdir -p results/$site/logs || true
+  cp site/* $ROOT/ansible/ceph || true
+  cp site_confs/${site}.yml site/group_vars/all
+  cp -r site/group_vars $ROOT/ansible/ceph/
+  #cp inventory/${nclients}client site/hosts
+  #cp inventory/${nclients}client $ROOT/ansible/ceph//hosts
+  cp site/hosts $ROOT/ansible/ceph/hosts
+
+  # cleanup and start ceph
+  $SRL_ANSIBLE cleanup.yml
+  $CEPH_ANSIBLE ceph.yml cephfs.yml
+  $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+  
+  # warmup and get baseline
+  for i in `seq 0 2`; do
+    ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
+  done
+  
+  ./ansible-playbook.sh -e site=$site collect.yml
+done

### 2017-12-14 experiments/run.sh

@@ -1,5 +1,6 @@
 #!/bin/bash
 set -ex
+
 cd baseline-durability
 ./run.sh
 cd -

### 2017-12-14 experiments/baseline-creates1/docker/cloudlab/install.sh
benchmark: re-run mechanisms and found problem - inaccurate durability, interfere, and merge results (#42) - fix networking problem - add multiple processes per node (#43)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,39 @@
+#!/bin/bash
+# Author: Michael Sevilla
+# Date: 10-8--17
+# Description: all this does is install Docker
+
+set -ex
+
+# the below is adapted from:
+# - https://docs.docker.com/engine/installation/linux/docker-ce/ubuntu/
+sudo apt-get update -y
+sudo apt-get install -y \
+    vim \
+    screen \
+    gdisk \
+    linux-image-extra-$(uname -r) \
+    linux-image-extra-virtual \
+    apt-transport-https \
+    ca-certificates \
+    curl \
+    software-properties-common
+
+curl -fsSL https://download.docker.com/linux/ubuntu/gpg | sudo apt-key add -
+sudo apt-key fingerprint 0EBFCD88
+sudo add-apt-repository \
+   "deb [arch=amd64] https://download.docker.com/linux/ubuntu \
+   $(lsb_release -cs) \
+   stable"
+
+sudo apt-get update -y
+sudo apt-get install -y docker-ce
+
+# setup an alias (for MPI related jobs)
+sudo ifconfig eth2 up || true
+ip=`ifconfig | grep 10.10.1. | awk '{print $2}' | sed 's/addr://g'`
+sudo ifconfig eth2:0 $ip up
+echo $HOSTNAME | sed "s/\./ /g" | awk '{print $1}' | xargs sudo hostname
+
+# pull the code
+git clone --recursive https://github.com/michaelsevilla/progFS-popper.git /tmp/progFS-popper

### 2017-12-14 experiments/baseline-creates1/visualize/plot.sh

@@ -1,0 +1,38 @@
+#!/bin/bash
+
+set -ex
+cd ../$1
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/node-4/nettotals/kbout/eth0.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbout/eth1.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbout/eth2.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbout/eth3.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbin/eth0.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbin/eth1.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbin/eth2.wsp" \
+   "tmp/graphite/whisper/node-4/nettotals/kbin/eth3.wsp" \
+  ; do
+  FNAME=`dirname $m | xargs basename`-`basename $m`.out
+
+  # dump trace
+  docker run \
+    -v `pwd`/tmp:/tmp \
+    --entrypoint=whisper-dump.py \
+    michaelsevilla/graphite \
+    $m > $FNAME
+
+  # clean trace to prepare for pandas
+  docker run \
+    -v `pwd`:/root \
+    --entrypoint=/bin/bash \
+    michaelsevilla/graphite \
+    -c "sed -i \"s/:/,/g\" /root/$FNAME"
+
+done
+
+cd -

### 2017-12-14 experiments/baseline-durability/run.sh

@@ -4,31 +4,33 @@ set -ex
 
 # setup the docker command
 ROOT=`dirname $PWD | xargs dirname`
-RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root -v $ROOT/ansible/srl:/popper/ansible/roles/srl/"
 ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
 CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
 SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
 
-for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
-  nclients=3
+for nclients in 1 5 10 15 18 20; do
+    cp inventory_cloudlab/${nclients}client site/hosts
+    cp site/hosts $ROOT/ansible/ceph/hosts
+  for site in "nojournal-cache" "journal210-cache" "journal120-cache" "journal30-cache"; do
 
-  # configure ceph and setup results directory
-  mkdir -p results/$site/logs || true
-  cp site/* $ROOT/ansible/ceph || true
-  cp site_confs/${site}.yml site/group_vars/all
-  cp -r site/group_vars site/roles/ceph-ansible/
-  cp inventory/${nclients}client site/hosts
-  cp inventory/${nclients}client site/roles/ceph-ansible/hosts
+    # configure ceph and setup results directory
+    mkdir -p results/$site/logs || true
+    cp site/* $ROOT/ansible/ceph || true
+    cp site_confs/${site}.yml site/group_vars/all
+    cp -r site/group_vars $ROOT/ansible/ceph/
 
-  # cleanup and start ceph
-  $SRL_ANSIBLE cleanup.yml
-  $CEPH_ANSIBLE ceph.yml cephfs.yml
-  $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
-  
-  # warmup and get baseline
-  for i in `seq 0 2`; do
-    ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
+    # cleanup and start ceph
+    $SRL_ANSIBLE cleanup.yml
[…]

### 2017-12-14 experiments/baseline-durability/tmp/results0/journal30-cache/plot.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
+   "tmp/graphite/whisper/issdm-12/meminfo/used.wsp" \
+   "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
+  ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+#for osd in 0 1 14 24 27 29 34 40; do
+#  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+#    ../../../../../visualize.sh ${m}
+#    sed -i "s/:/,/g" `basename ${m}`.out
+#    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+#  done
+#done
+

### 2017-12-14 experiments/baseline-durability/tmp/results0/nojournal-cache/plot.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+set -ex
+for i in `ls *.tar.gz`; do
+  tar xzf $i
+done
+
+# prepare as a csv
+for m in \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino+.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_mem/ino-.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_bottom.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_with_caps.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pin_tail.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_pinned.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds/inodes_top.wsp" \
+   "tmp/graphite/whisper/issdm-12/mds_server/handle_client_request_tput.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/user.wsp" \
+   "tmp/graphite/whisper/issdm-12/cputotals/sys.wsp" \
+   "tmp/graphite/whisper/issdm-12/nettotals/kbout.wsp" \
+    "tmp/graphite/whisper/issdm-12/meminfo/used.wsp"\
+  ; do
+  ../../../../../visualize.sh ${m}
+  sed -i "s/:/,/g" `basename ${m}`.out
+done
+
+for osd in 0 1 14 24 27 29 34 40; do
+  for m in "tmp/graphite/whisper/issdm-${osd}/disktotals/writekbs.wsp"; do
+    ../../../../../visualize.sh ${m}
+    sed -i "s/:/,/g" `basename ${m}`.out
+    mv `basename ${m}`.out `basename ${m}`-issdm-${osd}.out
+  done
+done
+

### 2017-12-14 experiments/cudele-mechanisms/run.sh

@@ -34,14 +34,19 @@ if [ ! -z $1 ]; then
   exit
 fi
 
+cp configs_$SITE/hosts hosts
 for run in 0 1 2; do
   for nfiles in 100000; do
-    mkdir results || true
-    ./teardown.sh
-    cp configs_$SITE/hosts hosts
-    $DOCKER -e nfiles=$nfiles ceph.yml monitor.yml /workloads/vapply.yml
+    for stream in "stream" "nostream"; do
+      mkdir -p results/${nfiles}/logs || true
+      cp configs_cloudlab/${stream}.yml ansible/group_vars/all
+      ./teardown.sh
+      $DOCKER -e nfiles=$nfiles -e stream=$stream ceph.yml /workloads/journal-rpcs.yml
+      ./teardown.sh
+      $DOCKER -e nfiles=$nfiles -e stream=$stream ceph.yml /workloads/journal-vapply.yml
+    done
   done
-  mv results results-$SITE-clients$clients-run$run-parmsweep
+  mv results results-$SITE-$stream-run$run
 done
 
 exit 0

### 2017-12-14 experiments/cudele-test/baseliner.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+# cleanup
+cd results; ls | grep -v visualize.ipynb | xargs rm; cd -
+set -ex
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root/site"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+
+# run the baseliner
+$RUN -v `pwd`:/root $ANSIBLE \
+  ../workloads/radosbench.yml \
+  ../workloads/netbench.yml \
+  ../workloads/osdbench.yml
+
+

### 2017-12-14 experiments/cudele-test/metawrites.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+if [ -z $1 ]; then
+  echo "Uh oh. Usage: $0 <number of files to create>"
+  exit 1
+fi
+
+# cleanup
+set -ex
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root/site"
+ANSIBLE="michaelsevilla/ansible --extra-vars "nfiles=$1" --forks 50 --skip-tags package-install,with_pkg"
+
+# run the benchmark
+$RUN -v `pwd`:/root $ANSIBLE ../workloads/metawrites.yml

### 2017-12-14 experiments/cudele-test/run.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+ROOT=`dirname $PWD | xargs dirname`
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root -v $ROOT/ansible/srl:/popper/ansible/roles/srl/"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+#for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache"; do
+for site in "nojournal-cache"; do
+  #nclients=3
+
+  # configure ceph and setup results directory
+  mkdir -p results/$site/logs || true
+  cp site/* $ROOT/ansible/ceph || true
+  cp site_confs/${site}.yml site/group_vars/all
+  cp -r site/group_vars $ROOT/ansible/ceph/
+  #cp inventory/${nclients}client site/hosts
+  #cp inventory/${nclients}client $ROOT/ansible/ceph//hosts
+  cp site/hosts $ROOT/ansible/ceph/hosts
+
+  # cleanup and start ceph
+  $SRL_ANSIBLE cleanup.yml
+  $CEPH_ANSIBLE ceph.yml cephfs.yml
+  $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+  
+  # warmup and get baseline
+  for i in 1; do
+    ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
+  done
+  
+  ./ansible-playbook.sh -e site=$site collect.yml
+done

### 2017-12-14 experiments/cudele-test/site/group_vars/ceph-stats_conf/ceph_perf_dump.py

@@ -1,0 +1,85 @@
+#! /usr/bin/python
+import os
+import sys
+import pickle
+import struct
+import socket
+import json
+import argparse
+import time
+
+# recurse down the json object and convert to paths
+def json_to_path(metrics, date, path, data):
+  if type(data) == dict:
+    for x in data:
+      json_to_path(metrics, date, path + "." + x, data[x])
+  else:
+    metrics[path] = (date, data)
+  return metrics
+
+# manage as dictionary so we can search it (for throughputs)
+#   [(name, (timestamp, value)), ...]
+def json_to_dict(date, data):
+  metrics = json_to_path({}, date, socket.gethostname(), data)
+  return metrics
+
+# convert a metric to throughput
+def throughput(name, pmetric, metric, interval):
+  p = pmetrics[name]
+  m = metrics[name]
+  t = float(m[0]) - float(p[0])
+  d = float(m[1]) - float(p[1])
+  tput = (m[0], d/interval)
+  print "throughput: tput=" + str(tput) + " m=" + str(m) + " p=" + str(p) + " time=" + str(t) + " diff=" + str(d)
+  return tput
+  
+# main
+parser = argparse.ArgumentParser(description='Parse Ceph perf counter dump and send to graphite')
+parser.add_argument('ip', metavar='ip', type=str, help='where the graphite collector daemon (carbon) lives')
+parser.add_argument('--port', metavar='p', type=int, default=2004, help='port of the graphite collector daemon (carbon)')
+parser.add_argument('--interval', metavar='i', type=int, default=1, help='port of the graphite collector daemon (carbon)')
+parser.add_argument('--jsonfile', metavar='f', type=str, default="/tmp/ceph_perf_dump.json", help='where the json dump is')
+parser.add_argument('--throughput', metavar='t', type=bool, default=True, help='whether to collect throughput')
+args = parser.parse_args()
+print "args:", args
+
+# connect to graphite
+s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+s.connect((args.ip, args.port))
+
+pmetrics = {}
+pdate = -1
+while 1:
+    time.sleep(args.interval)
+
+    # get the current time
+    f = os.popen('date +%s')
+    date = f.read().strip('\n')
+
+    # read the perf dump
+    os.popen("ceph --admin-daemon /var/run/ceph/ceph-mds* perf dump > " + args.jsonfile)
+    try:
+      with open(args.jsonfile) as f:
[…]

### 2017-12-15 experiments/baseline-durability/run.sh
results before cluster expires
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -9,27 +9,28 @@ ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
 CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
 SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
 
-for site in "nojournal-cache" "journal120-cache" "journal150-cache" "journal180-cache" "journal30-cache" "journal60-cache" "journal90-cache" "nojournal-cache"; do
-  #nclients=3
+for nclients in 1 5 10 15 18 20; do
+    cp inventory_cloudlab/${nclients}client site/hosts
+    cp site/hosts $ROOT/ansible/ceph/hosts
+  for site in "nojournal-cache" "journal210-cache" "journal120-cache" "journal30-cache"; do
 
-  # configure ceph and setup results directory
-  mkdir -p results/$site/logs || true
-  cp site/* $ROOT/ansible/ceph || true
-  cp site_confs/${site}.yml site/group_vars/all
-  cp -r site/group_vars $ROOT/ansible/ceph/
-  #cp inventory/${nclients}client site/hosts
-  #cp inventory/${nclients}client $ROOT/ansible/ceph//hosts
-  cp site/hosts $ROOT/ansible/ceph/hosts
+    # configure ceph and setup results directory
+    mkdir -p results/$site/logs || true
+    cp site/* $ROOT/ansible/ceph || true
+    cp site_confs/${site}.yml site/group_vars/all
+    cp -r site/group_vars $ROOT/ansible/ceph/
 
-  # cleanup and start ceph
-  $SRL_ANSIBLE cleanup.yml
-  $CEPH_ANSIBLE ceph.yml cephfs.yml
-  $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
-  
-  # warmup and get baseline
-  for i in `seq 0 2`; do
-    ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
+    # cleanup and start ceph
+    $SRL_ANSIBLE cleanup.yml
+    $CEPH_ANSIBLE ceph.yml cephfs.yml
+    $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+    
+    # warmup and get baseline
+    for i in `seq 0 2`; do
[…]

### 2017-12-20 experiments/baseline-creates1/run.sh
benchmark: more scalability results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -36,11 +36,13 @@ fi
 
 mkdir results-all || true
 
-for procs in 1 5 10 15 20 25 30 35; do
+#for procs in 1 5 10 15 20 25 30 35; do
+for procs in 25 30 35; do
   for clients in 1; do
     cp configs_$SITE/clients$clients hosts
     for job in "creates"; do
-      for log in "nolog" "log1" "log10" "log30" "log50"; do
+      #for log in "nolog" "log1" "log10" "log30" "log50"; do
+      for log in "nolog" "log1"; do
         cp configs_$SITE/all-$log ansible/group_vars/all 
         sudo rm -rf results || true; mkdir results
         $DOCKER -e processes_per_client=$procs cleanup.yml

### 2017-12-20 experiments/baseline-creates1/run.sh
benchmark: durability parameter sweeps
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -34,17 +34,21 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-for log in "log"; do
-  cp configs_$SITE/all-$log ansible/group_vars/all 
+mkdir results-all || true
+
+#for procs in 1 5 10 15 20 25 30 35; do
+for procs in 25 30 35; do
   for clients in 1; do
     cp configs_$SITE/clients$clients hosts
     for job in "creates"; do
-      for procs in 1 5 10 15 20 25 30 35; do
+      #for log in "nolog" "log1" "log10" "log30" "log50"; do
+      for log in "nolog" "log1"; do
+        cp configs_$SITE/all-$log ansible/group_vars/all 
         sudo rm -rf results || true; mkdir results
         $DOCKER -e processes_per_client=$procs cleanup.yml
         $DOCKER -e processes_per_client=$procs -e nfiles=98000 \
           ceph.yml monitor.yml /workloads/${job}.yml collect.yml
-        mv results results-$SITE-clients$clients-procs$procs-$log
+        mv results results-all/results-$SITE-clients$clients-procs$procs-$log
       done
     done
   done

### 2017-12-22 experiments/baseline-creates1/run.sh
benchmark: more durability results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -6,6 +6,7 @@ set -ex
 
 # if you know Ansible and Docker, the below should make sense
 # - we attach ceph-ansible to root because they expect us to be in that dir
+OUTP="results-multipleruns"
 SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
 ROOT=`dirname $PWD | xargs dirname`
 NETW="--net host -v $HOME/.ssh:/root/.ssh"
@@ -34,21 +35,38 @@ if [ ! -z $1 ]; then
   exit
 fi
 
-mkdir results-all || true
-
-#for procs in 1 5 10 15 20 25 30 35; do
-for procs in 25 30 35; do
+mkdir $OUTP || true
+for procs in 1 5 10 15; do
   for clients in 1; do
     cp configs_$SITE/clients$clients hosts
     for job in "creates"; do
-      #for log in "nolog" "log1" "log10" "log30" "log50"; do
-      for log in "nolog" "log1"; do
+      for log in "nolog" "log1" "log10" "log30" "log50"; do
         cp configs_$SITE/all-$log ansible/group_vars/all 
-        sudo rm -rf results || true; mkdir results
-        $DOCKER -e processes_per_client=$procs cleanup.yml
-        $DOCKER -e processes_per_client=$procs -e nfiles=98000 \
-          ceph.yml monitor.yml /workloads/${job}.yml collect.yml
-        mv results results-all/results-$SITE-clients$clients-procs$procs-$log
+        for run in 0 1 2; do
+          mkdir results || true
+          $DOCKER -e processes_per_client=$procs cleanup.yml
+          $DOCKER -e processes_per_client=$procs -e nfiles=98000 \
+            ceph.yml monitor.yml /workloads/${job}.yml collect.yml
+          mv results $OUTP/results-$SITE-clients$clients-procs$procs-$log-run$run
+        done
+      done
+    done
+  done
+done
+
+for procs in 20 25 30 35 40 45 50; do
+  for clients in 1; do
+    cp configs_$SITE/clients$clients hosts
+    for job in "creates"; do
+      for log in "nolog" "log1"; do
+      cp configs_$SITE/all-$log ansible/group_vars/all 
+        for run in 0 1 2; do
+          mkdir results || true
+          $DOCKER -e processes_per_client=$procs cleanup.yml
+          $DOCKER -e processes_per_client=$procs -e nfiles=98000 \
[…]

### 2017-12-22 paper/sections/3_posix_overheads.tex
paper: fix graphs and add baselines (#33, #34)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
-
 \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-compile/visualize/viz.ipynb}{source}]
@@
 of CephFS.}\label{fig:overhead-creates}
-
 \end{figure}
@@
       \centering
-      \includegraphics[width=0.95\linewidth]{graphs/slowdown-journal.png}
+      \includegraphics[width=1\linewidth]{graphs/slowdown-journal.png}
       \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-durability/visualize/viz.ipynb}{source}]
-      Effect of Journal} \label{fig:overhead-a}
+      effect of journal} \label{fig:overhead-a}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=0.98\linewidth]{graphs/slowdown-interfere-scale.png}
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-creates/visualize/viz.ipynb}{source}]
+      interference hurts variability}
+      \label{fig:overhead-b}
   \end{subfigure}
@@
       \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
-      Interference increases RPCs}
-      \label{fig:overhead-b}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1\linewidth]{graphs/slowdown-interfere-scale.png}
-      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-creates/visualize/viz.ipynb}{source}]
-      Interference hurts variability}
+      interference increases RPCs}
       \label{fig:overhead-c}
   \end{subfigure}
-  \caption{ The overhead of durability and strong consistency in CephFS.  (a)
-  shows the effect of different journal segment sizes, normalized to a setup
-  without a journal (``No Journal").  The journal is streamed into the object
-  store for fault tolerance. (b) and (c) show that when a second client
-  ``interferes", capabilities are revoked and metadata servers do more work.
-  \label{fig:overhead}}
+
+  \caption{\newcomment{The durability and strong consistency slowdown of the
+existing CephFS implementation increases as the number of clients scales.
+Results are normalized to 1 client that creates 100K files in isolation.  (a)
+shows the effect of journaling metadata updates; ``segment(s)" is the number
+of journal segments dispatched to disk at once. (b) shows the slowdown when
+another client interferes by creating files in all directories and (c)
+highlights the cause: }\oldcomment{The overhead of durability and strong
+consistency in CephFS. (a) shows the effect of different journal segment sizes,
+which are streamed into the object store for fault tolerance. (b) and (c) show
+that} when another client ``interferes", capabilities are revoked and metadata
+servers do more work.} \label{fig:overhead}
+
 \end{figure*}
@@
 resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to be large
-(on the order of MBs) which ensures (1) sequential writes into the object store
-and (2) the ability for daemons to trim redundant or irrelevant journal
-entries.  The journal is striped over objects where multiple journal updates
-can reside on the same object. There are two tunables for controlling the
-journal: the segment size and the number of segments that can be written in
-parallel. Unless the journal saturates memory or CPU resources, larger values
+WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to be large (on
+the order of MBs) which ensures (1) sequential writes into the object store and
+(2) the ability for daemons to trim redundant or irrelevant journal entries.
+The journal is striped over objects where multiple journal updates can reside
+on the same object. There are two tunables\newcomment{, related to groups of
+journal events called segments,} for controlling the journal: the segment size
+and the dispatch size ({\it i.e.} the number of segments that can be dispatched
+at once).  Unless the journal saturates memory or CPU resources, larger values
 for these tunables results in better performance.
@@
 % Effects on performance
-Figure~\ref{fig:overhead-a} shows the effect of journaling of with different
-segment sizes; the larger the segment size the bigger that the writes into the
-object store are. The trade-off for better performance is memory consumption
-because larger segments take up more space for buffering. When journaling is
-on, the metadata server periodically stops serving requests to flush ({\it
-i.e.} apply journal updates) to the metadata store.  The journal overhead is
-sufficient enough to slow down metadata throughput but not so much as to
-overwhelm the bandwidth of the object store.
+Figure~\ref{fig:overhead-a} shows the effect of journaling with different
+\oldcomment{segment sizes; the larger the segment size the bigger that the
+writes into the object store are}\newcomment{dispatch sizes, normalized to 1
+client that creates 100K files with journaling off (about 654 creates/sec). In
+this case a dispatch size of 10 degrades performance more because the metadata
+server is overloaded with requests and cannot spare cycles to manage concurrent
+segments.  Although the ``no journal" curve appears flat, it is actually a
+slowdown of about \(0.3\times\) per concurrent client; this slowdown is a
+result of the peak throughput of a single metadata server, which we found to be
+about 3000 operations per second}.  The trade-off for better performance is
+memory consumption because a larger \oldcomment{segment}\newcomment{dispatch}
+size uses more space for buffering.  When journaling is on, the metadata server
+periodically stops serving requests to flush ({\it i.e.} apply journal updates)
+to the metadata store.  The journal overhead is sufficient enough to slow down
+metadata throughput but not so much as to overwhelm the bandwidth of the object
+store.
 
@@
 % benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
-The benefits of caching the directory inode when creating files is shown in
-Figure~\ref{fig:overhead-b}. The colors show the behavior of the client for two
+\oldcomment{This degradation is shown in Figure~\ref{fig:overhead-b}, where we
+scale the number of clients and show the slowdown of the slowest client.  The
+results are normalized to a single isolated client without a metadata server
+journal and the error bars are the standard deviations of all client runtimes.}
+\newcomment{Figure~\ref{fig:overhead-b} shows the slowdown of maintaining
+strong consistency when scaling the number of clients.  We plot the slowdown of
+the slowest client, normalized to 1 client that creates 100K files (about 513
+creates/sec because the journal is turned back on).} For the ``interference"
+curve, each client creates files in private directories and at 30 seconds we
+launch another process that creates files in those directories. \oldcomment{18
+is}\newcomment{20 clients has an asterisk\* because} the maximum number of
+clients the metadata server can handle for this metadata-intensive workload
+\newcomment{is actually 18}; at higher client load, the metadata server
+complains about laggy and unresponsive requests.
+
+\oldcomment{The benefits of caching the directory inode when creating
+files}\newcomment{The cause for this slowdown} is shown in
+Figure~\ref{fig:overhead-c}. The colors show the behavior of the client for two
 different runs.  If only one client is creating files in a directory
@@
 \(y1\) axis because the metadata server can handle the extra load but
-performance suffers.  This degradation is shown in Figure~\ref{fig:overhead-c},
-where we scale the number of clients and show the slowdown of the slowest
-client.  The results are normalized to a single isolated client without a
-metadata server journal and the error bars are the standard deviations of all
-client runtimes.  For the ``interfere" bars, each client creates files in
-private directories and at 30 seconds we launch another process that creates
-files in those directories. 18 is the maximum number of clients the metadata
-server can handle for this metadata-intensive workload; at higher client load,
-the metadata server complains about laggy and unresponsive requests.
+performance suffers.
 

### 2017-12-22 paper/sections/6_results.tex
paper: fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
       \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
-      \caption{\texttt{create}: done in parallel on clients}
-      \label{fig:mergescale}
+      \caption{
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
+      \newcomment{parallel creates on clients}
+      }\label{fig:mergescale}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
-      \caption{Subtree semantics: block interference}
-      \label{fig:block-allow}
+      \caption{
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
+      \newcomment{block interference}
+      }\label{fig:block-allow}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
-      \caption{Syncing to global namespace}
-      \label{fig:slowdown-sync}
+      \caption{
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
+      \newcomment{syncing to global namespace}
+      }\label{fig:slowdown-sync}
   \end{subfigure}
-\caption{ The performance and features of Cudele. (a)
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
-shows the cost of merging client journals at the metadata server. Shipping and
-merging journals of updates scales better than RPCs because there are less
-messages and consistency/durability code paths are bypassed. (b)
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
-the allow/block API isolates directories from interfering clients. (c)
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
-is the slowdown of a single client syncing updates to the global namespace. The
+\caption{\oldcomment{ The performance and features of Cudele}\newcomment{Cudele
+performance}. (a) \newcomment{shows the speedup of decoupled namespaces over
+RPCs; \texttt{create} is the throughput of clients creating files in-parallel
+and writing updates locally; \texttt{create+merge} includes the time to merge
+updates at the metadata server. }\oldcomment{shows the cost of merging client
+journals at the metadata server. Shipping and merging journals of
+updates}\newcomment{Decoupled namespaces} scale better than RPCs because there
+are less messages and consistency/durability code paths are bypassed. (b) shows
+how the allow/block API isolates directories from interfering clients. (c) is
+the slowdown of a single client syncing updates to the global namespace. The
 inflection point is the trade-off of frequent updates vs. larger journal
-files.} \label{fig:use-cases}
+files.}
+\label{fig:use-cases}
 \end{figure*}

### 2017-12-22 experiments/cudele-mergescale/run.sh
benchmark: rerun merge scale with 1-20 clients
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -35,14 +35,14 @@ if [ ! -z $1 ]; then
 fi
 
 mkdir results || true
-#./teardown.sh
+./teardown.sh
 #$DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/generate_journals.yml
-for run in 3 4 5; do
-  for nclients in 40 30 20 10 1; do
+for run in 0 1 2; do
+  for nclients in 1 5 10 15 20; do
     ./teardown.sh
     cp configs_$SITE/hosts hosts
     $DOCKER -e nfiles=100000 -e nclients=$nclients ceph.yml /workloads/merge.yml #/workloads/clients${nclients}.yml
   done
-  mv results results-run$run
+  mv results results-rerun-run$run-blah
 done
 exit 0

### 2017-12-22 paper/sections/3_posix_overheads.tex

@@
 client that creates 100K files with journaling off (about 654 creates/sec). In
-this case a dispatch size of 10 degrades performance more because the metadata
-server is overloaded with requests and cannot spare cycles to manage concurrent
-segments.  Although the ``no journal" curve appears flat, it is actually a
-slowdown of about \(0.3\times\) per concurrent client; this slowdown is a
-result of the peak throughput of a single metadata server, which we found to be
-about 3000 operations per second}.  The trade-off for better performance is
-memory consumption because a larger \oldcomment{segment}\newcomment{dispatch}
-size uses more space for buffering.  When journaling is on, the metadata server
-periodically stops serving requests to flush ({\it i.e.} apply journal updates)
-to the metadata store.  The journal overhead is sufficient enough to slow down
-metadata throughput but not so much as to overwhelm the bandwidth of the object
-store.
+this case a dispatch size of 10 degrades performance the most because the
+metadata server is overloaded with requests and cannot spare cycles to manage
+concurrent segments.  Tuning and parameter sweeps show that a dispatch size of
+10 is the worst and that larger sizes approach a dispatch size of 1; for all
+future journal experiments we use a dispatch size of 40 which is a more
+realistic configuration.  Although the ``no journal" curve appears flat, it is
+actually a slowdown of about \(0.3\times\) per concurrent client; this slowdown
+is a result of the peak throughput of a single metadata server, which we found
+to be about 3000 operations per second}.  The trade-off for better performance
+is memory consumption because a larger
+\oldcomment{segment}\newcomment{dispatch} size uses more space for buffering.
+When journaling is on, the metadata server periodically stops serving requests
+to flush ({\it i.e.} apply journal updates) to the metadata store.  The journal
+overhead is sufficient enough to slow down metadata throughput but not so much
+as to overwhelm the bandwidth of the object store.
 

### 2017-12-22 paper/sections/6_results.tex

@@
 {\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
-mechanisms that rely on poorly scaling data structures have large slowdowns for
-the larger number of creates. For example, RPCs has a \(90\times\) slowdown
+mechanisms that rely on poorly scaling data structures \oldcomment{have large
+slowdowns}\newcomment{can slowdown performance by {\it orders of magnitude}}
+for the larger number of creates. For example, RPCs has a \(90\times\) slowdown
 because it relies on the internal CephFS directory data structures. It is a
@@
 
-\textbf{Takeaway}: measuring the mechanisms individually shows that their
+\oldcomment{\textbf{Takeaway}: measuring the mechanisms individually shows that their
 overheads and costs can differ {\it by orders of magnitude}. We also show that
 some mechanisms, like Nonvolatile Apply, are not worthwhile as currently
-implemented.
+implemented.}
 
-\subsection{Use Case 1: Creates in the Same Directory}
-\label{sec:use-case-1}
-
-\begin{figure*}[t]
+\begin{figure*}[tb]
   \centering
@@
 
+\subsection{Use Case 1: Creates in the Same Directory}
+\label{sec:use-case-1}
+
 First we show how we can build application-specific subtrees by composing
-mechanisms and that this approach results in more a scalable global namespace.
-We start with clients creating files in private directories because this
-workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
-ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
-sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
-But the workload also appears in cloud workloads~\cite{xiao:socc15-shardfs},
-where systems like Hadoop use the file abstraction to exchange work units to
-workers or to indicate when jobs
+mechanisms and that \oldcomment{this approach}\newcomment{coupling
+well-established techniques to specific applications} results in a more a
+scalable global namespace.  We start with clients creating files in private
+directories because this workload is heavily studied in
+HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
+zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
+checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
+cloud workloads~\cite{xiao:socc15-shardfs}, where systems like Hadoop use the
+file abstraction to exchange work units to workers or to indicate when jobs
 complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
@@
 
-\textbf{Takeaway}: we confirm the performance benefits of other
+\oldcomment{\textbf{Takeaway}: we confirm the performance benefits of other
 well-established research systems but, more importantly, we show that the
 Cudele mechanisms we propose are useful for building and evaluating different
-consistency/durability semantics.\\
+consistency/durability semantics.\\}
 
@@
 decoupled namespace strategy. For RPCs, we increase the number of clients
-making requests to the metadata server and for decoupled (the merge bars), we
-increase the number of concurrent journal merge requests. So for the 30 clients
-data point, we are measuring the operations per second for 30 client journals
-that land on the metadata server at the same time. The ``create \& merge" bar
-adds the time it takes for clients to generate the journal of events in
-parallel. The experiment is run three times but the standard deviation is too
-small to see.
+making requests to the metadata server and for decoupled we increase the number
+of concurrent clients operating in parallel. So for the 30 clients data point,
+we are measuring the operations per second for 30 client journals that land on
+the metadata server at the same time. The ``create + merge" bar adds the time
+it takes for clients to generate the journal of events in parallel. The
+experiment is run three times but the standard deviation is too small to see.
 
@@
 
-\textbf{Takeaway}: the ability to couple well-established techniques to
+\oldcomment{\textbf{Takeaway}: the ability to couple well-established techniques to
 specific applications allows the global namespace to scale further and perform
 better. In our case, RPCs overwhelm the system but decoupling/shipping a
-journal of updates improves scalability by 2\(\times\).
+journal of updates improves scalability by 2\(\times\).}
 
@@
 
-Next we show how Cudele can be programmed to block interfering clients using
-the same problematic workload from Figure~\ref{fig:overhead-b}. In this
-workload, clients create files in private directories and a separate client
-interferes by creating more files in each directory. This introduces false
-sharing and the metadata server revokes capabilities on directories touched by
-the interfering client. While HPC tries to avoid these situations with
-workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
-happens in distributed file systems when users unintentionally access
-directories in a shared file system.  We only scale to 15 clients because the
-performance variability of a nearly overloaded metadata server, in our case 18
-clients, is too high.
+Next we show how Cudele can be programmed to block interfering
+clients\oldcomment{ using the same problematic workload from
+Figure~\ref{fig:overhead-b}. In this workload, c}\newcomment{, which lets
+applications control isolation to get better and more reliable performance.
+C}lients create files in private directories while another client interferes by
+creating \oldcomment{more}\newcomment{1000} files in each directory. This
+introduces false sharing and the metadata server revokes capabilities on
+directories touched by the interfering client. While HPC tries to avoid these
+situations with workflows~\cite{zheng:pdsw2014-batchfs,
+zheng:pdsw2015-deltafs}, it still happens in distributed file systems when
+users unintentionally access directories in a shared file system.
+\oldcomment{We only scale to 15 clients because the performance variability of
+a nearly overloaded metadata server, in our case 18 clients, is too high.}
 
-Figure~\ref{fig:block-allow} plots the overhead of the slowest client and the
-error bars are the standard deviation of 3 runs.  ``Interfere (Fig 3c)" is the
-result from Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API
-to return \texttt{-EBUSY} to interfering clients; and ``Isolated" is the
-baseline performance without an interfering. Each subtree has RPCs and
-Stream enabled.  Results are normalized to the runtime of a single
-client creating files in a directory. Note that IndexFS has a similar behavior
-to ``interfere" with leases, except clients block.
+\newcomment{We switch back to clients doing RPCs per operation with journaling
+enabled, to mirror the setup from the problem presented in
+Figure~\ref{fig:overhead-b}.} Figure~\ref{fig:block-allow} plots the overhead
+of the slowest client \oldcomment{and the error bars are the standard deviation
+of 3 runs.  ``Interfere (Fig 3c)" is the result from
+Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API to return
+\texttt{-EBUSY} to interfering clients; and ``Isolated" is the baseline
+performance without an interfering.  Each subtree has RPCs and Stream enabled.
+Results are normalized to the runtime of a single client creating files in a
+directory.  Note that IndexFS has a similar behavior to ``interfere" with
+leases, except clients block.  }\newcomment{normalized to 1 client that creates
+100K files in isolation (about 513 creates/sec). ``Interference" and ``no
+interference" is the performance with and without an interfering client
+touching files in every directory, respectively. The goal is to explicitly
+isolate clients so that performance is similar to the ``no interfere" curve,
+which has lower slowdowns (on average, \(1.42\times\) per client compared to
+\(1.67\times\) per client for ``interfere") and less variability (on average, a
+standard deviation of 0.06 compared to 0.44 for ``interfere").  ``Block
+interference" uses the Cudele API to block interfering clients with
+\texttt{-EBUSY}. That curve shows slowdowns similar to ``interfere" for smaller
+clusters because the overhead to reject requests is more evident when the
+metadata server is underloaded; but the slowdown (\(1.34\times\) per client)
+and variability (0.09) look very similar to ``no interfere" for larger
+clusters. Clients can block interfering clients to get the same performance as
+isolated clients but there is a non-negligible overhead for rejecting requests
+when the metadata server is not operating at peak efficiency.}
 
-We draw three conclusions: (1) clients that use the API to block interfering
+\oldcomment{We draw three conclusions: (1) clients that use the API to block interfering
 clients  get the same performance as isolated clients, (2) there is a
@@
 need better and more reliable performance. Blocking updates is an effective way of
-controlling consistency.
+controlling consistency.}
 
@@
 The final use case shows how the API gives users fine-grained control of the
-consistency semantics.  The use case is that users often leverage the file
-system to check the progress of jobs using \texttt{ls} even though this
-operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
+consistency semantics\newcomment{ to support current practices and experimental
+workflows}. \oldcomment{The use case is that u}\newcomment{U}sers often
+leverage the file system to check the progress of jobs using \texttt{ls} even
+though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
 eshel:fast10-panache} . The number of files or size of the files is indicative
@@
 node. In this scenario, Cudele users will not see the progress of decoupled
-namespaces since their updates are not globally visible.  To help users
-judge the progress of their jobs, Cudele clients have a ``namespace sync" that
-sends batches of updates back to the global namespace at regular intervals.
+namespaces since their updates are not globally visible.  To help users judge
+the progress of their jobs, Cudele clients have a ``namespace sync" that sends
+batches of updates back to the global namespace at regular intervals.
 
@@
 Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
-client writing updates to a decoupled namespace and pausing to send updates to
-the metadata server. The error bars are the standard deviations of 3 runs. We
-scale the namespace sync interval to show the trade-off of frequently pausing
-or writing large logs of updates.  We use an idle core to log the updates and
-to do the network transfer. The client only pauses to fork off a background
-process, which is expensive as the address space needs to copied. The
-alternative is to pause the client completely and write the update to disk but
-since this implementation is limited by the speed of the disk, we choose the
-memory-to-memory copy of the fork approach.
+client writing \newcomment{1 million} updates to a decoupled namespace and
+pausing to send updates to the metadata server. The error bars are the standard
+deviations of 3 runs. We scale the namespace sync interval to show the
+trade-off of frequently pausing or writing large logs of updates.  We use an
+idle core to log the updates and to do the network transfer. The client only
+pauses to fork off a background process, which is expensive as the address
+space needs to copied. The alternative is to pause the client completely and
+write the update to disk but since this implementation is limited by the speed
+of the disk, we choose the memory-to-memory copy of the fork approach.
 
-\textbf{Takeaway}: syncing namespace updates has up to a 9\% overhead and can
-be tuned depending on the users preference but, more importantly, the API gives
-users fine-grain control of their consistency/durability to support current
-practices or experimental workflows.
+As expected, syncing namespace updates too frequently has the highest overhead
+(up to 9\% overhead if done every second). The optimal sync interval for
+performance is 10 seconds because larger intervals must write more updates to
+disk and network. For the 25 second interval, the client only pauses 3-4 times
+but each sync writes about 278 thousand updates at once, which is a 678MB
+journal.
+
+\oldcomment{\textbf{Takeaway}: Syncing namespace updates has up to a 9\%
+overhead and can be tuned depending on the users preference but, more
+importantly, the API gives users fine-grain control of their
+consistency/durability to support current practices or experimental workflows.}
 

### 2017-12-23 paper/sections/0_abstract.tex
paper: finish evaluation (#44, #38, #34, #33, #32, #27)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-23 paper/sections/0_abstract.tex
@@
 storage system.  By custom fitting a subtree to a create-heavy application, we
-show 8\(\times\) speedup and 2\(\times\) scale when compared to our baseline
-system.
+show similar speedups to related work \newcomment{(up to \(91.7\times\)
+speedup) but more importantly, our prototype can custom fit subtrees in the
+same namespace to applications common in large data centers, such as
+checkpoint-restart (\(91.7\times\), user home directories (within a 0.03
+standard deviation from optimal), and clients checking for partial results
+(only a 2\% overhead).}  \oldcomment{ and can scale to 2\(\times\) as many
+clients when compared to our baseline system.}
 

### 2017-12-23 paper/sections/1_introduction.tex
paper: finish evaluation (#44, #38, #34, #33, #32, #27)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 The results in this paper pave the way for such a system and confirm the
-assertions of ``clean-state" research that decouple namespaces; specifically
-that these techniques drastically improve performance (we see 104\(\times\) speed up).
-We go a step further by quantifying the costs of merging updates (7\(\times\)
-slow down) and maintaining durability (\(10\times\) slow down). In our
-prototype, we get an 8\(\times\) speedup and can scale to twice as many clients
-when we assign a more relaxed form of consistency and durability to a subtree
-with a create-heavy workload.  We use Ceph as a prototyping platform because it
-is used in cloud-based and data center systems and has a presence in
+assertions of ``clean-state" research of decoupled namespaces; specifically
+that these techniques drastically improve performance.  We go a step further by
+quantifying the costs of merging updates (3.37\(\times\) slow down) and
+maintaining durability (\(2.4\times\) slow down).  In our prototype,
+\newcomment{we also show how to assign subtree semantics for certain
+applications such as checkpoint-restart (\(91.7\times\) speedup if consistency
+is fully relaxed), user home directories (within a 0.03 standard deviation from
+optimal), and clients checking for partial results (only a 2\% overhead).}
+\oldcomment{we get an 8\(\times\) speedup and can scale to twice as many
+clients when we assign a more relaxed form of consistency and durability to a
+subtree with a create-heavy workload.}  We use Ceph as a prototyping platform
+because it is used in cloud-based and data center systems and has a presence in
 HPC~\cite{wang:pdsw13-ceph-hpc}.

### 2017-12-23 paper/sections/6_results.tex
paper: finish evaluation (#44, #38, #34, #33, #32, #27)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 can help the system scale further when subtrees are coupled to the correct type
-of application.  We scope the evaluation to one metadata server and scale the
-number of parallel clients. This type of analysis shows the capacity and
-performance of a metadata server with superior metadata protocols, which should
-be used to inform load balancing across a metadata cluster. Load balancing
-across a cluster of metadata servers with partitioning and replication can be
-explored with something like Mantle~\cite{sevilla:sc15-mantle}.
+of application.  \newcomment{These techniques have already proven to be
+effective and scalable in systems that specialize an entire namespace according
+to a single optimization strategy (e.g., Lustre, BatchFS, DeltaFS), so our
+evaluation explores the behavior of the consistency/durability mechanisms and
+their effect with/without isolation.}
+
+\newcomment{\indent We run experiments on the same cluster and graph the
+standard deviations for three runs (sometimes error bars are too small to see).
+We also normalize results to control for hardware differences and to make our
+results more generally applicable. We use a CloudLab cluster of 34 nodes
+connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.
+Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel release, version
+10.2.1, which was released in May 2016. All daemons run as a single process
+(the default setting for Ceph) and the Ceph
+components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
+object storage daemons, 1 metadata server, and 29 clients.} We scope the
+evaluation to one metadata server and scale the number of parallel
+clients\newcomment{ each doing 100K operations}\footnote{This is the maximum
+recommended size of a directory in CephFS; preliminary experiments with larger
+directory sizes show memory problems.}. This type of analysis shows the
+capacity and performance of a metadata server with superior metadata protocols,
+which should be used to inform load balancing across a metadata cluster. Load
+balancing across a cluster of metadata servers with partitioning and
+replication can be explored with something like
+Mantle~\cite{sevilla:sc15-mantle}.
+
+\newcomment{\indent To make our results reproducible, this paper adheres to The
+Popper Convention~\cite{jimenez_popper_2016}, so experiments presented here are
+available in the repository for this article.  Experiments can be examined in
+more detail, or even re-run, by visiting the \texttt{[source]} link next to
+each figure. That link points to a Jupyter notebook that shows the analysis and
+source code for that graph, which points to an experiment and its artifacts.
+The source code is available on the Cudele
+branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
+of our Ceph fork.} \addressesissue{1}
+
+%\newcomment{\indent We understand concerns about cluster size and network, especially
+%for systems aimed at HPC but our evaluation focuses on verifying the behavior
+%of different techniques in the same global namespace.
 
@@
 
-\newcomment{We run experiments on the same cluster and normalize results to
-control for hardware differences and to make our results more generally
-applicable. We use a CloudLab cluster of 34 nodes connected with 10Gbit
-ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.  Each node uses
-Ubuntu 14.04 and we develop on Ceph's Jewel release, version 10.2.1, which was
-released in May 2016. All daemons run as a single process (the default setting
-for Ceph) and the Ceph
-components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
-object storage daemons, 1 metadata server, and 29 clients.}
-
-\newcomment{We understand concerns about cluster size and network, especially
-for systems aimed at HPC but our evaluation focuses on verifying the behavior of
-different techniques in the same global namespace. These techniques have
-already proven to be effective and scalable in systems that specialize an
-entire namespace according to a single optimization strategy (e.g., Lustre,
-BatchFS, DeltaFS), so our evaluation explores the behavior of the
-consistency/durability mechanisms and their effect with/without isolation.}
-
-\newcomment{To make our results reproducible, this paper adheres to The Popper
-Convention~\cite{jimenez_popper_2016}, so experiments presented here are
-available in the repository for this article.  Experiments can be examined in
-more detail, or even re-run, by visiting the \texttt{[source]} link next to
-each figure. That link points to a Jupyter notebook that shows the analysis and
-source code for that graph, which points to an experiment and its artifacts.
-The source code is available on the Cudele
-branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
-of our Ceph fork.} \addressesissue{1}
-
 \subsection{Microbenchmarks}
@@
 [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
-\newcomment{Overhead of processing 100K events for each mechanism in
+\newcomment{Overhead of processing 100K create events for each mechanism in
 Table~\ref{table:mechanisms}. Results are normalized to the runtime of writing
@@
 
-Figure~\ref{fig:composable-mechanisms} shows the runtime of the Cudele
-mechanisms for a single client creating files in the same directory, normalized
-to the time it takes to write 100K file create updates to the client's
-in-memory journal ({\it i.e.} the Append Client Journal mechanism).
-Stream is an approximation of the overhead and is calculated by subtracting
-the runtime of the job with the journal turned off from the runtime with the
-journal turned on.  100K is the maximum recommended size of a directory in
-CephFS; preliminary experiments with larger directory sizes show memory
-problems.
+\oldcomment{Figure~\ref{fig:composable-mechanisms} shows the runtime of the
+Cudele mechanisms for a single client creating files in the same directory, }
+\newcomment{We measure the overhead of each mechanism by having 1 client create
+100K files in a directory for various subtree configurations.
+Figure~\ref{fig:composable-mechanisms} shows the time that it takes each Cudele
+mechanism to process all metadata events.  Results are} normalized to the time
+it takes to write 100K file create updates to the client's in-memory journal
+({\it i.e.} the Append Client Journal mechanism).  \newcomment{The first graph
+groups the consistency mechanisms, the second groups the durability mechanisms,
+and the third has compositions representing real-world systems.} Stream is an
+approximation of the overhead and is calculated by subtracting the runtime of
+the job with the journal turned off from the runtime with the journal turned
+on.  \oldcomment{100K is the maximum recommended size of a directory in CephFS;
+preliminary experiments with larger directory sizes show memory problems.}
 
 {\it Poorly Scaling Data Structures:} Despite doing the same amount of work,
-mechanisms that rely on poorly scaling data structures \oldcomment{have large
-slowdowns}\newcomment{can slowdown performance by {\it orders of magnitude}}
-for the larger number of creates. For example, RPCs has a \(90\times\) slowdown
-because it relies on the internal CephFS directory data structures. It is a
-well-known problem that directory data structures do not scale when creating
-files in the same directory~\cite{ren:sc2014-indexfs} and any mechanism that
-uses these data structures will experience similar slowdowns. Other mechanisms
-that write events to a journal) experience a much less drastic slowdown because
-the journal data structure does not need to be scanned for every operation.
-Events are written to the end of the journal without even checking the validity
-({\it e.g.}, if the file already exists for a create), which is another form of
+mechanisms that rely on poorly scaling data structures have large
+slowdowns\oldcomment{ for the larger number of creates}. For example, RPCs has
+a \oldcomment{\(90\times\)}\newcomment{\(18\times\)} slowdown because this
+technique relies on the internal directory data structures. It is a well-known
+problem that directory data structures do not scale when creating files in the
+same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
+structures will experience similar slowdowns. Other mechanisms that write
+events to a journal experience a much less drastic slowdown because the journal
+data structure does not need to be scanned for every operation.  Events are
+written to the end of the journal without even checking the validity ({\it
+e.g.}, if the file already exists for a create), which is another form of
 relaxed consistency because the file system assumes the application has
@@
 % RPCs vs. apply: calls to metadata server vs. RADOS
-{\it Overhead of RPCs:} RPCs is \(66\times\) slower than Volatile
-Apply because sending individual metadata updates over the network is costly.
-While RPCs sends a request for every file create, Volatile Apply
-writes all the updates to the in-memory journal and applies them to the
-in-memory data structures in the metadata server. While communicating the
-decoupled namespace directly to the metadata server is faster, communicating
-through the object store (Nonvolatile Apply) is \(10\times\) slower.
+{\it Overhead of RPCs:} RPCs is
+\oldcomment{\(66\times\)}\newcomment{\(19.9\times\)} slower than Volatile Apply
+because sending individual metadata updates over the network is costly.  While
+RPCs sends a request for every file create, Volatile Apply \oldcomment{writes
+all the updates to the in-memory journal and applies them}\newcomment{writes
+directly} to the in-memory data structures in the metadata server. While
+communicating the decoupled namespace directly to the metadata server
+\newcomment{with Volatile Apply} is faster, communicating through the object
+store \newcomment{with Nonvolatile Apply} is
+\oldcomment{\(10\times\)}\newcomment{\(78\times\)} slower.
 
-% TODO: why is apply so slow.  
-% apply: no CephFS changes, pulls/pushes same RADOS obj.
-% v_apply vs. apply/persist: communicating through RADOS
-{\it Overhead of Nonvolatile Apply:} The cost of Nonvolatile
-Apply is much larger than all the other mechanisms.  That mechanism was not
-implemented as part of Cudele -- it was a debugging and recovery tool packaged
-with CephFS. It works by iterating over the updates in the journal and pulling
-all objects that {\it may} be affected by the update.  This means that two
-objects are repeatedly pulled, updated, and pushed: the object that houses the
-experiment directory and the object that contains the root directory ({\it
-i.e.} \texttt{/}).  The cost of communicating through the object store is shown
-by comparing the runtime of Volatile Apply + Global Persist to
-Nonvolatile Apply. These two operations end up with the same final metadata
-state but using Nonvolatile Apply is clearly inferior.
+% TODO: why is apply so slow.  apply: no CephFS changes, pulls/pushes same
+% RADOS obj.  v_apply vs. apply/persist: communicating through RADOS
+{\it Overhead of Nonvolatile Apply:} The cost of Nonvolatile Apply is much
+larger than all the other mechanisms.  That mechanism was not implemented as
+part of Cudele -- it was a debugging and recovery tool packaged with CephFS. It
+works by iterating over the updates in the journal and pulling all objects that
+{\it may} be affected by the update.  This means that two objects are
+repeatedly pulled, updated, and pushed: the object that houses the experiment
+directory and the object that contains the root directory ({\it i.e.},
+\texttt{/}).  The cost of communicating through the object store is shown by
+comparing the runtime of Volatile Apply + Global Persist
+\newcomment{(\(1.3\times\))} to Nonvolatile Apply\newcomment{ (\(78\times\))}.
+These two operations end up with the same final metadata state but using
+Nonvolatile Apply is clearly inferior.
 
 % persist vs. save: one disk vs. many
-{\it Parallelism of the Object Store:} Comparing Local and Global
+{\it Parallelism of the Object Store:} \newcomment{Stream has the highest
+slowdown at \(2.4\times\) because the overhead of maintaining and streaming the
+journal is incurred by the metadata server}.  Comparing Local and Global
 Persist demonstrates the bandwidth advantages of storing the journal in a
-distributed object store. The Global Persist
-performance is \(1.5\times\) faster because the object store is leveraging the
-collective bandwidth of the disks in the cluster. This benefit comes from the
-object store itself but should be acknowledged when making decisions for the
-application; the bandwidth of the object store can help mitigate the overheads of
-globally persisting metadata updates.
-
-{\it Journal Size:} We measure the amount of storage per journal update to be
-about 2.5KB. The storage footprint scales linearly with the number of
-metadata creates and suggests that updates for a million files would be 2.5KB
-\(*\) 1 million files \(=\) 2.38GB. Transfer times for payloads of this size in
-most HPC/data center networks are reasonable.
+distributed object store. The Global Persist performance is \newcomment{only
+\(0.2\times\) slower than Local Persist because Global
+Persist}\oldcomment{\(1.5\times\) faster because the object store} is
+leveraging the collective bandwidth of the disks in the cluster. This benefit
+comes from the object store itself but should be acknowledged when making
+decisions for the application; the bandwidth of the object store can help
+mitigate the overheads of globally persisting metadata updates. The storage per
+journal update is about 2.5KB. So the storage footprint scales linearly with
+the number of metadata creates and suggests that updates for a million files
+would be \newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1 million files \(=\)
+2.38GB. Transfer times for payloads of this size in most HPC/data center
+networks are reasonable.}
 
@@
 
+\newcomment{{\it Composing Mechanisms}: The graph on the right of
+Figure~\ref{fig:composable-mechanisms} shows how applications can compose
+mechanisms together to get the consistency/durability guarantees they need in a
+global namespace.  We label the \(x\)-axis with systems that employ these
+semantics, as described in Figure~\ref{fig:subtree-policies}.  Again, the
+runtime is normalized to creating files in the client's in-memory journal.  We
+make no guarantees during execution of the mechanisms or when transitioning
+semantics -- the semantics are guaranteed {\it once the mechanism completes}.
+So if servers fail during a mechanism, metadata or data may be lost. This graph
+shows how we can build application-specific subtrees by composing mechanisms
+and that coupling well-established techniques to specific applications results
+in a more scalable global namespace.}
+
 \subsection{Use Case 1: Creates in the Same Directory}
@@
 
-First we show how we can build application-specific subtrees by composing
+\oldcomment{Next we show how we can build application-specific subtrees by composing
 mechanisms and that \oldcomment{this approach}\newcomment{coupling
 well-established techniques to specific applications} results in a more a
-scalable global namespace.  We start with clients creating files in private
-directories because this workload is heavily studied in
-HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
-zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
-checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
-cloud workloads~\cite{xiao:socc15-shardfs}, where systems like Hadoop use the
-file abstraction to exchange work units to workers or to indicate when jobs
+scalable global namespace.  }
+
+We start with clients creating files in private directories because this
+workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
+ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
+sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
+But the workload also appears in cloud workloads~\cite{xiao:socc15-shardfs},
+where systems like Hadoop use the file abstraction to exchange work units to
+workers or to indicate when jobs
 complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
@@
 services a flash crowd of creates across all directories as shown in
-Figure~\ref{fig:overhead-creates}.  The workload is clients creating 100K files
-in private directories in the same global namespace.
+Figure~\ref{fig:overhead-creates}. \oldcomment{ The workload is clients creating 100K files
+in private directories in the same global namespace. }
 
@@
 
-The graph on the right of Figure~\ref{fig:composable-mechanisms} shows how
+\oldcomment{The graph on the right of Figure~\ref{fig:composable-mechanisms} shows how
 applications can compose mechanisms together to get the consistency/durability
@@
 semantics are guaranteed {\it once the mechanism completes}. So if servers fail
-during a mechanism, metadata or data may be lost.
+during a mechanism, metadata or data may be lost.}
 
@@
 
-To show the contention at the metadata server, in Figure~\ref{fig:mergescale}
-we scale the number of clients for the RPC per operation strategy and the
-decoupled namespace strategy. For RPCs, we increase the number of clients
-making requests to the metadata server and for decoupled we increase the number
-of concurrent clients operating in parallel. So for the 30 clients data point,
-we are measuring the operations per second for 30 client journals that land on
-the metadata server at the same time. The ``create + merge" bar adds the time
-it takes for clients to generate the journal of events in parallel. The
-experiment is run three times but the standard deviation is too small to see.
+\oldcomment{To show the contention at the metadata server, }In
+Figure~\ref{fig:mergescale} we scale the number of clients \newcomment{each
+doing 100K file creates in their own directories for three types of subtrees:
+one with strong consistency and global durability (RPCs), one with invisible
+consistency and local durability (decoupled: create), and one with weak
+consistency and local durability (decoupled: create + merge). Results are
+normalized to 1 client that creates 100K files using RPCs (about 549
+creates/sec). As opposed to earlier graphs in
+Section~\S\ref{sec:posix-overheads} that plotted the throughput of the slowest
+client, Figure~\ref{fig:mergescale} plots the throughput from the perspective
+of the metadata server. Plotting this way is easier to understand because of
+how we normalize but the speedups over the RPC approach are the same, whether
+we look at the slowest client or not.}
+
+\newcomment{When the metadata server is operating at peak efficiency between 5
+and 20 clients, ``decoupled: merge + create" outperforms ``RPCs" by
+\(3.37\times\) because ``decoupled: merge + create" uses a relaxed form of
+consistency and leverages bulk updates just liked
+DeltaFS~\cite{zheng:pdsw2015-deltafs}.  The ``RPCs" and ``decoupled: merge +
+create" curves flatten out at a slowdown of \(4.5\times\) and \(15\times\),
+respectively, because more clients increases the load at the metadata server.
+Alternatively, the ``decoupled: create" curve scales linearly with the number
+of concurrent clients because clients operate in parallel and write updates
+locally. At 20 clients, we observe a \(91.7\times\) speedup for ``decoupled:
+create" over RPCs.}\oldcomment{ for the RPC per operation strategy and the
+decoupled namespace strategy.  For RPCs, scaling the number of clients
+increases the load on the metadata server and for decoupled we increase the
+number of concurrent clients operating in parallel.  Related work has pointed
+to on-disk metadata format as the primary factor for improved scalability; but
+our results show that the weakened consistency and durability enables the
+scalability improvements. The on-disk metadata formats in this experiment are
+the same for all schemes.}
+
+\newcomment{Another reason that} the performance of decoupled namespaces is
+better than RPCs \newcomment{is because decoupled namespaces (1)} place no
+restrictions on the validity of metadata inserted into the journal and
+\newcomment{(2)} avoid touching poorly scaling data structures. We can scale
+linearly if we weaken our consistency by never checking for the existence of
+files before creating files ({\it i.e.} only append \texttt{open()} requests to
+the journal).  When the updates are merged by the metadata server into the
+in-memory metadata store they never scan growing data structures.  Had we
+implemented the \oldcomment{Append Client Journal mechanism}\newcomment{client}
+to include \texttt{lookup()} commands before \texttt{open()} requests, we would
+have seen the poor scaling \newcomment{behavior} that we see with the RPCs
+mechanism.
+
+\newcomment{One final point about the ``decoupled: merge + create" curve is
+that the results in the graph are pessimistic. They model a scenario in which
+all client journals arrive at the same time.} So for the 20 clients data point,
+we are measuring the operations per second for 20 client journals that land on
+the metadata server at the same time. \newcomment{Had we added infrastructure
+to overlay journal arrivals or time client sync intervales, we could have
+scaled more closely to ``decoupled: create".}\oldcomment{The ``create + merge"
+bar adds the time it takes for clients to generate the journal of events in
+parallel. ALSO DISCUSS 3.7X!!}
 
@@
 
-This workload scales further with decoupled namespaces because the metadata
+\oldcomment{This workload scales further with decoupled namespaces because the metadata
 server is not overwhelmed with RPC requests. Compared to
@@
 double the capacity of a single metadata server using the RPC strategy to
-maintain strong consistency\footnote{\newcomment{Note that for the 20 clients
+maintain strong consistency Note that for the 20 clients
 bar in Figure~\ref{fig:mergescale}, we plot the throughput for using 18 clients
-(the maximum number of clients supported by RPCs)}}. Related work has pointed
-to on-disk metadata format as the primary factor for improved scalability; but
-our results show that the weakened consistency and durability enables the
-scalability improvements. The on-disk metadata formats in this experiment are
-the same for all schemes.
-
-The performance of decoupled namespaces is better than RPCs because we place no
-restrictions on the validity of metadata inserted into the journal and avoid
-touching poorly scaling data structures. We can scale linearly if we weaken our
-consistency by never checking for the existence of files before creating files
-({\it i.e.} only append \texttt{open()} requests to the journal).  When the
-updates are merged by the metadata server into the in-memory metadata store
-they never scan growing data structures.  Had we implemented the Append Client
-Journal mechanism to include \texttt{lookup()} commands before \texttt{open()}
-requests, we would have seen the poor scaling that we see with the RPCs
-mechanism.
-
-%% DeltaFS vs. BatchFS vs. POSIX IO: decoupled is faster
-%% (5-7x) vs. (90-104x) slower
-%% Does not scale?
-%% Strong Consistency > 10X slower than Weak Consistency 
-%{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
-%decoupled namespace strategy and shows up to a \(20\times\) speedup over the
-%traditional namespaces that use RPCs. Compared to the baseline the slowdown is
-%\(5-7\times\) for Strong Consistency, which emulates BatchFS and
-%\(90-104\times\) for Weak Consistency, which emulates DeltaFS.
-
-%{\it Durability \(<<\) Consistency:} The performance of these semantics
-%suggests that the overhead of consistency is much larger than the overhead of
-%durability. This conclusion should be stronger as we scale the number of files
-%because the cost of streaming the journal into the object store is constant.
-%
-% Weak Consistency benefit not due to metadata format
+(the maximum number of clients supported by RPCs). }
 
@@
 applications control isolation to get better and more reliable performance.
-C}lients create files in private directories while another client interferes by
+C}lients create 100K files in their own directories while another client interferes by
 creating \oldcomment{more}\newcomment{1000} files in each directory. This
@@
 return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
-negligible effect on performance.
+negligible effect on performance.}
 
-\textbf{Takeaway}: the API lets users isolate directories when applications
+\oldcomment{\textbf{Takeaway}: the API lets users isolate directories when applications
 need better and more reliable performance. Blocking updates is an effective way of
@@
 The final use case shows how the API gives users fine-grained control of the
-consistency semantics\newcomment{ to support current practices and experimental
+consistency semantics \newcomment{to support current practices and experimental
 workflows}. \oldcomment{The use case is that u}\newcomment{U}sers often
@@
 though this operation is notoriously heavy-weight~\cite{carns:ipdps09-pvfs,
-eshel:fast10-panache} . The number of files or size of the files is indicative
+eshel:fast10-panache}. The number of files or size of the files is indicative
 of the progress. This practice is not too different from systems that use the
@@
 % Does this implementation need to be described in the implementation?
+\newcomment{We configure a subtree as a decoupled namespace with invisible
+consistency, local durability, and partial updates enabled.}
 Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
-client writing \newcomment{1 million} updates to a decoupled namespace and
-pausing to send updates to the metadata server. The error bars are the standard
-deviations of 3 runs. We scale the namespace sync interval to show the
-trade-off of frequently pausing or writing large logs of updates.  We use an
-idle core to log the updates and to do the network transfer. The client only
-pauses to fork off a background process, which is expensive as the address
-space needs to copied. The alternative is to pause the client completely and
+client writing \newcomment{1 million} updates to the decoupled namespace and
+pausing to send updates to the metadata server. \oldcomment{The error bars are
+the standard deviations of 3 runs.} We scale the namespace sync interval to
+show the trade-off of frequently pausing or writing large logs of updates.  We
+use an idle core to log the updates and to do the network transfer. The client
+only pauses to fork off a background process, which is expensive as the address
+space needs to be copied. The alternative is to pause the client completely and
 write the update to disk but since this implementation is limited by the speed
@@
 (up to 9\% overhead if done every second). The optimal sync interval for
-performance is 10 seconds because larger intervals must write more updates to
-disk and network. For the 25 second interval, the client only pauses 3-4 times
-but each sync writes about 278 thousand updates at once, which is a 678MB
-journal.
+performance is 10 seconds, which only incurs 2\% overhead), because larger
+intervals must write more updates to disk and network. For the 25 second
+interval, the client only pauses 3-4 times but each sync writes about 278
+thousand updates at once, which is a 678MB journal.
 

### 2017-12-26 paper/cover_letter.tex
paper: wordsmithing
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-%Experimental Setup (R1, R2, R3)
-%- code R1
-%- servers, network, storage R1, R3
-\section*{Issue I: No Experimental Setup (Reviewers 1, 2, and 3)}
+%Use Cases R2, R4
+%- Spark discussion R2
+%- more emphasis on parallel and distributed computing themes R4
+\section*{Issue I: Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}
 
-We apologize for the omission of experimental setup and source code details. To
-provide a more comprehensive view of our experiments, we added text to the
-paper and made paper artifacts available online. For the experimental setup, we
-describe the cluster setup (hardware, software, etc.) in Section~{\S}V.  We
-have also made the infrastructure code available and added links after each
-figure to show exactly how experiments are run.  This infrastructure code
-contains scripts to deploy the system, run experiments, and gather results.
-This process follows the Popper
-Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
-aims to make research reproducible.  For the source code, we have added code
-snippets to the evaluation to concretize what was actually done and made the
-source code available online. A link is provided in a footnote in
-Section~{\S}V.
+To show the relevance of the use cases, we re-wrote Section~{\S}V-B to
+highlight examples from HPC and cloud workloads that would benefit from using
+Cudele. For HPC we focus on user home directories used for experiments and for
+the cloud we focus on the Hadoop and Spark runtimes (as suggested by Reviewer
+2). To connect the problematic use cases to Cudele, we added \textbf{Cudele
+setup} headers in Section~{\S}V-B to show which subtree semantics accomodate
+these workloads in the global namespace.  By showing these setups, we hope to
+demonstrate that different applications can use a file system in parallel with
+different consistency/durability semantics.  Although we do not actually run
+the workloads on our prototype, hopefully our changes show how the use cases,
+and the synthetic benchmarks we used to represent them, are indeed relevant to
+parallel and distributed computing.
+
+To align the paper to the themes of the conference, we changed Section~{\S}V-A
+and Figures 6a, 6b, and 6c to emphasize that decoupled namespaces facilitate
+client-driven parallelism. Clients detach subtrees form the global namespaces
+and do operations in-parallel to their local disk. We also hope that changes to
+the use cases above motivates the need for robust distributed storage systems
+that can handle today's applications.  So while the paper is storage-centric,
+we try to show that the workloads it supports are highly distributed and need a
+flexible solution like Cudele.
 
@@
 One confusing component of this issue is that we mixed future work with the
-contributions of this paper; to make the contributions more explicit we remove
-future work from the introduction and add it to the future work section in
-Section~{\S}VI. We also connect all numbers we cite in the introduction to the
-evaluation. To further clarify the evaluation, we:
+contributions of the paper; to make the contributions more explicit we remove
+future work from the introduction and add it to Section~{\S}VI; we do not
+attempt to validate or prove any of the statements about the benefits of
+changing consistency and durability properties of a subtree dynamically, such
+as saving resource provisioning costs.  To further clarify the evaluation, we:
 
@@
 
-  \item added raw numbers (as annotations) to the baselines in each figure.
-This helps the reader calculate the raw numbers instead of just relying on
-overhead and speedup numbers.
+  \item explicitly state the Cudele setups and subtree configurations in bold
+in Section~{\S}V-A. Hopefully this helps the reader keep track of use cases and
+demonstrates how Cudele might be able to help in these scenarios.
 
-  \item re-organized the section to make experiments more independent and
-removed cross-references. Experiments are now self-contained so the reader can
-see the effects of different API configurations and we do a better job of
-explaining how results are derived.
+and include baseline values in the
+text so readers can quickly calculate raw numbers from our speedup/slowdown
+graphs. We also make sure to connect all speedups from the introduction to the
+evaluation. We hope this clarifies confusion about what we are comparing to
+when we reference speedups and slowdowns.
 
-  \item removed the ``major takeaways". We deleted the headings but left the
-conclusions we make as transitions. We also add insights into the results by
-analyzing the raw numbers we observe in comparison to hardware capabilities.
+  \item re-organized the sections and removed cross-references. Experiments are
+now self-contained so the reader can see the effects of different API
+configurations and we do a better job of explaining how results are derived.
+
+  \item removed the ``major takeaways". We deleted the headings but moved the
+text to the beginning of each section to motivate why we are doing each
+experiment.  We also add insights into the results by analyzing the raw numbers
+we observe in comparison to hardware capabilities.
+
+  \item re-labeled all graphs with the same metric (throughput
+slowdown/speedup) and aligned Section~{\S}V with the graphs in Section~{\S}II.
+We hope this eliminates more of the confusion triggered by the
+cross-referencing we did, referenced by Reviewers 2 and 3.
 
@@
 
-%Use Cases R2, R4
-%- Spark discussion R2
-%- more emphasis on parallel and distributed computing themes R4
-\section*{Issue III: Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}
+%Experimental Setup (R1, R2, R3)
+%- code R1
+%- servers, network, storage R1, R3
+\section*{Issue III: No Experimental Setup (Reviewers 1, 2, and 3)}
 
-We add a section to the introduction that scopes the importance of storage for
-large scale runtimes and workflows. Parallel and distributed jobs (whether they
-are MPI-based, MapReduce-based, etc.) need parallel and distributed storage
-systems that keep up with the computation. Furthermore, we highlight the
-parallel and distributed computing themes and cast light on them from the
-context of our prototype. We also add the Spark discussion proposed by to the
-evaluation. The goal is to show how a system such as Cudele would benefit from
-more general runtimes. We also move the discussion about the potential for HDFS
-and CephFS subtrees in the same namespace and the cost of dynamically changing
-consistency and durability semantics on a subtree.
+We apologize for the omission of experimental setup and source code details. To
+provide a more comprehensive view of our experiments, we added text to the
+paper and made paper artifacts available online. For the experimental setup, we
+describe the cluster setup (hardware, software, etc.) in Section~{\S}V.  We
+have also made the infrastructure code available and added links after each
+figure to show exactly how experiments are run.  This infrastructure code
+contains scripts to deploy the system, run experiments, and gather results.
+This process follows the Popper
+Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
+aims to make research reproducible.  We also made the source code available
+online and a link is provided in a footnote in Section~{\S}V. Finally, we add
+details about which tools and classes we modified in Section~{\S}IV to address
+Reviewer 1's questions about the framework implementation.
 

### 2017-12-26 paper/sections/0_abstract.tex
paper: wordsmithing
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-23 paper/sections/0_abstract.tex
@@
 storage system.  By custom fitting a subtree to a create-heavy application, we
-show 8\(\times\) speedup and 2\(\times\) scale when compared to our baseline
-system.
+show similar speedups to related work \newcomment{but more importantly, our
+prototype can custom fit subtrees in the same namespace to applications common
+in large data centers, such as checkpoint-restart (91.7\(\times\), user home
+directories (within a 0.03 standard deviation from optimal), and clients
+checking for partial results (only a 2\% overhead).}  \oldcomment{ and can
+scale to 2\(\times\) as many clients when compared to our baseline system.}
 

### 2017-12-26 paper/sections/5_implementation.tex
paper: wordsmithing
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % how it works in CephFS
-\textbf{Metadata Store}: In CephFS, the metadata store is a data structure that represents the file
+\subsection{Metadata Store}
+
+In CephFS, the metadata store is a data structure that represents the file
 system namespace. This data structure is stored in two places: in memory ({\it
@@
 
-\textbf{Journal}: The journal is the second way that CephFS represents the file system namespace;
-it is a log of metadata updates that can materialize the namespace when the
-updates are replayed onto the metadata store. The journal is a ``pile system";
-writes are fast but reads are slow because state must be reconstructed.
-Specifically, reads are slow because there is more state to read, it is
-unorganized, and many of the updates may be redundant.
+\subsection{Journal}
+
+The journal is the second way that CephFS represents the
+file system namespace; it is a log of metadata updates that can materialize the
+namespace when the updates are replayed onto the metadata store. The journal is
+a ``pile system"; writes are fast but reads are slow because state must be
+reconstructed.  Specifically, reads are slow because there is more state to
+read, it is unorganized, and many of the updates may be redundant.
 
@@
 
-\textbf{Journal Tool}: The journal tool is used for disaster recovery and lets administrators view and
+\subsection{Journal Tool}
+
+The journal tool is used for disaster recovery and lets administrators view and
 modify the journal. It can read the journal, export the journal as a file,
@@
 
-\textbf{Inode Cache}: In CephFS, the inode cache reduces the number of RPCs between clients and
+\subsection{Inode Cache}
+
+In CephFS, the inode cache reduces the number of RPCs between clients and
 metadata servers. Without contention clients can resolve metadata reads locally
-thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For example, if a client or
-metadata server is not caching the directory inode, all creates within that
-directory will result in a lookup and a create request. If the directory inode
-is cached then only the create needs to be sent.  The size of the inode cache
-is configurable so as not to saturate the memory on the metadata server --
-inodes in CephFS are about 1400
+thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For
+example, if a client or metadata server is not caching the directory inode, all
+creates within that directory will result in a lookup and a create request. If
+the directory inode is cached then only the create needs to be sent.  The size
+of the inode cache is configurable so as not to saturate the memory on the
+metadata server -- inodes in CephFS are about 1400
 bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
@@
 
-\textbf{Large Inodes}: In CephFS, inodes already store policies, like how the file is striped
+\subsection{Large Inodes}
+
+In CephFS, inodes already store policies, like how the file is striped
 across the object store or for managing subtrees for load balancing. These

### 2017-12-26 paper/sections/6_results.tex
paper: wordsmithing
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 constructing semantics used by real systems, (2) the techniques can work
-side-by-side in the same namespace, and (3) the combination of these techniques
+side-by-side in the same namespace\oldcomment{, and (3) the combination of these techniques
 can help the system scale further when subtrees are coupled to the correct type
-of application.  \newcomment{These techniques have already proven to be
+of application}. \newcomment{These techniques have already proven to be
 effective and scalable in systems that specialize an entire namespace according
-to a single optimization strategy (e.g., Lustre, BatchFS, DeltaFS), so our
+to a single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS), so our
 evaluation explores the behavior of the consistency/durability mechanisms and
-their effect with/without isolation.}
+their effect with and without isolation.}
 
@@
 results more generally applicable. We use a CloudLab cluster of 34 nodes
-connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.
-Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel release, version
-10.2.1, which was released in May 2016. All daemons run as a single process
-(the default setting for Ceph) and the Ceph
+connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and 400GB
+SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel release,
+version 10.2.1, which was released in May 2016. All daemons run as a single
+process (the default setting for Ceph) and the Ceph
 components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
-object storage daemons, 1 metadata server, and 29 clients.} We scope the
-evaluation to one metadata server and scale the number of parallel
-clients\newcomment{ each doing 100K operations}\footnote{This is the maximum
+object storage daemons, 1 metadata server, and 20 clients.} We scope the
+evaluation to one metadata server and scale the number of parallel clients
+\newcomment{each doing 100K operations}\footnote{This is the maximum
 recommended size of a directory in CephFS; preliminary experiments with larger
-directory sizes show memory problems.}. This type of analysis shows the
-capacity and performance of a metadata server with superior metadata protocols,
-which should be used to inform load balancing across a metadata cluster. Load
-balancing across a cluster of metadata servers with partitioning and
-replication can be explored with something like
+directory sizes show memory problems.} \newcomment{We scale to 20 clients
+because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
+to saturate the resources of a single metadata server.} This type of analysis
+shows the capacity and performance of a metadata server with superior metadata
+protocols, which should be used to inform \oldcomment{load
+balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
+cluster. Load balancing across a cluster of metadata servers with partitioning
+and replication can be explored with something like
 Mantle~\cite{sevilla:sc15-mantle}.
@@
 Cudele mechanisms for a single client creating files in the same directory, }
-\newcomment{We measure the overhead of each mechanism by having 1 client create
+\newcomment{We measure the overhead of each Cudele mechanism by having 1 client create
 100K files in a directory for various subtree configurations.
@@
 mechanism to process all metadata events.  Results are} normalized to the time
-it takes to write 100K file create updates to the client's in-memory journal
+it takes to write \oldcomment{100K file create} updates to the client's in-memory journal
 ({\it i.e.} the Append Client Journal mechanism).  \newcomment{The first graph
 groups the consistency mechanisms, the second groups the durability mechanisms,
-and the third has compositions representing real-world systems.} Stream is an
+and the third has compositions representing real-world systems.} \oldcomment{Stream is an
 approximation of the overhead and is calculated by subtracting the runtime of
 the job with the journal turned off from the runtime with the journal turned
-on.  \oldcomment{100K is the maximum recommended size of a directory in CephFS;
+on.  100K is the maximum recommended size of a directory in CephFS;
 preliminary experiments with larger directory sizes show memory problems.}
@@
 slowdowns\oldcomment{ for the larger number of creates}. For example, RPCs has
-a \oldcomment{\(90\times\)}\newcomment{\(18\times\)} slowdown because this
-technique relies on the internal directory data structures. It is a well-known
+a \oldcomment{\(90\times\)}\newcomment{\(17.9\times\)} slowdown because this
+technique relies on internal directory data structures\newcomment{, which is a
+well-known problem~\cite{ren:sc2014-indexfs}}. \oldcomment{It is a well-known
 problem that directory data structures do not scale when creating files in the
 same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
-structures will experience similar slowdowns. Other mechanisms that write
+structures will experience similar slowdowns.} Other mechanisms that write
 events to a journal experience a much less drastic slowdown because the journal
@@
 
-% RPCs vs. apply: calls to metadata server vs. RADOS
-{\it Overhead of RPCs:} RPCs is
+% RPCs vs. apply: calls to metadata server vs. RADOS TODO: why is apply so
+% slow.  apply: no CephFS changes, pulls/pushes same RADOS obj.  v_apply vs.
+% apply/persist: communicating through RADOS
+{\it Overhead of \oldcomment{RPCs}\newcomment{Consistency}:} RPCs is
 \oldcomment{\(66\times\)}\newcomment{\(19.9\times\)} slower than Volatile Apply
@@
 store \newcomment{with Nonvolatile Apply} is
-\oldcomment{\(10\times\)}\newcomment{\(78\times\)} slower.
-
-% TODO: why is apply so slow.  apply: no CephFS changes, pulls/pushes same
-% RADOS obj.  v_apply vs. apply/persist: communicating through RADOS
-{\it Overhead of Nonvolatile Apply:} The cost of Nonvolatile Apply is much
-larger than all the other mechanisms.  That mechanism was not implemented as
-part of Cudele -- it was a debugging and recovery tool packaged with CephFS. It
-works by iterating over the updates in the journal and pulling all objects that
-{\it may} be affected by the update.  This means that two objects are
-repeatedly pulled, updated, and pushed: the object that houses the experiment
-directory and the object that contains the root directory ({\it i.e.},
-\texttt{/}).  The cost of communicating through the object store is shown by
-comparing the runtime of Volatile Apply + Global Persist
-\newcomment{(\(1.3\times\))} to Nonvolatile Apply\newcomment{ (\(78\times\))}.
-These two operations end up with the same final metadata state but using
-Nonvolatile Apply is clearly inferior.
+\oldcomment{\(10\times\)}\newcomment{\(78\times\)} slower.  \oldcomment{{\it
+Overhead of Nonvolatile Apply:} The cost of Nonvolatile Apply is much larger
+than all the other mechanisms.  That mechanism}\newcomment{Nonvolatile Apply}
+was not implemented as part of Cudele -- it was a debugging and recovery tool
+packaged with CephFS. It works by iterating over the updates in the journal and
+pulling all objects that {\it may} be affected by the update.  This means that
+two objects are repeatedly pulled, updated, and pushed: the object that houses
+the experiment directory and the object that contains the root directory ({\it
+i.e.}, \texttt{/}).  \oldcomment{The cost of communicating through the object
+store is shown by comparing the runtime of Volatile Apply + Global Persist to
+Nonvolatile Apply.  These two operations} \newcomment{Nonvolatile Apply
+(\(78\times\)) and composing Volatile Apply + Global Persist (\(1.3\times\))}
+end up with the same final metadata state but using Nonvolatile Apply is
+clearly inferior.
 
 % persist vs. save: one disk vs. many
-{\it Parallelism of the Object Store:} \newcomment{Stream has the highest
-slowdown at \(2.4\times\) because the overhead of maintaining and streaming the
-journal is incurred by the metadata server}.  Comparing Local and Global
-Persist demonstrates the bandwidth advantages of storing the journal in a
-distributed object store. The Global Persist performance is \newcomment{only
-\(0.2\times\) slower than Local Persist because Global
-Persist}\oldcomment{\(1.5\times\) faster because the object store} is
-leveraging the collective bandwidth of the disks in the cluster. This benefit
-comes from the object store itself but should be acknowledged when making
-decisions for the application; the bandwidth of the object store can help
-mitigate the overheads of globally persisting metadata updates. The storage per
-journal update is about 2.5KB. So the storage footprint scales linearly with
-the number of metadata creates and suggests that updates for a million files
-would be \newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1 million files \(=\)
-2.38GB. Transfer times for payloads of this size in most HPC/data center
-networks are reasonable.}
+{\it Parallelism of the Object Store:} \newcomment{ Stream\footnote{This
+measurement is an approximation of the overhead and is calculated by
+subtracting the runtime of the job with the journal turned off from the runtime
+with the journal turned on.} has the highest slowdown at \(2.4\times\)
+because the overhead of maintaining and streaming the journal is incurred by
+the metadata server}.  Comparing Local and Global Persist demonstrates the
+bandwidth advantages of storing the journal in a distributed object store. The
+Global Persist performance is \newcomment{only \(0.2\times\) slower than Local
+Persist because Global Persist}\oldcomment{\(1.5\times\) faster because the
+object store} is leveraging the collective bandwidth of the disks in the
+cluster. This benefit comes from the object store itself but should be
+acknowledged when making decisions for the application; the bandwidth of the
+object store can help mitigate the overheads of globally persisting metadata
+updates. The storage per journal update is about 2.5KB. So the storage
+footprint scales linearly with the number of metadata creates and suggests that
+updates for a million updates in a single journal would be
+\newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1 million files \(=\) 2.38GB.
+Transfer times for payloads of this size in most HPC/data center networks are
+reasonable.}
 
@@
 global namespace.  We label the \(x\)-axis with systems that employ these
-semantics, as described in Figure~\ref{fig:subtree-policies}.  Again, the
-runtime is normalized to creating files in the client's in-memory journal.  We
-make no guarantees during execution of the mechanisms or when transitioning
-semantics -- the semantics are guaranteed {\it once the mechanism completes}.
-So if servers fail during a mechanism, metadata or data may be lost. This graph
-shows how we can build application-specific subtrees by composing mechanisms
-and that coupling well-established techniques to specific applications results
-in a more scalable global namespace.}
+semantics, as described in Figure~\ref{fig:subtree-policies}.  We make no
+guarantees during execution of the mechanisms or when transitioning semantics
+-- the semantics are guaranteed {\it once the mechanism completes}.  So if
+servers fail during a mechanism, metadata or data may be lost. This graph shows
+how we can build application-specific subtrees by composing mechanisms and that
+coupling well-established techniques to specific applications results in a more
+scalable global namespace.}
 
-\subsection{Use Case 1: Creates in the Same Directory}
+\subsection{Use Cases}
+
+\newcomment{Next we present three uses cases: creates in the same directory,
+interfering clients, and read while writing. For each use case, we provide
+motivation from HPC and cloud workloads; specifically, we look at users using
+the file system in parallel to run large-scale experiments in HPC and parallel
+runtimes that use the file system, such as Hadoop and Spark. The synthetic
+benchmarks model scenarios where these workloads co-exist in a global namespace
+and we provide insight into how the workload benefits from Cudele.}
+
+\subsubsection{Creates in the Same Directory}
 \label{sec:use-case-1}
 
-\oldcomment{Next we show how we can build application-specific subtrees by composing
-mechanisms and that \oldcomment{this approach}\newcomment{coupling
-well-established techniques to specific applications} results in a more a
-scalable global namespace.  }
-
-We start with clients creating files in private directories because this
-workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
+\oldcomment{Next we show how we can build application-specific subtrees by
+composing mechanisms and that this approach of coupling well-established
+techniques to specific applications results in a more a scalable global
+namespace.} We start with clients creating files in private directories because
+this workload is heavily studied in HPC~\cite{weil:sc2004-dyn-metadata,
 ren:sc2014-indexfs, patil:fast2011-giga, zheng:pdsw2014-batchfs,
 sevilla:sc15-mantle}, mostly due to checkpoint-restart~\cite{bent_plfs_2009}.
-But the workload also appears in cloud workloads~\cite{xiao:socc15-shardfs},
-where systems like Hadoop use the file abstraction to exchange work units to
-workers or to indicate when jobs
-complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
-uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
-services a flash crowd of creates across all directories as shown in
-Figure~\ref{fig:overhead-creates}. \oldcomment{ The workload is clients creating 100K files
-in private directories in the same global namespace. }
+A more familiar example is uncompressing an archive ({\it e.g.}, \texttt{tar
+xzf}), where the file system services a flash crowd of creates across all
+directories as shown in Figure~\ref{fig:overhead-creates}.  But the workload
+also appears in cloud workloads: Hadoop/Spark use the file \newcomment{system
+to assign work units to workers and the performance is proportional to the
+open/create throughput of the underlying file system~\cite{xiao:socc15-shardfs,
+shi:vldb15-spark, shvachko:login2012-hdfs-scalability}; Spark queries have an
+O(\(n\)) complexity, where \(n\) is the number of partitions.  Big Data
+Benchmark jobs examined in~\cite{chaimov:hpdc16-spark} have on the order of
+15,000 file opens or creates just to start a single Spark query and the Lustre
+system they tested on did not handle creates well, showing up to a \(24\times\)
+slowdown compared to other metadata operations. Common approaches to solve
+these types of bottlenecks is to change the application behavior or to design a
+new file system, like BatchFS or DeltaFS, that uses one set of metadata
+optimizations for the entire namespace.} \oldcomment{ abstraction to exchange
+work units to workers or to indicate when jobs complete  The workload is
+clients creating 100K files in private directories in the same global
+namespace.}
+
+\newcomment{\textbf{Cudele setup}: accommodate these workloads in the global
+namespace by configuring three subtrees with the following semantics: one with
+strong consistency and global durability (RPCs), one with invisible consistency
+and local durability (decoupled: create), and one with weak consistency and
+local durability (decoupled: create + merge).}
 
@@
 
-\oldcomment{The graph on the right of Figure~\ref{fig:composable-mechanisms} shows how
-applications can compose mechanisms together to get the consistency/durability
-guarantees they need in a global namespace.  We label the \(x\)-axis with
-systems that employ these semantics, as described in
+\oldcomment{The graph on the right of Figure~\ref{fig:composable-mechanisms}
+shows how applications can compose mechanisms together to get the
+consistency/durability guarantees they need in a global namespace.  We label
+the \(x\)-axis with systems that employ these semantics, as described in
 Figure~\ref{fig:subtree-policies}.  Again, the runtime is normalized to
-creating files in the client's in-memory journal.  We make no guarantees
-during execution of the mechanisms or when transitioning semantics -- the
-semantics are guaranteed {\it once the mechanism completes}. So if servers fail
-during a mechanism, metadata or data may be lost.}
+creating files in the client's in-memory journal.  We make no guarantees during
+execution of the mechanisms or when transitioning semantics -- the semantics
+are guaranteed {\it once the mechanism completes}. So if servers fail during a
+mechanism, metadata or data may be lost.}
 
@@
 Figure~\ref{fig:mergescale} we scale the number of clients \newcomment{each
-doing 100K file creates in their own directories for three types of subtrees:
-one with strong consistency and global durability (RPCs), one with invisible
-consistency and local durability (decoupled: create), and one with weak
-consistency and local durability (decoupled: create + merge). Results are
-normalized to 1 client that creates 100K files using RPCs (about 549
-creates/sec). As opposed to earlier graphs in
-Section~\S\ref{sec:posix-overheads} that plotted the throughput of the slowest
-client, Figure~\ref{fig:mergescale} plots the throughput from the perspective
-of the metadata server. Plotting this way is easier to understand because of
-how we normalize but the speedups over the RPC approach are the same, whether
-we look at the slowest client or not.}
+doing 100K file creates in their own directories.  Results are normalized to 1
+client that creates 100K files using RPCs (about 549 creates/sec). As opposed
+to earlier graphs in Section~\S\ref{sec:posix-overheads} that plotted the
+throughput of the slowest client, Figure~\ref{fig:mergescale} plots the
+throughput of the total job ({\it i.e.} from the perspective of the metadata
+server). Plotting this way is easier to understand because of how we normalize
+but the speedups over the RPC approach are the same, whether we look at the
+slowest client or not.}
 
-\newcomment{When the metadata server is operating at peak efficiency between 5
-and 20 clients, ``decoupled: merge + create" outperforms ``RPCs" by
-\(3.37\times\) because ``decoupled: merge + create" uses a relaxed form of
-consistency and leverages bulk updates just liked
-DeltaFS~\cite{zheng:pdsw2015-deltafs}.  The ``RPCs" and ``decoupled: merge +
-create" curves flatten out at a slowdown of \(4.5\times\) and \(15\times\),
-respectively, because more clients increases the load at the metadata server.
-Alternatively, the ``decoupled: create" curve scales linearly with the number
-of concurrent clients because clients operate in parallel and write updates
-locally. At 20 clients, we observe a \(91.7\times\) speedup for ``decoupled:
-create" over RPCs.}\oldcomment{ for the RPC per operation strategy and the
-decoupled namespace strategy.  For RPCs, scaling the number of clients
+\newcomment{When the metadata server is operating at peak efficiency at
+20 clients, the performance of the ``RPCs" and ``decoupled: merge + create"
+subtrees is bottlenecked by the metadata server processing power, so the curves
+flatten out at a slowdown of \(4.5\times\) and \(15\times\), respectively.  On
+the other hand, the ``decoupled: create" subtree performance scales linearly
+with the number of concurrent clients because clients operate in parallel and
+write updates locally. At 20 clients, we observe a \(91.7\times\) speedup for
+``decoupled: create" over RPCs.}\oldcomment{ for the RPC per operation strategy
+and the decoupled namespace strategy.  For RPCs, scaling the number of clients
 increases the load on the metadata server and for decoupled we increase the
@@
 
-\newcomment{Another reason that} the performance of decoupled namespaces is
-better than RPCs \newcomment{is because decoupled namespaces (1)} place no
-restrictions on the validity of metadata inserted into the journal and
-\newcomment{(2)} avoid touching poorly scaling data structures. We can scale
-linearly if we weaken our consistency by never checking for the existence of
-files before creating files ({\it i.e.} only append \texttt{open()} requests to
-the journal).  When the updates are merged by the metadata server into the
-in-memory metadata store they never scan growing data structures.  Had we
-implemented the \oldcomment{Append Client Journal mechanism}\newcomment{client}
-to include \texttt{lookup()} commands before \texttt{open()} requests, we would
-have seen the poor scaling \newcomment{behavior} that we see with the RPCs
-mechanism.
+\newcomment{ ``Decoupled: merge + create" outperforms ``RPCs" by \(3.37\times\)
+because ``decoupled: merge + create" uses a relaxed form of consistency and
+leverages bulk updates just like DeltaFS~\cite{zheng:pdsw2015-deltafs}.
+Decoupled namespaces (1)} place no restrictions on the validity of
+metadata inserted into the journal ({\it e.g.}, never checking for the
+existence of files before creating files), \newcomment{(2)} avoid touching
+poorly scaling data structures\newcomment{, and (3) allow clients to batch
+events into bulk updates}. \oldcomment{We can scale linearly if we weaken our
+consistency by never checking for the existence of files before creating files
+({\it i.e.} only append \texttt{open()} requests to the journal).  When the
+updates are merged by the metadata server into the in-memory metadata store
+they never scan growing data structures.}  Had we implemented the
+\oldcomment{Append Client Journal mechanism}\newcomment{client to send
+updates one at a time and} to include \texttt{lookup()} commands before
+\texttt{open()} requests, we would have seen \newcomment{performance closer to
+the ``RPC" subtree}\oldcomment{the poor scaling that we see with the RPCs
+mechanism}.
 
@@
 
-\subsection{Use Case 2: Creates with Interfering Client}
+\subsubsection{Interfering Clients}
 
@@
 Figure~\ref{fig:overhead-b}. In this workload, c}\newcomment{, which lets
-applications control isolation to get better and more reliable performance.
-C}lients create 100K files in their own directories while another client interferes by
-creating \oldcomment{more}\newcomment{1000} files in each directory. This
-introduces false sharing and the metadata server revokes capabilities on
-directories touched by the interfering client. While HPC tries to avoid these
-situations with workflows~\cite{zheng:pdsw2014-batchfs,
-zheng:pdsw2015-deltafs}, it still happens in distributed file systems when
-users unintentionally access directories in a shared file system.
+applications control isolation to get better and more reliable performance.}
+Clients create 100K files in their own directories while another client
+interferes by creating \oldcomment{more}\newcomment{1000} files in each
+directory. \oldcomment{This}\newcomment{The workload} introduces false sharing
+and the metadata server revokes capabilities on directories touched by the
+interfering client. While HPC tries to avoid these situations with
+workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
+happens in distributed file systems when users unintentionally access
+directories in a shared file system.  \newcomment{ In the cloud, Spark and
+Hadoop stacks use HDFS, which lets users ignore this type of consistency completely
+by letting interfering clients read files opened for
+writing~\cite{hakimzadeh:dais14-hdfs-consistency}.}
+
+\newcomment{\textbf{Cudele setup}: enable global durability with Stream and
+strong consistency with RPCs to mirror the setup from the problem presented in
+Figure~\ref{fig:overhead-b}. We configure one subtree with an interfere policy
+of ``allow" and another with subtree with ``block" so \texttt{-EBUSY} is
+returned to interfering clients.  The former is the default behavior in file
+systems and the latter isolates performance from malicious clients.}
+
 \oldcomment{We only scale to 15 clients because the performance variability of
 a nearly overloaded metadata server, in our case 18 clients, is too high.}
+Figure~\ref{fig:block-allow} plots the overhead of the slowest
+client\oldcomment{and the error bars are the standard deviation of 3 runs.
+``Interfere (Fig 3c)" is the result from Figure~\ref{fig:overhead-b};
+``Interfere" uses the allow/block API to return \texttt{-EBUSY} to interfering
+clients; and ``Isolated" is the baseline performance without an interfering.
+Each subtree has RPCs and Stream enabled.  Results are normalized to the
+runtime of a single client creating files in a directory.  Note that IndexFS
+has a similar behavior to ``interfere" with leases, except clients block.
+}\newcomment{, normalized to 1 client that creates 100K files in isolation
+(about 513 creates/sec). ``Interference" and ``no interference" is the
+performance with and without an interfering client touching files in every
+directory, respectively. The goal is to explicitly isolate clients so that
+performance is similar to the ``no interfere" curve, which has lower slowdowns
+(on average, \(1.42\times\) per client compared to \(1.67\times\) per client
+for ``interfere") and less variability (on average, a standard deviation of
+0.06 compared to 0.44 for ``interfere").  ``Block interference" uses the Cudele
+API to block interfering clients and the slowdown (\(1.34\times\) per client)
+and variability (0.09) look very similar to ``no interfere" for larger
+clusters. For smaller clusters the overhead to reject requests is more evident
+when the metadata server is underloaded so the slowdowns are similar to
+``interfere".  Clients can block interfering clients to get the same
+performance as isolated clients but there is a non-negligible overhead for
+rejecting requests when the metadata server is not operating at peak
+efficiency.}
 
-\newcomment{We switch back to clients doing RPCs per operation with journaling
-enabled, to mirror the setup from the problem presented in
-Figure~\ref{fig:overhead-b}.} Figure~\ref{fig:block-allow} plots the overhead
-of the slowest client \oldcomment{and the error bars are the standard deviation
-of 3 runs.  ``Interfere (Fig 3c)" is the result from
-Figure~\ref{fig:overhead-b}; ``Interfere" uses the allow/block API to return
-\texttt{-EBUSY} to interfering clients; and ``Isolated" is the baseline
-performance without an interfering.  Each subtree has RPCs and Stream enabled.
-Results are normalized to the runtime of a single client creating files in a
-directory.  Note that IndexFS has a similar behavior to ``interfere" with
-leases, except clients block.  }\newcomment{normalized to 1 client that creates
-100K files in isolation (about 513 creates/sec). ``Interference" and ``no
-interference" is the performance with and without an interfering client
-touching files in every directory, respectively. The goal is to explicitly
-isolate clients so that performance is similar to the ``no interfere" curve,
-which has lower slowdowns (on average, \(1.42\times\) per client compared to
-\(1.67\times\) per client for ``interfere") and less variability (on average, a
-standard deviation of 0.06 compared to 0.44 for ``interfere").  ``Block
-interference" uses the Cudele API to block interfering clients with
-\texttt{-EBUSY}. That curve shows slowdowns similar to ``interfere" for smaller
-clusters because the overhead to reject requests is more evident when the
-metadata server is underloaded; but the slowdown (\(1.34\times\) per client)
-and variability (0.09) look very similar to ``no interfere" for larger
-clusters. Clients can block interfering clients to get the same performance as
-isolated clients but there is a non-negligible overhead for rejecting requests
-when the metadata server is not operating at peak efficiency.}
+\oldcomment{We draw three conclusions: (1) clients that use the API to block
+interfering clients  get the same performance as isolated clients, (2) there is
+a negligible effect on performance for the extra work the metadata server does
+to return \texttt{-EBUSY}, and (3) merging updates from the decoupled client
+has a negligible effect on performance.  \textbf{Takeaway}: the API lets users
+isolate directories when applications need better and more reliable
+performance. Blocking updates is an effective way of controlling consistency.}
 
-\oldcomment{We draw three conclusions: (1) clients that use the API to block interfering
-clients  get the same performance as isolated clients, (2) there is a
-negligible effect on performance for the extra work the metadata server does to
-return \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
-negligible effect on performance.}
-
-\oldcomment{\textbf{Takeaway}: the API lets users isolate directories when applications
-need better and more reliable performance. Blocking updates is an effective way of
-controlling consistency.}
-
-\subsection{Use Case 3: Read while Writing}
+\subsubsection{Read while Writing}
 
@@
 of the progress. This practice is not too different from systems that use the
-file system to manage the progress of jobs; Hadoop writes to temporary files,
-renames them when complete, and creates a ``DONE" file to indicate to the
-runtime that the task did not fail and should not be re-scheduled on another
-node. In this scenario, Cudele users will not see the progress of decoupled
-namespaces since their updates are not globally visible.  To help users judge
-the progress of their jobs, Cudele clients have a ``namespace sync" that sends
-batches of updates back to the global namespace at regular intervals.
+file system to manage the progress of jobs; Spark and Hadoop write to temporary
+files, renames them when complete, and creates a ``DONE" file to indicate to
+the runtime that the task did not fail and should not be re-scheduled on
+another node. \newcomment{Also, the browser interface lets users check the
+progress by querying the file system and returning a \% of job complete
+metric.}
 
 % Does this implementation need to be described in the implementation?
-\newcomment{We configure a subtree as a decoupled namespace with invisible
-consistency, local durability, and partial updates enabled.}
+\newcomment{\textbf{Cudele setup}: in this scenario, Cudele users will not see
+the progress of decoupled namespaces since their updates are not globally
+visible.  To provide the performance of decoupled namespaces and to help users
+judge the progress of their jobs, Cudele clients have a ``namespace sync" that
+sends batches of updates back to the global namespace at regular intervals.
+We configure a subtree as a decoupled namespace with invisible consistency, local
+durability, and partial updates enabled.}
+
 Figure~\ref{fig:slowdown-sync} shows the performance degradation of a single
@@
 interval, the client only pauses 3-4 times but each sync writes about 278
-thousand updates at once, which is a 678MB journal.
+thousand updates at once, which is a journal of size 678MB.
 

### 2017-12-28 paper/cover_letter.tex
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 2). To connect the problematic use cases to Cudele, we added \textbf{Cudele
-setup} headers in Section~{\S}V-B to show which subtree semantics accomodate
+setup} headers in Section~{\S}V-B to show which subtree semantics accommodate
 these workloads in the global namespace.  By showing these setups, we hope to
@@
 flexible solution like Cudele.
+
+We also clarified confusing terminology in a footnote in Section~{\S}I: we use
+the term administrator to refer to a person that configures subtrees and client
+to refer to a storage client or application that interacts with the metadata
+server. We avoid using the term ``user" except when referring to end-users that
+interact with home directories in a file system or the Hadoop/Spark runtime
+itself.
 

### 2017-12-28 paper/sections/0_abstract.tex
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-23 paper/sections/0_abstract.tex
@@
 (2) manually moving data across system boundaries.  We present a framework and
-API that lets clients specify their consistency/durability requirements and
-dynamically assign them to subtrees in the same namespace, allowing users to
-optimize subtrees over time and space for different workloads.  We confirm the
-performance benefits of techniques presented in related work but also explore
-new consistency/durability metadata designs, all integrated over the same
-storage system.  By custom fitting a subtree to a create-heavy application, we
-show 8\(\times\) speedup and can scale to 2\(\times\) as many clients when
-compared to our baseline system.
+API that lets \oldcomment{clients}\newcomment{administrators} specify their
+consistency/durability requirements and dynamically assign them to subtrees in
+the same namespace, allowing administrators to optimize subtrees over time and
+space for different workloads.  We confirm the performance benefits of
+techniques presented in related work but also explore new
+consistency/durability metadata designs, all integrated over the same storage
+system.  By custom fitting a subtree to a create-heavy application, we show
+similar speedups to related work \newcomment{but more importantly, our
+prototype can custom fit subtrees in the same namespace to applications common
+in large data centers, such as checkpoint-restart (91.7\(\times\), user home
+directories (within a 0.03 standard deviation from optimal), and
+\oldcomment{clients}\newcomment{users} checking for partial results (only a 2\%
+overhead).}  \oldcomment{ and can scale to 2\(\times\) as many clients when
+compared to our baseline system.}
 

### 2017-12-28 paper/sections/1_introduction.tex
paper: wordsmithing
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
-metadata server for every client is unreasonable. This problem is exacerbated
-by current hardware and software trends; for example, HPC architectures are
-transitioning from complex storage stacks with burst buffer, file system,
-object store, and tape tiers to more simplified stacks with just a burst buffer
-and object store~\cite{bent:login16-hpc-trends}. These types of trends put
-pressure on data access because more requests end up hitting the same layer and
-latencies cannot be hidden while data migrates across tiers.
+metadata server for every \newcomment{client\footnote{In this paper, ``client"
+is a storage client or application that interacts with the metadata server,
+``administrator" is a system administrator that configures the storage, and
+``end-users" interact with the file system, via home directories or runtimes.}}
+is unreasonable. This problem is exacerbated by current hardware and software
+trends: parallel runtimes like Hadoop and Spark use storage systems to
+implement important scheduling and timing semantics, and  \oldcomment{; for
+example, } HPC architectures are transitioning from complex storage stacks with
+burst buffer, file system, object store, and tape tiers to more simplified
+stacks with just a burst buffer and object
+store~\cite{bent:login16-hpc-trends}. These types of trends put pressure on
+data access because more requests from different nodes end up hitting the same
+layer in parallel and latencies cannot be hidden while data migrates across
+tiers.
 
@@
 \caption{Illustration of subtrees with different semantics co-existing in a
-global namespace.  For performance, clients can relax consistency on their
+global namespace.  For performance, clients \oldcomment{can} relax consistency on their
 subtree (HDFS) or decouple the subtree and move it locally (BatchFS, RAMDisk).
@@
 Our prototype system, Cudele, achieves this by exposing ``mechanisms" that
-users combine to specify their preferred semantics.  Cudele supports 3 forms
-of consistency (invisible, weak, and strong) and 3 degrees of durability (none,
-local, and global) giving the user a wide range of policies and optimizations
-that can be custom fit to an application. We make the following contributions:
+\oldcomment{users}\newcomment{administrators} combine to specify their
+preferred semantics.  Cudele supports 3 forms of consistency (invisible, weak,
+and strong) and 3 degrees of durability (none, local, and global) giving the
+\oldcomment{user}\newcomment{administrator} a wide range of policies and
+optimizations that can be custom fit to an application. We make the following
+contributions:
 
@@
 
-  \item A framework and API for assigning consistency/durability policies 
-  to subtrees in the file system namespace; this lets users navigate
-  the trade-offs of different metadata protocols on the same storage system.
+  \item A framework and API for assigning consistency/durability policies to
+  subtrees in the file system namespace; this lets
+  \oldcomment{users}\newcomment{administrators} navigate the trade-offs of
+  different metadata protocols on the same storage system.
 
@@
 
-  \item A prototype that lets users custom fit subtrees to applications
-  dynamically.
+  \item A prototype that lets \oldcomment{users}\newcomment{administrators}
+  custom fit subtrees to applications dynamically.
 
@@
 For example, the results of a Hadoop job do not need to be migrated into a Ceph
-file system (CephFS) for other processing; instead the user can change the
-semantics of the HDFS subtree into a CephFS subtree.  This may cause
-metadata/data movement to make things strongly consistent again but this is
-superior to moving all data across file system boundaries. Our prototype
-enables studies that adjust these semantics over {\it time and space}, where
-subtrees can change their semantics and migrate around the cluster without ever
-moving the data they reference.}
+file system (CephFS) for other processing; instead the
+\oldcomment{users}\newcomment{administrator} can change the semantics of the
+HDFS subtree into a CephFS subtree.  This may cause metadata/data movement to
+make things strongly consistent again but this is superior to moving all data
+across file system boundaries. Our prototype enables studies that adjust these
+semantics over {\it time and space}, where subtrees can change their semantics
+and migrate around the cluster without ever moving the data they reference.}
 
@@
 that these techniques drastically improve performance.  We go a step further by
-quantifying the costs of merging updates (3.37\(\times\) slow down) and
-maintaining durability (\(2.4\times\) slow down).  In our prototype,
+quantifying the costs of merging updates (3.37\(\times\) slowdown) and
+maintaining durability (\(2.4\times\) slowdown).  In our prototype,
 \newcomment{we also show how to assign subtree semantics for certain
@@
 is fully relaxed), user home directories (within a 0.03 standard deviation from
-optimal), and clients checking for partial results (only a 2\% overhead).}
+optimal), and end-users checking for partial results (only a 2\% overhead).}
 \oldcomment{we get an 8\(\times\) speedup and can scale to twice as many

### 2017-12-28 paper/sections/3_posix_overheads.tex
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{figure}[tb]
-\centering
+\begin{figure}[tb] \centering
 \includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
 \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-compile/visualize/viz.ipynb}{source}]
-Create-heavy workloads (such as \texttt{untar}) incur the highest disk,
-network, and CPU utilization because of the consistency and durability demands
-of CephFS.}\label{fig:overhead-creates}
+For the CephFS metadata server, create-heavy workloads ({\it e.g.},
+\texttt{untar}) incur the highest disk, network, and CPU utilization because of
+the consistency and durability demands of CephFS.}\label{fig:overhead-creates}
 \end{figure}
@@
   \centering
-  \begin{subfigure}[b]{.3\linewidth}
+  \begin{subfigure}[b]{.32\linewidth}
       \centering
@@
   \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
+  \begin{subfigure}[b]{.32\linewidth}
       \centering
@@
   \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
+  \begin{subfigure}[b]{.32\linewidth}
       \centering
-      \includegraphics[width=1.1\linewidth]{graphs/behavior-interfere.png}
+      \includegraphics[width=1.15\linewidth]{graphs/behavior-interfere.png}
       \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
@@
   \end{subfigure}
-
   \caption{\newcomment{The durability and strong consistency slowdown of the
 existing CephFS implementation increases as the number of clients scales.
-Results are normalized to 1 client that creates 100K files in isolation.  (a)
-shows the effect of journaling metadata updates; ``segment(s)" is the number
-of journal segments dispatched to disk at once. (b) shows the slowdown when
-another client interferes by creating files in all directories and (c)
-highlights the cause: }\oldcomment{The overhead of durability and strong
-consistency in CephFS. (a) shows the effect of different journal segment sizes,
-which are streamed into the object store for fault tolerance. (b) and (c) show
-that} when another client ``interferes", capabilities are revoked and metadata
-servers do more work.} \label{fig:overhead}
-
+Results in (a) and (b)  are normalized to 1 client that creates 100K files in
+isolation.  (a) shows the effect of journaling metadata updates; ``segment(s)"
+is the number of journal segments dispatched to disk at once. (b) shows the
+slowdown when another client interferes by creating files in all directories
+and (c) highlights the cause: }\oldcomment{The overhead of durability and
+strong consistency in CephFS. (a) shows the effect of different journal segment
+sizes, which are streamed into the object store for fault tolerance. (b) and
+(c) show that} when another client ``interferes", capabilities are revoked and
+metadata servers do more work.} \label{fig:overhead}
 \end{figure*}
@@
 % what is durability
-While durability is not specified by POSIX IO, users expect that files they
-create or modify survive failures.  We define three types of durability:
-global, local, and none.  Global durability means that the client or server can
-fail at any time and metadata will not be lost because it is ``safe" ({\it
-i.e.} striped or replicated across a cluster). Local durability means that
-metadata can be lost if the client or server stays down after a failure. None
-means that metadata is volatile and that the system provides no guarantees when
-clients or servers fail.  None is different than local durability because
-regardless of the type of failure, metadata will be lost when components die in a
-None configuration.
+While durability is not specified by POSIX IO,
+\oldcomment{users}\newcomment{administrators} expect that files they create or
+modify survive failures.  We define three types of durability: global, local,
+and none.  Global durability means that the client or server can fail at any
+time and metadata will not be lost because it is ``safe" ({\it i.e.} striped or
+replicated across a cluster). Local durability means that metadata can be lost
+if the client or server stays down after a failure. None means that metadata is
+volatile and that the system provides no guarantees when clients or servers
+fail.  None is different than local durability because regardless of the type
+of failure, metadata will be lost when components die in a None configuration.
 

### 2017-12-28 paper/sections/4_methodology.tex
paper: final version
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-In this section we describe our API and framework that lets users
-assign consistency and durability semantics to subtrees in the global
-namespace. A \textbf{mechanism} is an abstraction and basic building block for
-constructing consistency and durability guarantees. Cudele exposes these
-mechanisms and the user composes them together to construct \textbf{policies}.
-These policies are assigned to subtrees and they dictate how the file system
-should handle operations within that subtree.  Below, we describe the
-mechanisms (which are \underline{underlined}), the policies, and the API for assigning
-policies to subtrees.
+In this section we describe our API and framework that lets
+\oldcomment{users}\newcomment{administrators} assign consistency and durability
+semantics to subtrees in the global namespace. A \textbf{mechanism} is an
+abstraction and basic building block for constructing consistency and
+durability guarantees. Cudele exposes these mechanisms and the
+\oldcomment{users}\newcomment{administrator} composes them together to
+construct \textbf{policies}.  These policies are assigned to subtrees and they
+dictate how the file system should handle operations within that subtree.
+Below, we describe the mechanisms (which are \underline{underlined}), the
+policies, and the API for assigning policies to subtrees.
 
@@
 % describe table
-The spectrum of consistency and durability guarantees that users can construct
-is shown in Table~\ref{table:spectrum}. The columns are the different
-consistency semantics and the rows cover the spectrum of durability guarantees.
-For consistency: ``invisible" means the system does not handle merging updates
-into a global namespace and it is assumed that middleware or the application
-manages consistency lazily; ``weak" merges updates at some time in the future
-({\it e.g.}, when the system has time, when the number of updates reaches a
-certain threshold, when the client is done writing, etc.); and updates in
-``strong" consistency are seen immediately by all clients. For durability,
-``none" means that updates are volatile and will be lost on a failure. Stronger
-guarantees are made with ``local", which means updates will be retained if the
-client node recovers and reads the updates from local storage, and ``global",
-where all updates are always recoverable.
+The spectrum of consistency and durability guarantees that
+\oldcomment{users}\newcomment{administrators} can construct is shown in
+Table~\ref{table:spectrum}. The columns are the different consistency semantics
+and the rows cover the spectrum of durability guarantees.  For consistency:
+``invisible" means the system does not handle merging updates into a global
+namespace and it is assumed that middleware or the application manages
+consistency lazily; ``weak" merges updates at some time in the future ({\it
+e.g.}, when the system has time, when the number of updates reaches a certain
+threshold, when the client is done writing, etc.); and updates in ``strong"
+consistency are seen immediately by all clients. For durability, ``none" means
+that updates are volatile and will be lost on a failure. Stronger guarantees
+are made with ``local", which means updates will be retained if the client node
+recovers and reads the updates from local storage, and ``global", where all
+updates are always recoverable.
 
@@
 in Section~\S\ref{sec:microbenchmarks}).  POSIX IO-compliant systems like
-CephFS and IndexFS have global consistency and durability~\footnote{This is the
+CephFS and IndexFS have global consistency and durability\footnote{This is the
 normal case.  IndexFS also has bulk merge which would transition the system
@@
 we assign labels here based on the points emphasized in the papers.  To compose
-the mechanisms users inject which mechanisms to run and which to use in
-parallel using a domain specific language.  Although we can achieve all
-permutations of the different guarantees in Table~\ref{table:spectrum}, not all
-of them make sense. For example, it makes little sense to do \texttt{append
-client journal+RPCs} since both mechanisms do the same thing or
-\texttt{stream+local persist} since ``global" durability is stronger and has more
-overhead than ``local" durability. The cost of each mechanism and the semantics
-described above are quantified in Sections~\S\ref{sec:microbenchmarks}
-and~\S\ref{sec:use-case-1}.
+the mechanisms \oldcomment{users}\newcomment{administrators} inject which
+mechanisms to run and which to use in parallel using a domain specific
+language.  Although we can achieve all permutations of the different guarantees
+in Table~\ref{table:spectrum}, not all of them make sense. For example, it
+makes little sense to do \texttt{append client journal+RPCs} since both
+mechanisms do the same thing or \texttt{stream+local persist} since ``global"
+durability is stronger and has more overhead than ``local" durability. The cost
+of each mechanism and the semantics described above are quantified in
+Sections~\S\ref{sec:microbenchmarks} and~\S\ref{sec:use-case-1}.
 

### 2017-12-28 paper/sections/6_results.tex
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-Cudele lets users construct consistency/durability guarantees using
-well-established research techniques from other systems; so instead of
-evaluating the scalability and performance of the techniques themselves against
-other file systems, we show that (1) the mechanisms we propose are useful for
-constructing semantics used by real systems, (2) the techniques can work
-side-by-side in the same namespace\oldcomment{, and (3) the combination of these techniques
-can help the system scale further when subtrees are coupled to the correct type
-of application}. \newcomment{These techniques have already proven to be
-effective and scalable in systems that specialize an entire namespace according
-to a single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS), so our
+Cudele lets \oldcomment{users}\newcomment{administrators} construct
+consistency/durability guarantees using well-established research techniques
+from other systems; so instead of evaluating the scalability and performance of
+the techniques themselves against other file systems, we show that (1) the
+mechanisms we propose are useful for constructing semantics used by real
+systems, (2) the techniques can work side-by-side in the same
+namespace\oldcomment{, and (3) the combination of these techniques can help the
+system scale further when subtrees are coupled to the correct type of
+application}. \newcomment{These techniques have already proven to be effective
+and scalable in systems that specialize an entire namespace according to a
+single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS), so our
 evaluation explores the behavior of the consistency/durability mechanisms and
@@
 directories in a shared file system.  \newcomment{ In the cloud, Spark and
-Hadoop stacks use HDFS, which lets users ignore this type of consistency completely
+Hadoop stacks use HDFS, which lets clients ignore this type of consistency completely
 by letting interfering clients read files opened for
@@
 
-The final use case shows how the API gives users fine-grained control of the
+The final use case shows how the API gives
+\oldcomment{users}\newcomment{administrators} fine-grained control of the
 consistency semantics \newcomment{to support current practices and experimental
@@
 the runtime that the task did not fail and should not be re-scheduled on
-another node. \newcomment{Also, the browser interface lets users check the
+another node. \newcomment{Also, the browser interface lets Hadoop/Spark users check the
 progress by querying the file system and returning a \% of job complete
@@
 % Does this implementation need to be described in the implementation?
-\newcomment{\textbf{Cudele setup}: in this scenario, Cudele users will not see
+\newcomment{\textbf{Cudele setup}: in this scenario, Cudele end-users will not see
 the progress of decoupled namespaces since their updates are not globally
-visible.  To provide the performance of decoupled namespaces and to help users
+visible.  To provide the performance of decoupled namespaces and to help end-users
 judge the progress of their jobs, Cudele clients have a ``namespace sync" that
@@
 \newcomment{We have shown that our API and framework is an effective
-abstraction for letting users custom fit subtrees to applications.  It is more
-scalable than the current practice of mounting different storage systems in a
-global namespace because there is no need to provision dedicated storage
-clusters to applications or move data between these systems.  For example, the
-results of a Hadoop job do not need to be migrated into a Ceph file system
-(CephFS) for other processing; instead the user can change the semantics of the
-HDFS subtree into a CephFS subtree.  This may cause metadata/data movement to
-make things strongly consistent again but this is superior to moving all data
-across file system boundaries. Our prototype enables studies that adjust these
+abstraction for letting \oldcomment{users}\newcomment{administrators} custom
+fit subtrees to applications.  It is more scalable than the current practice of
+mounting different storage systems in a global namespace because there is no
+need to provision dedicated storage clusters to applications or move data
+between these systems.  For example, the results of a Hadoop job do not need to
+be migrated into a Ceph file system (CephFS) for other processing; instead the
+\oldcomment{user}\newcomment{administrator} can change the semantics of the HDFS
+subtree into a CephFS subtree.  This may cause metadata/data movement to make
+things strongly consistent again but this is superior to moving all data across
+file system boundaries. Our prototype enables studies that adjust these
 semantics over {\it time and space}, where subtrees can change their semantics

### 2017-12-28 paper/sections/7_conclusion.tex
paper: terminology, spark, scoping (#47, #36, #35)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 namespace is convenient but wastes resources and incurs data movement.  Cudele
-lets users assign consistency/durability guarantees to subtrees in the global
-namespace, which can be custom fit to the application. Using Cudele, we show
-how applications can co-exist and perform well in a global namespace and lay
-the groundwork for new systems that dynamically change consistency/durability
-guarantees to avoid provisioning dedicated storage clusters and moving large
-amounts of data.
+lets \oldcomment{users}\newcomment{administrators} assign
+consistency/durability guarantees to subtrees in the global namespace, which
+can be custom fit to the application. Using Cudele, we show how applications
+can co-exist and perform well in a global namespace and lay the groundwork for
+new systems that dynamically change consistency/durability guarantees to avoid
+provisioning dedicated storage clusters and moving large amounts of data.
 

### 2017-12-29 paper/cover_letter.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-12-26 paper/cover_letter.tex
@@
 \begin{document}
-\title{Cover Letter}
+\title{Cover Letter: Cudele}
 \maketitle
@@
 \textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
-feedback; Reviewer 1 and 3's comments are addressed in issues I and II,
+feedback; Reviewer 1 and 3's comments are addressed in issues II and III,
 Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
-comments are addressed in issues II and III.
+comments are addressed in issues I and II.
 
@@
 %- more emphasis on parallel and distributed computing themes R4
-\section*{Issue I: Synthetic and/or Irrelevant Use Cases (Reviewers 2 and 4)}
+\section*{Issue I: Irrelevant and/or Synthetic Use Cases (Reviewers 2 and 4)}
 
-To show the relevance of the use cases, we re-wrote Section~{\S}V-B to
-highlight examples from HPC and cloud workloads that would benefit from using
-Cudele. For HPC we focus on user home directories used for experiments and for
-the cloud we focus on the Hadoop and Spark runtimes (as suggested by Reviewer
-2). To connect the problematic use cases to Cudele, we added \textbf{Cudele
-setup} headers in Section~{\S}V-B to show which subtree semantics accomodate
-these workloads in the global namespace.  By showing these setups, we hope to
-demonstrate that different applications can use a file system in parallel with
-different consistency/durability semantics.  Although we do not actually run
-the workloads on our prototype, hopefully our changes show how the use cases,
-and the synthetic benchmarks we used to represent them, are indeed relevant to
+We re-wrote Section~{\S}V-B to highlight examples from HPC and cloud workloads
+that would benefit from using Cudele. For HPC we focus on user home directories
+used for experiments and for the cloud we focus on the Hadoop and Spark
+runtimes (as suggested by Reviewer 2).  We also added \textbf{In Cudele}
+headers in Section~{\S}V-B to show which subtree semantics accommodate each
+workload and to demonstrate how Cudele might be able to help in these
+scenarios.  By showing these setups, we hope to demonstrate that different
+applications can use a file system in parallel with different
+consistency/durability semantics.  Although we do not actually run the
+workloads on our prototype, hopefully our changes show how the use cases, and
+the synthetic benchmarks we used to represent them, are indeed relevant to
 parallel and distributed computing.
@@
 and Figures 6a, 6b, and 6c to emphasize that decoupled namespaces facilitate
-client-driven parallelism. Clients detach subtrees form the global namespaces
+client-driven parallelism, where clients detach subtrees form the global namespaces
 and do operations in-parallel to their local disk. We also hope that changes to
@@
 
-All reviewers note that the results and contributions we cite in the
-introduction are not validated, explained, or even mentioned in the evaluation.
-One confusing component of this issue is that we mixed future work with the
-contributions of the paper; to make the contributions more explicit we remove
-future work from the introduction and add it to Section~{\S}VI; we do not
-attempt to validate or prove any of the statements about the benefits of
-changing consistency and durability properties of a subtree dynamically, such
-as saving resource provisioning costs.  To further clarify the evaluation, we:
+To clarify the evaluation and eliminate some of the confusion due to our
+cross-referencing, we:
 
@@
 
-  \item explicitly state the Cudele setups and subtree configurations in bold
-in Section~{\S}V-A. Hopefully this helps the reader keep track of use cases and
-demonstrates how Cudele might be able to help in these scenarios.
+  \item connected performance gains from the abstract and Section~{\S}I to
+Section~{\S}V to clarify speedup/slowdown comparisons and to show how results
+are derived; we also re-labeled all graphs with the same metric (throughput
+slowdown/speedup) and aligned Section~{\S}V with the graphs in Section~{\S}II.
 
-and include baseline values in the
-text so readers can quickly calculate raw numbers from our speedup/slowdown
-graphs. We also make sure to connect all speedups from the introduction to the
-evaluation. We hope this clarifies confusion about what we are comparing to
-when we reference speedups and slowdowns.
+  \item made experiments self-contained in Sections~{\S}II and~{\S}V by
+removing cross-references and adding baseline values so readers can calculate
+raw numbers from our speedup/slowdown graphs.
 
-  \item re-organized the sections and removed cross-references. Experiments are
-now self-contained so the reader can see the effects of different API
-configurations and we do a better job of explaining how results are derived.
+  \item explicitly state Cudele setups and subtree configurations in bold in
+Section~{\S}V-B to guide the reader through use cases and the effects of
+different Cudele API configurations.
 
-  \item removed the ``major takeaways". We deleted the headings but moved the
-text to the beginning of each section to motivate why we are doing each
-experiment.  We also add insights into the results by analyzing the raw numbers
-we observe in comparison to hardware capabilities.
+  \item removed ``major takeaways" from Section~{\S}V and focused on
+insights into results by analyzing raw numbers in comparison to hardware
+speeds.
 
-  \item re-labeled all graphs with the same metric (throughput
-slowdown/speedup) and aligned Section~{\S}V with the graphs in Section~{\S}II.
-We hope this eliminates more of the confusion triggered by the
-cross-referencing we did, referenced by Reviewers 2 and 3.
+  \item removed future work from Section~{\S}I to make the contributions
+explicit. We do not attempt to validate or prove any of the statements about
+the benefits of changing consistency and durability properties of a subtree
+dynamically, such as saving resource provisioning costs.
+
+  \item clarified confusing terminology in Section~{\S}II: administrator refers
+to a person that configures subtrees and client refers to a storage client or
+application that interacts with the metadata server. We avoid using the term
+``user" except when referring to end-users that interact with home directories
+in a file system or the Hadoop/Spark runtime itself.
 
@@
 
-We apologize for the omission of experimental setup and source code details. To
-provide a more comprehensive view of our experiments, we added text to the
-paper and made paper artifacts available online. For the experimental setup, we
-describe the cluster setup (hardware, software, etc.) in Section~{\S}V.  We
-have also made the infrastructure code available and added links after each
-figure to show exactly how experiments are run.  This infrastructure code
-contains scripts to deploy the system, run experiments, and gather results.
-This process follows the Popper
+We apologize for the omission of experimental setup and source code details.
+We made artifacts available online for all experiments in the paper. For the
+experimental setup, we describe the cluster setup (hardware, software, etc.) in
+Section~{\S}V.  We have also made the infrastructure code available and added
+links after each figure to show exactly how experiments are run.  This
+infrastructure code contains scripts to deploy the system, run experiments, and
+gather results.  This process follows the Popper
 Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
 aims to make research reproducible.  We also made the source code available
-online and a link is provided in a footnote in Section~{\S}V. Finally, we add
+online and a link is provided in a reference in Section~{\S}V. Finally, we add
 details about which tools and classes we modified in Section~{\S}IV to address
@@
 
-\section*{Cosmetic Fixes}
-%Terminology
-%- users vs. clients R2
-\begin{itemize}
-  \item Reviewer 1: fixed user vs. client terminology in Section~{\S}III
-  \item Reviewer 1: clarified that the cost of dynamically changing semantics is future work Sections~{\S}III and ~{\S}V.E
-  \item Reviewer 1: added source code pointer in Section~{\S}IV
-  \item Reviewer 2: remove major takeways in Section~{\S}V
-  \item Reviewer 2: remove major cross-references in Section~{\S}V
-  \item Reviewer 3: quantified speedups with figure annotations in Section~{\S}V
-  \item Reviewers 1 and 3: add servers, network, and storage setups in Section~{\S}V
-  \item Reviewer 4: add new section ({\S}V.F) describing how Cudele would work with a system like Spark
-\end{itemize}
+%\section*{Cosmetic Fixes}
+%%Terminology
+%%- users vs. clients R2
+%\begin{itemize}
+%  \item Reviewer 1: fixed user vs. client terminology in Section~{\S}III
+%  \item Reviewer 1: clarified that the cost of dynamically changing semantics is future work Sections~{\S}III and ~{\S}V.E
+%  \item Reviewer 1: added source code pointer in Section~{\S}IV
+%  \item Reviewer 2: remove major takeways in Section~{\S}V
+%  \item Reviewer 2: remove major cross-references in Section~{\S}V
+%  \item Reviewer 3: quantified speedups with figure annotations in Section~{\S}V
+%  \item Reviewers 1 and 3: add servers, network, and storage setups in Section~{\S}V
+%  \item Reviewer 4: add new section ({\S}V.F) describing how Cudele would work with a system like Spark
+%\end{itemize}
 

### 2017-12-29 paper/sections/0_abstract.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-12-26 paper/sections/0_abstract.tex
@@
 HPC and data center scale application developers are abandoning POSIX IO
-because the file system metadata synchronization and serialization overheads of
+because file system metadata synchronization and serialization overheads of
 providing strong consistency and durability are too costly -- and often
@@
 stronger guarantees, forcing developers to re-write their applications or
-deploy them on a different system.  Users can mount multiple systems in the
-global namespace but this means (1) provisioning separate storage clusters and
-(2) manually moving data across system boundaries.  We present a framework and
-API that lets clients specify their consistency/durability requirements and
-dynamically assign them to subtrees in the same namespace, allowing users to
-optimize subtrees over time and space for different workloads.  We confirm the
-performance benefits of techniques presented in related work but also explore
-new consistency/durability metadata designs, all integrated over the same
-storage system.  By custom fitting a subtree to a create-heavy application, we
-show similar speedups to related work \newcomment{but more importantly, our
-prototype can custom fit subtrees in the same namespace to applications common
-in large data centers, such as checkpoint-restart (91.7\(\times\), user home
-directories (within a 0.03 standard deviation from optimal), and clients
-checking for partial results (only a 2\% overhead).}  \oldcomment{ and can
-scale to 2\(\times\) as many clients when compared to our baseline system.}
+deploy them on a different system.  We present a framework and API that lets
+\oldcomment{clients}\newcomment{administrators} specify their
+consistency/durability requirements and dynamically assign them to subtrees in
+the same namespace, allowing administrators to optimize subtrees over time and
+space for different workloads. \oldcomment{By custom fitting a subtree to a
+create-heavy application,} We show similar speedups to related work but more
+importantly, \oldcomment{our prototype can custom fit subtrees in the same
+namespace to applications common in large data centers, }\newcomment{ we show
+performance improvements when we custom fit subtree semantics to applications}
+such as checkpoint-restart (91.7\(\times\) speedup), user home directories
+(0.03 standard deviation from optimal), and
+\oldcomment{clients}\newcomment{users} checking for partial results (2\%
+overhead). \oldcomment{ and can scale to 2\(\times\) as many clients when
+compared to our baseline system.}
 
 \end{abstract}
+
+%Users can mount multiple systems in the
+%global namespace but this means (1) provisioning separate storage clusters and
+%(2) manually moving data across system boundaries.
+
+%We confirm the performance benefits of
+%techniques presented in related work but also explore new
+%consistency/durability metadata designs, all integrated over the same storage
+%system.

### 2017-12-29 paper/sections/1_introduction.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
-metadata server for every \newcomment{client\footnote{In this paper, ``client"
-is a storage client or application that interacts with the metadata server,
-``administrator" is a system administrator that configures the storage, and
-``end-users" interact with the file system, via home directories or runtimes.}}
-is unreasonable. This problem is exacerbated by current hardware and software
-trends: parallel runtimes like Hadoop and Spark use storage systems to
-implement important scheduling and timing semantics, and  \oldcomment{; for
-example, } HPC architectures are transitioning from complex storage stacks with
-burst buffer, file system, object store, and tape tiers to more simplified
-stacks with just a burst buffer and object
-store~\cite{bent:login16-hpc-trends}. These types of trends put pressure on
-data access because more requests from different nodes end up hitting the same
-layer in parallel and latencies cannot be hidden while data migrates across
-tiers.
+metadata server for every \newcomment{client} is unreasonable. This problem is
+exacerbated by current hardware and software trends: parallel runtimes like
+Hadoop and Spark use storage systems to implement important scheduling and
+timing semantics, and  \oldcomment{; for example, } HPC architectures are
+transitioning from complex storage stacks with burst buffer, file system,
+object store, and tape tiers to more simplified stacks with just a burst buffer
+and object store~\cite{bent:login16-hpc-trends}. These types of trends put
+pressure on data access because more requests from different nodes end up
+hitting the same layer in parallel and latencies cannot be hidden while data
+migrates across tiers.
 
@@
 \caption{Illustration of subtrees with different semantics co-existing in a
-global namespace.  For performance, clients \oldcomment{can} relax consistency on their
-subtree (HDFS) or decouple the subtree and move it locally (BatchFS, RAMDisk).
-Decoupled subtrees can relax durability for even better performance.
+global namespace.  For performance, clients \oldcomment{can} relax consistency/durability on their
+subtree ({\it e.g.}, HDFS) or decouple the subtree and move it locally ({\it e.g.}, BatchFS, RAMDisk).
 }\label{fig:subtree-policies} \end{figure}
@@
 % What did we do
+%\footnote{HDFS itself is not directly evaluated in this paper, although the
+%semantics and their performance is explored in~\S\ref{sec:use-case-1}.} 
 To address this problem, we present an API and framework that lets developers
@@
 applications optimized with techniques from different state-of-the-art
-architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
-paper, although the semantics and their performance is explored
-in~\S\ref{sec:use-case-1}.} subtree has weaker than strong consistency because
-it lets clients read files opened for
+architectures.  The HDFS subtree has weaker than strong consistency because it
+lets clients read files opened for
 writing~\cite{hakimzadeh:dais14-hdfs-consistency}, which means that not all
@@
 
-  \item A framework and API for assigning consistency/durability policies to
-  subtrees in the file system namespace; this lets
-  \oldcomment{users}\newcomment{administrators} navigate the trade-offs of
+  \item A framework/API for assigning consistency/durability policies to
+  subtrees in a global namespace; this lets
+  \oldcomment{users}\newcomment{administrators} navigate trade-offs of
   different metadata protocols on the same storage system.
 
-  \item This framework lets subtrees with different semantics co-exist in a
-  global namespace. We show how this scales further and performs better than
-  systems that use one strategy for the entire namespace .
+  \item
+
+  \oldcomment{This framework lets subtrees with different semantics co-exist in
+a global namespace. We show how this} \newcomment{We show how that letting
+different semantics co-exist in a global namespace scales further and performs
+better than systems that use one strategy.}
 

### 2017-12-29 paper/sections/3_posix_overheads.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \texttt{untar}) incur the highest disk, network, and CPU utilization because of
-the consistency and durability demands of CephFS.}\label{fig:overhead-creates}
+consistency/durabiliy demands.}\label{fig:overhead-creates}
 \end{figure}
@@
 CephFS, the file system that uses Ceph's object store ({\it i.e.} RADOS) to
-store its data and metadata. To show how the file system behaves under high
-metadata load we use a create-heavy workload.  During this process we
-discovered, based on the analysis and breakdown of costs, that durability and
-consistency have high overhead but we urge the reader to keep in mind that this
-file system is an implementation of one set of design decisions and our goal
-here is to highlight the effect that those decisions have on performance.
+store its data and metadata. \newcomment{In this paper, ``client" is a storage
+client or application that interacts with the metadata server, ``administrator"
+is a system administrator that configures the storage, and ``end-users"
+interact with the file system, via home directories or runtimes.} To show how
+the file system behaves under high metadata load we use a create-heavy
+workload.  During this process we discovered, based on the analysis and
+breakdown of costs, that durability and consistency have high overhead but we
+urge the reader to keep in mind that this file system is an implementation of
+one set of design decisions and our goal here is to highlight the effect that
+those decisions have on performance.
 

### 2017-12-29 paper/sections/4_methodology.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \centering
-\includegraphics[width=90mm]{figures/fig-decouple.png}
+%\includegraphics[width=50mm]{figures/fig-decouple.png}
+\includegraphics[width=1.0\linewidth]{figures/fig-decouple.png}
 \caption{Illustration of the \textbf{mechanisms} used by applications to build
-consistency/durability semantics. Descriptions are provided in
-Table~\ref{table:mechanisms} and the underlined words in
-Section~\S\ref{sec:the-cudelesfs-mechanisms}.}\label{fig:decouple}
+consistency/durability semantics. Descriptions are provided in by the
+underlined words in Section~\S\ref{sec:the-cudelesfs-mechanisms}.
+}\label{fig:decouple}
 \end{figure}
 
-\begin{table}
-\begin{tabular}{ r | l }
-  Mechanism         & Description \\\hline
-  RPCs              & round trip remote procedure calls \\
-  Stream            & stream journal into object store \\
-  Append Client Journal & events appended to in-memory journal \\
-  Volatile Apply & apply to metadata store in memory \\
-  Nonvolatile Apply    & apply to metadata store in obj store \\
-  Local Persist     & journal saved to client's disk \\
-  Global Persist    & journal saved in object store \\
-\end{tabular}
-\caption{Descriptions of the mechanisms. Example compositions are shown in Table~\ref{table:spectrum}.\label{table:mechanisms}} 
-\end{table}
+%\begin{table}
+%\begin{tabular}{ r | l }
+%  Mechanism         & Description \\\hline
+%  RPCs              & round trip remote procedure calls \\
+%  Stream            & stream journal into object store \\
+%  Append Client Journal & events appended to in-memory journal \\
+%  Volatile Apply & apply to metadata store in memory \\
+%  Nonvolatile Apply    & apply to metadata store in obj store \\
+%  Local Persist     & journal saved to client's disk \\
+%  Global Persist    & journal saved in object store \\
+%\end{tabular}
+%\caption{Descriptions of the mechanisms. Example compositions are shown in Table~\ref{table:spectrum}.\label{table:mechanisms}} 
+%\end{table}
 
@@
 Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
-which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
-description of what each mechanism does.  Decoupled clients use the
+which daemon(s) they are performed by. \oldcomment{Table~\ref{table:mechanisms} has a
+description of what each mechanism does.}  Decoupled clients use the
 \underline{Append Client Journal} mechanism to append metadata updates to a
@@
 in Section~\S\ref{sec:microbenchmarks}).  POSIX IO-compliant systems like
-CephFS and IndexFS have global consistency and durability\footnote{This is the
-normal case.  IndexFS also has bulk merge which would transition the system
-into ``weak consistency"}; DeltaFS uses ``invisible" consistency and ``local"
-durability and BatchFS uses ``weak" consistency and ``local" durability. These
-systems have other features that could push them into different semantics but
-we assign labels here based on the points emphasized in the papers.  To compose
-the mechanisms \oldcomment{users}\newcomment{administrators} inject which
-mechanisms to run and which to use in parallel using a domain specific
-language.  Although we can achieve all permutations of the different guarantees
-in Table~\ref{table:spectrum}, not all of them make sense. For example, it
-makes little sense to do \texttt{append client journal+RPCs} since both
-mechanisms do the same thing or \texttt{stream+local persist} since ``global"
-durability is stronger and has more overhead than ``local" durability. The cost
-of each mechanism and the semantics described above are quantified in
+CephFS and IndexFS have global consistency and durability\footnote{ IndexFS
+also has bulk merge which is a form of ``weak consistency"}; DeltaFS uses
+``invisible" consistency and ``local" durability and BatchFS uses ``weak"
+consistency and ``local" durability. These systems have other features that
+could push them into different semantics but we assign labels here based on the
+points emphasized in the papers.  To compose the mechanisms
+\oldcomment{users}\newcomment{administrators} inject which mechanisms to run
+and which to use in parallel using a domain specific language.  Although we can
+achieve all permutations of the different guarantees in
+Table~\ref{table:spectrum}, not all of them make sense. For example, it makes
+little sense to do \texttt{append client journal+RPCs} since both mechanisms do
+the same thing or \texttt{stream+local persist} since ``global" durability is
+stronger and has more overhead than ``local" durability. The cost of each
+mechanism and the semantics described above are quantified in
 Sections~\S\ref{sec:microbenchmarks} and~\S\ref{sec:use-case-1}.
@@
 would apply the policies in ``policies.yml".  The policies file supports the
-following parameters (default values are in parenthesis):
-
-\begin{itemize}
-
-  \item \texttt{consistency}: which consistency model to use (\texttt{RPCs})
-
-  \item \texttt{durability}: which durability model to use (\texttt{stream})
-
-  \item \texttt{allocated\_inodes}: the number of inodes to provision to the
-  decoupled namespace (100)
-
-  \item \texttt{interfere\_policy}: how to handle a request from another
-  client targeted at this subtree (\texttt{allow})
-
-\end{itemize}
-
-The ``Consistency" and ``Durability" parameters are compositions of mechanisms;
-they can be serialized (\(+\)) or run in parallel (\(|\|\)).  ``Allocated
-Inodes" is a way for the application to specify how many files it intends to
-create. It is a contract so that the file system can provision enough resources
-for the incumbent merge and so it can give valid inodes to other clients. The
-inodes can be used anywhere within the decoupled namespace ({i.e.} at any depth
-in the subtree).  ``Interfere Policy" has two settings: \texttt{block} and
-\texttt{allow}.  For \texttt{block}, any requests to this part of the namespace
-returns with ``Device is busy", which will spare the metadata server from
-wasting resources for updates that may get overwritten. If the application does
-not mind losing updates, for example it wants approximations for results that
-take too long to compute, it can select \texttt{allow}. In this case, metadata
-from the interfering client will be written and the computation from the
-decoupled namespace will take priority at merge time because the results are
-more accurate.  Given these default values decoupling the namespace with an
-empty policies file would give the application 100 inodes but the subtree would
-behave like the existing CephFS implementation.
+following parameters (default values are in parenthesis): which consistency
+model to use (\texttt{RPCs}), which durability model to use (\texttt{stream})
+number of inodes to provision to the decoupled namespace (100), which interfere
+policy to use, {\it i.e.} how to handle a request from another client
+targeted at this subtree (\texttt{allow}). The ``Consistency" and ``Durability"
+parameters are compositions of mechanisms; they can be serialized (\(+\)) or
+run in parallel (\(|\|\)).  ``Allocated Inodes" is a way for the application to
+specify how many files it intends to create. It is a contract so that the file
+system can provision enough resources for the incumbent merge and so it can
+give valid inodes to other clients. The inodes can be used anywhere within the
+decoupled namespace ({i.e.} at any depth in the subtree).  ``Interfere Policy"
+has two settings: \texttt{block} and \texttt{allow}.  For \texttt{block}, any
+requests to this part of the namespace returns with ``Device is busy", which
+will spare the metadata server from wasting resources for updates that may get
+overwritten. If the application does not mind losing updates, for example it
+wants approximations for results that take too long to compute, it can select
+\texttt{allow}. In this case, metadata from the interfering client will be
+written and the computation from the decoupled namespace will take priority at
+merge time because the results are more accurate.  Given these default values
+decoupling the namespace with an empty policies file would give the application
+100 inodes but the subtree would behave like the existing CephFS
+implementation.

### 2017-12-29 paper/sections/5_implementation.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 store data structure is structured as a tree of directory fragments making it
-easier to read and traverse.
-
-% how i used it
-\textbf{Cudele}: the RPCs mechanism uses the metadata store to
-service requests. Using code designed for recovery, Nonvolatile Apply and
-Volatile Apply replay updates onto the metadata store in memory and in the
-object store, respectively.  When the clients are ready to merge their updates
-back into the global namespace, they pass a binary file of  metadata updates to
-the metadata server.
+easier to read and traverse. \textbf{In Cudele} the RPCs mechanism uses the
+metadata store to service requests. Using code designed for recovery,
+Nonvolatile Apply and Volatile Apply replay updates onto the metadata store in
+memory and in the object store, respectively.  When the clients are ready to
+merge their updates back into the global namespace, they pass a binary file of
+metadata updates to the metadata server.
 
@@
 
-\subsection{Journal}
+\subsection{Journal Format and Journal Tool}
 
-The journal is the second way that CephFS represents the
-file system namespace; it is a log of metadata updates that can materialize the
-namespace when the updates are replayed onto the metadata store. The journal is
-a ``pile system"; writes are fast but reads are slow because state must be
-reconstructed.  Specifically, reads are slow because there is more state to
-read, it is unorganized, and many of the updates may be redundant.
-
-\textbf{Cudele}: the journal format is used by Stream, Append Client Journal,
-Local Persist, and Global Persist.  Stream is the default
-implementation for achieving global durability in CephFS but the rest of the
-mechanisms are implemented by writing with the journal format.  By writing with
-the same format, the metadata servers can read and use the recovery code to
-materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
-These implementations required no changes to CephFS because the metadata
-servers know how to read the events the library is writing.  By re-using the
-journal subsystem to implement the namespace decoupling, Cudele leverages the
-write/read optimized data structures, the formats for persisting events
-(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
-for replaying events onto the internal namespace data structures.
+The journal is the second way that CephFS represents the file system namespace;
+it is a log of metadata updates that can materialize the namespace when the
+updates are replayed onto the metadata store. The journal is a ``pile system";
+writes are fast but reads are slow because state must be reconstructed.
+Specifically, reads are slow because there is more state to read, it is
+unorganized, and many of the updates may be redundant.  \textbf{In Cudele} the
+journal format is used by Stream, Append Client Journal, Local Persist, and
+Global Persist.  Stream is the default implementation for achieving global
+durability in CephFS but the rest of the mechanisms are implemented by writing
+with the journal format.  By writing with the same format, the metadata servers
+can read and use the recovery code to materialize the updates from a client's
+decoupled namespace ({\it i.e.} merge).  These implementations required no
+changes to CephFS because the metadata servers know how to read the events the
+library is writing.  By re-using the journal subsystem to implement the
+namespace decoupling, Cudele leverages the write/read optimized data
+structures, the formats for persisting events (similar to TableFS's
+SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
+onto the internal namespace data structures.
 
@@
 %\label{sec:journal-tool}
-
-\subsection{Journal Tool}
 
@@
 store objects and replays the updates on the metadata store in the object
-store.
-
-\textbf{Cudele}: the external library the clients link into is based
+store.  \textbf{In Cudele} the external library the clients link into is based
 on the journal tool. It already had functions for importing, exporting, and
@@
 
-\subsection{Inode Cache}
+\subsection{Inode Cache and Large Inodes}
 
@@
 metadata server -- inodes in CephFS are about 1400
-bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
-The inode cache has code for manipulating inode numbers, such as pre-allocating
-inodes to clients.
-
-\textbf{Cudele}: Nonvolatile Apply uses the internal inode cache
-code to allocate inodes to clients that decouple parts of the namespace and to
-skip inodes used by the client at merge time.
+bytes~\cite{docs:cephinternalds}.  The inode cache has code for manipulating
+inode numbers, such as pre-allocating inodes to clients. \textbf{In
+Cudele},Nonvolatile Apply uses the internal inode cache code to allocate inodes
+to clients that decouple parts of the namespace and to skip inodes used by the
+client at merge time.
 
@@
 
-\subsection{Large Inodes}
-
-In CephFS, inodes already store policies, like how the file is striped
-across the object store or for managing subtrees for load balancing. These
-policies adhere to logical partitionings of metadata or data, like Ceph pools and file
+In CephFS, inodes already store policies, like how the file is striped across
+the object store or for managing subtrees for load balancing. These policies
+adhere to logical partitionings of metadata or data, like Ceph pools and file
 system namespace subtrees. To implement this, the namespace data structure has
 the ability to recursively apply policies to subtrees and to isolate subtrees
-from each other.
-
-\textbf{Cudele}: uses the large inodes to store consistency and
+from each other. \textbf{In Cudele} the large inodes store consistency and
 durability policies in the directory inode. This approach uses the File Type

### 2017-12-29 paper/sections/6_results.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\newcomment{\indent We run experiments on the same cluster and graph the
-standard deviations for three runs (sometimes error bars are too small to see).
-We also normalize results to control for hardware differences and to make our
-results more generally applicable. We use a CloudLab cluster of 34 nodes
-connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and 400GB
-SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel release,
-version 10.2.1, which was released in May 2016. All daemons run as a single
-process (the default setting for Ceph) and the Ceph
-components\footnote{http://docs.ceph.com/docs/jewel/rados/} are: 1 monitor, 3
+\newcomment{\indent We graph standard deviations for three runs (sometimes
+error bars are too small to see) and normalize results to make our results more
+generally applicable to different hardware. We use a CloudLab cluster of 34
+nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
+400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
+release, version 10.2.1, which was released in May 2016. All daemons run as a
+single process (and the Ceph components~\cite{docs:cephrados} are: 1 monitor, 3
 object storage daemons, 1 metadata server, and 20 clients.} We scope the
 evaluation to one metadata server and scale the number of parallel clients
-\newcomment{each doing 100K operations}\footnote{This is the maximum
-recommended size of a directory in CephFS; preliminary experiments with larger
-directory sizes show memory problems.} \newcomment{We scale to 20 clients
-because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
-to saturate the resources of a single metadata server.} This type of analysis
-shows the capacity and performance of a metadata server with superior metadata
-protocols, which should be used to inform \oldcomment{load
-balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
-cluster. Load balancing across a cluster of metadata servers with partitioning
-and replication can be explored with something like
-Mantle~\cite{sevilla:sc15-mantle}.
-
-\newcomment{\indent To make our results reproducible, this paper adheres to The
-Popper Convention~\cite{jimenez_popper_2016}, so experiments presented here are
-available in the repository for this article.  Experiments can be examined in
-more detail, or even re-run, by visiting the \texttt{[source]} link next to
-each figure. That link points to a Jupyter notebook that shows the analysis and
-source code for that graph, which points to an experiment and its artifacts.
-The source code is available on the Cudele
-branch\footnote{\newcomment{https://github.com/michaelsevilla/ceph/tree/cudele}}
-of our Ceph fork.} \addressesissue{1}
+\newcomment{each doing 100K operations because 100K is the maximum recommended
+size of a directory in CephFS. We scale to 20 clients because, as shown in
+Section~\S\ref{sec:posix-overheads}, 20 clients is enough to saturate the
+resources of a single metadata server.} This type of analysis shows the
+capacity and performance of a metadata server with superior metadata protocols,
+which should be used to inform \oldcomment{load balancing}\newcomment{metadata
+distribution} across a \oldcomment{metadata} cluster. Load balancing across a
+cluster of metadata servers with partitioning and replication can be explored
+with something like Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our
+results reproducible, this paper adheres to The Popper
+Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
+detail, or even re-run, by visiting the \texttt{[source]} link next to each
+figure.  The source code is available on the Cudele branch~\cite{cudele-src} of
+our Ceph fork.}
 
@@
 \newcomment{Overhead of processing 100K create events for each mechanism in
-Table~\ref{table:mechanisms}. Results are normalized to the runtime of writing
-events to client memory. The bars on the far right show the overhead of
-building consistency/durability semantics of real world systems.}
-\oldcomment{Performance of each mechanism (left) and building the consistency/durability
-semantics of real-world systems (right) for 100K files creates from a single
-client. Results are normalized to the runtime of writing events to the client's
-in-memory journal.}\label{fig:composable-mechanisms}}
+\oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
+normalized to the runtime of writing events to client memory. The far right
+graph shows the overhead of building semantics of real world systems.}
+\oldcomment{Performance of each mechanism (left) and building the
+consistency/durability semantics of real-world systems (right) for 100K files
+creates from a single client. Results are normalized to the runtime of writing
+events to the client's in-memory journal.}\label{fig:composable-mechanisms}}
 \end{figure}
@@
 % persist vs. save: one disk vs. many
-{\it Parallelism of the Object Store:} \newcomment{ Stream\footnote{This
-measurement is an approximation of the overhead and is calculated by
-subtracting the runtime of the job with the journal turned off from the runtime
-with the journal turned on.} has the highest slowdown at \(2.4\times\)
-because the overhead of maintaining and streaming the journal is incurred by
-the metadata server}.  Comparing Local and Global Persist demonstrates the
-bandwidth advantages of storing the journal in a distributed object store. The
-Global Persist performance is \newcomment{only \(0.2\times\) slower than Local
-Persist because Global Persist}\oldcomment{\(1.5\times\) faster because the
-object store} is leveraging the collective bandwidth of the disks in the
-cluster. This benefit comes from the object store itself but should be
-acknowledged when making decisions for the application; the bandwidth of the
-object store can help mitigate the overheads of globally persisting metadata
-updates. The storage per journal update is about 2.5KB. So the storage
-footprint scales linearly with the number of metadata creates and suggests that
-updates for a million updates in a single journal would be
-\newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1 million files \(=\) 2.38GB.
-Transfer times for payloads of this size in most HPC/data center networks are
-reasonable.}
+{\it Parallelism of the Object Store:}
+\newcomment{Stream, which is an approximation (journal on minus journal off) has
+the highest slowdown at \(2.4\times\) because the overhead of maintaining and
+streaming the journal is incurred by the metadata server}.  Comparing Local and
+Global Persist demonstrates the bandwidth advantages of storing the journal in
+a distributed object store. The Global Persist performance is \newcomment{only
+\(0.2\times\) slower than Local Persist because Global
+Persist}\oldcomment{\(1.5\times\) faster because the object store} is
+leveraging the collective bandwidth of the disks in the cluster. This benefit
+comes from the object store itself but should be acknowledged when making
+decisions for the application; the bandwidth of the object store can help
+mitigate the overheads of globally persisting metadata updates. The storage per
+journal update is about 2.5KB. So the storage footprint scales linearly with
+the number of metadata creates and suggests that updates for a million updates
+in a single journal would be \newcomment{2.38GB}\oldcomment{2.5KB \(*\) 1
+million files \(=\) 2.38GB.  Transfer times for payloads of this size in most
+HPC/data center networks are reasonable.}
 
@@
 leverages bulk updates just like DeltaFS~\cite{zheng:pdsw2015-deltafs}.
-Decoupled namespaces (1)} place no restrictions on the validity of
-metadata inserted into the journal ({\it e.g.}, never checking for the
-existence of files before creating files), \newcomment{(2)} avoid touching
-poorly scaling data structures\newcomment{, and (3) allow clients to batch
-events into bulk updates}. \oldcomment{We can scale linearly if we weaken our
-consistency by never checking for the existence of files before creating files
-({\it i.e.} only append \texttt{open()} requests to the journal).  When the
-updates are merged by the metadata server into the in-memory metadata store
-they never scan growing data structures.}  Had we implemented the
-\oldcomment{Append Client Journal mechanism}\newcomment{client to send
-updates one at a time and} to include \texttt{lookup()} commands before
-\texttt{open()} requests, we would have seen \newcomment{performance closer to
-the ``RPC" subtree}\oldcomment{the poor scaling that we see with the RPCs
-mechanism}.
-
-\newcomment{One final point about the ``decoupled: merge + create" curve is
-that the results in the graph are pessimistic. They model a scenario in which
+Decoupled namespaces (1)} place no restrictions on the validity of metadata
+inserted into the journal ({\it e.g.}, never checking for the existence of
+files before creating files), \newcomment{(2)} avoid touching poorly scaling
+data structures\newcomment{, and (3) allow clients to batch events into bulk
+updates}. \oldcomment{We can scale linearly if we weaken our consistency by
+never checking for the existence of files before creating files ({\it i.e.}
+only append \texttt{open()} requests to the journal).  When the updates are
+merged by the metadata server into the in-memory metadata store they never scan
+growing data structures.}  Had we implemented the \oldcomment{Append Client
+Journal mechanism}\newcomment{client to send updates one at a time and} to
+include \texttt{lookup()} commands before \texttt{open()} requests, we would
+have seen \newcomment{performance closer to the ``RPC" subtree}\oldcomment{the
+poor scaling that we see with the RPCs mechanism}. \newcomment{The ``decoupled:
+merge + create" curve is also pessimistic because it models a scenario in which
 all client journals arrive at the same time.} So for the 20 clients data point,
@@
 
-\subsection{Discussion and Future Work}
-
-\newcomment{We have shown that our API and framework is an effective
-abstraction for letting \oldcomment{users}\newcomment{administrators} custom
-fit subtrees to applications.  It is more scalable than the current practice of
-mounting different storage systems in a global namespace because there is no
-need to provision dedicated storage clusters to applications or move data
-between these systems.  For example, the results of a Hadoop job do not need to
-be migrated into a Ceph file system (CephFS) for other processing; instead the
-\oldcomment{user}\newcomment{administrator} can change the semantics of the HDFS
-subtree into a CephFS subtree.  This may cause metadata/data movement to make
-things strongly consistent again but this is superior to moving all data across
-file system boundaries. Our prototype enables studies that adjust these
-semantics over {\it time and space}, where subtrees can change their semantics
-and migrate around the cluster without ever moving the data they reference.}
-
-Cudele and the experiments we present here prompt many avenues for future work.
-First is quantifying the benefits of dynamically changing the semantics of a
-subtree from stronger to weaker guarantees (or vice versa), as described in the
-introduction. Second is embeddable policies, where child subtrees have
-specialized features but still maintain the guarantees  of their parent
-subtrees. We already show how applications can compose guarantees and control
-their performance, knobs that we did not have before, but embeddable policies
-are a way to compose the policies themselves.  For example, a RAMDisk subtree
-is POSIX IO-compliant but relaxes durability constraints, so it can reside
-under a POSIX IO subtree that is strongly consistent. These embeddable policies
-should be controlled and enforced by Cudele.  Finally, performance prediction
-benefits from our cost quantification of each mechanism. Applications can use
-Cudele to microbenchmark their components and software, similar to what we did
-in Section~\S\ref{sec:microbenchmarks}. Using those results, they can predict
-how much slower their system will be if they adopt stronger consistency or
-durability.  This is a form of co-design that takes a ``dirty-slate" approach
-but building just the guarantees the application needs from existing
-implementations.  This can also be a verification tool where performance that
-varies wildly from the predicted performance can be a red flag that something
-is wrong or that the bottleneck is not in the consistency or durability plane.
+%\subsection{Discussion and Future Work}
+%
+%\newcomment{We have shown that our API and framework is an effective
+%abstraction for letting \oldcomment{users}\newcomment{administrators} custom
+%fit subtrees to applications.
 

### 2017-12-29 paper/sections/7_conclusion.tex
paper: #41 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-23 paper/sections/7_conclusion.tex
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
 
-Relaxing consistency/durability semantics in the file system is a double-edged
+Relaxing consistency/durability semantics in file systems is a double-edged
 sword. While it performs and scales better, it alienates applications that rely
-on strong consistency and durability. Mounting other systems to the global
-namespace is convenient but wastes resources and incurs data movement.  Cudele
-lets users assign consistency/durability guarantees to subtrees in the global
-namespace, which can be custom fit to the application. Using Cudele, we show
-how applications can co-exist and perform well in a global namespace and lay
-the groundwork for new systems that dynamically change consistency/durability
-guarantees to avoid provisioning dedicated storage clusters and moving large
-amounts of data.
+on strong consistency and durability.  Cudele lets
+\oldcomment{users}\newcomment{administrators} assign consistency/durability
+guarantees to subtrees in the global namespace, which lets administrators
+custom fit semantics to application. We show how applications can co-exist and
+perform well in a global namespace and our prototype enables studies that
+adjust these semantics over {\it time and space}, where subtrees can change
+without ever moving the data they reference.
+
+%This approach is superior to mounting different storage systems in a global
+%namespace because there is no need to provision dedicated storage clusters to
+%applications or move data between these systems.  
+Cudele prompts many avenues for future work.  First is dynamically changing
+semantics of a subtree from stronger to weaker guarantees (or vice versa).  For
+example, the results of a Hadoop job do not need to be migrated into CephFS for
+other processing; instead the \oldcomment{user}\newcomment{administrator} can
+change the semantics of the HDFS subtree into a CephFS subtree, which may cause
+metadata/data movement to make ensure strong consistency. Second is embeddable
+policies, where child subtrees have specialized features but still maintain
+guarantees  of their parent subtrees.  For example, a RAMDisk subtree is POSIX
+IO-compliant but relaxes durability constraints, so it can reside under a POSIX
+IO subtree that is strongly consistent.
+
+%Finally, our prototype enables
+%performance prediction because it helps administrators quantify the costs of
+%each mechanism (similar to Section~\S\ref{sec:microbenchmarks}).
+
+%Applications can use Cudele to microbenchmark their components and software,
+%similar to what we did in
+
+%Using those
+%results, they can predict how much slower their system will be if they adopt
+%stronger consistency or durability.  This is a form of co-design that takes a
+%``dirty-slate" approach but building just the guarantees the application needs
+%from existing implementations.  
+%This can also be a verification tool where performance that varies wildly from
+%the predicted performance can be a red flag that something is wrong or that
+%the bottleneck is not in the consistency or durability plane.
 

### 2018-01-07 paper/cover_letter.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 We thank the reviewers for their hard work and genuinely helpful suggestions.
-In addition to this cover letter, we have posted a PDF online explicitly
-showing additions in \textcolor{red}{\textbf{red}} and deletions in
-\textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
-feedback; Reviewer 1 and 3's comments are addressed in issues II and III,
-Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
+In addition to this cover letter, we have posted a
+\href{https://github.com/michaelsevilla/cudele-popper/blob/master/paper/diff.pdf}{PDF
+online} explicitly showing additions in \textcolor{blue}{\textbf{blue}} and
+deletions in \textcolor{red}{\textbf{red}}. Three issues were common across
+reviewer feedback; Reviewer 1 and 3's comments are addressed in issues II and
+III, Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
 comments are addressed in issues I and II.
@@
 used for experiments and for the cloud we focus on the Hadoop and Spark
-runtimes (as suggested by Reviewer 2).  We also added \textbf{In Cudele}
+runtimes (as suggested by Reviewer 2).  We also added \textbf{Cudele Setup}
 headers in Section~{\S}V-B to show which subtree semantics accommodate each
@@
 and do operations in-parallel to their local disk. We also hope that changes to
-the use cases above motivates the need for robust distributed storage systems
+the use cases above motivate the need for robust distributed storage systems
 that can handle today's applications.  So while the paper is storage-centric,
@@
 infrastructure code contains scripts to deploy the system, run experiments, and
-gather results.  This process follows the Popper
-Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
-aims to make research reproducible.  We also made the source code available
-online and a link is provided in a reference in Section~{\S}V. Finally, we add
+gather results.  This process follows the \href{http://falsifiable.us/}{Popper
+Convention}, which aims to make research reproducible.  We also made the source
+code available online and a link is provided in Section~{\S}V. Finally, we add
 details about which tools and classes we modified in Section~{\S}IV to address
@@
 
-%\section*{Cosmetic Fixes}
+\section*{Additional Cosmetic Fixes}
 %%Terminology
 %%- users vs. clients R2
-%\begin{itemize}
-%  \item Reviewer 1: fixed user vs. client terminology in Section~{\S}III
-%  \item Reviewer 1: clarified that the cost of dynamically changing semantics is future work Sections~{\S}III and ~{\S}V.E
-%  \item Reviewer 1: added source code pointer in Section~{\S}IV
-%  \item Reviewer 2: remove major takeways in Section~{\S}V
-%  \item Reviewer 2: remove major cross-references in Section~{\S}V
-%  \item Reviewer 3: quantified speedups with figure annotations in Section~{\S}V
-%  \item Reviewers 1 and 3: add servers, network, and storage setups in Section~{\S}V
-%  \item Reviewer 4: add new section ({\S}V.F) describing how Cudele would work with a system like Spark
-%\end{itemize}
 
-\bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,paper}
+For Reviewer 1, we fixed user vs. client terminology in Section~{\S}III,
+clarified that the cost of dynamically changing semantics is future work in
+Sections~{\S}I and~{\S}III, and added source code pointers in Section~{\S}V.
+For Reviewer 2, we removed the major takeways headings and cross-references in
+Section~{\S}V. For Reviewer 3, we quantified speedups with figure annotations
+in Section~{\S}V. For Reviewers 1 and 3, we added servers, network, and storage
+setups in Section~{\S}V. For Reviewer 4 we added descriptions to each use case
+in Section~{\S}V describing how Cudele would work with Spark.
+
 \end{document}

### 2018-01-07 paper/sections/0_abstract.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 namespace to applications common in large data centers, }\newcomment{ we show
-performance improvements when we custom fit subtree semantics to applications}
+performance improvements when we custom fit subtree semantics to applications
 such as checkpoint-restart (91.7\(\times\) speedup), user home directories
-(0.03 standard deviation from optimal), and
-\oldcomment{clients}\newcomment{users} checking for partial results (2\%
-overhead). \oldcomment{ and can scale to 2\(\times\) as many clients when
-compared to our baseline system.}
+(0.03 standard deviation from optimal), and users checking for partial results
+(2\% \oldcomment{clients}\newcomment{users} overhead).} \oldcomment{ and can
+scale to 2\(\times\) as many clients when compared to our baseline system.}
 

### 2018-01-07 paper/sections/3_posix_overheads.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
-\caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-compile/visualize/viz.ipynb}{source}]
+\caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-compile/visualize/viz.ipynb}{source}]
 For the CephFS metadata server, create-heavy workloads ({\it e.g.},
@@
       \includegraphics[width=1\linewidth]{graphs/slowdown-journal.png}
-      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-durability/visualize/viz.ipynb}{source}]
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-durability/visualize/viz.ipynb}{source}]
       effect of journal} \label{fig:overhead-a}
@@
       \includegraphics[width=0.98\linewidth]{graphs/slowdown-interfere-scale.png}
-      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-creates/visualize/viz.ipynb}{source}]
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-creates/visualize/viz.ipynb}{source}]
       interference hurts variability}
@@
       \includegraphics[width=1.15\linewidth]{graphs/behavior-interfere.png}
-      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
+      \caption{[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/baseline-interfere/visualize/viz.ipynb}{source}]
       interference increases RPCs}

### 2018-01-07 paper/sections/6_results.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
-release, version 10.2.1, which was released in May 2016. All daemons run as a
-single process (and the Ceph components~\cite{docs:cephrados} are: 1 monitor, 3
-object storage daemons, 1 metadata server, and 20 clients.} We scope the
-evaluation to one metadata server and scale the number of parallel clients
-\newcomment{each doing 100K operations because 100K is the maximum recommended
-size of a directory in CephFS. We scale to 20 clients because, as shown in
-Section~\S\ref{sec:posix-overheads}, 20 clients is enough to saturate the
-resources of a single metadata server.} This type of analysis shows the
-capacity and performance of a metadata server with superior metadata protocols,
-which should be used to inform \oldcomment{load balancing}\newcomment{metadata
-distribution} across a \oldcomment{metadata} cluster. Load balancing across a
-cluster of metadata servers with partitioning and replication can be explored
-with something like Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our
-results reproducible, this paper adheres to The Popper
+release, version 10.2.1, which was released in May 2016. We use 1 monitor
+daemon, 3 object storage daemons, 1 metadata server daemon, and up to 20
+clients.}  We scope the evaluation to one metadata server and scale the number
+of parallel clients \newcomment{each doing 100K operations because 100K is the
+maximum recommended size of a directory in CephFS. We scale to 20 clients
+because, as shown in Section~\S\ref{sec:posix-overheads}, 20 clients is enough
+to saturate the resources of a single metadata server.} This type of analysis
+shows the capacity and performance of a metadata server with superior metadata
+protocols, which should be used to inform \oldcomment{load
+balancing}\newcomment{metadata distribution} across a \oldcomment{metadata}
+cluster. Load balancing across a cluster of metadata servers with partitioning
+and replication can be explored with something like
+Mantle~\cite{sevilla:sc15-mantle}.  \newcomment{To make our results
+reproducible, this paper adheres to The Popper
 Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
 detail, or even re-run, by visiting the \texttt{[source]} link next to each
-figure.  The source code is available on the Cudele branch~\cite{cudele-src} of
-our Ceph fork.}
+figure.  The source code is available on the Cudele branch~\cite{docs:cudele}
+of our Ceph fork.}
 
@@
 \caption{
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
 \newcomment{Overhead of processing 100K create events for each mechanism in
@@
       \caption{
-      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mergescale/visualize/viz.ipynb}{source}]
       \newcomment{parallel creates on clients}
@@
       \caption{
-      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-blockapi/visualize/viz.ipynb}{source}]
       \newcomment{block interference}
@@
       \caption{
-      [\href{https://github.com/michaelsevilla/cudele-popper/blob/revision/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
+      [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-partialreads/visualize/viz.ipynb}{source}]
       \newcomment{syncing to global namespace}

### 2018-01-07 paper/sections/7_conclusion.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \oldcomment{users}\newcomment{administrators} assign consistency/durability
-guarantees to subtrees in the global namespace, which lets administrators
-custom fit semantics to application. We show how applications can co-exist and
+guarantees to subtrees in the global namespace, resulting in 
+custom fit semantics for application. We show how applications can co-exist and
 perform well in a global namespace and our prototype enables studies that
@@
 %applications or move data between these systems.  
-Cudele prompts many avenues for future work.  First is dynamically changing
+Cudele prompts many avenues for future work.  First is to co-locate HPC
+workflows with real highly parallel runtimes from the cloud in the same
+namespace. This setup shows how Cudele reasonably incorporates both programming
+models (client-driven parallelism and user-defined workflows) at the same time
+and should show large performance gains.  Second is dynamically changing
 semantics of a subtree from stronger to weaker guarantees (or vice versa).  For
@@
 change the semantics of the HDFS subtree into a CephFS subtree, which may cause
-metadata/data movement to make ensure strong consistency. Second is embeddable
+metadata/data movement to ensure strong consistency. Third is embeddable
 policies, where child subtrees have specialized features but still maintain

### 2018-01-07 experiments/baseline-creates1/run.sh

@@ -1,74 +1,0 @@
-#!/bin/bash
-# This file should contain the series of steps that are required to execute 
-# the experiment. Any non-zero exit code will be interpreted as a failure
-# by the 'popper check' command.
-set -ex
-
-# if you know Ansible and Docker, the below should make sense
-# - we attach ceph-ansible to root because they expect us to be in that dir
-OUTP="results-multipleruns"
-SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
-ROOT=`dirname $PWD | xargs dirname`
-NETW="--net host -v $HOME/.ssh:/root/.ssh"
-DIRS="-v `pwd`:/popper \
-      -v $ROOT/ansible/ceph:/root \
-      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
-      -w /root "
-ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
-      -v `pwd`/hosts:/etc/ansible/hosts \
-      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
-      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
-CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
-      -v `pwd`/ansible/monitor.yml:/root/monitor.yml \
-      -v `pwd`/ansible/cleanup.yml:/root/cleanup.yml \
-      -v `pwd`/ansible/collect.yml:/root/collect.yml"
-WORK="-v `pwd`/ansible/workloads:/workloads"
-ARGS="--forks 50 --skip-tags package-install,with_pkg"
-VARS="-e @/popper/vars.yml \
-      -e @/popper/ansible/vars.yml \
-      -i /etc/ansible/hosts"
-DOCKER="docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK michaelsevilla/ansible $ARGS $VARS"
-
-# debug mode
-if [ ! -z $1 ]; then
-  docker run -it --rm $NETW $DIRS $ANSB $CODE $WORK --entrypoint=ansible michaelsevilla/ansible $VARS $@
-  exit
-fi
-
-mkdir $OUTP || true
-for procs in 1 5 10 15; do
-  for clients in 1; do
-    cp configs_$SITE/clients$clients hosts
-    for job in "creates"; do
-      for log in "nolog" "log1" "log10" "log30" "log50"; do
[…]

### 2018-01-07 experiments/baseline-durability-initial/run.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+ROOT=`dirname $PWD | xargs dirname`
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root -v $ROOT/ansible/srl:/popper/ansible/roles/srl/"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+for nclients in 1 5 10 15 18 20; do
+    cp inventory_cloudlab/${nclients}client site/hosts
+    cp site/hosts $ROOT/ansible/ceph/hosts
+  for site in "nojournal-cache" "journal210-cache" "journal120-cache" "journal30-cache"; do
+
+    # configure ceph and setup results directory
+    mkdir -p results/$site/logs || true
+    cp site/* $ROOT/ansible/ceph || true
+    cp site_confs/${site}.yml site/group_vars/all
+    cp -r site/group_vars $ROOT/ansible/ceph/
+
+    # cleanup and start ceph
+    $SRL_ANSIBLE cleanup.yml
+    $CEPH_ANSIBLE ceph.yml cephfs.yml
+    $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+    
+    # warmup and get baseline
+    for i in `seq 0 2`; do
+      ./ansible-playbook.sh -e site=$site -e nfiles=100000 ../workloads/creates.yml
+    done
+    
+    ./ansible-playbook.sh -e site=$site collect.yml
+  done
+  mv results results-${nclients}clients-g1-10G
+done

### 2018-01-07 experiments/baseline-durability/run.sh

@@ -1,36 +1,74 @@
 #!/bin/bash
-
+# This file should contain the series of steps that are required to execute 
+# the experiment. Any non-zero exit code will be interpreted as a failure
+# by the 'popper check' command.
 set -ex
 
-# setup the docker command
+# if you know Ansible and Docker, the below should make sense
+# - we attach ceph-ansible to root because they expect us to be in that dir
+OUTP="results-multipleruns"
+SITE=`cat vars.yml | grep "site: " | grep -v "#" | awk '{print $2}'`
 ROOT=`dirname $PWD | xargs dirname`
-RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root -v $ROOT/ansible/srl:/popper/ansible/roles/srl/"
-ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
-CEPH_ANSIBLE="$RUN -v $ROOT/ansible/ceph:/root $ANSIBLE"
-SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+NETW="--net host -v $HOME/.ssh:/root/.ssh"
+DIRS="-v `pwd`:/popper \
+      -v $ROOT/ansible/ceph:/root \
+      -v $ROOT/ansible/srl:/popper/ansible/roles/srl \
+      -w /root "
+ANSB="-v `pwd`/ansible/group_vars/:/root/group_vars \
+      -v `pwd`/hosts:/etc/ansible/hosts \
+      -v `pwd`/ansible/ansible.cfg:/etc/ansible/ansible.cfg \
+      -e ANSIBLE_CONFIG=/etc/ansible/ansible.cfg"
+CODE="-v `pwd`/ansible/ceph.yml:/root/ceph.yml \
+      -v `pwd`/ansible/monitor.yml:/root/monitor.yml \
+      -v `pwd`/ansible/cleanup.yml:/root/cleanup.yml \
+      -v `pwd`/ansible/collect.yml:/root/collect.yml"
+WORK="-v `pwd`/ansible/workloads:/workloads"
+ARGS="--forks 50 --skip-tags package-install,with_pkg"
+VARS="-e @/popper/vars.yml \
+      -e @/popper/ansible/vars.yml \
+      -i /etc/ansible/hosts"
[…]

### 2018-01-07 paper/cover_letter.tex
misc touch ups to address #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-12-29 paper/cover_letter.tex
@@
 In addition to this cover letter, we have posted a PDF online explicitly
-showing additions in \textcolor{red}{\textbf{red}} and deletions in
-\textcolor{blue}{\textbf{blue}}. Three issues were common across reviewer
+showing additions in \textcolor{blue}{\textbf{blue}} and deletions in
+\textcolor{red}{\textbf{red}}. Three issues were common across reviewer
 feedback; Reviewer 1 and 3's comments are addressed in issues II and III,
@@
 infrastructure code contains scripts to deploy the system, run experiments, and
-gather results.  This process follows the Popper
-Convention\footnote{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which
-aims to make research reproducible.  We also made the source code available
-online and a link is provided in a reference in Section~{\S}V. Finally, we add
-details about which tools and classes we modified in Section~{\S}IV to address
-Reviewer 1's questions about the framework implementation.
+gather results.  This process follows the \href{Popper
+Convention}{http://falsifiable.us/}~\cite{jimenez_popper_2016}, which aims to
+make research reproducible.  We also made the source code available online and
+a link is provided in Section~{\S}V. Finally, we add details about which tools
+and classes we modified in Section~{\S}IV to address Reviewer 1's questions
+about the framework implementation.
 

### 2018-01-07 paper/sections/5_implementation.tex
misc touch ups to address #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 metadata server -- inodes in CephFS are about 1400
-bytes~\cite{docs:cephinternalds}.  The inode cache has code for manipulating
-inode numbers, such as pre-allocating inodes to clients. \textbf{In
-Cudele},Nonvolatile Apply uses the internal inode cache code to allocate inodes
+bytes~\cite{docs:cephinternals}.  The inode cache has code for manipulating
+inode numbers, such as pre-allocating inodes to clients. \textbf{ In
+Cudele}, Nonvolatile Apply uses the internal inode cache code to allocate inodes
 to clients that decouple parts of the namespace and to skip inodes used by the

### 2018-01-08 paper/sections/0_abstract.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 unnecessary -- for their applications.  Unfortunately, designing file systems
-with weaker consistency or durability excludes applications that rely on
+with weaker consistency or durability semantics excludes applications that rely on
 stronger guarantees, forcing developers to re-write their applications or

### 2018-01-08 paper/sections/1_introduction.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 semantics in the file system because weaker guarantees are sufficient for their
-applications. For example, many batch style jobs do not need the strong
+applications. For example, many HPC batch jobs do not need the strong
 consistency that the file system provides, so
@@
 applications optimized with techniques from different state-of-the-art
-architectures.  The HDFS subtree has weaker than strong consistency because it
-lets clients read files opened for
-writing~\cite{hakimzadeh:dais14-hdfs-consistency}, which means that not all
-updates are immediately seen by all clients; the BatchFS and RAMDisk subtrees
-are decoupled from the global namespace and have similar consistency/durability
-behavior to those systems; and the POSIX IO subtree retains the rigidity of
-POSIX IO's strong consistency.  Subtrees without policies inherit the
-consistency/durability semantics of the parent and future work will examine
-embeddable or inheritable policies.
+architectures.  The BatchFS and RAMDisk subtrees are decoupled from the global
+namespace and have similar consistency/durability behavior to those systems;
+the HDFS subtree has weaker than strong consistency because it lets clients
+read files opened for writing~\cite{hakimzadeh:dais14-hdfs-consistency}, which
+means that not all updates are immediately seen by all clients; and the POSIX
+IO subtree retains the rigidity of POSIX IO's strong consistency.  Subtrees
+without policies inherit the consistency/durability semantics of the parent and
+future work will examine embeddable or inheritable policies.
 
@@
 % Results
-The results in this paper pave the way for such a system and confirm the
-assertions of ``clean-state" research of decoupled namespaces; specifically
-that these techniques drastically improve performance.  We go a step further by
-quantifying the costs of merging updates (3.37\(\times\) slowdown) and
-maintaining durability (\(2.4\times\) slowdown).  In our prototype,
-\newcomment{we also show how to assign subtree semantics for certain
-applications such as checkpoint-restart (\(91.7\times\) speedup if consistency
-is fully relaxed), user home directories (within a 0.03 standard deviation from
-optimal), and end-users checking for partial results (only a 2\% overhead).}
-\oldcomment{we get an 8\(\times\) speedup and can scale to twice as many
-clients when we assign a more relaxed form of consistency and durability to a
-subtree with a create-heavy workload.}  We use Ceph as a prototyping platform
-because it is used in cloud-based and data center systems and has a presence in
-HPC~\cite{wang:pdsw13-ceph-hpc}.
+The results in this paper confirm the assertions of ``clean-state" research of
+decoupled namespaces; specifically that these techniques drastically improve
+performance.  We go a step further by quantifying the costs of merging updates
+(3.37\(\times\) slowdown) and maintaining durability (\(2.4\times\) slowdown).
+In our prototype, \newcomment{we also show how to assign subtree semantics for
+certain applications such as checkpoint-restart (\(91.7\times\) speedup if
+consistency is fully relaxed), user home directories (within a 0.03 standard
+deviation from optimal), and end-users checking for partial results (only a 2\%
+overhead).} \oldcomment{we get an 8\(\times\) speedup and can scale to twice as
+many clients when we assign a more relaxed form of consistency and durability
+to a subtree with a create-heavy workload.}  We use Ceph as a prototyping
+platform because it is used in cloud-based and data center systems and has a
+presence in HPC~\cite{wang:pdsw13-ceph-hpc}.
 

### 2018-01-08 paper/sections/3_posix_overheads.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 is a system administrator that configures the storage, and ``end-users"
-interact with the file system, via home directories or runtimes.} To show how
+interact with the file system via home directories or runtimes.} To show how
 the file system behaves under high metadata load we use a create-heavy
@@
 
-Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
-the Linux kernel in a CephFS mount.  The \texttt{untar} phase, which is
+%The RPCs are also serialized because the
+%metadata server is single threaded; although na\"{\i}ve, this design is common
+%because of the complexity of multi-threaded metadata
+%servers~\cite{konstantinos:pdsw2014-lustre-metadata}.  
+In this section we quantify the costs of consistency and durability in CephFS.
+At the end of each subsection we compare the approach to ``decoupled
+namespaces", the technique in related work that detaches subtrees from the
+global namespace to relax consistency/durability guarantees. We use a
+create-heavy workload for this study because it has high resource utilization,
+as shown by the trace of compiling the Linux kernel in a CephFS mount in
+Figure~\ref{fig:overhead-creates}.  The \texttt{untar} phase, which is
 characterized by many creates, has the highest resource usage (combined CPU,
-network, and disk) because of the number of RPCs needed for consistency and
-durability. The RPCs are also serialized because the metadata server is single
-threaded; although na\"{\i}ve, this design is common because of the complexity
-of multi-threaded metadata
-servers~\cite{konstantinos:pdsw2014-lustre-metadata}.  In this section, we
-quantify the costs of consistency and durability in CephFS.  At the end of each
-subsection we compare the approach to ``decoupled namespaces", the technique in
-related work that detaches subtrees from the global namespace to relax
-consistency/durability guarantees.
+network, and disk) on the metadata server because of the number of RPCs needed
+for consistency and durability.
 
@@
 sizes, which are streamed into the object store for fault tolerance. (b) and
-(c) show that} when another client ``interferes", capabilities are revoked and
+(c) show that} when another client interferes, capabilities are revoked and
 metadata servers do more work.} \label{fig:overhead}
@@
 at once).  Unless the journal saturates memory or CPU resources, larger values
-for these tunables results in better performance.
+for these tunables result in better performance.
 
@@
 \oldcomment{segment}\newcomment{dispatch} size uses more space for buffering.
-When journaling is on, the metadata server periodically stops serving requests
+\oldcomment{When journaling is on, the metadata server periodically stops serving requests
 to flush ({\it i.e.} apply journal updates) to the metadata store.  The journal
 overhead is sufficient enough to slow down metadata throughput but not so much
-as to overwhelm the bandwidth of the object store.
+as to overwhelm the bandwidth of the object store.}
 
@@
 
-\textbf{CephFS Design}: Capabilities keep metadata strongly
-consistent. To reduce the number of RPCs needed for consistency, clients can
-obtain capabilities for reading and writing inodes, as well as caching reads,
-buffering writes, changing the file size, and performing lazy IO.
+% inode cache - reduces RPCs (lookups for create, readdirs for stats)
 
-% inode cache - reduces RPCs (lookups for create, readdirs for stats)
-To keep track of the read caching and write buffering capabilities, the clients
-and metadata servers agree on the state of each inode using an inode cache.  If
-a client has the directory inode cached it can do metadata writes ({\it e.g.},
+\textbf{CephFS Design}: Capabilities keep metadata strongly consistent. To
+reduce the number of RPCs needed for consistency, clients can obtain
+capabilities for reading and writing inodes, as well as caching reads,
+buffering writes, changing the file size, and performing lazy IO.  To keep
+track of the read caching and write buffering capabilities, the clients and
+metadata servers agree on the state of each inode using an inode cache.  If a
+client has the directory inode cached it can do metadata writes ({\it e.g.},
 create) with a single RPC. If the client is not caching the directory inode
@@
 different runs.  If only one client is creating files in a directory
-(``isolated" curve on \(y1\) axis) then that client can lookup the existence of
+(``no interference" curve on \(y1\) axis) then that client can lookup the existence of
 new files locally before issuing a create request to the metadata server. If
@@
 inode transitions out of read caching and the first client must send
-\texttt{lookup()}s to the metadata server (``interfere" curve on \(y2\) axis).
-These extra requests increase the throughput of the ``interfere" curve on the
+\texttt{lookup()}s to the metadata server (``interference" curve on \(y2\) axis).
+These extra requests increase the throughput of the ``interference" curve on the
 \(y1\) axis because the metadata server can handle the extra load but

### 2018-01-08 paper/sections/4_methodology.tex
paper: misc touch ups to address #40 and #48
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \caption{Illustration of the \textbf{mechanisms} used by applications to build
-consistency/durability semantics. Descriptions are provided in by the
+consistency/durability semantics. Descriptions are provided by the
 underlined words in Section~\S\ref{sec:the-cudelesfs-mechanisms}.
@@
 abstraction and basic building block for constructing consistency and
-durability guarantees. Cudele exposes these mechanisms and the
-\oldcomment{users}\newcomment{administrator} composes them together to
+durability guarantees.\oldcomment{Cudele exposes these mechanisms and the
+users}\newcomment{ The administrator} composes these mechanisms together to
 construct \textbf{policies}.  These policies are assigned to subtrees and they
@@
 \subsubsection{Mechanisms Used for Durability} \underline{Stream} is one of the
-mechanisms used by default in CephFS.  Using existing configuration settings in
-Ceph we can turn Stream on and off.  If it is off, then the metadata servers
-will not save journals in the object store. For \underline{Local Persist}, clients write
-serialized log events to a file on local disk and for \underline{Global Persist},
-clients push the journal into the object store. The overheads for both Local
-Persist and Global Persist is the write bandwidth of the local disk and
-object store, respectively.  These persist mechanisms are part of the library that
-links into the client.
+mechanisms used by default in CephFS.  Using existing configuration settings we
+can turn Stream on and off.  If it is off, then the metadata servers will not
+save journals in the object store. For \underline{Local Persist}, clients write
+serialized log events to a file on local disk and for \underline{Global
+Persist}, clients push the journal into the object store. The overheads for
+both Local Persist and Global Persist is the write bandwidth of the local disk
+and object store, respectively.  These persist mechanisms are part of the
+library that links into the client.
 
@@
 Existing, state-of-the-art systems in HPC can be represented by the cells in
-Table~\ref{table:spectrum} (we construct the semantics for these systems later
-in Section~\S\ref{sec:microbenchmarks}).  POSIX IO-compliant systems like
-CephFS and IndexFS have global consistency and durability\footnote{ IndexFS
-also has bulk merge which is a form of ``weak consistency"}; DeltaFS uses
-``invisible" consistency and ``local" durability and BatchFS uses ``weak"
-consistency and ``local" durability. These systems have other features that
-could push them into different semantics but we assign labels here based on the
-points emphasized in the papers.  To compose the mechanisms
+Table~\ref{table:spectrum} POSIX IO-compliant systems like CephFS and IndexFS
+have global consistency and durability\footnote{ IndexFS also has bulk merge
+which is a form of ``weak consistency"}; DeltaFS uses ``invisible" consistency
+and ``local" durability and BatchFS uses ``weak" consistency and ``local"
+durability. These systems have other features that could push them into
+different semantics but we assign labels here based on the points emphasized in
+the papers.  To compose the mechanisms
 \oldcomment{users}\newcomment{administrators} inject which mechanisms to run
@@
 mechanism and the semantics described above are quantified in
-Sections~\S\ref{sec:microbenchmarks} and~\S\ref{sec:use-case-1}.
+Sections~\S\ref{sec:microbenchmarks}.
 
 % talk of eventual consistency
-The consistency and durability properties in Table~\ref{table:spectrum} are not
-guaranteed until all mechanisms in the cell are complete. The compositions
-should be considered atomic and there are no guarantees while transitioning
-between policies. For example, updates are not deemed to have ``global"
-durability until they are safely saved in the object store. If a failure occurs
-during Global Persist or if we inject a new policy that changes a subtree from
-Local Persist to Global Persist, Cudele makes no guarantee until the mechanisms
-are complete. Despite this, production systems that use Cudele should state
-up-front what the transition guarantees are for subtrees.
+In our prototype, the consistency and durability properties in
+Table~\ref{table:spectrum} are not guaranteed until all mechanisms in the cell
+are complete. The compositions should be considered atomic and there are no
+guarantees while transitioning between policies. For example, updates are not
+deemed to have ``global" durability until they are safely saved in the object
+store. If a failure occurs during Global Persist or if we inject a new policy
+that changes a subtree from Local Persist to Global Persist, Cudele makes no
+guarantee until the mechanisms are complete. Despite this, production systems
+that use Cudele should state up-front what the transition guarantees are for
+subtrees. This is not a limitation of our approach; it just lead to the
+simplest implementation.
 
@@
 following parameters (default values are in parenthesis): which consistency
-model to use (\texttt{RPCs}), which durability model to use (\texttt{stream})
-number of inodes to provision to the decoupled namespace (100), which interfere
+model to use (\texttt{RPCs}), which durability model to use (\texttt{stream}), 
+number of inodes to provision to the decoupled namespace (100), and which interfere
 policy to use, {\it i.e.} how to handle a request from another client
@@
 give valid inodes to other clients. The inodes can be used anywhere within the
-decoupled namespace ({i.e.} at any depth in the subtree).  ``Interfere Policy"
+decoupled namespace ({\it i.e.} at any depth in the subtree).  ``Interfere Policy"
 has two settings: \texttt{block} and \texttt{allow}.  For \texttt{block}, any

### 2018-01-08 paper/sections/5_implementation.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 We use a programmable storage approach~\cite{sevilla:eurosys17} to design
-Cudele; namely, we try to leverage components inside Ceph to inherit the
+Cudele; namely, we try to leverage components inside CephFS to inherit the
 robustness and correctness of the internal subsystems. Using this `dirty-slate'
@@
 Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
-system itself.  In this section, we first describe a Ceph internal subsystem or
+system itself.  In this section, we first describe a CephFS internal subsystem or
 component and then we show how we use it in Cudele.
@@
 metadata store to service requests. Using code designed for recovery,
-Nonvolatile Apply and Volatile Apply replay updates onto the metadata store in
+Volatile Apply and Nonvolatile Apply replay updates onto the metadata store in
 memory and in the object store, respectively.  When the clients are ready to
@@
 the ability to recursively apply policies to subtrees and to isolate subtrees
-from each other. \textbf{In Cudele} the large inodes store consistency and
-durability policies in the directory inode. This approach uses the File Type
-interface from the Malacology programmable store
+from each other. \textbf{In Cudele} the large inodes also store consistency and
+durability policies. This approach uses the File Type
+interface from the Malacology programmable storage
 system~\cite{sevilla:eurosys17} and it tells clients how to access the

### 2018-01-08 paper/sections/6_results.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-01-07 paper/sections/6_results.tex
@@
 mechanisms we propose are useful for constructing semantics used by real
-systems, (2) the techniques can work side-by-side in the same
-namespace\oldcomment{, and (3) the combination of these techniques can help the
+systems and (2) the techniques can work side-by-side in the same
+namespace for common use cases\oldcomment{, and (3) the combination of these techniques can help the
 system scale further when subtrees are coupled to the correct type of
@@
 and scalable in systems that specialize an entire namespace according to a
-single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS), so our
-evaluation explores the behavior of the consistency/durability mechanisms and
-their effect with and without isolation.}
+single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS).}
+%, so our
+%evaluation explores the behavior of the consistency/durability mechanisms and
+%their effect with and without isolation.}
 
@@
 it takes to write \oldcomment{100K file create} updates to the client's in-memory journal
-({\it i.e.} the Append Client Journal mechanism).  \newcomment{The first graph
+({\it i.e.} the Append Client Journal mechanism), which is about 11K creates/sec.  \newcomment{The first graph
 groups the consistency mechanisms, the second groups the durability mechanisms,
@@
 the experiment directory and the object that contains the root directory ({\it
-i.e.}, \texttt{/}).  \oldcomment{The cost of communicating through the object
+i.e.} \texttt{/}).  \oldcomment{The cost of communicating through the object
 store is shown by comparing the runtime of Volatile Apply + Global Persist to
@@
 {\it Parallelism of the Object Store:}
-\newcomment{Stream, which is an approximation (journal on minus journal off) has
+\newcomment{Stream, which is an approximation (journal on minus journal off), has
 the highest slowdown at \(2.4\times\) because the overhead of maintaining and
@@
 servers fail during a mechanism, metadata or data may be lost. This graph shows
-how we can build application-specific subtrees by composing mechanisms and that
-coupling well-established techniques to specific applications results in a more
-scalable global namespace.}
+how we can build application-specific subtrees by composing mechanisms and the performance of
+coupling well-established techniques to specific applications over the same file system.}
+
+%results in a more
+%scalable global namespace.}
 
@@
 open/create throughput of the underlying file system~\cite{xiao:socc15-shardfs,
-shi:vldb15-spark, shvachko:login2012-hdfs-scalability}; Spark queries have an
-O(\(n\)) complexity, where \(n\) is the number of partitions.  Big Data
-Benchmark jobs examined in~\cite{chaimov:hpdc16-spark} have on the order of
-15,000 file opens or creates just to start a single Spark query and the Lustre
-system they tested on did not handle creates well, showing up to a \(24\times\)
-slowdown compared to other metadata operations. Common approaches to solve
-these types of bottlenecks is to change the application behavior or to design a
-new file system, like BatchFS or DeltaFS, that uses one set of metadata
-optimizations for the entire namespace.} \oldcomment{ abstraction to exchange
-work units to workers or to indicate when jobs complete  The workload is
-clients creating 100K files in private directories in the same global
-namespace.}
+shi:vldb15-spark, shvachko:login2012-hdfs-scalability}; Big Data Benchmark jobs
+examined in~\cite{chaimov:hpdc16-spark} have on the order of 15,000 file opens
+or creates just to start a single Spark query and the Lustre system they tested
+on did not handle creates well, showing up to a \(24\times\) slowdown compared
+to other metadata operations. Common approaches to solve these types of
+bottlenecks is to change the application behavior or to design a new file
+system, like BatchFS or DeltaFS, that uses one set of metadata optimizations
+for the entire namespace.} \oldcomment{ abstraction to exchange work units to
+workers or to indicate when jobs complete  The workload is clients creating
+100K files in private directories in the same global namespace.}
 
@@
 Figure~\ref{fig:overhead-b}. We configure one subtree with an interfere policy
-of ``allow" and another with subtree with ``block" so \texttt{-EBUSY} is
+of ``allow" and another subtree with ``block" so \texttt{-EBUSY} is
 returned to interfering clients.  The former is the default behavior in file
@@
 directory, respectively. The goal is to explicitly isolate clients so that
-performance is similar to the ``no interfere" curve, which has lower slowdowns
-(on average, \(1.42\times\) per client compared to \(1.67\times\) per client
-for ``interfere") and less variability (on average, a standard deviation of
-0.06 compared to 0.44 for ``interfere").  ``Block interference" uses the Cudele
-API to block interfering clients and the slowdown (\(1.34\times\) per client)
-and variability (0.09) look very similar to ``no interfere" for larger
-clusters. For smaller clusters the overhead to reject requests is more evident
-when the metadata server is underloaded so the slowdowns are similar to
-``interfere".  Clients can block interfering clients to get the same
-performance as isolated clients but there is a non-negligible overhead for
+performance is similar to the ``no interference" curve, which has lower
+slowdowns (on average, \(1.42\times\) per client compared to \(1.67\times\) per
+client for ``interfere") and less variability (on average, a standard deviation
+of 0.06 compared to 0.44 for ``interfere").  ``Block interference" uses the
+Cudele API to block interfering clients and the slowdown (\(1.34\times\) per
+client) and variability (0.09) look very similar to ``no interference" for a
+larger number of clients. For smaller clusters the overhead to reject requests
+is more evident when the metadata server is underloaded so the slowdowns are
+similar to ``interference".  We conclude that administrators can block interfering clients to get the same
+performance as isolated scenarios but there is a non-negligible overhead for
 rejecting requests when the metadata server is not operating at peak
@@
 \oldcomment{users}\newcomment{administrators} fine-grained control of the
-consistency semantics \newcomment{to support current practices and experimental
-workflows}. \oldcomment{The use case is that u}\newcomment{U}sers often
+consistency semantics \newcomment{to support current practices and scientific
+workflows in HPC}. \oldcomment{The use case is that u}\newcomment{U}sers often
 leverage the file system to check the progress of jobs using \texttt{ls} even
@@
 eshel:fast10-panache}. The number of files or size of the files is indicative
-of the progress. This practice is not too different from systems that use the
-file system to manage the progress of jobs; Spark and Hadoop write to temporary
+of the progress. This practice is not too different from cloud systems that use the
+file system to manage the progress of jobs; Spark/Hadoop writes to temporary
 files, renames them when complete, and creates a ``DONE" file to indicate to
 the runtime that the task did not fail and should not be re-scheduled on
-another node. \newcomment{Also, the browser interface lets Hadoop/Spark users check the
+another node. \newcomment{Also, the browser interface lets Hadoop/Spark users check
 progress by querying the file system and returning a \% of job complete
@@
 (up to 9\% overhead if done every second). The optimal sync interval for
-performance is 10 seconds, which only incurs 2\% overhead), because larger
+performance is 10 seconds, which only incurs 2\% overhead, because larger
 intervals must write more updates to disk and network. For the 25 second

### 2018-01-08 paper/sections/7_conclusion.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Relaxing consistency/durability semantics in file systems is a double-edged
-sword. While it performs and scales better, it alienates applications that rely
+sword. While the technique performs and scales better, it alienates applications that rely
 on strong consistency and durability.  Cudele lets
@@
 guarantees to subtrees in the global namespace, resulting in 
-custom fit semantics for application. We show how applications can co-exist and
+custom fit semantics for applications. We show how applications can co-exist and
 perform well in a global namespace and our prototype enables studies that
@@
 and should show large performance gains.  Second is dynamically changing
-semantics of a subtree from stronger to weaker guarantees (or vice versa).  For
-example, the results of a Hadoop job do not need to be migrated into CephFS for
+semantics of a subtree from stronger to weaker guarantees (or vice versa). This
+reduces data movement across storage cluster and file system boundaries so
+the results of a Hadoop job do not need to be migrated into CephFS for
 other processing; instead the \oldcomment{user}\newcomment{administrator} can
@@
 IO-compliant but relaxes durability constraints, so it can reside under a POSIX
-IO subtree that is strongly consistent.
+IO subtree alongside a globally durable subtree.
 

### 2018-01-08 paper/cover_letter.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: typos and #40
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-01-07 paper/cover_letter.tex
@@
 \begin{document}
-\title{Cover Letter: Cudele}
+\title{Cover Letter: Cudele (Paper 306)}
 \maketitle
@@
 
-For Reviewer 1, we fixed user vs. client terminology in Section~{\S}III,
+For Reviewer 1, we fixed user/client terminology in Section~{\S}III,
 clarified that the cost of dynamically changing semantics is future work in
 Sections~{\S}I and~{\S}III, and added source code pointers in Section~{\S}V.
-For Reviewer 2, we removed the major takeways headings and cross-references in
+For Reviewer 2, we removed the major takeaways headings and cross-references in
 Section~{\S}V. For Reviewer 3, we quantified speedups with figure annotations

### 2018-01-08 paper/sections/6_results.tex
paper: typos and #40
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the metadata server at the same time. \newcomment{Had we added infrastructure
-to overlay journal arrivals or time client sync intervales, we could have
+to overlay journal arrivals or time client sync intervals, we could have
 scaled more closely to ``decoupled: create".}\oldcomment{The ``create + merge"

### 2018-01-09 paper/cover_letter.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 reviewer feedback; Reviewer 1 and 3's comments are addressed in issues II and
-III, Reviewer 2's comments are addressed in issues I and III, and Reviewer 4's
+III, Reviewer 2's comments are addressed in all issues, and Reviewer 4's
 comments are addressed in issues I and II.
@@
 and Figures 6a, 6b, and 6c to emphasize that decoupled namespaces facilitate
-client-driven parallelism, where clients detach subtrees form the global namespaces
+client-driven parallelism, where clients detach subtrees form the global namespace
 and do operations in-parallel to their local disk. We also hope that changes to
@@
 
-  \item explicitly state Cudele setups and subtree configurations in bold in
-Section~{\S}V-B to guide the reader through use cases and the effects of
-different Cudele API configurations.
+%  \item explicitly state Cudele setups and subtree configurations in bold in
+%Section~{\S}V-B to guide the reader through use cases and the effects of
+%different Cudele API configurations.
 
@@
 
-  \item clarified confusing terminology in Section~{\S}II: administrator refers
+  \item clarified confusing terminology in Section~{\S}I: administrator refers
 to a person that configures subtrees and client refers to a storage client or
@@
 We apologize for the omission of experimental setup and source code details.
-We made artifacts available online for all experiments in the paper. For the
-experimental setup, we describe the cluster setup (hardware, software, etc.) in
-Section~{\S}V.  We have also made the infrastructure code available and added
-links after each figure to show exactly how experiments are run.  This
-infrastructure code contains scripts to deploy the system, run experiments, and
-gather results.  This process follows the \href{http://falsifiable.us/}{Popper
-Convention}, which aims to make research reproducible.  We also made the source
-code available online and a link is provided in Section~{\S}V. Finally, we add
-details about which tools and classes we modified in Section~{\S}IV to address
-Reviewer 1's questions about the framework implementation.
+For the experimental setup, we describe the cluster setup (hardware, software,
+etc.) in Section~{\S}V.  We have also made the infrastructure code available
+and added links after each figure to show exactly how experiments are run.
+This infrastructure code contains scripts to deploy the system, run
+experiments, and gather results.  This process follows the
+\href{http://falsifiable.us/}{Popper Convention}, which aims to make research
+reproducible.  We also made the source code available online and a link is
+provided in Section~{\S}V. Finally, we add details about which tools and
+code bases we modified in Section~{\S}IV to address Reviewer 1's questions about
+the framework implementation.
 
@@
 Sections~{\S}I and~{\S}III, and added source code pointers in Section~{\S}V.
-For Reviewer 2, we removed the major takeaways headings and cross-references in
+For Reviewer 2, we removed the major takeaway headings and cross-references in
 Section~{\S}V. For Reviewer 3, we quantified speedups with figure annotations
 in Section~{\S}V. For Reviewers 1 and 3, we added servers, network, and storage
-setups in Section~{\S}V. For Reviewer 4 we added descriptions to each use case
+setups in Section~{\S}V. For Reviewer 4, we added descriptions to each use case
 in Section~{\S}V describing how Cudele would work with Spark.

### 2018-01-09 paper/sections/0_abstract.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 unnecessary -- for their applications.  Unfortunately, designing file systems
-with weaker consistency or durability semantics excludes applications that rely on
-stronger guarantees, forcing developers to re-write their applications or
+with weaker consistency or durability semantics excludes applications that rely
+on stronger guarantees, forcing developers to re-write their applications or
 deploy them on a different system.  We present a framework and API that lets
@@
 (0.03 standard deviation from optimal), and users checking for partial results
-(2\% \oldcomment{clients}\newcomment{users} overhead).} \oldcomment{ and can
-scale to 2\(\times\) as many clients when compared to our baseline system.}
+(2\% overhead).} \oldcomment{ and can scale to 2\(\times\) as many clients when
+compared to our baseline system.}
 

### 2018-01-09 paper/sections/1_introduction.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
-metadata server for every \newcomment{client} is unreasonable. This problem is
-exacerbated by current hardware and software trends: parallel runtimes like
-Hadoop and Spark use storage systems to implement important scheduling and
-timing semantics, and  \oldcomment{; for example, } HPC architectures are
-transitioning from complex storage stacks with burst buffer, file system,
-object store, and tape tiers to more simplified stacks with just a burst buffer
-and object store~\cite{bent:login16-hpc-trends}. These types of trends put
-pressure on data access because more requests from different nodes end up
-hitting the same layer in parallel and latencies cannot be hidden while data
-migrates across tiers.
+metadata server for every \newcomment{client}\footnote{In this paper, ``client"
+is a storage client or application that interacts with the metadata server,
+``administrator" is a system administrator that configures the storage, and
+``end-users" interact with the file system via home directories or runtimes.  }
+is unreasonable. This problem is exacerbated by current hardware and software
+trends: parallel runtimes like Hadoop and Spark use storage systems to
+implement important scheduling and timing semantics and  \oldcomment{; for
+example, } HPC architectures are transitioning from complex storage stacks with
+burst buffer, file system, object store, and tape tiers to more simplified
+stacks with just a burst buffer and object
+store~\cite{bent:login16-hpc-trends}. These types of trends put pressure on
+data access because more requests from different nodes end up hitting the same
+layers in parallel\oldcomment{and latencies cannot be hidden while data
+migrates across tiers}.
 
@@
 bent:slides-twotiers}. Then the file system can change its internal structure
-to optimize performance. For example, the file system could enter a mode that
-prevents other clients from interfering with the decoupled directory.  This
-delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
-durability improves performance and scalability by avoiding the costs of remote
-procedure calls (RPCs), synchronization, false sharing, and serialization.
-While the performance benefits of decoupling the namespace are obvious,
-applications that rely on the file system's guarantees must be deployed on an
-entirely different system or re-written to coordinate strong
-consistency/durability themselves.
+to optimize performance. For example, the file system could enter a mode where
+clients with decoupled directories perform operations locally and bulk merge their
+updates at completion. This delayed merge ({\it i.e.} a form of eventual
+consistency) and relaxed durability improves performance and scalability by
+avoiding the costs of remote procedure calls (RPCs), synchronization, false
+sharing, and serialization.  While the performance benefits of decoupling the
+namespace are obvious, applications that rely on the file system's guarantees
+must be deployed on an entirely different system or re-written to coordinate
+strong consistency/durability themselves.
 
@@
   \oldcomment{This framework lets subtrees with different semantics co-exist in
-a global namespace. We show how this} \newcomment{We show how that letting
+a global namespace. We show how this} \newcomment{We show that letting
 different semantics co-exist in a global namespace scales further and performs
@@
 % Results
-The results in this paper confirm the assertions of ``clean-state" research of
+The results in this paper confirm the assertions of ``clean-slate" research of
 decoupled namespaces; specifically that these techniques drastically improve
-performance.  We go a step further by quantifying the costs of merging updates
-(3.37\(\times\) slowdown) and maintaining durability (\(2.4\times\) slowdown).
-In our prototype, \newcomment{we also show how to assign subtree semantics for
-certain applications such as checkpoint-restart (\(91.7\times\) speedup if
-consistency is fully relaxed), user home directories (within a 0.03 standard
-deviation from optimal), and end-users checking for partial results (only a 2\%
-overhead).} \oldcomment{we get an 8\(\times\) speedup and can scale to twice as
-many clients when we assign a more relaxed form of consistency and durability
-to a subtree with a create-heavy workload.}  We use Ceph as a prototyping
-platform because it is used in cloud-based and data center systems and has a
-presence in HPC~\cite{wang:pdsw13-ceph-hpc}.
+performance.  We go a step further by quantifying the costs of traditional file
+system approaches to maintaining consistency (3.37\(\times\) slowdown) and
+durability (\(2.4\times\) slowdown).  In our prototype, \newcomment{we also
+show the benefits of assigning subtree semantics to certain applications such
+as checkpoint-restart (\(91.7\times\) speedup if consistency is fully relaxed),
+user home directories (within a 0.03 standard deviation from optimal), and
+end-users checking for partial results (only a 2\% overhead).} \oldcomment{we
+get an 8\(\times\) speedup and can scale to twice as many clients when we
+assign a more relaxed form of consistency and durability to a subtree with a
+create-heavy workload.}  We use Ceph as a prototyping platform because it is
+used in cloud-based and data center systems and has a presence in
+HPC~\cite{wang:pdsw13-ceph-hpc}.
 

### 2018-01-09 paper/sections/3_posix_overheads.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \texttt{untar}) incur the highest disk, network, and CPU utilization because of
-consistency/durabiliy demands.}\label{fig:overhead-creates}
+consistency/durability demands.}\label{fig:overhead-creates}
 \end{figure}
@@
 In our examination of the overheads of POSIX IO we benchmark and analyze
-CephFS, the file system that uses Ceph's object store ({\it i.e.} RADOS) to
-store its data and metadata. \newcomment{In this paper, ``client" is a storage
-client or application that interacts with the metadata server, ``administrator"
-is a system administrator that configures the storage, and ``end-users"
-interact with the file system via home directories or runtimes.} To show how
-the file system behaves under high metadata load we use a create-heavy
-workload.  During this process we discovered, based on the analysis and
+CephFS, the file system that uses Ceph's object store (RADOS) to
+store its data/metadata and a metadata server cluster to service client requests
+more quickly.  During this process we discovered, based on the analysis and
 breakdown of costs, that durability and consistency have high overhead but we
@@
 one set of design decisions and our goal here is to highlight the effect that
-those decisions have on performance.
+those decisions have on performance.  At the end of each subsection we compare
+the approach to ``decoupled namespaces", the technique in related work that
+detaches subtrees from the global namespace to relax consistency/durability
+guarantees.
 
@@
 %servers~\cite{konstantinos:pdsw2014-lustre-metadata}.
-In this section we quantify the costs of consistency and durability in CephFS.
-At the end of each subsection we compare the approach to ``decoupled
-namespaces", the technique in related work that detaches subtrees from the
-global namespace to relax consistency/durability guarantees. We use a
-create-heavy workload for this study because it has high resource utilization,
-as shown by the trace of compiling the Linux kernel in a CephFS mount in
-Figure~\ref{fig:overhead-creates}.  The \texttt{untar} phase, which is
+
+We use a create-heavy workload for this study because it has high resource
+utilization, as shown by the trace of compiling the Linux kernel in a CephFS
+mount in Figure~\ref{fig:overhead-creates}.  The \texttt{untar} phase, which is
 characterized by many creates, has the highest resource usage (combined CPU,
@@
 client that creates 100K files with journaling off (about 654 creates/sec). In
-this case a dispatch size of 10 degrades performance the most because the
+this case a dispatch size of 30 degrades performance the most because the
 metadata server is overloaded with requests and cannot spare cycles to manage
@@
 merge batches of metadata operations into the global namespaces when the job
-completes.  In BatchFS the merge is delayed by the application using an API to
+completes.  In BatchFS, the merge is delayed by the application using an API to
 switch between asynchronous and synchronous mode. The merge itself is explicitly
 managed by the application but future work looks at more automated
-methodologies. In DeltaFS snapshots of the metadata subtrees stays on the client
+methodologies. In DeltaFS, snapshots of the metadata subtrees stays on the client
 machines; there is no ground truth and consistent namespaces are constructed

### 2018-01-09 paper/sections/4_methodology.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 abstraction and basic building block for constructing consistency and
-durability guarantees.\oldcomment{Cudele exposes these mechanisms and the
-users}\newcomment{ The administrator} composes these mechanisms together to
+durability guarantees. \oldcomment{Cudele exposes these mechanisms and the
+users}\newcomment{The administrator} composes these mechanisms together to
 construct \textbf{policies}.  These policies are assigned to subtrees and they
@@
 
-\subsubsection{Mechanisms Used for Durability} \underline{Stream} is one of the
-mechanisms used by default in CephFS.  Using existing configuration settings we
-can turn Stream on and off.  If it is off, then the metadata servers will not
-save journals in the object store. For \underline{Local Persist}, clients write
-serialized log events to a file on local disk and for \underline{Global
-Persist}, clients push the journal into the object store. The overheads for
-both Local Persist and Global Persist is the write bandwidth of the local disk
-and object store, respectively.  These persist mechanisms are part of the
-library that links into the client.
+\subsubsection{Mechanisms Used for Durability} \underline{Stream}, the default
+setting in CephFS, saves a journal of metadata updates in the object store.
+Using existing configuration settings we can turn Stream on and off.  For
+\underline{Local Persist}, clients write serialized log events to a file on
+local disk and for \underline{Global Persist}, clients push the journal into
+the object store. The overheads for both Local Persist and Global Persist is
+the write bandwidth of the local disk and object store, respectively.  These
+persist mechanisms are part of the library that links into the client.
 
@@
 Existing, state-of-the-art systems in HPC can be represented by the cells in
-Table~\ref{table:spectrum} POSIX IO-compliant systems like CephFS and IndexFS
+Table~\ref{table:spectrum}. POSIX IO-compliant systems like CephFS and IndexFS
 have global consistency and durability\footnote{ IndexFS also has bulk merge
@@
 (msevilla/mydir, policies.yml) would decouple the path ``msevilla/mydir" and
-would apply the policies in ``policies.yml".  The policies file supports the
-following parameters (default values are in parenthesis): which consistency
-model to use (\texttt{RPCs}), which durability model to use (\texttt{stream}), 
-number of inodes to provision to the decoupled namespace (100), and which interfere
-policy to use, {\it i.e.} how to handle a request from another client
-targeted at this subtree (\texttt{allow}). The ``Consistency" and ``Durability"
-parameters are compositions of mechanisms; they can be serialized (\(+\)) or
-run in parallel (\(|\|\)).  ``Allocated Inodes" is a way for the application to
-specify how many files it intends to create. It is a contract so that the file
-system can provision enough resources for the incumbent merge and so it can
-give valid inodes to other clients. The inodes can be used anywhere within the
-decoupled namespace ({\it i.e.} at any depth in the subtree).  ``Interfere Policy"
-has two settings: \texttt{block} and \texttt{allow}.  For \texttt{block}, any
-requests to this part of the namespace returns with ``Device is busy", which
-will spare the metadata server from wasting resources for updates that may get
-overwritten. If the application does not mind losing updates, for example it
-wants approximations for results that take too long to compute, it can select
-\texttt{allow}. In this case, metadata from the interfering client will be
-written and the computation from the decoupled namespace will take priority at
-merge time because the results are more accurate.  Given these default values
-decoupling the namespace with an empty policies file would give the application
-100 inodes but the subtree would behave like the existing CephFS
-implementation.
+would apply the policies in ``policies.yml".
+
+The policies file supports the following parameters (default values are in
+parenthesis): which consistency model to use (\texttt{RPCs}), which durability
+model to use (\texttt{stream}), number of inodes to provision to the decoupled
+namespace (100), and which interfere policy to use, {\it i.e.} how to handle a
+request from another client targeted at this subtree (\texttt{allow}). The
+``Consistency" and ``Durability" parameters are compositions of mechanisms;
+they can be serialized (\(+\)) or run in parallel (\(||\)).  ``Allocated
+Inodes" is a way for the application to specify how many files it intends to
+create. It is a contract so that the file system can provision enough resources
+for the incumbent merge and so it can give valid inodes to other clients. The
+inodes can be used anywhere within the decoupled namespace ({\it i.e.} at any
+depth in the subtree).  ``Interfere Policy" has two settings: \texttt{block}
+and \texttt{allow}.  For \texttt{block}, any requests to this part of the
+namespace returns with ``Device is busy", which will spare the metadata server
+from wasting resources for updates that may get overwritten. If the application
+does not mind losing updates, for example it wants approximations for results
+that take too long to compute, it can select \texttt{allow}. In this case,
+metadata from the interfering client will be written and the computation from
+the decoupled namespace will take priority at merge time because the results
+are more accurate.  Given these default values decoupling the namespace with an
+empty policies file would give the application 100 inodes but the subtree would
+behave like the existing CephFS implementation.

### 2018-01-09 paper/sections/5_implementation.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Cudele; namely, we try to leverage components inside CephFS to inherit the
-robustness and correctness of the internal subsystems. Using this `dirty-slate'
+robustness and correctness of the internal subsystems. Using this ``dirty-slate"
 approach, we only had to implement 4 of the 6 mechanisms from
@@
 store data structure is structured as a tree of directory fragments making it
-easier to read and traverse. \textbf{In Cudele} the RPCs mechanism uses the
-metadata store to service requests. Using code designed for recovery,
+easier to read and traverse. \textbf{In Cudele,} the RPCs mechanism uses the
+in-memory metadata store to service requests. Using code designed for recovery,
 Volatile Apply and Nonvolatile Apply replay updates onto the metadata store in
-memory and in the object store, respectively.  When the clients are ready to
-merge their updates back into the global namespace, they pass a binary file of
-metadata updates to the metadata server.
+memory and in the object store, respectively.
+
+%When the clients are ready to
+%merge their updates back into the global namespace, they pass a binary file of
+%metadata updates to the metadata server.
 
@@
 Specifically, reads are slow because there is more state to read, it is
-unorganized, and many of the updates may be redundant.  \textbf{In Cudele} the
+unorganized, and many of the updates may be redundant.  \textbf{In Cudele,} the
 journal format is used by Stream, Append Client Journal, Local Persist, and
@@
 store objects and replays the updates on the metadata store in the object
-store.  \textbf{In Cudele} the external library the clients link into is based
+store.  \textbf{In Cudele,} the external library the clients link into is based
 on the journal tool. It already had functions for importing, exporting, and
@@
 inode numbers, such as pre-allocating inodes to clients. \textbf{ In
-Cudele}, Nonvolatile Apply uses the internal inode cache code to allocate inodes
+Cudele,} Nonvolatile Apply uses the internal inode cache code to allocate inodes
 to clients that decouple parts of the namespace and to skip inodes used by the
@@
 the ability to recursively apply policies to subtrees and to isolate subtrees
-from each other. \textbf{In Cudele} the large inodes also store consistency and
+from each other. \textbf{In Cudele,} the large inodes also store consistency and
 durability policies. This approach uses the File Type

### 2018-01-09 paper/sections/6_results.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 reproducible, this paper adheres to The Popper
-Convention~\cite{jimenez_popper_2016}, so experiments can be examined in more
+Convention~\cite{jimenez_popper_2016} so experiments can be examined in more
 detail, or even re-run, by visiting the \texttt{[source]} link next to each
-figure.  The source code is available on the Cudele branch~\cite{docs:cudele}
+figure.  The source code for Cudele is available on a branch~\cite{docs:cudele}
 of our Ceph fork.}
@@
 mechanism to process all metadata events.  Results are} normalized to the time
-it takes to write \oldcomment{100K file create} updates to the client's in-memory journal
+it takes to write \oldcomment{100K file create}updates to the client's in-memory journal
 ({\it i.e.} the Append Client Journal mechanism), which is about 11K creates/sec.  \newcomment{The first graph
@@
 returned to interfering clients.  The former is the default behavior in file
-systems and the latter isolates performance from malicious clients.}
+systems and the latter isolates performance from interfering clients.}
 
@@
 slowdowns (on average, \(1.42\times\) per client compared to \(1.67\times\) per
-client for ``interfere") and less variability (on average, a standard deviation
-of 0.06 compared to 0.44 for ``interfere").  ``Block interference" uses the
+client for ``interference") and less variability (on average, a standard deviation
+of 0.06 compared to 0.44 for ``interference").  ``Block interference" uses the
 Cudele API to block interfering clients and the slowdown (\(1.34\times\) per
@@
 the runtime that the task did not fail and should not be re-scheduled on
-another node. \newcomment{Also, the browser interface lets Hadoop/Spark users check
+another node. \newcomment{So the browser interface lets Hadoop/Spark users check
 progress by querying the file system and returning a \% of job complete

### 2018-01-09 paper/sections/7_conclusion.tex
paper: typos and revision 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 workflows with real highly parallel runtimes from the cloud in the same
-namespace. This setup shows how Cudele reasonably incorporates both programming
+namespace. This setup would show how Cudele reasonably incorporates both programming
 models (client-driven parallelism and user-defined workflows) at the same time

### 2018-01-09 paper/sections/1_introduction.tex
paper: final revision for round 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 is unreasonable. This problem is exacerbated by current hardware and software
-trends: parallel runtimes like Hadoop and Spark use storage systems to
-implement important scheduling and timing semantics and  \oldcomment{; for
-example, } HPC architectures are transitioning from complex storage stacks with
-burst buffer, file system, object store, and tape tiers to more simplified
-stacks with just a burst buffer and object
+trends; for example,  HPC architectures are transitioning from complex storage
+stacks with burst buffer, file system, object store, and tape tiers to more
+simplified stacks with just a burst buffer and object
 store~\cite{bent:login16-hpc-trends}. These types of trends put pressure on
 data access because more requests from different nodes end up hitting the same
-layers in parallel\oldcomment{and latencies cannot be hidden while data
-migrates across tiers}.
+software layers in parallel and latencies cannot be hidden while data migrates
+across tiers.
 
@@
 %semantics and their performance is explored in~\S\ref{sec:use-case-1}.} 
-To address this problem, we present an API and framework that lets developers
+To address this problem, we present an API and framework that lets administrators
 dynamically control the consistency and durability guarantees for subtrees in

### 2018-02-22 paper/sections/1_introduction.tex
paper: camera ready template and squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 global namespace.  For performance, clients \oldcomment{can} relax consistency/durability on their
-subtree ({\it e.g.}, HDFS) or decouple the subtree and move it locally ({\it e.g.}, BatchFS, RAMDisk).
+subtree ({\it e.g.}, HDFS) or decouple the subtree and move it locally ({\it e.g.}, BatchFS, RAMDisk).\vspace{-3ex}
 }\label{fig:subtree-policies} \end{figure}

### 2018-02-22 paper/sections/2_related_work.tex
paper: camera ready template and squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-% General
-The bottlenecks associated with accessing POSIX IO file system metadata are not
-limited to HPC workloads and the same challenges that plagued these systems for
-years are finding their way into the cloud. Workloads that deal with many small
-files ({\it e.g.}, log processing and database
-queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
-simultaneous clients ({\it e.g.}, MapReduce
-jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
-the metadata service. The biggest challenge is that whenever a file is touched
-the client must access the file's metadata and maintaining a file system
-namespace imposes small, frequent accesses on the underlying storage
-system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
-system metadata is a well-known problem and solutions for scaling data IO do
-not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
-abad:ucc2012-mimesis, alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. 
+%% General
+%The bottlenecks associated with accessing POSIX IO file system metadata are not
+%limited to HPC workloads and the same challenges that plagued these systems for
+%years are finding their way into the cloud. Workloads that deal with many small
+%files ({\it e.g.}, log processing and database
+%queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
+%simultaneous clients ({\it e.g.}, MapReduce
+%jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
+%the metadata service. The biggest challenge is that whenever a file is touched
+%the client must access the file's metadata and maintaining a file system
+%namespace imposes small, frequent accesses on the underlying storage
+%system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
+%system metadata is a well-known problem and solutions for scaling data IO do
+%not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
+%abad:ucc2012-mimesis, alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. 
 %There are two approaches for improving the performance of metadata access.
@@
 % Non POSIX IO
-High performance computing has unique requirements for file systems ({\it
-e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
-make relaxing POSIX IO sensible.  BatchFS assumes the application coordinates
+High performance computing has unique requirements for file systems and
+well-defined workloads that make relaxing POSIX IO sensible.
+BatchFS~\cite{zheng:pdsw2014-batchfs} assumes the application coordinates
 accesses to the namespace, so the clients can batch local operations and merge
-with a global namespace image lazily. Similarly, DeltaFS eliminates RPC
-traffic using subtree snapshots for non-conflicting workloads and middleware
-for conflicting workloads. MarFS gives users the ability to lock
-``project directories" and allocate GPFS clusters for demanding metadata
-workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
-tier; applications lock the namespace so that metadata can be accessed more quickly.
-Unfortunately, decoupling the namespace has costs: (1) merging metadata state
-back into the global namespace is slow; (2) failures are local to the failing
-node; and (3) the systems are not backwards compatible.
+with a global namespace image lazily. Similarly,
+DeltaFS~\cite{zheng:pdsw2015-deltafs} eliminates RPC traffic using subtree
+snapshots for non-conflicting workloads and middleware for conflicting
+workloads. MarFS~\cite{grider:pdsw2015-marfs} gives users the ability to lock
+``project directories" and allocate clusters to demanding metadata
+workloads.  Unfortunately, decoupling the namespace has costs: (1) merging
+metadata state back into the global namespace is slow; (2) failures are local
+to the failing node; and (3) the systems are not backwards compatible.
 
@@
 For (1), state-of-the-art systems manage consistency in non-traditional ways:
-IndexFS maintains the global namespace but blocks operations from other clients
+IndexFS~\cite{ren:sc2014-indexfs} maintains the global namespace but blocks operations from other clients
 until the first client drops the lease, BatchFS does operations on a snapshot
@@
 on-disk structures (for BatchFS and DeltaFS this is the SSTables used in
-TableFS) can be recovered. In other words, the clients have state that cannot
+TableFS~\cite{ren:atc2013-tablefs}) can be recovered. In other words, the clients have state that cannot
 be recovered if the node stays failed and any progress will be lost. This

### 2018-02-22 paper/sections/3_posix_overheads.tex
paper: camera ready template and squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \texttt{untar}) incur the highest disk, network, and CPU utilization because of
-consistency/durability demands.}\label{fig:overhead-creates}
+consistency/durability demands.\vspace{-3ex}}\label{fig:overhead-creates}
 \end{figure}
@@
 (c) show that} when another client interferes, capabilities are revoked and
-metadata servers do more work.} \label{fig:overhead}
+metadata servers do more work.\vspace{-3ex}} \label{fig:overhead}
 \end{figure*}
@@
 \textbf{CephFS Design}: A journal of metadata updates that streams into the
-resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to be large (on
+resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} 
+the metadata journal is designed to be large (on
 the order of MBs) which ensures (1) sequential writes into the object store and

### 2018-02-22 paper/sections/4_methodology.tex
paper: camera ready template and squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\begin{figure}[tb]
+\begin{figure*}[tb]
   \centering
-%\includegraphics[width=50mm]{figures/fig-decouple.png}
-\includegraphics[width=1.0\linewidth]{figures/fig-decouple.png}
-\caption{Illustration of the \textbf{mechanisms} used by applications to build
-consistency/durability semantics. Descriptions are provided by the
+  \begin{subfigure}[b]{.32\linewidth}
+    \centering
+    \includegraphics[width=0.7\linewidth]{figures/fig-decouple0.png}
+    \caption{Cudele \textbf{mechanisms}
+    }\label{fig:decouple}
+  \end{subfigure}
+  \begin{subfigure}[b]{.32\linewidth}
+    \centering
+    \scalebox{0.7}{
+    \begin{tabular}{ l | l | l | l }
+      C \(\rightarrow\) &&& \\
+      D \(\downarrow\)  & invisible             & weak                  & strong  \\\hline
+      none              & append client journal & append client journal & RPCs    \\
+                        &                       & +volatile apply       &         \\
+                        &                       &                       &         \\\hdashline
+      local             & append client journal & append client journal & RPCs    \\
+                        & +local persist        & +local persist        & +local  \\
+                        &                       & +volatile apply       & persist \\
+                        &                       &                       &         \\\hdashline
+      global            & append client journal & append client journal & RPCs    \\
+                        & +global persist       & +global persist       & +stream \\
+                        &                       & +volatile apply       &         \\
+    \end{tabular}}
+    \caption{building guarantees from mechanisms
+    }\label{table:spectrum}
+  \end{subfigure}
+  \begin{subfigure}[b]{.32\linewidth}
+    \includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
+    \caption{
+    [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
+    \textbf{mechanism} performance
+    }\label{fig:composable-mechanisms}
+  \end{subfigure}
+  \caption{Cudele mechanism performance.
+    (a) is an illustration of the \textbf{mechanisms} used by applications to build
+    semantics. Descriptions are provided by the
     underlined words in Section~\S\ref{sec:the-cudelesfs-mechanisms}.
-}\label{fig:decouple}
-\end{figure}
+    (b) shows how users can explore the consistency (C) and
+    durability (D) spectrum by composing Cudele mechanisms.
+    \newcomment{(c) shows the overhead of processing 100K create events for each mechanism in
+    \oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
+    normalized to the runtime of writing events to client memory. The far right
+    graph shows the cost of building real world systems.}
+    \oldcomment{Performance of each mechanism (left) and building the
+    consistency/durability semantics of real-world systems (right) for 100K files
+    creates from a single client. Results are normalized to the runtime of writing
+    events to the client's in-memory journal.}\vspace{-3ex}}
+\end{figure*}
 
@@
 
-\begin{table}[t]
-\begin{center}
-\begin{tabular}{ l | l | l | l }
-  C \(\rightarrow\) &&& \\  
-  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
-  none                       & append client journal            & append client journal          & RPCs    \\
-                             &                   & +volatile apply &         \\\hdashline
-  local                      & append client journal            & append client journal          & RPCs    \\
-                             & +local persist    & +local persist  & +local  \\
-                             &                   & +volatile apply &  persist\\\hdashline
-  global                     & append client journal            & append client journal          & RPCs    \\
-                             & +global persist   & +global persist & +stream \\
-                             &                   & +volatile apply &         \\
-\end{tabular}
-
-\caption{Users can explore the consistency (C) and
-durability (D) spectrum by composing Cudele mechanisms. 
-\label{table:spectrum}}
-\end{center}
-\end{table}
-
 % describe table

### 2018-02-22 paper/sections/6_results.tex
paper: final revision for round 2
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: camera ready template and squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 system scale further when subtrees are coupled to the correct type of
-application}. \newcomment{These techniques have already proven to be effective
-and scalable in systems that specialize an entire namespace according to a
-single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS).}
+application}.
+%\newcomment{These techniques have already proven to be effective
+%and scalable in systems that specialize an entire namespace according to a
+%single optimization strategy ({\it e.g.}, Lustre, BatchFS, DeltaFS).}
 %, so our
@@
 reproducible, this paper adheres to The Popper
-Convention~\cite{jimenez_popper_2016} so experiments can be examined in more
+Convention~\cite{jimenez:ipdpsw17-popper} so experiments can be examined in more
 detail, or even re-run, by visiting the \texttt{[source]} link next to each
@@
 \label{sec:microbenchmarks}
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
-\caption{
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
-\newcomment{Overhead of processing 100K create events for each mechanism in
-\oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
-normalized to the runtime of writing events to client memory. The far right
-graph shows the overhead of building semantics of real world systems.}
-\oldcomment{Performance of each mechanism (left) and building the
-consistency/durability semantics of real-world systems (right) for 100K files
-creates from a single client. Results are normalized to the runtime of writing
-events to the client's in-memory journal.}\label{fig:composable-mechanisms}}
-\end{figure}
+%\begin{figure}[tb]
+%\centering
+%\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
+%\caption{
+%[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/experiments/cudele-mechanisms/visualize/viz.ipynb}{source}]
+%\newcomment{Overhead of processing 100K create events for each mechanism in
+%\oldcomment{Table~\ref{table:mechanisms}}\newcomment{Figure~\ref{fig:decouple}},
+%normalized to the runtime of writing events to client memory. The far right
+%graph shows the overhead of building semantics of real world systems.}
+%\oldcomment{Performance of each mechanism (left) and building the
+%consistency/durability semantics of real-world systems (right) for 100K files
+%creates from a single client. Results are normalized to the runtime of writing
+%events to the client's in-memory journal.}\vspace{-3ex}\label{fig:composable-mechanisms}}
+%\end{figure}
 
@@
 inflection point is the trade-off of frequent updates vs. larger journal
-files.}
+files.\vspace{-3ex}}
 \label{fig:use-cases}
@@
 to assign work units to workers and the performance is proportional to the
-open/create throughput of the underlying file system~\cite{xiao:socc15-shardfs,
-shi:vldb15-spark, shvachko:login2012-hdfs-scalability}; Big Data Benchmark jobs
+open/create throughput of the underlying file system~\cite{shi:vldb15-spark}; Big Data Benchmark jobs
 examined in~\cite{chaimov:hpdc16-spark} have on the order of 15,000 file opens
@@
 only pauses to fork off a background process, which is expensive as the address
-space needs to be copied. The alternative is to pause the client completely and
-write the update to disk but since this implementation is limited by the speed
-of the disk, we choose the memory-to-memory copy of the fork approach.
+space needs to be copied.  As expected, syncing namespace updates too
+frequently has the highest overhead (up to 9\% overhead if done every second).
+The optimal sync interval for performance is 10 seconds, which only incurs 2\%
+overhead, because larger intervals must write more updates to disk and network.
+For the 25 second interval, the client only pauses 3-4 times but each sync
+writes about 278 thousand updates at once, which is a journal of size 678MB.
 
-As expected, syncing namespace updates too frequently has the highest overhead
-(up to 9\% overhead if done every second). The optimal sync interval for
-performance is 10 seconds, which only incurs 2\% overhead, because larger
-intervals must write more updates to disk and network. For the 25 second
-interval, the client only pauses 3-4 times but each sync writes about 278
-thousand updates at once, which is a journal of size 678MB.
+%The alternative is to pause the client completely and
+%write the update to disk but since this implementation is limited by the speed
+%of the disk, we choose the memory-to-memory copy of the fork approach.
 

### 2018-02-22 paper/sections/7_conclusion.tex
paper: camera ready template and squeeze to 10 pgs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-01-08 paper/sections/7_conclusion.tex
@@
 adjust these semantics over {\it time and space}, where subtrees can change
-without ever moving the data they reference.
+without moving data they reference.`
 
@@
 %applications or move data between these systems.  
-Cudele prompts many avenues for future work.  First is to co-locate HPC
-workflows with real highly parallel runtimes from the cloud in the same
-namespace. This setup shows how Cudele reasonably incorporates both programming
-models (client-driven parallelism and user-defined workflows) at the same time
-and should show large performance gains.  Second is dynamically changing
-semantics of a subtree from stronger to weaker guarantees (or vice versa). This
-reduces data movement across storage cluster and file system boundaries so
-the results of a Hadoop job do not need to be migrated into CephFS for
-other processing; instead the \oldcomment{user}\newcomment{administrator} can
-change the semantics of the HDFS subtree into a CephFS subtree, which may cause
-metadata/data movement to ensure strong consistency. Third is embeddable
-policies, where child subtrees have specialized features but still maintain
-guarantees of their parent subtrees.  For example, a RAMDisk subtree is POSIX
-IO-compliant but relaxes durability constraints, so it can reside under a POSIX
-IO subtree alongside a globally durable subtree.
+\textbf{Future Work}: First is to co-locate HPC workflows with real highly
+parallel runtimes from the cloud in the same namespace. This setup would show
+how Cudele reasonably incorporates both programming models at the same time.
+Second is dynamically changing semantics of a subtree from stronger to weaker
+guarantees. This reduces data movement across storage cluster and file system
+boundaries so the results of a Hadoop job do not need to be migrated into
+CephFS for other processing; instead the
+\oldcomment{user}\newcomment{administrator} can change the semantics of the
+HDFS subtree into a CephFS subtree. Third is embeddable policies, where child
+subtrees have specialized features but still maintain guarantees of the
+parent subtree.  For example, a RAMDisk subtree is POSIX IO-compliant but
+relaxes durability, so it can reside under a POSIX IO subtree
+alongside a globally durable subtree.\vspace{-1ex}
 

### 2018-02-28 paper/sections/2_related_work.tex
paper: typos and acknowledgements
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-10-23 paper/sections/2_related_work.tex
@@
 % Non POSIX IO
-High performance computing has unique requirements for file systems ({\it
-e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
-make relaxing POSIX IO sensible.  BatchFS assumes the application coordinates
-accesses to the namespace, so the clients can batch local operations and merge
-with a global namespace image lazily. Similarly, DeltaFS eliminates RPC
-traffic using subtree snapshots for non-conflicting workloads and middleware
-for conflicting workloads. MarFS gives users the ability to lock
-``project directories" and allocate GPFS clusters for demanding metadata
-workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
-tier; applications lock the namespace so that metadata can be accessed more quickly.
-Unfortunately, decoupling the namespace has costs: (1) merging metadata state
-back into the global namespace is slow; (2) failures are local to the failing
-node; and (3) the systems are not backwards compatible.
+High performance computing has unique requirements for file systems and
+well-defined workloads that make relaxing POSIX IO sensible.  BatchFS assumes
+the application coordinates accesses to the namespace, so the clients can batch
+local operations and merge with a global namespace image lazily. Similarly,
+DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
+workloads and middleware for conflicting workloads.
+MarFS~\cite{grider:pdsw2015-marfs} gives users the ability to lock ``project
+directories" and allocate clusters for demanding metadata workloads.
+TwoTiers~\cite{bent:slides-twotiers} eliminates high-latencies by storing
+metadata in a flash tier; applications lock the namespace so that metadata can
+be accessed more quickly.  Unfortunately, decoupling the namespace has costs:
+(1) merging metadata state back into the global namespace is slow; (2) failures
+are local to the failing node; and (3) the systems are not backwards
+compatible.
 

### 2018-02-28 paper/sections/6_results.tex
paper: typos and acknowledgements
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-01-09 paper/sections/6_results.tex
@@
 open/create throughput of the underlying file system~\cite{xiao:socc15-shardfs,
-shi:vldb15-spark, shvachko:login2012-hdfs-scalability}; Big Data Benchmark jobs
-examined in~\cite{chaimov:hpdc16-spark} have on the order of 15,000 file opens
-or creates just to start a single Spark query and the Lustre system they tested
-on did not handle creates well, showing up to a \(24\times\) slowdown compared
-to other metadata operations. Common approaches to solve these types of
+shi:vldb15-spark}; Big Data Benchmark jobs examined
+in~\cite{chaimov:hpdc16-spark} have on the order of 15,000 file opens or
+creates just to start a single Spark query and the Lustre system they tested on
+did not handle creates well, showing up to a \(24\times\) slowdown compared to
+other metadata operations. Common approaches to solve these types of
 bottlenecks is to change the application behavior or to design a new file

### 2018-02-28 paper/sections/3_posix_overheads.tex
paper: typos, reviewer comments, and  deltafs clarifications
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2018-01-09 paper/sections/3_posix_overheads.tex
@@
 
-\textbf{Comparison to decoupled namespaces}: For BatchFS, if a client fails
-when it is writing to the local log-structured merge tree (implemented as an
-SSTable~\cite{ren:atc2013-tablefs}) then unwritten metadata operations are
-lost. For DeltaFS, if the client fails then, on restart, the computation does
-the work again -- since the snapshots of the namespace are never globally
-consistent and there is no ground truth.  On the server side, BatchFS and
-DeltaFS use IndexFS~\cite{ren:sc2014-indexfs}. IndexFS writes metadata to
-SSTables, which initially reside in memory but are later flushed to the
-underlying distributed file system.
+\textbf{Comparison to decoupled namespaces}: For BatchFS and DeltaFS, if the
+client fails and stays down then computation must be done again because the
+metadata updates were never globally consistent.  Both systems write metadata
+to log-structured merge trees (implemented as
+SSTables~\cite{ren:atc2013-tablefs}) so updates initially reside in memory but
+are later flushed to the underlying file system. As a result, after a client
+crash, updates that have not been written safely to disk may be lost.  On the
+server side, BatchFS and DeltaFS use IndexFS~\cite{ren:sc2014-indexfs}.
+IndexFS writes metadata to SSTables as well but writes to lower tiers are
+globally persisted in a distributed file system.
 

### 2018-02-28 paper/sections/6_results.tex
paper: typos, reviewer comments, and  deltafs clarifications
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 generally applicable to different hardware. We use a CloudLab cluster of 34
-nodes connected with 10Gbit ethernet, each with 16 2.4 GHz CPUs, 128GB RAM, and
+nodes connected with 10Gbit ethernet, each with 2 2.4 GHz CPUs (16 total cores), 128GB RAM, and
 400GB SSDs.  Each node uses Ubuntu 14.04 and we develop on Ceph's Jewel
@@
 bottlenecks is to change the application behavior or to design a new file
-system, like BatchFS or DeltaFS, that uses one set of metadata optimizations
+system that uses one set of metadata optimizations
 for the entire namespace.} \oldcomment{ abstraction to exchange work units to
@@
 because ``decoupled: merge + create" uses a relaxed form of consistency and
-leverages bulk updates just like DeltaFS~\cite{zheng:pdsw2015-deltafs}.
+leverages bulk updates just like DeltaFS.
 Decoupled namespaces (1)} place no restrictions on the validity of metadata
